---
title: "Build Your Cohort"
output: html_document
author: 'Dabriel Zimmerman'
date: "`r Sys.Date()`"
---

Script written by Dabriel and updated by Margaret to get data from 2025-03 Clinical Scans release.

Prelude to this is to make sure pandas and numpy are installed. Run `pip install pandas, numpy` in a terminal
```{r setup r, include=FALSE}

# install.packages("librarian")

librarian::shelf(ggplot2, readr, dplyr, data.table, MetBrewer, reticulate, gtsummary, purrr, stringr, skimr)

basic <-  theme(axis.line = element_line(colour = "black"),
                # panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                panel.border = element_blank(),
                plot.title = element_text(size = 15, 
                                          face = 'bold'),
                axis.title = element_text(size = 10, 
                                            face = 'bold'),
                panel.background = element_blank(),
                text = element_text(size = 14),
                legend.title = element_text(size = 10, 
                                            face = 'bold'),
                axis.text.x = element_text(size = 10),
                legend.text = element_text(size = 10))
```

```{python setup py, include=False}
import re, os, os.path as op, glob, pandas as pd, numpy as np
import warnings
warnings.simplefilter(action='ignore', category=[FutureWarning,DeprecationWarning])

df = pd.read_csv("/mnt/isilon/bgdlab_processing/releases_clinical/2025_03_release/BIDS/participants.tsv", sep = "\t")
tot_subs = df.subject_id.nunique()
tot_ses = df.session_id.nunique()
print(f"This release contains {tot_subs} subjects and {tot_ses} sessions.")
```

# Create a cohort with filters

Getting all SLIP participants

## Set cohort filters

-   Age range is inclusive
-   Year of scan is inclusive

```{python set filters}
long = False          # Get subjects with longitudinal data
slip_only = True      # Specify false to include sessions with grades < 1. Default is sessions matching SLIP criteria
age_min = 0           # Specify min age in days (365.25 * age years).
age_max = None        # Specify max age in days (365.25 * age years), or leave it as is (25 years old), or replace with None
sex = None            # Specify 'Male' or 'Female'
project = None # Specify a project listed in the projects.json
standardMPR = False   # Specify True to only get sessions where there is a CHOP standard 3T MPR (used in SLIP models)
highResT2 = False     # Specify True to only get sessions where there is a high res 3T T2w
hasMPR = False        # Specify True to only get sessions where there is an MPRAGE
hasT2 = False         # Specify True to only get sessions that have a T2w
gestAge = False       # Specify True to get sessions where there is gestational age
manufacturer = None   # Specify manufacturer 'Siemens' or 'GE'
mfs = None            # Specify magnetic field strength '3T'or 1.5T'
dsn = None            # Specify device serial number(s) using options listed in scanners.json. Single value as string or list of strings
hasBW = None          # Specify True to only get sessions where birth weight is listed
hasBL = None          # Specify True to only get sessions where birth length is listed
rsample = 0           # Replace with an integer to get an n random sample of a larger cohort
sample_group = None   # Specify a group to sample with a column name from the data dictionary
yos_min = 2000        # Specify with a year minimum for session year
yos_max = None        # Specify with a year maximum for session year
   
```

### Set groups

```{python set groups}
set_groups = False                   # Specify True to create groups based on a study variable
# grouping_var = 'year_of_scan'       # Specify study variable present in the data dictionary
# group1_val = '<=2006'               # Specify the filter for group 1
# group2_val = '>2006'               # Specify the filter for group 2
# group1_name = 'old'                 # Specify the name for group 1
# group2_name = 'new'                 # Specify the name for group 2
```

## Filter the release to get your cohort data

```{python filtering}
filters = [
    None if not slip_only else f"(df['useable']==True",
    None if not age_min > 0 else f"&(df['age_at_scan']>={age_min}",
    None if not age_max or age_max is False else f"&(df['age_at_scan']<={age_max}",
    None if not sex or sex is False else f"&(df['sex']=={sex}",
    None if not project or project is False else f"&(df['project'].str.contains('{project}')==True",
    None if not standardMPR or standardMPR is False else f"&(df['standardMPR']=={standardMPR}",
    None if not highResT2 or highResT2 is False else f"&(df['highResT2']=={highResT2}",
    None if not hasMPR or hasMPR is False else f"&(df['hasMPR']=={hasMPR}",
    None if not hasT2 or hasT2 is False else f"&(df['hasT2']=={hasT2}",
    None if not gestAge or gestAge is False else f"&(df['gestational_age'].notna()",
    None if not manufacturer or manufacturer is False else f"&(df['manufacturer']=={manufacturer}",
    None if not mfs or mfs is False else f"&(df['magnetic_field_strength']=={mfs}",
    None if not dsn or dsn is False else f"&(df['device_serial_number']=={dsn}" if isinstance(dsn, str) else f"&(df['device_serial_number'].isin{dsn}",
    None if not hasBW or hasBW is False else f"&(df['birth_weight_kg'].notna()",
    None if not hasBL or hasBL is False else f"&(df['birth_length_cm'].notna()",
    None if not yos_min or yos_min is False else f"&(df['year_of_scan']>=yos_min",
    None if not yos_max or yos_max is False else f"&(df['year_of_scan']<=yos_max"
  ]
  
filterStr = ")".join(filter(bool, filters)) + ")"
filterStr = "filtered = df[" + re.sub(r'^&','',filterStr) + "]"
    
print(filterStr)
exec(filterStr)
myDf = filtered.copy()

# Add session counter
myDf['session_count'] = myDf.groupby('subject_id', sort=False)['subject_id'].transform('size')
    
if long is True:
  myDf = myDf[myDf['session_count']>1]
    
n_subs = myDf.subject_id.nunique()
n_ses = myDf.session_id.nunique()
    
print(f"After filtering data, {n_subs} subjects with {n_ses} sessions remain.")
```

```{r set groups, echo=FALSE}
has_groups <- py$set_groups

if (has_groups == TRUE){
  myDf <- py$myDf |>
  mutate(group = case_when(eval(parse(text=print(sprintf("%s %s", py$grouping_var,
                              py$group1_val)))) ~ py$group1_name,
                           eval(parse(text=print(sprintf("%s %s", py$grouping_var,
                              py$group2_val)))) ~ py$group2_name))
} else {
  myDf <- py$myDf
}
```
At this point you now have a dataframe **myDf** containing the information for subjects with data matching your criteria. It includes (also see the participants.json for full data dictionary):

## Grading relevant columns

-   `avg_grade` average grade a session was given by the graders listed in `graders`
-   `graded` Boolean value representing whether a session has been graded by at least 2 people (or 1 person and the NLP models) and the session was not given a 0 by anyone
-   `useable` Boolean value representing if a session can be considered **SLIP**
-   `project` Lists all projects a session belongs to. Multiple projects can be listed where a '; ' is used to separate them. If project was a specified filter then it will only contain the specified project.

## Metrics

-   `weight_at_enc` Weight in kg of subject at encounter close to or on the same date as the MRI scan
-   `height_at_enc` Height in cm of subject at encounter close to or on the same date as the MRI scan
-   `hc_at_enc` Head circumference in cm of subject at encounter close to or on the same date as the MRI scan
-   `scan_hc_measure_age_diff` Difference in age between scan visit and visit where metrics (height, weight, head circumference) were taken.

## Some Quick Stats/Viz
```{r demo table}
myDf_edited <- myDf |>
    mutate(age_at_scan = (age_at_scan/365.25)*12,
           hasMPR = ifelse(hasMPR==TRUE,TRUE,FALSE),
           isLong = ifelse(session_count > 1, 
                            TRUE, 
                            FALSE))

if (has_groups == TRUE){
  demo_tbl <- myDf_edited |>
    select(group, age_at_scan, 
         sex, race, ethnicity,
         hasMPR, isLong) |>
    tbl_summary(
      by = group,
      include = c(age_at_scan, sex, 
                race, ethnicity,
                hasMPR, isLong),
      label = list(age_at_scan = "Age at MRI Scan (months)",
                 sex = "Sex",
                 race = "Race",
                 ethnicity = "Ethnicity",
                 hasMPR = "Has MPRAGE",
                 isLong = "Longtidunal"),
      statistic = list(age_at_scan ~ "{median} ({p25}, {p75})",
                     sex ~ "{n}",
                     race ~ "{n}",
                     ethnicity ~ "{n}",
                     hasMPR ~ "{n}",
                     isLong ~ "{n}")) |>
    add_overall() |>
    modify_header(label ~ "**Characteristic**") |>
    modify_footnote_header("Median (IQR) or n", 
                         columns = all_stat_cols()) |>
    modify_caption("**Cohort Characteristics**")
} else {
  
  demo_tbl <- myDf_edited |>
    select(age_at_scan, 
         sex, race, ethnicity,
         hasMPR, isLong) |>
    tbl_summary(
      include = c(age_at_scan,sex, 
                race, ethnicity,
                hasMPR, isLong),
      label = list(age_at_scan = "Age at MRI Scan (months)",
                 sex = "Sex",
                 race = "Race",
                 ethnicity = "Ethnicity",
                 hasMPR = "Has MPRAGE",
                 isLong = "Longtidunal"),
      statistic = list(age_at_scan ~ "{median} ({p25}, {p75})",
                     sex ~ "{n}",
                     race ~ "{n}",
                     ethnicity ~ "{n}",
                     hasMPR ~ "{n}",
                     isLong ~ "{n}")) |>
    modify_header(label ~ "**Characteristic**") |>
    modify_footnote_header("Median (IQR) or n", 
                         columns = all_stat_cols()) |>
    modify_caption("**Cohort Characteristics**")
}

demo_tbl

```

```{r age/sex plot}
myDf |>
  ggplot(aes(x=age_at_scan)) +
    geom_histogram(aes(fill = sex)) +
    ggtitle("Distribution of Age at Scan By Sex") +
    basic + xlab("Age (Days)") + ylab("Sessions") + labs(fill = 'Sex')
```
```{r year of scan plot}
myDf |>
  ggplot(aes(x=year_of_scan)) +
    geom_histogram(aes(fill = sex)) +
    ggtitle("Distribution of Year at Scan By Sex") +
    basic + xlab("Year") + ylab("Sessions") + labs(fill = 'Sex')
```

## Add IDPs


```{r}
deriv_aparc <- fread("/mnt/isilon/bgdlab_processing/releases_clinical/2025_03_release/derivatives/recon-all-clinical/fs7.4.1-clinical-reconall-aparc.csv", fill=TRUE)
#names(deriv_aparc)

deriv_global <- fread("/mnt/isilon/bgdlab_processing/releases_clinical/2025_03_release/derivatives/recon-all-clinical/fs7.4.1-clinical-reconall-global.csv", fill=TRUE)
#names(deriv_global)

deriv_qc <- fread("/mnt/isilon/bgdlab_processing/releases_clinical/2025_03_release/derivatives/recon-all-clinical/fs7.4.1-clinical-reconall-qc.csv", fill=TRUE)
#names(deriv_qc)

deriv_ss <- fread("/mnt/isilon/bgdlab_processing/releases_clinical/2025_03_release/derivatives/recon-all-clinical/fs7.4.1-clinical-reconall-ss.csv", fill=TRUE)
#names(deriv_ss)
```

remove curvature measures and merge

```{r}
deriv_aparc_filt <- deriv_aparc %>%
  dplyr::select(-matches("^(MeanCurv|NumVert|GausCurv|FoldInd|CurvInd|ThickStd)")) %>%
  mutate(session_id = str_extract(scan_id, "ses-.*?ageDays"),
         scan_seq = str_extract(scan_id, "[^_]+$")) %>%
  mutate(sub_ses = paste(subject_id, session_id, "_"))
nrow(deriv_aparc_filt) == n_unique(deriv_aparc_filt$sub_ses)

deriv_ss_filt <- deriv_ss %>%
  dplyr::select(-starts_with("ctx-"))

#merge
deriv_all <- reduce(list(deriv_aparc_filt, deriv_global, deriv_qc, deriv_ss_filt), full_join)  %>%
  rename_with(~ str_replace(.x, " ", "_")) %>%
 mutate(QC.RCA = do.call(pmin, c(across(starts_with("QC")), na.rm = TRUE))) #using Lena's method for combining

table(deriv_all$scan_seq)
#skim(deriv_all)

#filter to match df
sample_full <- inner_join(myDf, deriv_all)
#stopifnot(nrow(myDf) == nrow(sample))
skim(sample_full)
nrow(sample_full) > n_unique(sample_full$session_id)

sample_full %>%
  group_by(session_id) %>%
  summarise(n_T1 = sum(scan_seq=="T1w"))

sample_full %>%
  filter(session_id == "ses-256387570729procId000249ageDays")

sample_full %>%
  ggplot() +
  geom_histogram(aes(x=QC.RCA))
```
### Infants
For infants < 1 year, take median IDP value across sequences w/in session

```{r}
lh_idps <- names(sample_full)[grep(".lh.|left_", names(sample_full))]
rh_idps <- names(sample_full)[grep(".rh.|right_", names(sample_full))]
vent <- names(sample_full)[grep("ventricle", names(sample_full))]
glob <- c("eTIV", "BrainSegVol", "BrainSegVolNotVent", "BrainSegVolNotVentSurf", "CerebralWhiteMatterVol", "CortexVol", "MaskVol", "SubCortGrayVol", "SupraTentorialVol", "SupraTentorialVolNotVent", "SupraTentorialVolNotVentVox", "TotalGrayVol", "VentricleChoroidVol", "lhCerebralWhiteMatterVol", "lhCortexVol", "rhCerebralWhiteMatterVol", "rhCortexVol", "total_intracranial", "brain-stem", "csf", "TCV", "cGMV", "sGMV", "CerebellumVolume", "WMV", "Ventricles")
idp_cols <- c(lh_idps, rh_idps, glob, vent)

setdiff(names(sample_full), idp_cols)

samp_inf <- sample_full %>%
  dplyr::filter(age_at_scan <= 365.25 & QC.RCA > 0.65) %>%
  group_by(subject_id, session_id) %>%
  summarise(participant_id = unique(participant_id),
            subject_id = unique(subject_id),
            session_id = unique(session_id),
            sex = unique(sex),
            sub_ses = unique(sub_ses),
            device_serial_number = unique(device_serial_number),
            age_at_scan = unique(age_at_scan),
            adjusted_age_in_days = unique(adjusted_age_in_days),
            across(all_of(idp_cols), \(x) median(x, na.rm = TRUE)),
            across(all_of(starts_with("QC")), \(x) min(x, na.rm = TRUE)),
            .groups = "drop")
stopifnot(n_unique(samp_inf$session_id) == nrow(samp_inf))
skim(samp_inf)
```

### Adults
For all subjects > 1 year, take IDP value from T1w MPRAGE

```{r}
samp_mpr <- sample_full %>%
  dplyr::filter(age_at_scan > 365.25 & scan_seq == "T1w") %>%
  dplyr::filter(str_detect(scan_id, 'MPRAGE')) %>%
  dplyr::filter(QC.RCA > 0.65) #filter now to be consistent
n_unique(samp_mpr$session_id) == nrow(samp_mpr)
  
samp_no_mpr <- sample_full %>%
  dplyr::filter(age_at_scan > 365.25 & scan_seq == "T1w" & QC.RCA > 0.65) %>%
  dplyr::filter(!str_detect(scan_id, 'MPRAGE')) %>%
  dplyr::filter(!(session_id %in% samp_mpr$session_id)) %>%
  group_by(subject_id, session_id) %>%
  summarise(participant_id = unique(participant_id),
            subject_id = unique(subject_id),
            session_id = unique(session_id),
            sex = unique(sex),
            sub_ses = unique(sub_ses),
            device_serial_number = unique(device_serial_number),
            age_at_scan = unique(age_at_scan),
            adjusted_age_in_days = unique(adjusted_age_in_days),
            across(all_of(idp_cols), \(x) median(x, na.rm = TRUE)),
            across(all_of(starts_with("QC")), \(x) min(x, na.rm = TRUE)),
            .groups = "drop")
stopifnot(n_unique(samp_no_mpr$session_id) == nrow(samp_no_mpr))
```

```{r}
samp_final <- reduce(list(samp_inf, samp_mpr, samp_no_mpr), full_join)
```
```{r}
ggplot(samp_final) +
  geom_histogram(aes(x=(age_at_scan/365.25), fill=sex))
```

```{r write csv}
output <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/data/slip_202503_release.csv"
write_csv(samp_final, output)
```