---
title: "CV Model Testing"
output: html_document
date: "2025-01-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, splitstackshape, gamlss, gamlssTools)
```

Testing CV models

##Split Samples

```{r}
df <- fread("./data/SBC_LBCC_CN_cleaned.csv")
```

Splitting data into 50-50 train and testing, stratified by site and sex. Using this [reference](https://stackoverflow.com/questions/74573270/stratified-data-splitting-in-r).

```{r}
table(df$study_site, df$sexMale)
```

going to drop small sites (<10 subjects per sex) for now, but alternatives could be grouping small sitest together? (go back and include sites with <5 total subjs)

```{r}
valid_sites <- df %>%
  mutate(sex=as.factor(sexMale)) %>%
  group_by(study_site, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(study_site) %>%
  summarise(
    valid = all(c("0", "1") %in% sex) && all(n[match(c("0", "1"), sex, nomatch = 0)] >= 10),
    .groups = "drop"
  ) %>%
  filter(valid) %>%
  pull(study_site)

df_filt <- df %>%
  filter(study_site %in% valid_sites)

nrow(df) - nrow(df_filt)

ggplot(df) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))
ggplot(df_filt) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))

#table(df_filt$study_site, df_filt$sexMale)
```

Still have almost 100k subjects, proceeding to divide into A and B samples

```{r}
#add sex factor var
df_filt <- df_filt %>% mutate(sex=as.factor(sexMale))

split_samples <- splitstackshape::stratified(df_filt, group=c("study_site", "sex"), size=0.5, bothSets = TRUE)

samp_A <- split_samples$SAMP1
samp_B <- split_samples$SAMP2

ggplot(samp_A) +
  geom_histogram(aes(x=age_days, fill=sex))
ggplot(samp_B) +
  geom_histogram(aes(x=age_days, fill=sex))
```

##Fit Model
Fit a test model on one brain phenotype using `samp_A` as training data.
Going to fit site as a random effect, but alternative would be ComBatLS (separately w/in each sample). Also ignoring euler for now

```{r}
mod_samp_A <- samp_A %>% 
  dplyr::select(INDEX.ID, lh.SA.middletemporal, sexMale, sexMale_x_logAge, logAge_days, study, study_site, fs_version) %>%
  mutate(study_site = as.factor(study_site)) %>%
  na.omit()
```

Fit full model

```{r}
m1 <- gamlss(lh.SA.middletemporal ~ sexMale + pb(sexMale_x_logAge, control=pb.control(order=5)) + pb(logAge_days, control=pb.control(order=5)) + random(study_site) + fs_version,
             sigma.formula = ~ sexMale + pb(sexMale_x_logAge, control=pb.control(order=5)) + pb(logAge_days, control=pb.control(order=5)) + random(study_site) + fs_version,
             family=GG(),
             data=mod_samp_A,
             control=gamlss.control(n.cyc=200))
```

```{r}
make_centile_fan(m1, mod_samp_A, "logAge_days", color_var="sexMale", x_axis="log_lifespan_fetal")
```

testing higher penalty
```{r}
m2 <- gamlss(lh.SA.middletemporal ~ sexMale + pb(sexMale_x_logAge, control=pb.control(order=4)) + pb(logAge_days, control=pb.control(order=4)) + random(study_site) + fs_version,
             sigma.formula = ~ sexMale + pb(sexMale_x_logAge, control=pb.control(order=4)) + pb(logAge_days, control=pb.control(order=4)) + random(study_site) + fs_version,
             family=GG(),
             data=mod_samp_A,
             control=gamlss.control(n.cyc=200))
```

```{r}
make_centile_fan(m2, mod_samp_A, "logAge_days", color_var="sexMale", x_axis="log_lifespan_fetal")
```

```{r}
m3 <- gamlss(lh.SA.middletemporal ~ sexMale + pb(sexMale_x_logAge, control=pb.control(order=3)) + pb(logAge_days, control=pb.control(order=3)) + random(study_site) + fs_version,
             sigma.formula = ~ sexMale + pb(sexMale_x_logAge, control=pb.control(order=3)) + pb(logAge_days, control=pb.control(order=3)) + random(study_site) + fs_version,
             family=GG(),
             data=mod_samp_A,
             control=gamlss.control(n.cyc=200))
```

```{r}
make_centile_fan(m3, mod_samp_A, "logAge_days", color_var="sexMale", x_axis="log_lifespan_fetal")
```