---
title: "Format Figures"
output: html_document
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, cowplot, sf)

devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
#remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

#FIGURE 1: Age-by-Sex Effects

Sample A exemplar curve(s)

```{r}
df <- fread("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/data/cv_sample_B_dfs/sGMV_totalFALSE_logPhenoFALSE_logAgeTRUE.csv", stringsAsFactors = TRUE, na.strings = "")

test_mod <- readRDS("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_B_test/global_vols_totalFALSE_logAgeTRUE_pbmods/model_objs/sGMV_mu_pbAge_BCTo_test_full_mod.rds")
test_mod$call$data <- "df"
test_mod$call$family <- test_mod$family[[1]]

sim_df <- sim_data(df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
vars_of_interest <- c("sexMale_x_logAge", "logAge_days", "sexMale")
resid_terms <- setdiff(list_predictors(test_mod), vars_of_interest)

sGMV_fan <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="logAge_days",
                 color_var="sexMale",
                 get_peaks=TRUE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df,
                 remove_point_effect = resid_terms,
                 label_centiles = "legend",
                 x_axis="log_lifespan_fetal",
                 point_color_manual = c('0' = "#F4A15BFF", '1' = "#8CB3D1FF"),
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_blank()) +
    labs(title=paste("sGMV validation model"),
         x ="Age (years)",
         y="Subcortical Gray Matter Volume (mm3)") +
    guides(fill=FALSE) +
    geom_vline(xintercept=log(280, base=10))
```
```{r}
age_by_sex_plt <- readRDS("figs/ggobjs/agebysex_B.rds")

age_by_sex_plt_wrap <- plot_grid(
  age_by_sex_plt[[2]], 
  age_by_sex_plt[[3]],
  ncol = 1,
  rel_heights = c(5,1.05),
  align = "v",
  axis = "lr"
)
```

try to combine

```{r}
sGMV_fan <- plot_grid(NULL, sGMV_fan, NULL, ncol=1, rel_heights=c(1,6,1))

fig1 <- plot_grid(sGMV_fan,
                  age_by_sex_plt_wrap, 
                  age_by_sex_plt[[4]],
                  nrow=1, ncol=3, 
                  rel_widths=c(3,2,.4), 
                  labels=c("A", "B", ""))

ggsave("figs/fig1.png", fig1, width=11, height=8, dpi=600, bg = "white")
```



#FIGURE 2: Sex Effects Controlling for Total Size

Sample A exemplar curve(s)

```{r}
df <- fread("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/data/cv_sample_B_dfs/sGMV_totalTRUE_logPhenoFALSE_logAgeTRUE.csv", stringsAsFactors = TRUE, na.strings = "")

test_mod <- readRDS("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_B_test/global_vols_totalTRUE_logAgeTRUE_pbmods/model_objs/sGMV_fsmu_pbAge_totalboth_test_full_mod.rds")
test_mod$call$data <- "df"
test_mod$call$family <- test_mod$family[[1]]
list_predictors(test_mod)

sim_df1 <- sim_data(df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
vars_of_interest <- c("sexMale_x_logAge", "logAge_days", "sexMale", "TBV")
resid_terms <- setdiff(list_predictors(test_mod), vars_of_interest)

#manually set tbv to zero rather than mean
sim_df1$'1'$TBV <- 0
sim_df1$'0'$TBV <- 0

sGMV_fan1 <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="logAge_days",
                 color_var="sexMale",
                 get_peaks=TRUE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df1,
                 label_centiles = "legend",
                 show_points = FALSE,
                 x_axis="log_lifespan_fetal",
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  theme(legend.position= "right", plot.title = element_blank()) +
    labs(x ="Age (years)",
         y="sGMV (mm3)") +
    guides(fill=FALSE) +
    geom_vline(xintercept=log(280, base=10))

sGMV_legend <- cowplot::get_legend(sGMV_fan1, legend="right")
sGMV_fan1 <- sGMV_fan1 + theme(legend.position="none")
```

```{r}
sim_df2 <- sim_data(df, "TBV", factor_var="sexMale")
#manually set age to zero rather than mean
sim_df2$'1'$logAge_days <- 0
sim_df2$'0'$logAge_days <- 0
sim_df2$'1'$sexMale_x_logAge <- 0
sim_df2$'0'$sexMale_x_logAge <- 0

sGMV_fan2 <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="TBV",
                 color_var="sexMale",
                 get_peaks=FALSE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df2,
                 label_centiles = "legend",
                 show_points = FALSE,
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  scale_x_continuous(labels = scales::label_number(scale = 0.001, suffix = "k")) +
  theme(legend.position= "none", plot.title = element_blank()) +
    labs(x ="Total Brain Volume (mm3)",
         y="sGMV (mm3)") +
    guides(fill=FALSE)
```

```{r}
#predict centiles
desiredCentiles_reord <- sort(c(0.05, 0.25, 0.5, 0.75, 0.95), decreasing = TRUE)
centile_dfs <- centile_predict(gamlssModel = test_mod, 
                               sim_data_list = sim_df2, 
                               x_var = "TBV", 
                               desiredCentiles = desiredCentiles_reord,
                               df = df,
                               average_over = FALSE)
  
names(centile_dfs) <- sub("fanCentiles_", "", names(centile_dfs)) #drop prefix
merged_centile_df <- bind_rows(centile_dfs, .id = "sexMale")

long_centile_df <- merged_centile_df %>%
        tidyr::gather(id.vars, values, !any_of(c("sexMale", "TBV")))
  
map_thickness <- function(x){
    if (x == 0.5){
      return(1.75)
    } else if (x < 0.1 || x > 0.9){
      return (0.25)
    } else if (x < 0.25 || x > 0.75){
      return (0.5)
    } else {
      return(1)
    }
  }

centile_linewidth <- sapply(desiredCentiles_reord, map_thickness)
names(centile_linewidth) <- unique(long_centile_df$id.vars)

#resid out linear effects
fit <- lm(values ~ TBV, data = long_centile_df)
long_centile_df$value_resid <- residuals(fit)

sGMV_fan_resid <- ggplot() +
        geom_line(data = long_centile_df,
                mapping = aes(x = TBV, y = value_resid,
                group = interaction(id.vars, sexMale),
                color = sexMale,
                linewidth = id.vars)) +
      scale_linewidth_manual(values = centile_linewidth,
                             breaks = names(centile_linewidth),
                             guide = "legend",
                             name = "Percentile",
                             label = scales::percent(as.numeric(gsub("cent_|deriv_", 
                                                                   "", 
                                                                   unique(long_centile_df$id.vars))))) +
  scale_x_continuous(labels = scales::label_number(scale = 0.001, suffix = "k")) +
  theme_linedraw() +
  theme(legend.position="none") +
  labs(x="Total Brain Volume (mm3)", y="Residual sGMV (mm3)") +
  scale_color_manual(values=c('0' = "#c05f0d", '1' = "#38688D"))
```

combine
```{r}
sGMV_fan <- plot_grid(sGMV_fan1, 
                      plot_grid(sGMV_fan2, sGMV_fan_resid,
                      ncol=2, 
                      nrow=1,
                      rel_widths=c(1,1),
                      align = "h",
                      labels=c("B", "C"),
                      label_size = 12),
                      ncol=1,
                      nrow=2,
                      labels=c("A", ""),
                      label_size=12)
                      
#add legend
sGMV_fan <- plot_grid(sGMV_fan, sGMV_legend, ncol = 2, rel_widths = c(1, .2))
```

brain plots
```{r}
sex_plt <- readRDS("figs/ggobjs/full_plt_tot_sex_B.rds")
sex_plt[[1]] <- sex_plt[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sex_plt_wrap <- plot_grid(
  plot_grid(
  sex_plt[[1]],
  sex_plt[[2]], 
  sex_plt[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
), sex_plt[[4]],
ncol = 2,
rel_widths = c(1, .2))
```


sex dominance
```{r}
sex_dom <- readRDS("figs/ggobjs/cent_dom_B.rds")

sex_dom[[1]] <- sex_dom[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())
```

```{r}
sig_sex_dom <- readRDS("figs/ggobjs/sig_dom_B.rds")

sig_sex_dom[[1]] <- sig_sex_dom[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sex_dom_wrap <- plot_grid(
  #median
  plot_grid(
  sex_dom[[1]],
  sex_dom[[2]], 
  sex_dom[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
),
#sigma
plot_grid(
  sig_sex_dom[[1]],
  sig_sex_dom[[2]], 
  sig_sex_dom[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
),
ncol = 2,
labels=c("E", "F"),
label_size = 12,
align="vh")

#add legend
sex_dom_wrap2 <- plot_grid(sex_dom_wrap, sex_dom[[4]],
                           ncol = 2,
                           rel_widths = c(1, .2))
```
try to combine

```{r}
fig2 <- plot_grid(sGMV_fan,
                  NULL,
                  sex_plt_wrap,
                  NULL,
                  sex_dom_wrap2,
                  nrow=1, ncol=5, 
                  rel_widths=c(1.8,.07,1,.05,2.3), 
                  labels=c("", "D", ""),
                  label_size = 12)

ggsave("figs/fig2.png", fig2, width=19, height=8, dpi=600, bg = "white")
```


#FIGURE 3: Sex Effects Over the Lifespan

median group
```{r}
cent_group <- readRDS("figs/ggobjs/cent_groups_brain_B.rds")

cent_group[[1]] <- cent_group[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

cent_group_wrap <- plot_grid(
  cent_group[[1]],
  cent_group[[2]], 
  cent_group[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
)
```

median group trajectories

```{r}
cent_traj_group <- readRDS("figs/ggobjs/cent_groups_B.rds")
```

```{r}
cent_med_group <- readRDS("figs/ggobjs/cent_groups_median_B.rds") +
  theme(legend.position = "none")
```

deriv group
```{r}
deriv_group <- readRDS("figs/ggobjs/deriv_groups_brain_B.rds")

deriv_group[[1]] <- deriv_group[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

deriv_group_wrap <- plot_grid(
  deriv_group[[1]],
  deriv_group[[2]], 
  deriv_group[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
)
```

deriv group trajectories

```{r}
deriv_med_group <- readRDS("figs/ggobjs/deriv_groups_median_B.rds") +
  theme(legend.position = "none")
```

combine
```{r}
grp_plt <- plot_grid(
  plot_grid(
    NULL,
  cent_group_wrap,
  cent_med_group,
  NULL,
  deriv_group_wrap,
  deriv_med_group,
  ncol=6,
  rel_widths=c(.1, 1, 2, .1, 1, 2),
  labels=c("B", "", "C", "D", "", "E"),
  align="h"
),
deriv_group[[4]],
nrow=2,
rel_heights=c(1, .2)
)
ggsave("figs/fig3.png", grp_plt, width=17, height=9, dpi=600, bg = "white")
```

#FIGURE 4: Case-Control Differences

```{r}
dx_bxpltB <- readRDS("figs/ggobjs/dx_bxpltB.rds")
dx_brain_plt <- readRDS("figs/ggobjs/dx_brain_plt.rds")
```

combine

```{r}
dx_plt <- plot_grid(
  dx_bxpltB,
  NULL,
  dx_brain_plt,
  ncol=3,
  rel_widths=c(1.2,.06, 4),
  labels=c("A", "B", "")
)
ggsave("figs/fig4.png", dx_plt, width=21, height=6.5, dpi=600, bg = "white")
```

#FIGURE 5: sex-differences across the lifespan

```{r}
cent_plt_pct <- readRDS("figs/ggobjs/sGMV_B_trajectories.rds")
```
