---
title: "Format Figures"
output: h
tml_document
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, cowplot, sf)

devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

base_path<- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/"
# install.packages("remotes")
#remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

```{r}
#reformat x to keep birth label
reformat_x <- function(x_values, simplified=FALSE){
  tickLabels <- c()
  tickMarks <- c()
  if(simplified==TRUE){
    for (year in c(0, 25, 50, 75, 100)) {
     tickMarks <- append(tickMarks, year * 365.25 + 280)
    }
    tickLabels <- c("Birth", "25", "50", "75", "100")
  } else {
    for (year in seq(0, 100, by = 10)) {
      tickMarks <- append(tickMarks, year * 365.25 + 280)
    }
    tickLabels <- c("Birth", "10", "20", "30", "40", "50", "60", "70", "80", "90", "100")
  }
  
  unit_lab <- "(years)"
  list(scale_x_continuous(breaks = tickMarks, labels = tickLabels), labs(x = paste("Age at Scan", unit_lab)))
}

```

#FIGURE 1: Age-by-Sex Effects

Sample B exemplar curve(s)

```{r}
df <- fread(paste0(base_path, "data/cv_sample_B_dfs/sGMV_totalFALSE_logAgeTRUE.csv"), stringsAsFactors = TRUE, na.strings = "")

test_mod <- readRDS(Sys.glob(paste0(base_path, "cv_sample_B_test/global_vols_totalFALSE_logAgeTRUE_pbmods/model_objs/sGMV_*_test_full_mod.rds")))
test_mod$call$data <- "df"
test_mod$call$family <- test_mod$family[[1]]

sim_df <- sim_data(df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
vars_of_interest <- c("sexMale_x_logAge", "logAge_days", "sexMale")
resid_terms <- setdiff(list_predictors(test_mod), vars_of_interest)

sGMV_fan <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="logAge_days",
                 color_var="sexMale",
                 get_peaks=TRUE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df,
                 remove_point_effect = resid_terms,
                 label_centiles = "legend",
                 x_axis="log_lifespan_fetal",
                 point_color_manual = c('0' = "#F4A15BFF", '1' = "#8CB3D1FF"),
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_blank()) +
    labs(title=paste("sGMV validation model"),
         x ="Age (years)",
         y="Subcortical Gray Matter Volume (mm3)") +
    guides(fill=FALSE) +
    geom_vline(xintercept=log(280, base=10))
```
```{r}
age_by_sex_plt <- readRDS("figs/ggobjs/agebysex_B.rds")

age_by_sex_plt_wrap <- plot_grid(
  age_by_sex_plt[[2]], 
  age_by_sex_plt[[3]],
  ncol = 1,
  rel_heights = c(5,1.05),
  align = "v",
  axis = "lr"
)
```

try to combine

```{r}
sGMV_fan <- plot_grid(NULL, sGMV_fan, NULL, ncol=1, rel_heights=c(1,6,1))

fig1 <- plot_grid(sGMV_fan,
                  age_by_sex_plt_wrap, 
                  age_by_sex_plt[[4]],
                  nrow=1, ncol=3, 
                  rel_widths=c(3,2,.4), 
                  labels=c("A", "B", ""))

ggsave("figs/fig1.png", fig1, width=11, height=8, dpi=600, bg = "white")
```



#FIGURE 2: Sex Effects Controlling for Total Size

Sample B exemplar curve(s)

```{r}
df <- fread(paste0(base_path, "data/cv_sample_B_dfs/sGMV_totalTRUE_logAgeTRUE.csv"), stringsAsFactors = TRUE, na.strings = "")

test_mod <- readRDS(Sys.glob(paste0(base_path, "cv_sample_B_test/global_vols_totalTRUE_logAgeTRUE_pbmods/model_objs/sGMV_*_test_full_mod.rds")))

test_mod$call$data <- "df"
test_mod$call$family <- test_mod$family[[1]]
list_predictors(test_mod)

sim_df1 <- sim_data(df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
vars_of_interest <- c("sexMale_x_logAge", "logAge_days", "sexMale", "TBV")
resid_terms <- setdiff(list_predictors(test_mod), vars_of_interest)

#manually set tbv to zero rather than mean
sim_df1$'1'$TBV <- 0
sim_df1$'0'$TBV <- 0

sGMV_fan_tbv <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="logAge_days",
                 color_var="sexMale",
                 get_peaks=TRUE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df1,
                 label_centiles = "legend",
                 show_points = FALSE,
                 x_axis="log_lifespan_fetal",
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  theme(legend.position= "bottom", 
        plot.title = element_blank(), 
        legend.title = element_blank(),
        legend.box.spacing = unit(0, "pt")) +
    labs(x ="Age (years)",
         y="sGMV (mm3)") +
    guides(fill=FALSE) +
    geom_vline(xintercept=log(280, base=10))
```


```{r}
#read in file
traject <- fread(paste0(base_path, "cv_sample_B_test/global_vols_totalTRUE_logAgeTRUE_pbmods/cent_csvs/sGMV_sexdiffs.csv"), stringsAsFactors = TRUE, na.strings = "")

#mean sex diff trajectory
sgmv_sexdiff_plt <- traject %>%
  ggplot() +
  geom_hline(yintercept=0, color="gray", linetype = "dashed") +
  geom_line(aes(x=logAge_days, y=centile_M_minus_F), color="#216936FF") +
  theme_linedraw() +
  theme(legend.position = "none") +
  format_x_axis("log_lifespan_fetal", traject$logAge_days) +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  labs(y="Sex-Difference in\nMedian Trajectory")

#variance sex diff trajectory
sgmv_var_plt <- traject %>%
  ggplot() +
  geom_hline(yintercept=0, color="gray", linetype = "dashed") +
  geom_line(aes(x=logAge_days, y=logcv_M_div_F), color="#216936FF") +
  theme_linedraw() +
  theme(legend.position = "none") +
  format_x_axis("log_lifespan_fetal", traject$logAge_days) +
  labs(y="Sex-Difference in\nVariance",
       x ="Age (years)")
```

combine trajectory plots

```{r}
sGMV_fan <- plot_grid(sGMV_fan_tbv, 
                      plot_grid(sgmv_sexdiff_plt, sgmv_var_plt,
                      ncol=1, 
                      nrow=2,
                      rel_widths=c(1,1),
                      align = "hv",
                      labels=c("B", "C"),
                      label_size = 12,
                      hjust=-.1,
                      vjust=-1),
                      ncol=1,
                      rel_heights=c(3,2),
                      labels=c("A", ""),
                      label_size=12)
```


brain plots
```{r}
sex_plt <- readRDS("figs/ggobjs/full_plt_tot_sex_B.rds")
sex_plt[[1]] <- sex_plt[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sex_plt_wrap <- plot_grid(
  sex_plt[[1]],
  sex_plt[[2]], 
  sex_plt[[3]],
  sex_plt[[4]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93, .7),
  align = "v",
  axis = "lr"
)
```


sex dominance
```{r}
sex_dom <- readRDS("figs/ggobjs/cent_dom_B.rds")
sex_dom[[1]] <- sex_dom[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sig_sex_dom <- readRDS("figs/ggobjs/sig_dom_B.rds")
sig_sex_dom[[1]] <- sig_sex_dom[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())
```

```{r}
#median
#try adding legend label
legend_labeled <- ggdraw(sex_dom[[4]]) +
  draw_label("Female",   x = 0.18, y = 0.5, size = 9, hjust = 0.5, fontface = "italic") +
  draw_label("Male", x = 0.7, y = 0.5, size = 9, hjust = 0.5, fontface = "italic")

sex_dom_median <- plot_grid(
  sex_dom[[1]],
  sex_dom[[2]], 
  sex_dom[[3]],
  legend_labeled,
  ncol = 1,
  rel_heights = c(.85, 4.5, .93, .7),
  align = "v",
  axis = "lr"
)

#sigma
#try adding legend label
sig_legend_labeled <- ggdraw(sig_sex_dom[[4]]) +
  draw_label("Female",   x = 0.18, y = 0.5, size = 9, hjust = 0.5, fontface = "italic") +
  draw_label("Male", x = 0.7, y = 0.5, size = 9, hjust = 0.5, fontface = "italic")

sex_dom_sigma <- plot_grid(
  sig_sex_dom[[1]],
  sig_sex_dom[[2]], 
  sig_sex_dom[[3]],
  sig_legend_labeled,
  ncol = 1,
  rel_heights = c(.85, 4.5, .93,.7),
  align = "v",
  axis = "lr"
)
```
try to combine

```{r}
fig2 <- plot_grid(sGMV_fan,
                  NULL,
                  sex_plt_wrap,
                  NULL,
                  sex_dom_median,
                  NULL,
                  sex_dom_sigma,
                  nrow=1, ncol=7, 
                  rel_widths=c(1.8,.05,1,.05,1,.05,1), 
                  labels=c("", "C", "", "E","", "F", ""),
                  label_size = 12)

ggsave("figs/fig2.png", fig2, width=19, height=11, dpi=600, bg = "white")
```


#FIGURE 3: Sex Effects Over the Lifespan

barplot

```{r}
barplot <- readRDS("figs/ggobjs/cent_groups_barplt_B.rds")
legend <- get_legend(barplot, "bottom")
barplot_final <- barplot + theme(legend.position="none")
```

median group trajectories

```{r}
cent_traj_group <- readRDS("figs/ggobjs/cent_groups_traj_B.rds") + 
  labs(y="Std. Median Sex Difference")
```


median group
```{r}
cent_group <- readRDS("figs/ggobjs/cent_groups_brain_B.rds")

cent_group[[1]] <- cent_group[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

cent_group_wrap <- plot_grid(
  cent_group[[1]],
  cent_group[[2]], 
  cent_group[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
)
```

medoid trajectories

```{r}
kmed_list <- readRDS(paste0(base_path,"figs/ggobjs/median_kmedoid_clust_B.rds"))
elbow <- kmed_list[[2]]
medoids <- kmed_list[[3]]

medoid_df <- kmed_list[[1]] %>%
  filter(pheno %in% medoids) %>%
  mutate(cluster_label = paste("Cluster", cluster, "Medoid Trajectory"))

medoid_plt <- ggplot(medoid_df) +
  geom_hline(yintercept=0, color="gray20", linetype = "dashed") +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z, color=as.factor(cluster), group=pheno)) +
  facet_wrap(~cluster_label, nrow=2) +
  theme_minimal() +
  scale_color_manual(values=c("2"="#FA9A16", 
                              "1" ="#1676FA")) +
  theme(legend.position = "none") +
  reformat_x(medoid_df$age_days) +
  labs(y="Sex Difference in Median", x="Age (years)")
```


```{r}
#update colors, etc
elbow <- elbow + theme(plot.title=element_blank()) +
  labs(y="Average\nSilhouette Width", 
       x="Number of Clusters")
```

combining

```{r}
top_down <- plot_grid(
  plot_grid(
  #barplot and trajectories
  plot_grid(
    barplot_final, 
    cent_traj_group, 
    nrow=2, 
    rel_heights = c(2,3),
    labels = c("A", "B")),
  #brain plot
    cent_group_wrap,
    labels = c("", "C"),
    ncol=2, rel_widths = c(3,1.3)),
  #finally add shared legend
  cent_group[[4]],
  nrow=2, 
  rel_heights = c(10,1)
)

data_driven <- plot_grid(
  elbow,
  medoid_plt,
  nrow=2,
  rel_heights = c(1,2),
  labels = c("D", "E")
)  

med_plt <-plot_grid(
  top_down, 
  data_driven,
  ncol=2,
  rel_widths = c(4.3,1.6))

ggsave("figs/fig3.png", med_plt, width=16, height=8.5, dpi=600, bg = "white")
```


#FIGURE 5: Case-Control Differences

```{r}
#brain plot
dx_brain_plt <- readRDS("figs/ggobjs/dx_brain_plt.rds")

#global pheno boxplots
dx_bxplt <- readRDS("figs/ggobjs/dx_bxplt.rds")

#change bar plots
cent_disagree_plt <- readRDS("figs/ggobjs/cent_disagree_bar_plt.rds")
```
try combining

```{r}
upper_row <- plot_grid(
  NULL,
  plot_grid(
    dx_brain_plt[[1]],
    dx_brain_plt[[2]],
    align = "hv",
    axis= c("lr"),
    nrow=2,
    rel_heights = c(5, 1.1)
  ),
  dx_brain_plt[[3]],
  ncol=3,
  rel_widths=c(0.3, 20,1)
)

lower_row <- plot_grid(
  dx_bxplt,
  cent_disagree_plt,
  nrow=1, ncol=2,
  rel_widths=c(2.5,1),
  labels = c("B", "C")
)

fig5 <- plot_grid(
  upper_row,
  lower_row, 
  ncol = 1,
  rel_heights = c(1.15,1),
  labels=c("A","")
)

ggsave("figs/fig5.png", fig4, width=17, height=13, dpi=600, bg = "white")

```

