---
title: "Format Figures"
output: html_document
date: "2026-02-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, cowplot, sf)

devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

base_path<- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/"
# install.packages("remotes")
#remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

#FIGURE 1: Age-by-Sex Effects

Sample A exemplar curve(s)

```{r}
df <- fread(paste0(base_path, "data/cv_sample_B_dfs/sGMV_totalFALSE_logPhenoFALSE_logAgeTRUE.csv"), stringsAsFactors = TRUE, na.strings = "")

test_mod <- readRDS(paste0(base_path, "cv_sample_B_test/global_vols_totalFALSE_logAgeTRUE_pbmods/model_objs/sGMV_mu_pbAge_BCTo_test_full_mod.rds"))
test_mod$call$data <- "df"
test_mod$call$family <- test_mod$family[[1]]

sim_df <- sim_data(df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
vars_of_interest <- c("sexMale_x_logAge", "logAge_days", "sexMale")
resid_terms <- setdiff(list_predictors(test_mod), vars_of_interest)

sGMV_fan <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="logAge_days",
                 color_var="sexMale",
                 get_peaks=TRUE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df,
                 remove_point_effect = resid_terms,
                 label_centiles = "legend",
                 x_axis="log_lifespan_fetal",
                 point_color_manual = c('0' = "#F4A15BFF", '1' = "#8CB3D1FF"),
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  theme(legend.position = "bottom", plot.title = element_blank()) +
    labs(title=paste("sGMV validation model"),
         x ="Age (years)",
         y="Subcortical Gray Matter Volume (mm3)") +
    guides(fill=FALSE) +
    geom_vline(xintercept=log(280, base=10))
```
```{r}
age_by_sex_plt <- readRDS("figs/ggobjs/agebysex_B.rds")

age_by_sex_plt_wrap <- plot_grid(
  age_by_sex_plt[[2]], 
  age_by_sex_plt[[3]],
  ncol = 1,
  rel_heights = c(5,1.05),
  align = "v",
  axis = "lr"
)
```

try to combine

```{r}
sGMV_fan <- plot_grid(NULL, sGMV_fan, NULL, ncol=1, rel_heights=c(1,6,1))

fig1 <- plot_grid(sGMV_fan,
                  age_by_sex_plt_wrap, 
                  age_by_sex_plt[[4]],
                  nrow=1, ncol=3, 
                  rel_widths=c(3,2,.4), 
                  labels=c("A", "B", ""))

ggsave("figs/fig1.png", fig1, width=11, height=8, dpi=600, bg = "white")
```



#FIGURE 2: Sex Effects Controlling for Total Size

Sample A exemplar curve(s)

```{r}
df <- fread(paste0(base_path, "data/cv_sample_B_dfs/sGMV_totalTRUE_logPhenoFALSE_logAgeTRUE.csv"), stringsAsFactors = TRUE, na.strings = "")

test_mod <- readRDS(paste0(base_path, "cv_sample_B_test/global_vols_totalTRUE_logAgeTRUE_pbmods/model_objs/sGMV_fsmu_pbAge_totalboth_test_full_mod.rds"))
test_mod$call$data <- "df"
test_mod$call$family <- test_mod$family[[1]]
list_predictors(test_mod)

sim_df1 <- sim_data(df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
vars_of_interest <- c("sexMale_x_logAge", "logAge_days", "sexMale", "TBV")
resid_terms <- setdiff(list_predictors(test_mod), vars_of_interest)

#manually set tbv to zero rather than mean
sim_df1$'1'$TBV <- 0
sim_df1$'0'$TBV <- 0

sGMV_fan_tbv <- make_centile_fan(gamlssModel=test_mod,
                 df=df,
                 x_var="logAge_days",
                 color_var="sexMale",
                 get_peaks=TRUE,
                 desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                 sim_data_list = sim_df1,
                 label_centiles = "legend",
                 show_points = FALSE,
                 x_axis="log_lifespan_fetal",
                 color_manual = c('0' = "#c05f0d", '1' = "#38688D"),
                 color_name = "Sex",
                 color_labels = c("Female", "Male"))  +
    theme_linedraw() +
  theme(legend.position= "bottom", plot.title = element_blank()) +
    labs(x ="Age (years)",
         y="sGMV (mm3)") +
    guides(fill=FALSE) +
    geom_vline(xintercept=log(280, base=10))
```


```{r}
#read in file
traject <- fread(paste0(base_path, "cv_sample_B_test/global_vols_totalTRUE_logAgeTRUE_pbmods/cent_csvs/sGMV_sexdiffs.csv"), stringsAsFactors = TRUE, na.strings = "")

#mean sex diff trajectory
sgmv_sexdiff_plt <- traject %>%
  ggplot() +
  geom_hline(yintercept=0, color="gray", linetype = "dashed") +
  geom_line(aes(x=logAge_days, y=centile_M_minus_F), color="#216936FF") +
  theme_linedraw() +
  theme(legend.position = "none") +
  format_x_axis("log_lifespan_fetal", traject$logAge_days) +
  theme(legend.position = "none", axis.title.x = element_blank()) +
  labs(y="Sex-Difference in\nMedian Trajectory")

#variance sex diff trajectory
sgmv_var_plt <- traject %>%
  ggplot() +
  geom_hline(yintercept=0, color="gray", linetype = "dashed") +
  geom_line(aes(x=logAge_days, y=logcv_M_div_F), color="#216936FF") +
  theme_linedraw() +
  theme(legend.position = "none") +
  format_x_axis("log_lifespan_fetal", traject$logAge_days) +
  labs(y="Sex-Difference in\nVariance",
       x ="Age (years)")
```

combine trajectory plots

```{r}
sGMV_fan <- plot_grid(sGMV_fan_tbv, 
                      plot_grid(sgmv_sexdiff_plt, sgmv_var_plt,
                      ncol=1, 
                      nrow=2,
                      rel_widths=c(1,1),
                      align = "hv",
                      labels=c("B", "C"),
                      label_size = 12,
                      hjust=-.1,
                      vjust=-1),
                      ncol=1,
                      rel_heights=c(3,2),
                      labels=c("A", ""),
                      label_size=12)
```


brain plots
```{r}
sex_plt <- readRDS("figs/ggobjs/full_plt_tot_sex_B.rds")
sex_plt[[1]] <- sex_plt[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sex_plt_wrap <- plot_grid(
  plot_grid(
  sex_plt[[1]],
  sex_plt[[2]], 
  sex_plt[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
), sex_plt[[4]],
ncol = 2,
rel_widths = c(1, .2))
```


sex dominance
```{r}
sex_dom <- readRDS("figs/ggobjs/cent_dom_B.rds")

sex_dom[[1]] <- sex_dom[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())
```

```{r}
sig_sex_dom <- readRDS("figs/ggobjs/sig_dom_B.rds")

sig_sex_dom[[1]] <- sig_sex_dom[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sex_dom_wrap <- plot_grid(
  #median
  plot_grid(
  sex_dom[[1]],
  sex_dom[[2]], 
  sex_dom[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
),
#sigma
plot_grid(
  sig_sex_dom[[1]],
  sig_sex_dom[[2]], 
  sig_sex_dom[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
),
ncol = 2,
labels=c("E", "F"),
label_size = 12,
align="vh")

#add legend
sex_dom_wrap2 <- plot_grid(sex_dom_wrap, sex_dom[[4]],
                           ncol = 2,
                           rel_widths = c(1, .2))
```
try to combine

```{r}
fig2 <- plot_grid(sGMV_fan,
                  NULL,
                  sex_plt_wrap,
                  NULL,
                  sex_dom_wrap2,
                  nrow=1, ncol=5, 
                  rel_widths=c(1.8,.07,1,.05,2.3), 
                  labels=c("", "D", ""),
                  label_size = 12)

ggsave("figs/fig2.png", fig2, width=19, height=8, dpi=600, bg = "white")
```


#FIGURE 3: Sex Effects Over the Lifespan

median group
```{r}
cent_group <- readRDS("figs/ggobjs/cent_groups_brain_B.rds")

cent_group[[1]] <- cent_group[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

cent_group_wrap <- plot_grid(
  cent_group[[1]],
  cent_group[[2]], 
  cent_group[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
)
```

median group trajectories

```{r}
cent_traj_group <- readRDS("figs/ggobjs/cent_groups_traj_B.rds") + theme(strip.text.x = element_blank(), strip.background.x = element_blank())
```

sigma group

```{r}
sigma_group <- readRDS("figs/ggobjs/sig_groups_brain_B.rds")

sigma_group[[1]] <- sigma_group[[1]] + theme(strip.text.x = element_blank(), strip.background.x = element_blank())

sigma_group_wrap <- plot_grid(
  sigma_group[[1]],
  sigma_group[[2]], 
  sigma_group[[3]],
  ncol = 1,
  rel_heights = c(.85, 4.5, .93),
  align = "v",
  axis = "lr"
)
```


```{r}
sigma_traj_group <- readRDS("figs/ggobjs/sig_groups_traj_B.rds") +
  theme(legend.position = "none") + theme(strip.text.x = element_blank(), strip.background.x = element_blank())
```

combine
```{r}
grp_plt <- plot_grid(
  plot_grid(
    NULL,
  cent_group_wrap,
  cent_traj_group,
  NULL,
  sigma_group_wrap,
  sigma_traj_group,
  ncol=6,
  rel_widths=c(.1, 1.3, 2, .1, 1.3, 2),
  labels=c("A", "", "B", "C", "", "D"),
  align="h",
  axis="tb"
),
cent_group[[4]],
nrow=2,
rel_heights=c(1, .06)
)
ggsave("figs/fig3.png", grp_plt, width=17, height=9, dpi=600, bg = "white")
```

#FIGURE 4: Case-Control Differences

```{r}
dx_bxpltB <- readRDS("figs/ggobjs/dx_bxpltB.rds")
dx_brain_plt <- readRDS("figs/ggobjs/dx_brain_plt.rds")
```

combine

```{r}
dx_plt <- plot_grid(
  dx_bxpltB,
  NULL,
  dx_brain_plt,
  ncol=3,
  rel_widths=c(1.2,.06, 4),
  labels=c("A", "B", "")
)
ggsave("figs/fig4.png", dx_plt, width=21, height=6.5, dpi=600, bg = "white")
```

#FIGURE 5: sex-differences across the lifespan

```{r}
cent_plt_pct <- readRDS("figs/ggobjs/sGMV_B_trajectories.rds")
```
