---
title: "Spline Testing"
output: html_document
date: "2024-12-18"
---


```{r warning=FALSE}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, gamlss, gamlssTools, gamlss.add, mgcv, gamlss2)
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
source("code/gamlss_fit_funs.R")

# install.packages("devtools")
devtools::install_github("BGDlab/gamlssTools", force=TRUE)

library(gamlssTools)

devtools::install_github("deepankardatta/blandr")
library(blandr)
```

##Get gamlss() formula
GMV
load standard gamlss brain chart

```{r}
brainchart_path <- "braincharts/global_vols_totalFALSE_logPhenoTRUE_logAgeTRUE_pbmods/model_objs/"
gmv_gamlss <- readRDS(paste0(brainchart_path, "GMV_brainchart.rds"))
gmv_gamlss$mu.formula
gmv_gamlss$sigma.formula
gmv_gamlss$nu.formula
```


```{r}
#separate smooths for male and female - basic pb model
gmv_formula_pb <- formula(GMV ~ pb(logAge_days, by=sex) + sex + s(study_site, bs='re') + fs_version_GM | 
                pb(logAge_days, by=sex) + sex + s(study_site, bs='re') | 
                1)

#s defaults
gmv_formula_s<- formula(GMV ~ s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') + fs_version_GM | 
                         s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') |
                         1)

#s - k=10
gmv_formula_ps10<- formula(GMV ~ s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
gmv_formula_ps50<- formula(GMV ~ s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
gmv_formula_ps100<- formula(GMV ~ s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
gmv_formula_ps500<- formula(GMV ~ s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=10
gmv_formula_tp10<- formula(GMV ~ s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
gmv_formula_tp50<- formula(GMV ~ s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
gmv_formula_tp100<- formula(GMV ~ s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
gmv_formula_tp500<- formula(GMV ~ s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
gmv_formula_tp20<- formula(GMV ~ s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#### cr splines

#s - k=10
gmv_formula_cr10<- formula(GMV ~ s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
gmv_formula_cr20<- formula(GMV ~ s(logAge_days, by=sex, bs="cr", k=20) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
gmv_formula_cr50<- formula(GMV ~ s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
gmv_formula_cr100<- formula(GMV ~ s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
gmv_formula_cr500<- formula(GMV ~ s(logAge_days, by=sex, bs="", k=500) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="cr", k=500) + sex + s(study_site, bs='re') |
                         1)
```

WMV

```{r}
wmv_gamlss <- readRDS(paste0(brainchart_path, "WMV_brainchart.rds"))
wmv_gamlss$mu.formula
wmv_gamlss$sigma.formula
wmv_gamlss$nu.formula
```

```{r}
#separate smooths for male and female - basic pb model
wmv_formula_pb <- formula(WMV ~ pb(logAge_days, by=sex) + sex + s(study_site, bs='re') | 
                pb(logAge_days, by=sex) + sex + s(study_site, bs='re') | 
                1)

#s defaults
wmv_formula_s<- formula(WMV ~ s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') | 
                         s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') |
                         1)

#s - k=10
wmv_formula_ps10<- formula(WMV ~ s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
wmv_formula_ps50<- formula(WMV ~ s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
wmv_formula_ps100<- formula(WMV ~ s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
wmv_formula_ps500<- formula(WMV ~ s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=10
wmv_formula_tp10<- formula(WMV ~ s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
wmv_formula_tp50<- formula(WMV ~ s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
wmv_formula_tp100<- formula(WMV ~ s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
wmv_formula_tp500<- formula(WMV ~ s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') |
                         1)
#s - k=20
wmv_formula_tp20<- formula(WMV ~ s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#### cr splines

#s - k=10
wmv_formula_cr10<- formula(WMV ~ s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
wmv_formula_cr20<- formula(WMV ~ s(logAge_days, by=sex, bs="cr", k=20) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
wmv_formula_cr50<- formula(WMV ~ s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
wmv_formula_cr100<- formula(WMV ~ s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
wmv_formula_cr500<- formula(WMV ~ s(logAge_days, by=sex, bs="", k=500) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="cr", k=500) + sex + s(study_site, bs='re') |
                         1)
```


CSF

```{r}
csf_gamlss <- readRDS(paste0(brainchart_path, "CSF_brainchart.rds"))
csf_gamlss$mu.formula
csf_gamlss$sigma.formula
csf_gamlss$nu.formula
```

```{r}
#separate smooths for male and female - basic pb model
csf_formula_pb <- formula(CSF ~ pb(logAge_days, by=sex) + sex + s(study_site, bs='re') | 
                pb(logAge_days, by=sex) + sex + s(study_site, bs='re') | 
                1)

#s defaults
csf_formula_s<- formula(CSF ~ s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') | 
                         s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') |
                         1)

#s - k=10
csf_formula_ps10<- formula(CSF ~ s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
csf_formula_ps50<- formula(CSF ~ s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
csf_formula_ps100<- formula(CSF ~ s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
csf_formula_ps500<- formula(CSF ~ s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=10
csf_formula_tp10<- formula(CSF ~ s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
csf_formula_tp50<- formula(CSF ~ s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
csf_formula_tp100<- formula(CSF ~ s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
csf_formula_tp500<- formula(CSF ~ s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
csf_formula_tp20<- formula(CSF ~ s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#### cr splines

#s - k=10
csf_formula_cr10<- formula(CSF ~ s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
csf_formula_cr20<- formula(CSF ~ s(logAge_days, by=sex, bs="cr", k=20) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
csf_formula_cr50<- formula(CSF ~ s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') |
                         s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
csf_formula_cr100<- formula(CSF ~ s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
csf_formula_cr500<- formula(CSF ~ s(logAge_days, by=sex, bs="", k=500) + sex + s(study_site, bs='re') | s(logAge_days, by=sex, bs="cr", k=500) + sex + s(study_site, bs='re') |
                         1)
```

sGMV

```{r}
sgmv_gamlss <- readRDS(paste0(brainchart_path, "sGMV_brainchart.rds"))
sgmv_gamlss$mu.formula
sgmv_gamlss$sigma.formula
sgmv_gamlss$nu.formula
```

```{r}
#separate smooths for male and female - basic pb model
sgmv_formula_pb <- formula(sGMV ~ pb(logAge_days, by=sex) + sex + s(study_site, bs='re') + fs_version_GM | 
                pb(logAge_days, by=sex) + sex + s(study_site, bs='re') | 
                1)

#s defaults
sgmv_formula_s<- formula(sGMV ~ s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') + fs_version_GM | 
                         s(logAge_days, by=sex, bs="ps") + sex + s(study_site, bs='re') |
                         1)

#s - k=10
sgmv_formula_ps10<- formula(sGMV ~ s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="ps", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
sgmv_formula_ps50<- formula(sGMV ~ s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="ps", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
sgmv_formula_ps100<- formula(sGMV ~ s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="ps", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
sgmv_formula_ps500<- formula(sGMV ~ s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="ps", k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=10
sgmv_formula_tp10<- formula(sGMV ~ s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
sgmv_formula_tp50<- formula(sGMV ~ s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
sgmv_formula_tp100<- formula(sGMV ~ s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
sgmv_formula_tp500<- formula(sGMV ~ s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, k=500) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
sgmv_formula_tp20<- formula(sGMV ~ s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#### cr splines

#s - k=10
sgmv_formula_cr10<- formula(sGMV ~ s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="cr", k=10) + sex + s(study_site, bs='re') |
                         1)

#s - k=20
sgmv_formula_cr20<- formula(sGMV ~ s(logAge_days, by=sex, bs="cr", k=20) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, k=20) + sex + s(study_site, bs='re') |
                         1)

#s - k=50
sgmv_formula_cr50<- formula(sGMV ~ s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') + fs_version_GM |
                         s(logAge_days, by=sex, bs="cr", k=50) + sex + s(study_site, bs='re') |
                         1)

#s - k=100
sgmv_formula_cr100<- formula(sGMV ~ s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="cr", k=100) + sex + s(study_site, bs='re') |
                         1)

#s - k=500
sgmv_formula_cr500<- formula(sGMV ~ s(logAge_days, by=sex, bs="", k=500) + sex + s(study_site, bs='re') + fs_version_GM | s(logAge_days, by=sex, bs="cr", k=500) + sex + s(study_site, bs='re') |
                         1)
```


#Write Config File

```{r}
clean_formula <- function(x){
  as.character(x)[3]
}
```

```{r, eval=FALSE}
pheno_list <- rep(c("GMV", "WMV", "CSF", "sGMV"), 16)
config <- data.frame(ct = seq(1:64),
                    pheno = c(pheno_list),
                     form = c(clean_formula(gmv_formula_pb), #regular formula
                              clean_formula(wmv_formula_pb), 
                              clean_formula(csf_formula_pb), 
                              clean_formula(sgmv_formula_pb), 
                              clean_formula(gmv_formula_s),
                              clean_formula(wmv_formula_s), 
                              clean_formula(csf_formula_s),
                              clean_formula(sgmv_formula_s),
                              clean_formula(gmv_formula_ps10),
                              clean_formula(wmv_formula_ps10), 
                              clean_formula(csf_formula_ps10), 
                              clean_formula(sgmv_formula_ps10),
                              clean_formula(gmv_formula_ps50),
                              clean_formula(wmv_formula_ps50), 
                              clean_formula(csf_formula_ps50), 
                              clean_formula(sgmv_formula_ps50),
                              clean_formula(gmv_formula_ps100),
                              clean_formula(wmv_formula_ps100), 
                              clean_formula(csf_formula_ps100), 
                              clean_formula(sgmv_formula_ps100),
                              clean_formula(gmv_formula_ps500), 
                              clean_formula(wmv_formula_ps500), 
                              clean_formula(csf_formula_ps500), 
                              clean_formula(sgmv_formula_ps500),
                              clean_formula(gmv_formula_tp10),
                              clean_formula(wmv_formula_tp10), 
                              clean_formula(csf_formula_tp10), 
                              clean_formula(sgmv_formula_tp10),
                              clean_formula(gmv_formula_tp50),
                              clean_formula(wmv_formula_tp50), 
                              clean_formula(csf_formula_tp50), 
                              clean_formula(sgmv_formula_tp50),
                              clean_formula(gmv_formula_tp100),
                              clean_formula(wmv_formula_tp100), 
                              clean_formula(csf_formula_tp100), 
                              clean_formula(sgmv_formula_tp100),
                              clean_formula(gmv_formula_tp500), 
                              clean_formula(wmv_formula_tp500), 
                              clean_formula(csf_formula_tp500), 
                              clean_formula(sgmv_formula_tp500),
                              clean_formula(gmv_formula_tp20),
                              clean_formula(wmv_formula_tp20), 
                              clean_formula(csf_formula_tp20), 
                              clean_formula(sgmv_formula_tp20),
                              clean_formula(gmv_formula_cr10),
                              clean_formula(wmv_formula_cr10), 
                              clean_formula(csf_formula_cr10), 
                              clean_formula(sgmv_formula_cr10),
                              clean_formula(gmv_formula_cr20),
                              clean_formula(wmv_formula_cr20), 
                              clean_formula(csf_formula_cr20), 
                              clean_formula(sgmv_formula_cr20),
                              clean_formula(gmv_formula_cr50),
                              clean_formula(wmv_formula_cr50), 
                              clean_formula(csf_formula_cr50), 
                              clean_formula(sgmv_formula_cr50),
                              clean_formula(gmv_formula_cr100),
                              clean_formula(wmv_formula_cr100), 
                              clean_formula(csf_formula_cr100), 
                              clean_formula(sgmv_formula_cr100),
                              clean_formula(gmv_formula_cr500), 
                              clean_formula(wmv_formula_cr500), 
                              clean_formula(csf_formula_cr500),
                              clean_formula(sgmv_formula_cr500)
                              ),
                     name = c("pb", "pb", "pb", "pb", 
                              "s", "s", "s", "s",
                              "ps10", "ps10", "ps10", "ps10",
                              "ps50", "ps50", "ps50", "ps50",
                              "ps100", "ps100", "ps100", "ps100",
                              "ps500", "ps500", "ps500", "ps500",
                              "tp10", "tp10", "tp10", "tp10",
                              "tp50", "tp50", "tp50", "tp50",
                              "tp100", "tp100", "tp100", "tp100",
                              "tp500", "tp500", "tp500", "tp500",
                              "tp20", "tp20", "tp20", "tp20",
                              "cr10", "cr10", "cr10", "cr10",
                              "cr20", "cr20", "cr20", "cr20",
                              "cr50", "cr50", "cr50", "cr50",
                              "cr100", "cr100", "cr100", "cr100",
                              "cr500", "cr500", "cr500", "cr500"))
stopifnot(nrow(config) == 64)
write.table(config, file = "code/config_files/gamlss2_test_config.txt", 
            append = TRUE, sep = "\t", row.names=FALSE, col.names=FALSE, quote=c(3))
```
 
#Prep Dataframes

```{r}
#GMV
gmv_df <- fread("data/pheno_dfs_totalFALSE/GMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
gmv_df <- gmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 1, "M", "F")),
           GMV = unscale(GMV))
# gmv_gamlss$call$data <- "df"
# gmv_gamlss$call$family <- gmv_gamlss$family[[1]]
gmv_sim <- sim_data(gmv_df, "logAge_days", factor_var="sex")

#sGMV
sgmv_df <- fread("data/pheno_dfs_totalFALSE/sGMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
sgmv_df <- sgmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 1, "M", "F")),
           sGMV = unscale(sGMV))
# sgmv_gamlss$call$data <- "df"
# sgmv_gamlss$call$family <- sgmv_gamlss$family[[1]]
sgmv_sim <- sim_data(sgmv_df, "logAge_days", factor_var="sex")

#WMV
wmv_df <- fread("data/pheno_dfs_totalFALSE/WMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
wmv_df <- wmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 1, "M", "F")),
           WMV = unscale(WMV))
# wmv_gamlss$call$data <- "df"
# wmv_gamlss$call$family <- wmv_gamlss$family[[1]]
wmv_sim <- sim_data(wmv_df, "logAge_days", factor_var="sex")

#CSF
csf_df <- fread("data/pheno_dfs_totalFALSE/CSF_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
csf_df <- csf_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 1, "M", "F")),
           CSF = unscale(CSF))
# csf_gamlss$call$data <- "df"
# csf_gamlss$call$family <- csf_gamlss$family[[1]]
csf_sim <- sim_data(csf_df, "logAge_days", factor_var="sex")
```

#Basic mgcv gams

need to play around more with family options

```{r}
gmv_gam <- mgcv::gam(GMV ~ s(logAge_days, by=sex) + sex + s(study_site, bs='re') + fs_version_GM, data=gmv_df)
plot(gmv_gam)

sgmv_gam <- mgcv::gam(sGMV ~ s(logAge_days, by=sex) + sex + s(study_site, bs='re') + fs_version_GM, data=sgmv_df)
plot(sgmv_gam)

wmv_gam <- mgcv::gam(WMV ~ s(logAge_days, by=sex) + sex + s(study_site, bs='re') + fs_version_GM, data=wmv_df)
plot(wmv_gam)

csf_gam <- mgcv::gam(CSF ~ s(logAge_days, by=sex) + sex + s(study_site, bs='re') + fs_version_GM, data=csf_df)
plot(csf_gam)
```

Look at distributions

```{r}
fitDist(gmv_df$GMV, type="realplus")
fitDist(sgmv_df$sGMV, type="realplus")
fitDist(wmv_df$WMV, type="realplus")
fitDist(csf_df$CSF, type="realplus")
```

#gamlss2 BCTo with s(bs="ps")
default k=10

###GMV

```{r}
gmv1 <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_sBCTo.rds")
gmv1$formula
list_predictors(gmv1)
```

compare effective degrees of freedom
```{r}
gmv1$df

gmv1$fitted.specials$mu$`s(logAge_days):sexF`$edf
gmv1$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(gmv1$specials[[1]]$knots)
length(gmv1$specials[[2]]$knots)
```
Centile fan plots

fit centile fan plots
```{r}
make_centile_fan(gmv1, gmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = gmv_sim)
```

residualizing points look weird, gives outliers?

worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv1), n.inter=5)
cent_cdf(gmv1, gmv_df, group = "sex")
```

###sGMV

```{r}
sgmv1 <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_sBCTo.rds")
sgmv1$formula
list_predictors(sgmv1)
```

compare effective degrees of freedom
```{r}
sgmv1$df

sgmv1$fitted.specials$mu$`s(logAge_days):sexF`$edf
sgmv1$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(sgmv1$specials[[1]]$knots)
length(sgmv1$specials[[2]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(sgmv1, sgmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = sgmv_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv1), n.inter=6)
cent_cdf(sgmv1, sgmv_df, group = "sex")
```

###WMV

```{r}
wmv1 <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_sBCTo.rds")
wmv1$formula
list_predictors(wmv1)
```

compare effective degrees of freedom
```{r}
wmv1$df

wmv1$fitted.specials$mu$`s(logAge_days):sexF`$edf
wmv1$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(wmv1$specials[[1]]$knots)
length(wmv1$specials[[2]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(wmv1, wmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = wmv_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv1), n.inter=6)
cent_cdf(wmv1, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf1 <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_sBCTo.rds")
csf1$formula
list_predictors(csf1)
```

compare effective degrees of freedom
```{r}
csf1$df

csf1$fitted.specials$mu$`s(logAge_days):sexF`$edf
csf1$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(csf1$specials[[1]]$knots)
length(csf1$specials[[2]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(csf1, csf_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = csf_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf1), n.inter=6)
cent_cdf(csf1, csf_df, group = "sex")
```

try CSF with BCT (mu link identity)

```{r}
csf_bct <- gamlss2(CSF ~ s(logAge_days, by = sex, bs = "ps") + sex + s(study_site, 
    bs = "re") | s(logAge_days, by = sex, bs = "ps") + sex + 
    s(study_site, bs = "re") | 1 | 1,
    data=csf_df,
    family="BCT")
```

#gamlss2 BCTo with s(bs="tp", k=20)

###GMV

```{r}
gmv2 <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_tp20BCTo.rds")
gmv2$formula
list_predictors(gmv2)
```

compare effective degrees of freedom
```{r}
gmv2$df

gmv2$fitted.specials$mu$`s(logAge_days):sexF`$edf
gmv2$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(gmv2$specials[[1]]$knots) #not sure how to 
length(gmv2$specials[[2]]$xp)
```
Centile fan plots

fit centile fan plots
```{r}
make_centile_fan(gmv2, gmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = gmv_sim)
```


worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv2), n.inter=6)
cent_cdf(gmv2, gmv_df, group = "sex")
```

###sGMV

```{r}
sgmv2 <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_tp20BCTo.rds")
sgmv2$formula
list_predictors(sgmv2)
```

compare effective degrees of freedom
```{r}
sgmv2$df

sgmv2$fitted.specials$mu$`s(logAge_days):sexF`$edf
sgmv2$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(sgmv2$specials[[1]]$knots)
length(sgmv2$specials[[2]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(sgmv2, sgmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = sgmv_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv2), n.inter=6)
cent_cdf(sgmv2, sgmv_df, group = "sex")
```

###WMV

```{r}
wmv2 <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_tp20BCTo.rds")
wmv2$formula
list_predictors(wmv2)
```

compare effective degrees of freedom
```{r}
wmv2$df

wmv2$fitted.specials$mu$`s(logAge_days):sexF`$edf
wmv2$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(wmv2$specials[[1]]$knots)
length(wmv2$specials[[2]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(wmv2, wmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = wmv_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv2), n.inter=6)
cent_cdf(wmv2, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf2 <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_tp20BCTo.rds")
csf2$formula
list_predictors(csf2)
```

compare effective degrees of freedom
```{r}
csf2$df

csf2$fitted.specials$mu$`s(logAge_days):sexF`$edf
csf2$fitted.specials$mu$`s(logAge_days):sexM`$edf

length(csf2$specials[[1]]$knots)
length(csf2$specials[[2]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(csf2, csf_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = csf_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf2), n.inter=6)
cent_cdf(csf2, csf_df, group = "sex")
```

#gamlss2 BCTo with pb() - s(bs="ps", k=20)

###GMV

```{r}
gmv3 <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_pbBCTo.rds")
gmv3$formula
list_predictors(gmv3)
```

compare effective degrees of freedom
```{r}
gmv3$df

gmv3$fitted.specials$mu$`pb2(logAge_days):sexF`$edf
gmv3$fitted.specials$mu$`pb2(logAge_days):sexM`$edf

length(gmv3$specials$`pb2(logAge_days):sexF`$knots)
length(gmv3$specials$`pb2(logAge_days):sexM`$knots)
```
Centile fan plots

fit centile fan plots
```{r}
make_centile_fan(gmv3, gmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = gmv_sim)
```

wiggly!

worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv3), n.inter=5)
cent_cdf(gmv3, gmv_df, group = "sex")
```

###sGMV

```{r}
sgmv3 <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_pbBCTo.rds")
sgmv3$formula
list_predictors(sgmv3)
```

compare effective degrees of freedom
```{r}
sgmv3$df

sgmv3$fitted.specials$mu$`pb2(logAge_days):sexF`$edf
sgmv3$fitted.specials$mu$`pb2(logAge_days):sexM`$edf

length(sgmv3$specials[[2]]$knots)
length(sgmv3$specials[[3]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(sgmv3, sgmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = sgmv_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv3), n.inter=6)
cent_cdf(sgmv3, sgmv_df, group = "sex")
```

###WMV

```{r}
wmv3 <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_pbBCTo.rds")
wmv3$formula
list_predictors(wmv3)
```

compare effective degrees of freedom
```{r}
wmv3$df

wmv3$fitted.specials$mu$`pb2(logAge_days):sexF`$edf
wmv3$fitted.specials$mu$`pb2(logAge_days):sexM`$edf

length(wmv3$specials[[2]]$knots)
length(wmv3$specials[[3]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(wmv3, wmv_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = wmv_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv3), n.inter=6)
cent_cdf(wmv3, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf3 <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_pbBCTo.rds")
csf3$formula
list_predictors(csf3)
```

compare effective degrees of freedom
```{r}
csf3$df

csf3$fitted.specials$mu$`pb2(logAge_days):sexF`$edf
csf3$fitted.specials$mu$`pb2(logAge_days):sexM`$edf

length(csf3$specials[[2]]$knots)
length(csf3$specials[[3]]$knots)
```

fit centile fan plots
```{r}
make_centile_fan(csf3, csf_df, "logAge_days", "sex",
                 get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend",
                 remove_point_effect = "study_site",
                 sim_data_list = csf_sim)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf3), n.inter=6)
cent_cdf(csf3, csf_df, group = "sex")
```


#gamlss2 with non-log-scaled age


###GMV

```{r}
gmv_gamlss2x <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
gmv_gamlss2x$formula
list_predictors(gmv_gamlss2x)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(gmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
gmv_df_unscaledx <- gmv_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(gmv_gamlss2x, gmv_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE, 
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```


worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df_unscaledx$age_days, resid=resid(gmv_gamlss2x), n.inter=4)
cent_cdf(gmv_gamlss2x, gmv_df_unscaledx, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2x <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
sgmv_gamlss2x$formula
list_predictors(sgmv_gamlss2x)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(sgmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
sgmv_df_unscaledx <- sgmv_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(sgmv_gamlss2x, sgmv_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df_unscaledx$age_days, resid=resid(sgmv_gamlss2x), n.inter=4)
cent_cdf(sgmv_gamlss2x, sgmv_df_unscaledx, group = "sexMale")
```


###WMV

```{r}
wmv_gamlss2x <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
wmv_gamlss2x$formula
list_predictors(wmv_gamlss2x)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(wmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
wmv_df_unscaledx <- wmv_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(wmv_gamlss2x, wmv_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df_unscaledx$age_days, resid=resid(wmv_gamlss2x), n.inter=4)
cent_cdf(wmv_gamlss2x, wmv_df_unscaledx, group = "sexMale")
```


###CSF

```{r}
csf_gamlss2x <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
csf_gamlss2x$formula
list_predictors(csf_gamlss2x)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
csf_df_unscaledx <- csf_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(csf_gamlss2x, csf_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df_unscaledx$age_days, resid=resid(csf_gamlss2x), n.inter=4)
cent_cdf(csf_gamlss2x, csf_df_unscaledx, group = "sexMale")
```

#gamlss2 models with increased k in mu pb(logAge)

###GMV

```{r}
gmv_gamlss2k <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
gmv_gamlss2k$formula
list_predictors(gmv_gamlss2k)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2k$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
gmv_gamlss2k$fitted.specials$mu$`pb2(sexMale_x_logAge)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_fan2k <- make_centile_fan(gmv_gamlss2k, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = gmv_sim,
                             y_scale=unscale)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2k$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2k), n.inter=4)
cent_cdf(gmv_gamlss2k, gmv_df, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2k <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
sgmv_gamlss2k$formula
list_predictors(sgmv_gamlss2k)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2k$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_fan2k <- make_centile_fan(sgmv_gamlss2k, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = sgmv_sim,
                             y_scale=unscale)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2k$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sgmv centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2k), n.inter=4)
cent_cdf(sgmv_gamlss2k, sgmv_df, group = "sexMale")
```

###WMV

```{r}
wmv_gamlss2k <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
wmv_gamlss2k$formula
list_predictors(wmv_gamlss2k)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2k$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_fan2k <- make_centile_fan(wmv_gamlss2k, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = wmv_sim,
                             y_scale=unscale)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2k$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("wmv centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2k), n.inter=4)
cent_cdf(wmv_gamlss2k, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf_gamlss2k <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
csf_gamlss2k$formula
list_predictors(csf_gamlss2k)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2k$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_fan2k <- make_centile_fan(csf_gamlss2k, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = csf_sim,
                             y_scale=unscale)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2k$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("csf centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2k), n.inter=4)
cent_cdf(csf_gamlss2k, csf_df, group = "sexMale")
```



#gamlss2 models with adaptive splines

###GMV

```{r}
gmv_gamlss2ad <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
gmv_gamlss2ad$formula
list_predictors(gmv_gamlss2ad)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2ad$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
gmv_gamlss2ad$fitted.specials$mu$`s(sexMale_x_logAge)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_fan2ad <- make_centile_fan(gmv_gamlss2ad, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "darkblue",
                            sim_data_list = gmv_sim)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2ad$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2ad), n.inter=4)
cent_cdf(gmv_gamlss2ad, gmv_df, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2ad <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
sgmv_gamlss2ad$formula
list_predictors(sgmv_gamlss2ad)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2ad$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_fan2ad <- make_centile_fan(sgmv_gamlss2ad, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "navy",
                            sim_data_list = sgmv_sim)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2ad$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sgmv centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2ad), n.inter=4)
cent_cdf(sgmv_gamlss2ad, sgmv_df, group = "sexMale")
```
cent_cdf() function seems to not work with adaptive splines

###WMV

```{r}
wmv_gamlss2ad <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
wmv_gamlss2ad$formula
list_predictors(wmv_gamlss2ad)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2ad$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_fan2ad <- make_centile_fan(wmv_gamlss2ad, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "navy",
                            sim_data_list = wmv_sim)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2ad$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("wmv centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2ad), n.inter=4)
cent_cdf(wmv_gamlss2ad, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf_gamlss2ad <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
csf_gamlss2ad$formula
list_predictors(csf_gamlss2ad)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2ad$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_fan2ad <- make_centile_fan(csf_gamlss2ad, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "navy",
                            sim_data_list = csf_sim)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2ad$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("csf centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2ad), n.inter=4)
cent_cdf(csf_gamlss2ad, csf_df, group = "sexMale")
```



#gamlss2 models with age-by-sex factor splines

###GMV

```{r}
gmv_gamlss2fac <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
gmv_gamlss2fac$formula
list_predictors(gmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2fac$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
gmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_df <- gmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(gmv_gamlss2fac, gmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(gmv_gamlss, gmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2fac), n.inter=4)
cent_cdf(gmv_gamlss2fac, gmv_df, group = "sex")
```

###sGMV

```{r}
sgmv_gamlss2fac <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
sgmv_gamlss2fac$formula
list_predictors(sgmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2fac$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
sgmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_df <- sgmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(sgmv_gamlss2fac, sgmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(sgmv_gamlss, sgmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2fac), n.inter=4)
cent_cdf(sgmv_gamlss2fac, sgmv_df, group = "sex")
```
cent_cdf() function seems to not work with adaptive splines

###WMV

```{r}
wmv_gamlss2fac <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
wmv_gamlss2fac$formula
list_predictors(wmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2fac$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
wmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_df <- wmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(wmv_gamlss2fac, wmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(wmv_gamlss, wmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2fac), n.inter=4)
cent_cdf(wmv_gamlss2fac, wmv_df, group = "sex")
```

###CSF

```{r}
csf_gamlss2fac <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
csf_gamlss2fac$formula
list_predictors(csf_gamlss2fac)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2fac$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
csf_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_df <- csf_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(csf_gamlss2fac, csf_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(csf_gamlss, csf_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2fac), n.inter=4)
cent_cdf(csf_gamlss2fac, csf_df, group = "sex")
```

#gamlss2 models with age-by-sex factor splines

###GMV

```{r}
gmv_gamlss2fac <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
gmv_gamlss2fac$formula
list_predictors(gmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2fac$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
gmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
make_centile_fan(gmv_gamlss2fac, gmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(gmv_gamlss, gmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2fac), n.inter=4)
cent_cdf(gmv_gamlss2fac, gmv_df, group = "sex")
```

###sGMV

```{r}
sgmv_gamlss2fac <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
sgmv_gamlss2fac$formula
list_predictors(sgmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2fac$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
sgmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_df <- sgmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(sgmv_gamlss2fac, sgmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(sgmv_gamlss, sgmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2fac), n.inter=4)
cent_cdf(sgmv_gamlss2fac, sgmv_df, group = "sex")
```
cent_cdf() function seems to not work with adaptive splines

###WMV

```{r}
wmv_gamlss2fac <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
wmv_gamlss2fac$formula
list_predictors(wmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2fac$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
wmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_df <- wmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(wmv_gamlss2fac, wmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(wmv_gamlss, wmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2fac), n.inter=4)
cent_cdf(wmv_gamlss2fac, wmv_df, group = "sex")
```

###CSF

```{r}
csf_gamlss2fac <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
csf_gamlss2fac$formula
list_predictors(csf_gamlss2fac)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2fac$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
csf_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_df <- csf_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(csf_gamlss2fac, csf_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(csf_gamlss, csf_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2fac), n.inter=4)
cent_cdf(csf_gamlss2fac, csf_df, group = "sex")
```

