---
title: "Spline Testing"
output: html_document
date: "2024-12-18"
---


```{r warning=FALSE}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, gamlss, gamlssTools, gamlss.add, mgcv, gamlss2)

source("code/gamlss_fit_funs.R")
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
# install.packages("devtools")
devtools::install_github("BGDlab/gamlssTools", force=TRUE)

library(gamlssTools)

devtools::install_github("deepankardatta/blandr")
library(blandr)
```

##Get gamlss() formula
GMV
load standard gamlss brain chart

```{r}
brainchart_path <- "braincharts/global_vols_totalFALSE_logPhenoTRUE_logAgeTRUE_pbmods/model_objs/"
gmv_gamlss <- readRDS(paste0(brainchart_path, "GMV_brainchart.rds"))
gmv_gamlss$mu.formula
gmv_gamlss$sigma.formula
gmv_gamlss$nu.formula
```

pass to gamlss2 using new pb() defaults and mgcv random - no longer letting me fit with `random()`?

```{r}
gmv_formula <- formula(GMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                1)

gmv_formula_fix <- formula(GMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                1)
```

WMV

```{r}
wmv_gamlss <- readRDS(paste0(brainchart_path, "WMV_brainchart.rds"))
wmv_gamlss$mu.formula
wmv_gamlss$sigma.formula
wmv_gamlss$nu.formula
```

```{r}
wmv_formula <- formula(WMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

wmv_formula_fix <- formula(WMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(logAge_days))
```

CSF

```{r}
csf_gamlss <- readRDS(paste0(brainchart_path, "CSF_brainchart.rds"))
csf_gamlss$mu.formula
csf_gamlss$sigma.formula
csf_gamlss$nu.formula
```

```{r}
csf_formula <- formula(CSF ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

csf_formula_fix <- formula(CSF ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(logAge_days))
```

sGMV

```{r}
sgmv_gamlss <- readRDS(paste0(brainchart_path, "sGMV_brainchart.rds"))
sgmv_gamlss$mu.formula
sgmv_gamlss$sigma.formula
sgmv_gamlss$nu.formula
```

```{r}
sgmv_formula <- formula(sGMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

sgmv_formula_fix <- formula(sGMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(logAge_days))
```

#Write Config File

```{r. eval=FALSE}
config <- data.frame(ct = seq(1:8),
                      pheno = c("GMV", "WMV", "CSF", "sGMV", "GMV", "WMV", "CSF", "sGMV"),
                     form = c(as.character(gmv_formula)[3], 
                              as.character(wmv_formula)[3], 
                              as.character(csf_formula)[3], 
                              as.character(sgmv_formula)[3], 
                              as.character(gmv_formula_fix)[3], 
                              as.character(wmv_formula_fix)[3], 
                              as.character(csf_formula_fix)[3],
                              as.character(sgmv_formula_fix)[3]),
                     rand = c("rand", "rand", "rand", "rand", "fixed", "fixed", "fixed", "fixed"))
stopifnot(nrow(config) == 8)
write.table(config, file = "code/config_files/gamlss2_test_config.txt", 
            append = TRUE, sep = "\t", row.names=FALSE, col.names=FALSE, quote=c(3))
```
#Read in gamlss2() mods

```{r}
#models with random effects for site
gmv_gamlss2 <- readRDS("GMV_gamlss2_testmod_rand.rds")
wmv_gamlss2 <- readRDS("WMV_gamlss2_testmod_rand.rds")
sgmv_gamlss2 <- readRDS("sGMV_gamlss2_testmod_rand.rds")
csf_gamlss2 <- readRDS("CSF_gamlss2_testmod_rand.rds")

#models with fixed effects for site
gmv_gamlss2f <- readRDS("GMV_gamlss2_testmod_fixed.rds")
wmv_gamlss2f <- readRDS("WMV_gamlss2_testmod_fixed.rds")
sgmv_gamlss2f <- readRDS("sGMV_gamlss2_testmod_fixed.rds")
csf_gamlss2f <- readRDS("CSF_gamlss2_testmod_fixed.rds")

#unscaled models -random 
gmv_gamlss2_un <- readRDS("GMV_gamlss2_testmod_rand_unscaled.rds")
wmv_gamlss2_un <- readRDS("WMV_gamlss2_testmod_rand_unscaled.rds")
sgmv_gamlss2_un <- readRDS("sGMV_gamlss2_testmod_rand_unscaled.rds")
csf_gamlss2_un <- readRDS("CSF_gamlss2_testmod_rand_unscaled.rds")
```


#Compare Centiles

###GMV

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2$df
gmv_gamlss2f$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2$fitted.specials$mu$`pb2(logAge_days)`$edf
gmv_gamlss2f$fitted.specials$mu$`pb2(logAge_days)`$edf
```

```{r}
#prep dataframes
gmv_df <- fread("data/pheno_dfs_totalFALSE/GMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
df <- gmv_df
gmv_gamlss$call$data <- "df"
gmv_gamlss$call$family <- gmv_gamlss$family[[1]]

#make sure models fit correctly
stopifnot(setequal(list_predictors(gmv_gamlss),list_predictors(gmv_gamlss2)))
stopifnot(setequal(list_predictors(gmv_gamlss),list_predictors(gmv_gamlss2f)))
```

fit centile fan plots
```{r}
gmv_sim <- sim_data(gmv_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

gmv_fan1 <- make_centile_fan(gmv_gamlss, df, "logAge_days", "sexMale",
                             get_peaks = FALSE,
                             x_axis="log_lifespan_fetal",
                            show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = gmv_sim,
                             y_scale=unscale)
gmv_fan2 <- make_centile_fan(gmv_gamlss2, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "lightblue",
                            sim_data_list = gmv_sim,
                             y_scale=unscale)

gmv_fan3 <- make_centile_fan(gmv_gamlss2f, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "orange",
                            sim_data_list = gmv_sim,
                             y_scale=unscale)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2$layers, gmv_fan3$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 (blue) vs gamlss2 with fixed effects (orange) by sex")
```

troubleshooting
```{r}
centile_dfs <- centile_predict(gamlssModel = gmv_gamlss2f, 
                               sim_df_list = gmv_sim, 
                               x_var = "logAge_days",
                               df = gmv_df)

```

```{r}
gmv_df2 <- gmv_df %>%
  mutate(GMV = unscale(GMV))
 make_centile_fan(gmv_gamlss2_un, gmv_df2, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend")
```

```{r}
cent_cdf(gmv_gamlss, df)
cent_cdf(gmv_gamlss2, df)
cent_cdf(gmv_gamlss2f, df)
```

how much do a given subjects' centiles change?

```{r}
gmv_gamlss_scores <- pred_og_centile(gmv_gamlss, gmv_df, get.std.scores = TRUE)
colnames(gmv_gamlss_scores) <- paste(colnames(gmv_gamlss_scores), "gamlss", sep = "_")
gmv_gamlss2_scores <- pred_og_centile(gmv_gamlss2, gmv_df, get.std.scores = TRUE)
colnames(gmv_gamlss2_scores) <- paste(colnames(gmv_gamlss2_scores), "gamlss2", sep = "_")

gmv_scores_df <- Reduce(cbind, list(gmv_df, gmv_gamlss_scores, gmv_gamlss2_scores))
stopifnot(sum(is.na(gmv_scores_df)) == 0)
```
visualize

```{r}
blandr.draw(gmv_scores_df$centile_gamlss, gmv_scores_df$centile_gamlss2)
blandr.draw(gmv_scores_df$std_score_gamlss, gmv_scores_df$std_score_gamlss2)

ggplot(gmv_scores_df) +
  geom_abline() +
  geom_point(aes(x=centile_gamlss, y=centile_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2")

ggplot(gmv_scores_df) +
  geom_abline() +
  geom_point(aes(x=std_score_gamlss, y=std_score_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2 z-scores")
```

### WMV
```{r}
wmv_gamlss$call$data <- "df"
wmv_gamlss$call$family <- wmv_gamlss$family[[1]]
stopifnot(setequal(list_predictors(wmv_gamlss),list_predictors(wmv_gamlss2)))
```

```{r}
wmv_gamlss2$formula
```

```{r}
wmv_df <- fread("data/pheno_dfs_totalFALSE/WMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
wmv_sim <- sim_data(wmv_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

wmv_fan1 <- make_centile_fan(wmv_gamlss, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = wmv_sim,
                             x_axis="log_lifespan_fetal",
                             y_scale=unscale)
wmv_fan2 <- make_centile_fan(wmv_gamlss2, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "lightblue",
                             x_axis = "log_lifespan_fetal",
                            sim_data_list = wmv_sim,
                             y_scale=unscale)
wmv_fan3 <- make_centile_fan(wmv_gamlss2f, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "orange",
                             x_axis = "log_lifespan_fetal",
                            sim_data_list = wmv_sim,
                             y_scale=unscale)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2$layers, wmv_fan3$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("WMV centiles from gamlss (green) vs \n gamlss2 (blue) and gamlss2 with fixed effects (orange) by sex")

wmv_fan1 + facet_wrap(~as.factor(sexMale)) + theme_bw()
wmv_fan2 + facet_wrap(~as.factor(sexMale)) + theme_bw()
wmv_fan3 + facet_wrap(~as.factor(sexMale)) + theme_bw()
```

```{r}
wmv_df2 <- wmv_df %>%
  mutate(WMV = unscale(WMV))
 make_centile_fan(wmv_gamlss2_un, wmv_df2, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend")
```

how much do a given subjects' centiles change?

```{r}
wmv_gamlss_scores <- pred_og_centile(wmv_gamlss, wmv_df, get.std.scores = TRUE)
colnames(wmv_gamlss_scores) <- paste(colnames(wmv_gamlss_scores), "gamlss", sep = "_")
wmv_gamlss2_scores <- pred_og_centile(wmv_gamlss2, wmv_df, get.std.scores = TRUE)
colnames(wmv_gamlss2_scores) <- paste(colnames(wmv_gamlss2_scores), "gamlss2", sep = "_")

wmv_scores_df <- Reduce(cbind, list(wmv_df, wmv_gamlss_scores, wmv_gamlss2_scores))
stopifnot(sum(is.na(wmv_scores_df)) == 0)
```
visualize

```{r}
blandr.draw(wmv_scores_df$centile_gamlss, wmv_scores_df$centile_gamlss2)
blandr.draw(wmv_scores_df$std_score_gamlss, wmv_scores_df$std_score_gamlss2)

ggplot(wmv_scores_df) +
  geom_abline() +
  geom_point(aes(x=centile_gamlss, y=centile_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2")

ggplot(wmv_scores_df) +
  geom_abline() +
  geom_point(aes(x=std_score_gamlss, y=std_score_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2 z-scores")
```

###sGMV
```{r}
sgmv_gamlss$call$data <- "df"
sgmv_gamlss$call$family <- sgmv_gamlss$family[[1]]

stopifnot(setequal(list_predictors(sgmv_gamlss),list_predictors(sgmv_gamlss2)))
```

```{r}
#read data ad plot
sgmv_df <- fread("data/pheno_dfs_totalFALSE/sGMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
sgmv_sim <- sim_data(sgmv_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

sgmv_fan1 <- make_centile_fan(sgmv_gamlss, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = sgmv_sim,
                             x_axis="log_lifespan_fetal",
                             y_scale=unscale)
sgmv_fan2 <- make_centile_fan(sgmv_gamlss2, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "lightblue",
                            sim_data_list = sgmv_sim,
                            x_axis="log_lifespan_fetal",
                             y_scale=unscale)

sgmv_fan3 <- make_centile_fan(sgmv_gamlss2f, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "orange",
                             x_axis = "log_lifespan_fetal",
                            sim_data_list = sgmv_sim,
                             y_scale=unscale)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2$layers, sgmv_fan3$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sGMV centiles from gamlss (green) vs gamlss2 (blue) by sex")
```
look more at gamlss2 model

```{r}
cent_cdf(sgmv_gamlss2, sgmv_df, group="sexMale")

make_centile_fan(sgmv_gamlss2, sgmv_df, "logAge_days", "sexMale", 
                             label_centiles = "legend",
                             sim_data_list = sgmv_sim,
                 remove_point_effect = c("fs_version_GM"), #can't remove point effects from random effect with s(re)
                 x_axis = "log_lifespan_fetal",
                             y_scale=unscale) +
  theme_bw()
```

```{r}
sgmv_df2 <- sgmv_df %>%
  mutate(sGMV = unscale(sGMV))

make_centile_fan(sgmv_gamlss2_un, sgmv_df2, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend")
```

how much do a given subjects' centiles change?

```{r}
sgmv_gamlss_scores <- pred_og_centile(sgmv_gamlss, sgmv_df, get.std.scores = TRUE)
colnames(sgmv_gamlss_scores) <- paste(colnames(sgmv_gamlss_scores), "gamlss", sep = "_")
sgmv_gamlss2_scores <- pred_og_centile(sgmv_gamlss2, sgmv_df, get.std.scores = TRUE)
colnames(sgmv_gamlss2_scores) <- paste(colnames(sgmv_gamlss2_scores), "gamlss2", sep = "_")

sgmv_scores_df <- Reduce(cbind, list(sgmv_df, sgmv_gamlss_scores, sgmv_gamlss2_scores))
stopifnot(sum(is.na(sgmv_scores_df)) == 0)
```
visualize

```{r}
blandr.draw(sgmv_scores_df$centile_gamlss, sgmv_scores_df$centile_gamlss2)
blandr.draw(sgmv_scores_df$std_score_gamlss, sgmv_scores_df$std_score_gamlss2)

ggplot(sgmv_scores_df) +
  geom_abline() +
  geom_point(aes(x=centile_gamlss, y=centile_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2")

ggplot(sgmv_scores_df) +
  geom_abline() +
  geom_point(aes(x=std_score_gamlss, y=std_score_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2 z-scores")
```


###CSF

```{r}
csf_gamlss$call$data <- "df"
csf_gamlss$call$family <- csf_gamlss$family[[1]]

stopifnot(setequal(list_predictors(csf_gamlss),list_predictors(csf_gamlss2)))
```


```{r}
csf_df <- fread("data/pheno_dfs_totalFALSE/CSF_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
csf_sim <- sim_data(csf_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

csf_fan1 <- make_centile_fan(csf_gamlss, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = csf_sim,
                             x_axis = "log_lifespan_fetal",
                             y_scale=unscale)
csf_fan2 <- make_centile_fan(csf_gamlss2, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "lightblue",
                            sim_data_list = csf_sim,
                            x_axis = "log_lifespan_fetal",
                             y_scale=unscale)
csf_fan3 <- make_centile_fan(csf_gamlss2f, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                             label_centiles = "legend", color_manual = "orange",
                             x_axis = "log_lifespan_fetal",
                            sim_data_list = sgmv_sim,
                             y_scale=unscale)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2$layers, csf_fan3$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("CSF centiles from gamlss (green) vs \ngamlss2 (blue) vs gamlss2 with fixed site effects (orange) by sex")
```
```{r}
csf_df2 <- csf_df %>%
  mutate(CSF = unscale(CSF))

make_centile_fan(csf_gamlss2_un, csf_df2, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "label")
```

how much do a given subjects' centiles change?

```{r}
csf_gamlss_scores <- pred_og_centile(csf_gamlss, csf_df, get.std.scores = TRUE)
colnames(csf_gamlss_scores) <- paste(colnames(csf_gamlss_scores), "gamlss", sep = "_")
csf_gamlss2_scores <- pred_og_centile(csf_gamlss2, csf_df, get.std.scores = TRUE)
colnames(csf_gamlss2_scores) <- paste(colnames(csf_gamlss2_scores), "gamlss2", sep = "_")

csf_scores_df <- Reduce(cbind, list(csf_df, csf_gamlss_scores, csf_gamlss2_scores))
stopifnot(sum(is.na(csf_scores_df)) == 0)
```
visualize

```{r}
blandr.draw(csf_scores_df$centile_gamlss, csf_scores_df$centile_gamlss2)
blandr.draw(csf_scores_df$std_score_gamlss, csf_scores_df$std_score_gamlss2)

ggplot(csf_scores_df) +
  geom_abline() +
  geom_point(aes(x=centile_gamlss, y=centile_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2")

ggplot(csf_scores_df) +
  geom_abline() +
  geom_point(aes(x=std_score_gamlss, y=std_score_gamlss2, color=logAge_days), alpha=.6) +
  ggtitle("gamlss vs gamlss2 z-scores")
```

