---
title: "Spline Testing"
output: html_document
date: "2024-12-18"
---


```{r warning=FALSE}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, gamlss, gamlssTools, gamlss.add, mgcv, gamlss2)
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
source("code/gamlss_fit_funs.R")

# install.packages("devtools")
devtools::install_github("BGDlab/gamlssTools", force=TRUE)

library(gamlssTools)

devtools::install_github("deepankardatta/blandr")
library(blandr)
```

##Get gamlss() formula
GMV
load standard gamlss brain chart

```{r}
brainchart_path <- "braincharts/global_vols_totalFALSE_logPhenoTRUE_logAgeTRUE_pbmods/model_objs/"
gmv_gamlss <- readRDS(paste0(brainchart_path, "GMV_brainchart.rds"))
gmv_gamlss$mu.formula
gmv_gamlss$sigma.formula
gmv_gamlss$nu.formula
```

pass to gamlss2 using new pb() defaults and mgcv random - no longer letting me fit with `random()`?

```{r}
#original formula converted to gamlss2
gmv_formula <- formula(GMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                1)

#fixed effects for site
gmv_formula_fix <- formula(GMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                1)

#increase k
gmv_formula_k30 <- formula(GMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3, k=30) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                1)

#adaptive splines
gmv_formula_ad <- formula(GMV ~ s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') | 
                1)

#age_days (not log scaled)
gmv_formula_day <- formula(GMV ~ pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') | 
                1)

#separate smooths for male and female
gmv_formula_by <- formula(GMV ~ s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') + fs_version_GM | 
                s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') | 
                1)
```

WMV

```{r}
wmv_gamlss <- readRDS(paste0(brainchart_path, "WMV_brainchart.rds"))
wmv_gamlss$mu.formula
wmv_gamlss$sigma.formula
wmv_gamlss$nu.formula
```

```{r}
wmv_formula <- formula(WMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

wmv_formula_fix <- formula(WMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(logAge_days))

wmv_formula_k30 <- formula(WMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3, k=30) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

wmv_formula_ad <- formula(WMV ~ s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') | 
                s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

wmv_formula_day<- formula(WMV ~ pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(age_days))

wmv_formula_by <- formula(WMV ~ s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') | 
                s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') | 
                pb(logAge_days))
```

CSF

```{r}
csf_gamlss <- readRDS(paste0(brainchart_path, "CSF_brainchart.rds"))
csf_gamlss$mu.formula
csf_gamlss$sigma.formula
csf_gamlss$nu.formula
```

```{r}
csf_formula <- formula(CSF ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

csf_formula_fix <- formula(CSF ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(logAge_days))

csf_formula_k30 <- formula(CSF ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3, k=30) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

csf_formula_ad <- formula(CSF ~ s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') | 
                s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

csf_formula_day<- formula(CSF ~ pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(age_days))

csf_formula_by <- formula(CSF ~ s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') | 
                s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') | 
                pb(logAge_days))
```

sGMV

```{r}
sgmv_gamlss <- readRDS(paste0(brainchart_path, "sGMV_brainchart.rds"))
sgmv_gamlss$mu.formula
sgmv_gamlss$sigma.formula
sgmv_gamlss$nu.formula
```

```{r}
sgmv_formula <- formula(sGMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM| 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

sgmv_formula_fix <- formula(sGMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site + fs_version_GM| 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + study_site | 
                pb(logAge_days))

sgmv_formula_k30 <- formula(sGMV ~ pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3, k=30) + sexMale + s(study_site, bs='re') + fs_version_GM| 
                pb(sexMale_x_logAge, m=3) + pb(logAge_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

sgmv_formula_ad <- formula(sGMV ~ s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                s(sexMale_x_logAge, bs="ad", k=20, m=3) + s(logAge_days, bs="ad", k=20, m=3) + sexMale + s(study_site, bs='re') | 
                pb(logAge_days))

sgmv_formula_day<- formula(sGMV ~ pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') + fs_version_GM | 
                pb(sexMale_x_age, m=3) + pb(age_days, m=3) + sexMale + s(study_site, bs='re') | 
                pb(age_days))

sgmv_formula_by <- formula(sGMV ~ s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') + fs_version_GM | 
                s(logAge_days, by=sex, bs="ps", k=20, m=3) + sex + s(study_site, bs='re') | 
                pb(logAge_days))
```

#Write Config File

```{r}
clean_formula <- function(x){
  as.character(x)[3]
}
```

```{r, eval=FALSE}
pheno_list <- rep(c("GMV", "WMV", "CSF", "sGMV"), 7)
config <- data.frame(ct = seq(1:(7*4)),
                    pheno = c(pheno_list),
                     form = c(clean_formula(gmv_formula), #regular formula
                              clean_formula(wmv_formula), 
                              clean_formula(csf_formula), 
                              clean_formula(sgmv_formula), 
                              clean_formula(gmv_formula_fix), #fixed site effects
                              clean_formula(wmv_formula_fix), 
                              clean_formula(csf_formula_fix),
                              clean_formula(sgmv_formula_fix),
                              clean_formula(gmv_formula), #un-log-scale pheno
                              clean_formula(wmv_formula), 
                              clean_formula(csf_formula), 
                              clean_formula(sgmv_formula),
                              clean_formula(gmv_formula_day), #un-log-scale age
                              clean_formula(wmv_formula_day), 
                              clean_formula(csf_formula_day), 
                              clean_formula(sgmv_formula_day),
                              clean_formula(gmv_formula_k30), #increase k
                              clean_formula(wmv_formula_k30), 
                              clean_formula(csf_formula_k30), 
                              clean_formula(sgmv_formula_k30),
                              clean_formula(gmv_formula_ad), #adaptive splines on unscaled pheno
                              clean_formula(wmv_formula_ad), 
                              clean_formula(csf_formula_ad), 
                              clean_formula(sgmv_formula_ad),
                              clean_formula(gmv_formula_by), #age by sex
                              clean_formula(wmv_formula_by), 
                              clean_formula(csf_formula_by), 
                              clean_formula(sgmv_formula_by)),
                     name = c("rand", "rand", "rand", "rand", 
                              "fixed", "fixed", "fixed", "fixed",
                              "rand", "rand", "rand", "rand",
                              "rand", "rand", "rand", "rand",
                              "k30", "k30", "k30", "k30",
                              "adapt", "adapt", "adapt", "adapt",
                              "by", "by", "by", "by"),
                    log_pheno= c(TRUE, TRUE, TRUE, TRUE,
                                 TRUE, TRUE, TRUE, TRUE,
                                 FALSE, FALSE, FALSE, FALSE,
                                 TRUE, TRUE, TRUE, TRUE,
                                 TRUE, TRUE, TRUE, TRUE,
                                 FALSE, FALSE, FALSE, FALSE,
                                 TRUE, TRUE, TRUE, TRUE),
                    log_age = c(TRUE, TRUE, TRUE, TRUE,
                                TRUE, TRUE, TRUE, TRUE,
                                TRUE, TRUE, TRUE, TRUE,
                                FALSE, FALSE, FALSE, FALSE,
                                TRUE, TRUE, TRUE, TRUE,
                                TRUE, TRUE, TRUE, TRUE,
                                TRUE, TRUE, TRUE, TRUE))
stopifnot(nrow(config) == 7*4)
write.table(config, file = "code/config_files/gamlss2_test_config.txt", 
            append = TRUE, sep = "\t", row.names=FALSE, col.names=FALSE, quote=c(3))
```
 
#Prep Dataframes

```{r}
#GMV
gmv_df <- fread("data/pheno_dfs_totalFALSE/GMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
gmv_gamlss$call$data <- "df"
gmv_gamlss$call$family <- gmv_gamlss$family[[1]]
gmv_sim <- sim_data(gmv_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

#sGMV
sgmv_df <- fread("data/pheno_dfs_totalFALSE/sGMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
sgmv_gamlss$call$data <- "df"
sgmv_gamlss$call$family <- sgmv_gamlss$family[[1]]
sgmv_sim <- sim_data(sgmv_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

#WMV
wmv_df <- fread("data/pheno_dfs_totalFALSE/WMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
wmv_gamlss$call$data <- "df"
wmv_gamlss$call$family <- wmv_gamlss$family[[1]]
wmv_sim <- sim_data(wmv_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

#CSF
csf_df <- fread("data/pheno_dfs_totalFALSE/CSF_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE)
csf_gamlss$call$data <- "df"
csf_gamlss$call$family <- csf_gamlss$family[[1]]
csf_sim <- sim_data(csf_df, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
```


*testing quick and dirty "harmonization" using `resid_data()`

```{r}
gmv_resid <- resid_data(gmv_gamlss, gmv_df, gmv_df, "study_site")
fwrite(gmv_resid, "data/test_GMV_resid.csv")

sgmv_resid <- resid_data(sgmv_gamlss, sgmv_df, sgmv_df, "study_site")
fwrite(sgmv_resid, "data/test_sGMV_resid.csv")

wmv_resid <- resid_data(wmv_gamlss, wmv_df, wmv_df, "study_site")
fwrite(wmv_resid, "data/test_WMV_resid.csv")

csf_resid <- resid_data(csf_gamlss, csf_df, csf_df, "study_site")
fwrite(csf_resid, "data/test_CSF_resid.csv")
```

**All models have smooths penalizing 3rd order (m=3), random site effects and fit on log-scaled age and phenotypes unless otherwise specified**

#Basic gamlss2 models

###GMV

```{r}
gmv_gamlss2 <- readRDS("gamlss2_test_mods/GMV_gamlss2_testmod_rand.rds")
gmv_gamlss2$formula
list_predictors(gmv_gamlss2)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2$fitted.specials$mu$`pb2(logAge_days)`$edf
gmv_gamlss2$fitted.specials$mu$`pb2(sexMale_x_logAge)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_fan1 <- make_centile_fan(gmv_gamlss, gmv_df, "logAge_days", "sexMale",
                             get_peaks = FALSE,
                             x_axis="log_lifespan_fetal",
                            show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = gmv_sim,
                             y_scale=unscale)
gmv_fan2 <- make_centile_fan(gmv_gamlss2, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "blue",
                            sim_data_list = gmv_sim,
                             y_scale=unscale)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 (blue)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2), n.inter=4)
cent_cdf(gmv_gamlss2, gmv_df, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2 <- readRDS("gamlss2_test_mods/sGMV_gamlss2_testmod_rand.rds")
sgmv_gamlss2$formula
list_predictors(sgmv_gamlss2)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_fan1 <- make_centile_fan(sgmv_gamlss, sgmv_df, "logAge_days", "sexMale",
                             get_peaks = FALSE,
                             x_axis="log_lifespan_fetal",
                            show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = sgmv_sim,
                             y_scale=unscale)
sgmv_fan2 <- make_centile_fan(sgmv_gamlss2, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "blue",
                            sim_data_list = sgmv_sim,
                             y_scale=unscale)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sgmv centiles from gamlss (green) vs gamlss2 (blue)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2), n.inter=4)
cent_cdf(sgmv_gamlss2, sgmv_df, group = "sexMale")
```

###WMV

```{r}
wmv_gamlss2 <- readRDS("gamlss2_test_mods/WMV_gamlss2_testmod_rand.rds")
wmv_gamlss2$formula
list_predictors(wmv_gamlss2)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_fan1 <- make_centile_fan(wmv_gamlss, wmv_df, "logAge_days", "sexMale",
                             get_peaks = FALSE,
                             x_axis="log_lifespan_fetal",
                            show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = wmv_sim,
                             y_scale=unscale)
wmv_fan2 <- make_centile_fan(wmv_gamlss2, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "blue",
                            sim_data_list = wmv_sim,
                             y_scale=unscale)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("wmv centiles from gamlss (green) vs gamlss2 (blue)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2), n.inter=4)
cent_cdf(wmv_gamlss2, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf_gamlss2 <- readRDS("gamlss2_test_mods/CSF_gamlss2_testmod_rand.rds")
csf_gamlss2$formula
list_predictors(csf_gamlss2)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_fan1 <- make_centile_fan(csf_gamlss, csf_df, "logAge_days", "sexMale",
                             get_peaks = FALSE,
                             x_axis="log_lifespan_fetal",
                            show_points = FALSE, 
                             label_centiles = "legend", color_manual = "darkgreen",
                             sim_data_list = csf_sim,
                             y_scale=unscale)
csf_fan2 <- make_centile_fan(csf_gamlss2, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "blue",
                            sim_data_list = csf_sim,
                             y_scale=unscale)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("csf centiles from gamlss (green) vs gamlss2 (blue)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2), n.inter=4)
cent_cdf(csf_gamlss2, csf_df, group = "sexMale")
```



#gamlss2 with fixed site effects

###GMV

```{r}
gmv_gamlss2f <- readRDS("gamlss2_test_mods/GMV_gamlss2_testmod_fixed.rds")
gmv_gamlss2f$formula
list_predictors(gmv_gamlss2f)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2f$df

#look at smooth age effect in mu specifically
stopifnot(gmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2f$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_fan2f <- make_centile_fan(gmv_gamlss2f, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "orange",
                            sim_data_list = gmv_sim,
                             y_scale=unscale)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2f$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs \ngamlss2 with fixed site effects (orange)")
```
worm plots, check actual centiles

```{r, eval=FALSE}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2f), n.inter=4)
cent_cdf(gmv_gamlss2f, gmv_df, group = "sexMale")
```

**fixed effect model is messed up, returning negatives**

###sGMV

```{r}
sgmv_gamlss2f <- readRDS("gamlss2_test_mods/sGMV_gamlss2_testmod_fixed.rds")
sgmv_gamlss2f$formula
list_predictors(sgmv_gamlss2f)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2f$df

#look at smooth age effect in mu specifically
stopifnot(sgmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2f$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_fan2f <- make_centile_fan(sgmv_gamlss2f, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "orange",
                            sim_data_list = sgmv_sim,
                             y_scale=unscale)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2f$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sgmv centiles from gamlss (green) vs gamlss2 \nwith fixed site effects (orange)")
```
worm plots, check actual centiles

```{r, eval=FALSE}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2f), n.inter=4)
cent_cdf(sgmv_gamlss2f, sgmv_df, group = "sexMale")
```
**also fails**


###WMV

```{r}
wmv_gamlss2f <- readRDS("gamlss2_test_mods/WMV_gamlss2_testmod_fixed.rds")
wmv_gamlss2f$formula
list_predictors(wmv_gamlss2f)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2f$df

#look at smooth age effect in mu specifically
stopifnot(wmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2f$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_fan2f <- make_centile_fan(wmv_gamlss2f, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "orange",
                            sim_data_list = wmv_sim,
                             y_scale=unscale)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2f$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("wmv centiles from gamlss (green) vs gamlss2 \nwith fixed site (orange)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2f), n.inter=4)
cent_cdf(wmv_gamlss2f, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf_gamlss2f <- readRDS("gamlss2_test_mods/CSF_gamlss2_testmod_fixed.rds")
csf_gamlss2f$formula
list_predictors(csf_gamlss2f)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_fan2f <- make_centile_fan(csf_gamlss2f, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "orange",
                            sim_data_list = csf_sim,
                             y_scale=unscale)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2f$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("csf centiles from gamlss (green) vs gamlss2 \nwith fixed site (orange)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2f), n.inter=4)
cent_cdf(csf_gamlss2f, csf_df, group = "sexMale")
```



#gamlss2 with non-log-scaled y

have-been fitting models to log(pheno) to try to ensure that low, early-life values are captured. undoing this scaling and fitting to see if that helps with excessive wigginess

###GMV

```{r}
gmv_gamlss2y <- readRDS("gamlss2_test_mods/GMV_gamlss2_testmod_rand_unscaled.rds")
gmv_gamlss2y$formula
list_predictors(gmv_gamlss2y)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2y$df

#look at smooth age effect in mu specifically
stopifnot(gmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2y$fitted.specials$mu$`pb2(logAge_days)`$edf
```

looks like it's using way fewer dfs

Centile fan plots

fit centile fan plots
```{r}
gmv_df_unscaled <- gmv_df %>%
  mutate(GMV = unscale(GMV))
gmv_fan2y <- make_centile_fan(gmv_gamlss2y, gmv_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "purple",
                            sim_data_list = gmv_sim)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2y$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 \nfit on non-log-scaled pheno (purple)")
```
compare against data points

```{r}
make_centile_fan(gmv_gamlss2y, gmv_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend",
                            sim_data_list = gmv_sim)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df_unscaled$logAge_days, resid=resid(gmv_gamlss2y), n.inter=4)
cent_cdf(gmv_gamlss2y, gmv_df_unscaled, group = "sexMale")
```

very bad worm plots

###sGMV

```{r}
sgmv_gamlss2y <- readRDS("gamlss2_test_mods/sGMV_gamlss2_testmod_rand_unscaled.rds")
sgmv_gamlss2y$formula
list_predictors(sgmv_gamlss2y)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2y$df

#look at smooth age effect in mu specifically
stopifnot(sgmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2y$fitted.specials$mu$`pb2(logAge_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
sgmv_df_unscaled <- sgmv_df %>%
  mutate(sGMV = unscale(sGMV))
sgmv_fan2y <- make_centile_fan(sgmv_gamlss2y, sgmv_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "purple",
                            sim_data_list = sgmv_sim)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2y$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sGMV centiles from gamlss (green) vs gamlss2 \nfit on non-log-scaled pheno (purple)")
```
compare against data points

```{r}
make_centile_fan(sgmv_gamlss2y, sgmv_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend",
                            sim_data_list = sgmv_sim)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df_unscaled$logAge_days, resid=resid(sgmv_gamlss2y), n.inter=4)
cent_cdf(sgmv_gamlss2y, sgmv_df_unscaled, group = "sexMale")
```


###WMV

```{r}
wmv_gamlss2y <- readRDS("gamlss2_test_mods/WMV_gamlss2_testmod_rand_unscaled.rds")
wmv_gamlss2y$formula
list_predictors(wmv_gamlss2y)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2y$df

#look at smooth age effect in mu specifically
stopifnot(wmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2y$fitted.specials$mu$`pb2(logAge_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
wmv_df_unscaled <- wmv_df %>%
  mutate(WMV = unscale(WMV))
wmv_fan2y <- make_centile_fan(wmv_gamlss2y, wmv_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "purple",
                            sim_data_list = wmv_sim)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2y$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("WMV centiles from gamlss (green) vs gamlss2 \nfit on non-log-scaled pheno (purple)")
```
compare against data points

```{r}
make_centile_fan(wmv_gamlss2y, wmv_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend",
                            sim_data_list = wmv_sim)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df_unscaled$logAge_days, resid=resid(wmv_gamlss2y), n.inter=4)
cent_cdf(wmv_gamlss2y, wmv_df_unscaled, group = "sexMale")
```


###CSF

```{r}
csf_gamlss2y <- readRDS("gamlss2_test_mods/CSF_gamlss2_testmod_rand_unscaled.rds")
csf_gamlss2y$formula
list_predictors(csf_gamlss2y)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2y$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2y$fitted.specials$mu$`pb2(logAge_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
csf_df_unscaled <- csf_df %>%
  mutate(CSF = unscale(CSF))
csf_fan2y <- make_centile_fan(csf_gamlss2y, csf_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "purple",
                            sim_data_list = csf_sim)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2y$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("CSF centiles from gamlss (green) vs gamlss2 \nfit on non-log-scaled pheno (purple)")
```
compare against data points

```{r}
make_centile_fan(csf_gamlss2y, csf_df_unscaled, "logAge_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend",
                            sim_data_list = csf_sim)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df_unscaled$logAge_days, resid=resid(csf_gamlss2y), n.inter=4)
cent_cdf(csf_gamlss2y, csf_df_unscaled, group = "sexMale")
```



#gamlss2 with non-log-scaled age


###GMV

```{r}
gmv_gamlss2x <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
gmv_gamlss2x$formula
list_predictors(gmv_gamlss2x)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(gmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
gmv_df_unscaledx <- gmv_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(gmv_gamlss2x, gmv_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE, 
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```


worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df_unscaledx$age_days, resid=resid(gmv_gamlss2x), n.inter=4)
cent_cdf(gmv_gamlss2x, gmv_df_unscaledx, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2x <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
sgmv_gamlss2x$formula
list_predictors(sgmv_gamlss2x)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(sgmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
sgmv_df_unscaledx <- sgmv_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(sgmv_gamlss2x, sgmv_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df_unscaledx$age_days, resid=resid(sgmv_gamlss2x), n.inter=4)
cent_cdf(sgmv_gamlss2x, sgmv_df_unscaledx, group = "sexMale")
```


###WMV

```{r}
wmv_gamlss2x <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
wmv_gamlss2x$formula
list_predictors(wmv_gamlss2x)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(wmv_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
wmv_df_unscaledx <- wmv_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(wmv_gamlss2x, wmv_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df_unscaledx$age_days, resid=resid(wmv_gamlss2x), n.inter=4)
cent_cdf(wmv_gamlss2x, wmv_df_unscaledx, group = "sexMale")
```


###CSF

```{r}
csf_gamlss2x <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeFALSE_rand.rds")
csf_gamlss2x$formula
list_predictors(csf_gamlss2x)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2x$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2x$fitted.specials$mu$`pb2(age_days)`$edf
```

Centile fan plots

fit centile fan plots
```{r}
csf_df_unscaledx <- csf_df %>%
    mutate(age_days=un_log(logAge_days)) %>%
    mutate(sexMale_x_age = ifelse(sexMale == 0, 0, age_days))

make_centile_fan(csf_gamlss2x, csf_df_unscaledx, "age_days", "sexMale", 
                             get_peaks = FALSE,
                              x_axis="lifespan_fetal",
                             label_centiles = "legend",
                 y_scale=unscale)
```

worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df_unscaledx$age_days, resid=resid(csf_gamlss2x), n.inter=4)
cent_cdf(csf_gamlss2x, csf_df_unscaledx, group = "sexMale")
```

#gamlss2 models with increased k in mu pb(logAge)

###GMV

```{r}
gmv_gamlss2k <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
gmv_gamlss2k$formula
list_predictors(gmv_gamlss2k)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2k$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
gmv_gamlss2k$fitted.specials$mu$`pb2(sexMale_x_logAge)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_fan2k <- make_centile_fan(gmv_gamlss2k, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = gmv_sim,
                             y_scale=unscale)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2k$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2k), n.inter=4)
cent_cdf(gmv_gamlss2k, gmv_df, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2k <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
sgmv_gamlss2k$formula
list_predictors(sgmv_gamlss2k)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2k$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_fan2k <- make_centile_fan(sgmv_gamlss2k, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = sgmv_sim,
                             y_scale=unscale)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2k$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sgmv centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2k), n.inter=4)
cent_cdf(sgmv_gamlss2k, sgmv_df, group = "sexMale")
```

###WMV

```{r}
wmv_gamlss2k <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
wmv_gamlss2k$formula
list_predictors(wmv_gamlss2k)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2k$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_fan2k <- make_centile_fan(wmv_gamlss2k, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = wmv_sim,
                             y_scale=unscale)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2k$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("wmv centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2k), n.inter=4)
cent_cdf(wmv_gamlss2k, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf_gamlss2k <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeTRUE_k30.rds")
csf_gamlss2k$formula
list_predictors(csf_gamlss2k)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2k$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2k$fitted.specials$mu$`pb2(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_fan2k <- make_centile_fan(csf_gamlss2k, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "maroon",
                            sim_data_list = csf_sim,
                             y_scale=unscale)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2k$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("csf centiles from gamlss (green) vs gamlss2 with k=30 (maroon)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2k), n.inter=4)
cent_cdf(csf_gamlss2k, csf_df, group = "sexMale")
```



#gamlss2 models with adaptive splines

###GMV

```{r}
gmv_gamlss2ad <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
gmv_gamlss2ad$formula
list_predictors(gmv_gamlss2ad)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2ad$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
gmv_gamlss2ad$fitted.specials$mu$`s(sexMale_x_logAge)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_fan2ad <- make_centile_fan(gmv_gamlss2ad, gmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "darkblue",
                            sim_data_list = gmv_sim)

gmv_plot_combined <- gmv_fan1
gmv_plot_combined$layers <- c(gmv_fan1$layers, gmv_fan2ad$layers)

gmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("GMV centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2ad), n.inter=4)
cent_cdf(gmv_gamlss2ad, gmv_df, group = "sexMale")
```

###sGMV

```{r}
sgmv_gamlss2ad <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
sgmv_gamlss2ad$formula
list_predictors(sgmv_gamlss2ad)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2ad$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_fan2ad <- make_centile_fan(sgmv_gamlss2ad, sgmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "navy",
                            sim_data_list = sgmv_sim)

sgmv_plot_combined <- sgmv_fan1
sgmv_plot_combined$layers <- c(sgmv_fan1$layers, sgmv_fan2ad$layers)

sgmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("sgmv centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2ad), n.inter=4)
cent_cdf(sgmv_gamlss2ad, sgmv_df, group = "sexMale")
```
cent_cdf() function seems to not work with adaptive splines

###WMV

```{r}
wmv_gamlss2ad <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
wmv_gamlss2ad$formula
list_predictors(wmv_gamlss2ad)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2ad$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_fan2ad <- make_centile_fan(wmv_gamlss2ad, wmv_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "navy",
                            sim_data_list = wmv_sim)

wmv_plot_combined <- wmv_fan1
wmv_plot_combined$layers <- c(wmv_fan1$layers, wmv_fan2ad$layers)

wmv_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("wmv centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2ad), n.inter=4)
cent_cdf(wmv_gamlss2ad, wmv_df, group = "sexMale")
```

###CSF

```{r}
csf_gamlss2ad <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoFALSE_logAgeTRUE_adapt.rds")
csf_gamlss2ad$formula
list_predictors(csf_gamlss2ad)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2ad$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2ad$fitted.specials$mu$`s(logAge_days)`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_fan2ad <- make_centile_fan(csf_gamlss2ad, csf_df, "logAge_days", "sexMale", 
                             get_peaks = FALSE, show_points = FALSE, 
                              x_axis="log_lifespan_fetal",
                             label_centiles = "legend", color_manual = "navy",
                            sim_data_list = csf_sim)

csf_plot_combined <- csf_fan1
csf_plot_combined$layers <- c(csf_fan1$layers, csf_fan2ad$layers)

csf_plot_combined + facet_wrap(~as.factor(sexMale)) + theme_bw() + ggtitle("csf centiles from gamlss (green) vs gamlss2 with adaptive splines (navy)")
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2ad), n.inter=4)
cent_cdf(csf_gamlss2ad, csf_df, group = "sexMale")
```



#gamlss2 models with age-by-sex factor splines

###GMV

```{r}
gmv_gamlss2fac <- readRDS("gamlss2_test_mods/GMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
gmv_gamlss2fac$formula
list_predictors(gmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
gmv_gamlss$df.fit
gmv_gamlss2fac$df

#look at smooth age effect in mu specifically
gmv_gamlss$mu.coefSmo[[2]]$name
gmv_gamlss$mu.coefSmo[[2]]$edf
gmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
gmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
gmv_df <- gmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(gmv_gamlss2fac, gmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(gmv_gamlss, gmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=gmv_df$logAge_days, resid=resid(gmv_gamlss2fac), n.inter=4)
cent_cdf(gmv_gamlss2fac, gmv_df, group = "sex")
```

###sGMV

```{r}
sgmv_gamlss2fac <- readRDS("gamlss2_test_mods/sGMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
sgmv_gamlss2fac$formula
list_predictors(sgmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
sgmv_gamlss$df.fit
sgmv_gamlss2fac$df

#look at smooth age effect in mu specifically
sgmv_gamlss$mu.coefSmo[[2]]$name
sgmv_gamlss$mu.coefSmo[[2]]$edf
sgmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
sgmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
sgmv_df <- sgmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(sgmv_gamlss2fac, sgmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(sgmv_gamlss, sgmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=sgmv_df$logAge_days, resid=resid(sgmv_gamlss2fac), n.inter=4)
cent_cdf(sgmv_gamlss2fac, sgmv_df, group = "sex")
```
cent_cdf() function seems to not work with adaptive splines

###WMV

```{r}
wmv_gamlss2fac <- readRDS("gamlss2_test_mods/WMV_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
wmv_gamlss2fac$formula
list_predictors(wmv_gamlss2fac)
```

compare effective degrees of freedom
```{r}
wmv_gamlss$df.fit
wmv_gamlss2fac$df

#look at smooth age effect in mu specifically
wmv_gamlss$mu.coefSmo[[2]]$name
wmv_gamlss$mu.coefSmo[[2]]$edf
wmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
wmv_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
wmv_df <- wmv_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(wmv_gamlss2fac, wmv_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(wmv_gamlss, wmv_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=wmv_df$logAge_days, resid=resid(wmv_gamlss2fac), n.inter=4)
cent_cdf(wmv_gamlss2fac, wmv_df, group = "sex")
```

###CSF

```{r}
csf_gamlss2fac <- readRDS("gamlss2_test_mods/CSF_gamlss2_logPhenoTRUE_logAgeTRUE_by.rds")
csf_gamlss2fac$formula
list_predictors(csf_gamlss2fac)
```

compare effective degrees of freedom
```{r}
csf_gamlss$df.fit
csf_gamlss2fac$df

#look at smooth age effect in mu specifically
stopifnot(csf_gamlss$mu.coefSmo[[2]]$name == "logAge_days")
csf_gamlss$mu.coefSmo[[2]]$edf
csf_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexM`$edf
csf_gamlss2fac$fitted.specials$mu$`s(logAge_days):sexF`$edf
```
Centile fan plots

fit centile fan plots
```{r}
csf_df <- csf_df %>%
    mutate(sex = as.factor(ifelse(sexMale == 0, "M", "F")))

make_centile_fan(csf_gamlss2fac, csf_df, "logAge_days", "sex",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)

make_centile_fan(csf_gamlss, csf_df, "logAge_days", "sexMale",
                  get_peaks = FALSE,
                 x_axis="log_lifespan_fetal",
                 label_centiles = "legend", 
                 y_scale = unscale)
```
worm plots, check actual centiles

```{r}
wp.taki(xvar=csf_df$logAge_days, resid=resid(csf_gamlss2fac), n.inter=4)
cent_cdf(csf_gamlss2fac, csf_df, group = "sex")
```

