#Curation

Script for cleaning and preparing LBCC data. 


```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel, ggpubr)
```

load data - prelim version of LBCC 3.0, waiting on Lena to look into missing sex info
```{r}
df <- fread("/mnt/isilon/bgdlab_processing/LBCC_releases/v3_Penn/v3_0_0_penn.csv", na.strings = c("", NA))
names(df)
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$site))
sum(is.na(df$dx))
sum(is.na(df$session))
sum(is.na(df$study_session))

hist(df$MAD.Euler)
```

write out lists of phenotypes by category
```{r}
names(df)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF")
write_lines(glob_vol_list, "./pheno_lists/global_vols.txt")
saveRDS(glob_vol_list, "./pheno_lists/global_vols.rds")

#cortical gmv
gmv_list <- names(df)[grep("GM\\.", names(df))]
stopifnot(length(gmv_list) == 68)
write_lines(gmv_list, "./pheno_lists/cortical_vols.txt")
saveRDS(gmv_list, "./pheno_lists/cortical_vols.rds")

#subcortical gmv
subcort <- names(df)[grep("SUBC\\.", names(df))]
subcort_list <- subcort[-grep("Cerebellum|WM|Vent|choroid|vessel|hypointensities|CSF", subcort)] #drop WM, cerebellum, etc
subcort_list <- subcort_list[-grep("Lentiform", subcort_list)] #drop segments only in infants
subcort_list <- subcort_list[-grep("SUBC.CC", subcort_list)]
length(subcort_list)
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")

#cortical thickness
ct_list <- names(df)[grep("CT\\.", names(df))]
stopifnot(length(ct_list) == 68)
write_lines(ct_list, "./pheno_lists/cortical_thickness.txt")
saveRDS(ct_list, "./pheno_lists/cortical_thickness.rds")

#cortical SA
sa_list <- names(df)[grep("SA\\.", names(df))]
stopifnot(length(sa_list) == 68)
write_lines(sa_list, "./pheno_lists/cortical_surf.txt")
saveRDS(sa_list, "./pheno_lists/cortical_surf.rds")
```

## Data Cleaning

remove intersex subj, dummy-code sex and sex*age, add totals for CT and SA

```{r}
df_filt <- df %>%
  dplyr::filter(sex != "intersex") %>%
  mutate(sexMale = ifelse(sex=="male", 1, 0),
         sexMale_x_age = ifelse(sex=="male", age_days, 0),
         logAge_days = log(age_days, base=10)) %>%
  mutate(logAge_days = ifelse(is.infinite(logAge_days),0.001, logAge_days)) %>% #make sure no inf values
  mutate(sexMale_x_logAge = sexMale * logAge_days, #add log-scaled ages
         total.CT = rowSums(across(all_of(ct_list)), na.rm=FALSE),
         total.SA = rowSums(across(all_of(sa_list)), na.rm=FALSE),
         study = as.factor(study),
         study_site = as.factor(study_site))
```

keep only columns we need (phenotypes to model, covariates)

```{r}
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)
covar_list <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age",
                "fs_version", "fs_version_SA", "fs_version_CT", "fs_version_GM",
                "age_days", "logAge_days", "sexMale_x_logAge", "site", "country", "study_site", 
                "QC.SS", "QC.RCA", "SurfaceHoles", 
                "eTIV", "TBV", "total.CT", "total.SA")

df_select <- df_filt %>%
  mutate(TBV = WMV + sGMV + GMV) %>%
  dplyr::select(all_of(c(pheno_list, covar_list)))

skim(df_select)
```

look at IDPs=0
```{r}
df_zeros <- df_select %>%
  filter(if_any(any_of(pheno_list), ~ . == 0))

ggplot(df_zeros) +
  geom_histogram(aes(x=age_days, fill=study))

df_zeros %>%
  filter(age_days < (280+365)) %>%
  group_by(study) %>%
  summarise(study = unique(study),
            n=n())

dhcp <- df_select %>%
  filter(study=="dHCP-fetal")
skim(dhcp)

df_zeros %>%
  filter(age_days <2500) %>% nrow()

df_zeros_infant <- df_zeros %>%
  filter(age_days <= 365.25 + 280) %>%
  droplevels()
nrow(df_zeros_infant)
```

looks like infant/fetal data have a lot of missing values
looking if the rest of the IDPs seem to be in line with other studies of that age

```{r, eval=FALSE}
df_infant <- df_select %>%
  filter(age_days < 365.25 + 280) %>%
  droplevels()

for (pheno in pheno_list){
  plt <- ggplot() +
    geom_point(aes(x=df_infant$age_days, y=df_infant[[pheno]], color=df_infant$study)) +
    ggtitle(pheno)
  
  ggsave(paste0("./figs/", pheno, ".png"), plt)
}

```

Going to remove phenotypes with 0s from infants (age_days <= 265+280) and remove ALL phenos from non-infant subjects with 0 phenos

```{r}
df_infant_nozero <- df_select %>%
  dplyr::filter(age_days <= (365.25 + 280)) %>%
  mutate(across(all_of(pheno_list), ~ifelse(.x==0, NA, .x)))

skim(df_infant_nozero)

stopifnot(df_infant_nozero %>%
    filter(if_any(any_of(pheno_list), ~ . == 0)) %>% 
      nrow() == 0)

df_adult_nozero <- df_select %>%
  dplyr::filter(age_days > (365.25 + 280)) %>%
  filter(if_all(pheno_list, ~ is.na(.) | . != 0))

stopifnot(df_adult_nozero %>%
    filter(if_any(any_of(pheno_list), ~ . == 0)) %>% 
      nrow() ==0)

df_nozero <- rbind(df_infant_nozero, df_adult_nozero)
nrow(df_select) - nrow(df_nozero)
unique(df_infant_nozero$dx)
```

### Remove Psychosis Studies
set aside selected psychosis studies
```{r}
psych_dx_list <- c(unique(df$dx)[grepl("scz|sz|psych|schizo|cos", unique(df$dx), ignore.case=T)], "CN")
psych_dx_list <- psych_dx_list[ !psych_dx_list == 'Psychological over-eating or binge-eating']

unique(df_nozero$study)
df %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(study, dx) %>%
  summarise(count = n())

scz_study_list <- c("ALSPAC", "AMP-SCZ", "BHRCS", "COBRE_FS6_part1", "Dublin_FS6", "HCP-EarlyPsychosis", "MCI", "Maastricht_FS6",
                    "NAPLS3", "NDAR-CCDP", "NDAR-PsychiatricCerebralBlood", "NDAR-PsychosisOMS", "NDAR-SCZAging", "NDAR-SCZEmot",
                    "NDAR-SCZGlu", "NDAR-SocialBrainSchizoAutism", "NDAR_BSNIP", "NDAR_PARDIP", "NDAR_SCZIowa", "OpenNeuro_LA5c",
                    "SRPBS1600", "TOPSY", "NDAR_NIH-IRP") #NDAR_NIH-IRP = controls for Armin's COS dataset

psychosis_df <- df_nozero %>%
  dplyr::filter(study %in% scz_study_list) %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

ggplot(psychosis_df) +
  geom_histogram(aes(x=(age_days/365.25), fill=dx))

psychosis_df %>%
  group_by(dx) %>%
  summarise(count=n())
```

write out csv - *not filtering/removing duplicate scans because planning to add more data!*
```{r}
fwrite(psychosis_df, file="./data/vpenn_scz_studies.csv")
```

FOR MORE SCZ STUDIES TO BE ADDED, SEE NDAR PROPOSAL GOOGLE DOC 

remove psychosis studies from main dataset so there's no data leakage
```{r}
df_nopsych <- df_nozero %>%
  dplyr::filter(!(study %in% scz_study_list))
```


### Scan QC filtering
```{r}
qc_df <- df_nopsych %>%
  mutate(QC.SS = ifelse(is.na(QC.SS), FALSE, TRUE),
         QC.RCA = ifelse(is.na(QC.RCA), FALSE, TRUE),
         SurfaceHoles = ifelse(is.na(SurfaceHoles), FALSE, TRUE))

table(qc_df$fs_version, qc_df$QC.SS)
table(qc_df$fs_version, qc_df$QC.RCA)

study_mad <- df_nopsych %>%
  drop_na(SurfaceHoles) %>%
  group_by(study) %>%
  summarize(study_euler_mad = mad(SurfaceHoles),
            study_euler_median = median(SurfaceHoles)) %>%
  mutate(euler_threshold = study_euler_median+(4*study_euler_mad))

left_join(df_nopsych, study_mad) %>%
  mutate(scan_pass = ifelse(SurfaceHoles <= euler_threshold, "yes", "no")) %>%
  group_by(study) %>%
  summarise(fail_percent = (sum(scan_pass == "no") / n())*100) %>%
  na.omit() %>%
  summarise(max_fail = max(fail_percent),
            min_fail=min(fail_percent),
            mean_fail=mean(fail_percent))

left_join(df_nopsych, study_mad) %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=SurfaceHoles)) +
  geom_point(aes(x=study, y=(study_euler_median+4*study_euler_mad)), color="red") +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

df_qc_filt <- left_join(df_nopsych, study_mad) %>%
  mutate(has_qc = ifelse(is.na(SurfaceHoles) & is.na(QC.SS) & is.na(QC.RCA),
                        FALSE, TRUE)) %>%
  dplyr::filter(SurfaceHoles <= euler_threshold | QC.SS < 0.65 | QC.RCA < 0.65 | has_qc == FALSE) #using SS recommended threshold

nrow(df_nopsych) - nrow(df_qc_filt)
```


### Final Filtering

filter to just controls, single scan per subject, drop sites with fewer than 5 ppl

```{r}
unique(df_qc_filt$dx)[grep("cn|control|none", unique(df_nopsych$dx), ignore.case=TRUE)]

df_clean <- df_qc_filt %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(age_days <= (101*365.25)) %>%
  dplyr::filter(dx == "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup() %>%
  group_by(study_site) %>%
  filter(n_distinct(INDEX.ID) >=5) %>% #remove sites with < 5 ppl
  ungroup()

#look at completeness for infant/toddler data
df_clean %>%
  dplyr::filter(age_days <= (365.25*3 + 280)) %>%
  skim()
#some phenotypes around 50%


df_clean %>%
  dplyr::filter(age_days <= (365.25 + 280)) %>%
  skim()
#some as low as 25%

nrow(df_clean)

```

write out full dataframe
```{r}
fwrite(df_clean, "./data/v3_CN_cleaned.csv")
```

###Split for CV Modeling

Splitting data into 50-50 train and testing, stratified by site and sex. Using this [reference](https://stackoverflow.com/questions/74573270/stratified-data-splitting-in-r).

```{r, eval=FALSE}
table(df_clean$study_site, df_clean$sexMale)
```

going to drop small sites (<10 subjects per sex) for now, but alternatives could be grouping small sitest together? (go back and include sites with <5 total subjs)

```{r}
valid_sites <- df_clean %>%
  mutate(sex=as.factor(sexMale)) %>%
  group_by(study_site, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(study_site) %>%
  summarise(
    valid = all(c("0", "1") %in% sex) && all(n[match(c("0", "1"), sex, nomatch = 0)] >= 10),
    .groups = "drop"
  ) %>%
  filter(valid) %>%
  pull(study_site)

df_clean_sites <- df_clean %>%
  filter(study_site %in% valid_sites)

nrow(df_clean) - nrow(df_clean_sites)

ggplot(df_clean) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))
ggplot(df_clean_sites) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))

#table(df_filt$study_site, df_filt$sexMale)
```

Still have almost 100k subjects, proceeding to divide into A and B samples

```{r}
set.seed(5678)
#add sex factor var
df_clean_sites <- df_clean_sites %>% mutate(sex=as.factor(sexMale))

split_samples <- splitstackshape::stratified(df_clean_sites, group=c("study_site", "sex"), size=0.5, bothSets = TRUE)

samp_A <- split_samples$SAMP1
samp_B <- split_samples$SAMP2

ggplot(samp_A) +
  geom_histogram(aes(x=age_days, fill=sex))
ggplot(samp_B) +
  geom_histogram(aes(x=age_days, fill=sex))

samp_A$sample <- "A"
samp_B$sample <- "B"
full_df <- rbind(samp_A, samp_B) %>%
  unite(sample_sex, c(sex, sample))

ggplot(full_df)+
  geom_violin(aes(x=sample_sex, y=age_days))
pairwise.wilcox.test(full_df$age_days, full_df$sample_sex, p.adjust.method = "none")
ggplot(full_df)+
  geom_violin(aes(x=sample_sex, y=SurfaceHoles))
pairwise.wilcox.test(full_df$SurfaceHoles, full_df$sample_sex, p.adjust.method = "none")
```
pretty good distributions of ages and Euler within sex, no significant differences between samples 

save
```{r}
fwrite(split_samples$SAMP1, "./data/cv_sample_A.csv")
fwrite(split_samples$SAMP2, "./data/cv_sample_B.csv")
```

###Subdivide by Feature Category

Write out separate csvs for each pheno category. These will be further subdivided into separate CSVs with maximally complete data in order to ComBat features (`subjobs_datasplit.sh` and `split_data_for_combat.R`), so for each feature (target feature to ComBat) keep as many other features as possible (as priors) without including any NAs.

```{r}
#save out QC separately since these would only be used for sensitivity analyses
df_covars <- df_clean %>%
  dplyr::select(all_of(covar_list))
fwrite(df_covars, file="./data/v3_CN_covars.csv")

covar_subset <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age", 
                  "age_days", "logAge_days", "sexMale_x_logAge", "study_site")

#global vols
df_global <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, glob_vol_list, 
                         "fs_version", "fs_version_SA", "fs_version_CT", "fs_version_GM", 
                         "eTIV", "TBV"))) %>%
  filter_at(glob_vol_list, any_vars(complete.cases(.)))
fwrite(df_global, "./data/csvs_by_pheno/global_vols.csv")

#regional gmv
df_gmv <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, gmv_list, "fs_version_GM", "eTIV")))%>%
  filter_at(gmv_list, any_vars(complete.cases(.)))
fwrite(df_gmv, "./data/csvs_by_pheno/cortical_vols.csv")

#regional sa
df_sa <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, sa_list, "fs_version_SA", "total.SA")))%>%
  filter_at(sa_list, any_vars(complete.cases(.)))
fwrite(df_sa, "./data/csvs_by_pheno/cortical_surf.csv")

#regional ct
df_ct <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, ct_list, "fs_version_CT", "total.CT")))%>%
  filter_at(ct_list, any_vars(complete.cases(.)))
fwrite(df_ct, "./data/csvs_by_pheno/cortical_thickness.csv")

#subcortical vols
df_sub <- df_clean %>%
  dplyr::select(any_of(c(covar_subset, subcort_list, "fs_version_GM", "eTIV")))%>%
  filter_at(vars(any_of(subcort_list)), any_vars(complete.cases(.)))
fwrite(df_sub, "./data/csvs_by_pheno/subcortical_vols.csv")
```

# Visualize Data

### Euler QC
looking at associations between Euler and demographics (in control sample and held-out psychosis sample)

```{r}
euler_df <- df_clean %>%
  filter(!is.na(SurfaceHoles)) %>%
  mutate(sexMale = as.factor(sexMale))

euler_df %>%
  ggplot(aes(x=sexMale, y=SurfaceHoles)) +
  geom_boxplot() +
  stat_compare_means()

euler_df %>%
  ggplot(aes(x=age_days, y=SurfaceHoles)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")
```

Try filtering more?
```{r}
names(euler_df)

euler_df_strict <- euler_df %>%
  mutate(euler_threshold = study_euler_median + (3.5*study_euler_mad)) %>%
  dplyr::filter(SurfaceHoles <= euler_threshold)

euler_df_strict %>%
  ggplot(aes(x=age_days, y=SurfaceHoles)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")

nrow(euler_df) - nrow(euler_df_strict)
```
loose a lot of subjects, doesn't seem to affect relationship

```{r}
euler_df_strict %>%
  ggplot(aes(x=age_days, y=eTIV)) +
  geom_point() +
  geom_hline(yintercept=4553071)
```

Look at psychosis sample
```{r}
psych_df_eu <- psychosis_df %>%
   filter(!is.na(SurfaceHoles)) %>%
  mutate(sexMale = as.factor(sexMale))
```

```{r}
unique(psych_df_eu$dx)
#collapsing all non-CN diagnoses for now
psych_df_eu <- psych_df_eu %>%
  mutate(dx_sum = case_when(dx=="CN" ~ "CN",
                            is.na(dx) ~ NA,
                            TRUE ~ "PS"))
psych_df_eu %>%
  ggplot(aes(x=sexMale, y=SurfaceHoles, fill=dx_sum)) +
  geom_boxplot() +
  stat_compare_means()

psych_df_eu %>%
  ggplot(aes(x=age_days, y=SurfaceHoles, fill=dx_sum, color=dx_sum)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")
```

### Global Phenotypes
```{r}
ggplot(df_global) + 
  geom_point(aes(x=age_days, y=GMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=WMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=sGMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=CSF,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))
```
```{r}
ggplot(df_global, aes(x=log(age_days), y=eTIV)) + 
  geom_point(aes(color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x))

ggplot(df_global, aes(x=age_days, y=eTIV)) + 
  geom_point(aes(color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x))

df_global <- df_global %>%
  mutate(TBV = WMV + sGMV + GMV)
ggplot(df_global, aes(x=log(age_days), y=TBV)) + 
  geom_point(aes(color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x))
```

### Cortical volumes
```{r}
ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.lingual,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.parsorbitalis,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.fusiform,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

#compare to original data

ggplot(df) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal, color=sex), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))
```

###Cortical Thickness
```{r}
ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

#compare to full dataset
ggplot(df_clean) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.parahippocampal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.caudalanteriorcingulate, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.transversetemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Cortical SA
```{r}
ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.pericalcarine, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.lateralorbitofrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.entorhinal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
df_sa %>%
  dplyr::filter(rh.SA.rostralmiddlefrontal < 3000 & age_days > 2500) %>%
  group_by(fs_version_SA, study) %>%
  summarise(n = n())

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=lh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)
```

values seem low, re-check post-ComBat

###Subcortical Volumes
```{r}
ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Hippocampus, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Accumbens.area, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Amygdala, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Right.Thalamus.Proper, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###ICV scaling

```{r}
df_clean %>%
  filter(eTIV < 3e6) %>%
  ggplot(aes(x=eTIV, y=SUBC.Left.Amygdala)) +
  geom_point(alpha=0.6, color="gray") +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x, bs = "cs", k=5)) +
  theme_bw()
```

###Visualize Batch Effects

using simple PCA
```{r}
df_phenos <- df_clean %>% dplyr::select(all_of(c("study_site", "study", ct_list, gmv_list, sa_list, subcort_list))) %>% na.omit()

pc_phenos <- prcomp(df_phenos[,-(1:2)], center=TRUE, scale=TRUE)

as.data.frame(pc_phenos$x) %>% 
  mutate(site = df_phenos$study_site) %>%
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(aes(color=site), alpha=0.4)+
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none")

as.data.frame(pc_phenos$x) %>% 
  mutate(site = df_phenos$study) %>%
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(aes(color=site), alpha=0.4)+
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none")
```
