#Curation

Script for cleaning and preparing data. Using dataset compiled by Lena, tentatively to be LBCC release v 5.0

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel, ggpubr)
```

load data
```{r}
df <- fread("./data/240820-L.R.LBCC.csv")
names(df)
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$site))
sum(is.na(df$dx))
sum(is.na(df$session))

hist(df$MAD.Euler)
```

```{r}
#look at missing country info - should be findable
df %>%
  dplyr::filter(is.na(country)) %>%
  group_by(study) %>%
  summarise(n = n())

#adding back country data based on OG nature paper

df <- df %>%
  dplyr::mutate(country = case_when(is.na(country) & (study == "SYS-Adolescents" | study == "SYS-Adults") ~ "Canada",
                                    is.na(country) & study == "TOPSY" ~ "Canada",
                                    is.na(country) & study == "dHCP-fetal" ~ "UK",
                                    TRUE ~ country))
```

##Calculate Global Vols

first from freesurfer-derived outputs (rename)

```{r}
#cortical GMV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'TotalGM'] <- 'GMV'

#sGMV
names(df)[grepl("subc", names(df), ignore.case=T)]
names(df)[names(df) == 'SubCortGM'] <- 'sGMV'

#WMV
names(df)[grepl("white", names(df), ignore.case=T)]
names(df)[names(df) == 'CerebralWhiteMatterVol'] <- 'WMV'

#CSF (ventricles)
names(df)[grepl("brainseg", names(df), ignore.case=T)]
df <- df %>%
  mutate(CSF = BrainSegVol - BrainSegVolNotVent)

#eTIV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'EstimatedTotalIntraCranialVol'] <- 'eTIV'
```

look at global SA and CT
```{r}
df %>% 
  dplyr::select(starts_with("HEMI")) %>%
  skim()

#complete! sum across hemispheres
df <- df %>%
  mutate(total.CT = HEMI.lh.CT + HEMI.rh.CT,
         total.SA = HEMI.lh.SA + HEMI.rh.SA) %>%
  mutate(across(c(total.CT, total.SA), ~ na_if(., 0)))
```

then from synthseg scans
```{r}
#add ID for merging scans
df <- df %>% mutate(merge_id = row_number())

df_synthseg <- df %>%
  dplyr::filter(fs_version == "synthseg-robust")

#confirm all missing
stopifnot(sum(is.na(df_synthseg$GMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$sGMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$WMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$CSF)) == nrow(df_synthseg))

#find appropriate cols to sum
df_synthseg %>%
  dplyr::select(all_of(c(contains("white"), contains("cort"), contains("hemi"), contains("ven")))) %>%
  skim()

df_synthseg %>%
  dplyr::select(contains("SUBC")) %>%
  skim()

ss_subcort_list <- c("SUBC.Left.Thalamus.Proper", "SUBC.Right.Thalamus.Proper", 
                     "SUBC.Left.Caudate", "SUBC.Right.Caudate",
                     "SUBC.Left.Putamen", "SUBC.Right.Putamen",
                     "SUBC.Left.Pallidum", "SUBC.Right.Pallidum",
                     "SUBC.Left.Hippocampus", "SUBC.Right.Hippocampus",
                     "SUBC.Left.Amygdala", "SUBC.Right.Amygdala",
                     "SUBC.Left.Accumbens.area", "SUBC.Right.Accumbens.area")

#calculate
df_glob <- df %>%
  mutate(GMV = ifelse(fs_version == "synthseg-robust", 
                      rowSums(across(c(rhCortexVol, lhCortexVol)), na.rm=FALSE),
                      GMV),
         WMV = ifelse(fs_version == "synthseg-robust",
                      rowSums(across(c(rhCerebralWhiteMatterVol, lhCerebralWhiteMatterVol)), na.rm=FALSE),
                      WMV),
         CSF = ifelse(fs_version == "synthseg-robust", Ventricles.all, CSF),
         sGMV = ifelse(fs_version == "synthseg-robust",
                       rowSums(across(all_of(ss_subcort_list)), na.rm=FALSE),
                       sGMV),
         eTIV = ifelse(fs_version == "synthseg-robust" & is.na(eTIV),
                       total.intracranial, eTIV)
         )
stopifnot(nrow(df_glob) == nrow(df))
```

plot for comparisons
```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

sum(is.na(df$GMV)) - sum(is.na(df_glob$GMV))
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280) +
  facet_wrap(~fs_version)
```

```{r}
df %>%
  dplyr::filter(sGMV > (1*10^5)) %>%
  group_by(fs_version, study) %>%
  summarise(n =n())
```

SYS-Adolescents looks way off on sGMV, removing this phenotype

```{r}
df_glob_nosys <- df_glob %>%
  mutate(sGMV = ifelse(study == "SYS-Adolescents", NA, sGMV))
```


```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

CSF has some outliers but overall looks good, going to check with Lena but proceed for now

```{r}
ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=total.CT,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=total.SA,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

check for any more big groups of missing global volumes
```{r}
df_glob %>%
  dplyr::filter(is.na(GMV&sGMV&WMV&CSF)) %>%
  group_by(fs_version) %>%
  summarise(n = n())
```

## Data Cleaning
remove intersex subj, dummy-code sex and sex*age, combine Euler and SS QC metrics (because of missingness)
also fix age calculations for `OpenNeuro-PaingenPlacebo` study - study reported years, accidentally calculated as months

```{r}
df_filt <- df_glob_nosys %>%
  dplyr::filter(sex != "intersex") %>%
  mutate(age_days = ifelse(study != "OpenNeuro-PaingenPlacebo", age_days,
                           ((age_days-280)*(12/365.25))*365.25+280)) %>% #convert back from days to 'months', then re-convert years to days
  mutate(sexMale = ifelse(sex=="male", 1, 0),
         sexMale_x_age = ifelse(sex=="male", age_days, 0),
         QC_Metric = ifelse(fs_version == "synthseg-robust", QC.SS, SurfaceHoles),
         logAge_days = log(age_days, base=10)) %>% #only keep QC.SS for scans processed with synthseg]
  mutate(sexMale_x_logAge = sexMale * logAge_days) #add log-scaled ages
```


write out lists of phenotypes by category
```{r}
#names(df_filt)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF","eTIV", "total.CT", "total.SA")
write_lines(glob_vol_list, "./pheno_lists/global_vols.txt")
saveRDS(glob_vol_list, "./pheno_lists/global_vols.rds")

#cortical gmv
gmv_list <- names(df_filt)[grep("GM\\.", names(df))]
stopifnot(length(gmv_list) == 68)
write_lines(gmv_list, "./pheno_lists/cortical_vols.txt")
saveRDS(gmv_list, "./pheno_lists/cortical_vols.rds")

#subcortical gmv
subcort <- names(df_filt)[grep("SUBC\\.Right|SUBC\\.Left", names(df))]
subcort_list <- subcort[-grep("Cerebellum|WM|Vent|choroid|vessel", subcort)] #drop WM, cerebellum, etc
length(subcort_list)
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")

#cortical thickness
ct_list <- names(df_filt)[grep("CT\\.", names(df))]
stopifnot(length(ct_list) == 68)
write_lines(ct_list, "./pheno_lists/cortical_thickness.txt")
saveRDS(ct_list, "./pheno_lists/cortical_thickness.rds")

#cortical SA
sa_list <- names(df_filt)[grep("SA\\.", names(df))]
stopifnot(length(sa_list) == 68)
write_lines(sa_list, "./pheno_lists/cortical_surf.txt")
saveRDS(sa_list, "./pheno_lists/cortical_surf.rds")
```

keep only columns we need (phenotypes to model, covariates)

```{r}
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)
covar_list <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age",
                "age_days", "logAge_days", "sexMale_x_logAge", "site", "country", "study_site", "QC_Metric", "fs_version", "eTIV")

df_select <- df_filt %>%
  dplyr::select(all_of(c(pheno_list, covar_list, "QC.SS", "SurfaceHoles")))

skim(df_select)
```

look at IDPs=0
```{r}
df_zeros <- df_select %>%
  filter(if_any(any_of(pheno_list), ~ . == 0 & !is.na(.))) %>%
  filter(if_any(all_of(pheno_list), ~ !is.na(.)))

ggplot(df_zeros) +
  geom_histogram(aes(x=age_days))

df_zeros %>%
  filter(age_days <2500) %>%
  summarise(study = unique(study))

df_zeros %>%
  filter(age_days <2500) %>% nrow()
```

looks like dHCP-fetal has a lot of missing values and one IDP with many zeros (`lh.SA.cuneus`)
looking if the rest of the IDPs seem to be in line with other studies of that age

```{r}
df_infant <- df_select %>%
  filter(age_days < 365.25 + 280) %>%
  droplevels()

ggplot(df_infant) +
  geom_point(aes(x=age_days, y=rh.SA.temporalpole, color=study))

ggplot(df_infant) +
  geom_point(aes(x=age_days, y=lh.SA.inferiorparietal, color=study))

ggplot(df_infant) +
  geom_point(aes(x=age_days, y=lh.GM.cuneus, color=study))

ggplot(df_infant) +
  geom_point(aes(x=age_days, y=rh.CT.parsorbitalis, color=study))

ggplot(df_infant) +
  geom_point(aes(x=age_days, y=rh.CT.cuneus, color=study))

ggplot(df_infant) +
  geom_point(aes(x=age_days, y=rh.GM.cuneus, color=study))

# df_infant %>%
#   filter(study=="dHCP-fetal") %>%
#   skim()
```

looks like `lh.SA.cuneus` is getting segmented as `lh.SA.parsorbitalis`. Going to remove both phenotypes from dHCP-fetal and remove all subjects from other studies with zero values
```{r}
zero_ppts <- df_zeros %>% 
  dplyr::filter(study !="dHCP-fetal") %>%
  select(INDEX.ID) %>%
  as.list() %>%
  unname() %>%
  unlist()

df_select_nozero <- df_select %>%
  mutate(lh.SA.cuneus = ifelse(study=="dHCP-fetal", NA, lh.SA.cuneus),
         lh.SA.parsorbitalis = ifelse(study=="dHCP-fetal", NA, lh.SA.parsorbitalis),
         rh.SA.cuneus = ifelse(study=="dHCP-fetal", NA, rh.SA.cuneus),
         rh.SA.parsorbitalis = ifelse(study=="dHCP-fetal", NA, rh.SA.parsorbitalis)) %>%
  dplyr::filter(!(INDEX.ID %in% zero_ppts))
```

### Remove Psychosis Studies
set aside selected psychosis studies
```{r}
psych_dx_list <- c(unique(df$dx)[grepl("scz|sz|psych|schizo|cos", unique(df$dx), ignore.case=T)], "CN")

unique(df_select$study)
df %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(study, dx) %>%
  summarise(count = n())

scz_study_list <- c("ALSPAC", "AMP-SCZ", "BHRCS", "COBRE_FS6_part1", "Dublin_FS6", "HCP-EarlyPsychosis", "MCI", "Maastricht_FS6",
                    "NAPLS3", "NDAR-CCDP", "NDAR-PsychiatricCerebralBlood", "NDAR-PsychosisOMS", "NDAR-SCZAging", "NDAR-SCZEmot",
                    "NDAR-SCZGlu", "NDAR-SocialBrainSchizoAutism", "NDAR_BSNIP", "NDAR_PARDIP", "NDAR_SCZIowa", "OpenNeuro_LA5c",
                    "SRPBS1600", "TOPSY", "NDAR_NIH-IRP") #NDAR_NIH-IRP = controls for Armin's COS dataset

psychosis_df <- df_select_nozero %>%
  dplyr::filter(study %in% scz_study_list) %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

ggplot(psychosis_df) +
  geom_histogram(aes(x=(age_days/365.25), fill=dx))

psychosis_df %>%
  group_by(dx) %>%
  summarise(count=n())
```

write out csv - *not filtering/removing duplicate scans because planning to add more data!*
```{r}
fwrite(psychosis_df, file="./data/240820-L.R.LBCC_scz_studies.csv")
```

FOR MORE SCZ STUDIES TO BE ADDED, SEE NDAR PROPOSAL GOOGLE DOC 

remove psychosis studies from main dataset so there's no data leakage
```{r}
df_nopsych <- df_select_nozero %>%
  dplyr::filter(!(study %in% scz_study_list))
```


look into subjects missing dx
```{r}
no_dx <- df_nopsych %>% dplyr::filter(is.na(dx))
ggplot(no_dx) +
  geom_histogram(aes(x=age_days)) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

no_dx %>%
  group_by(study) %>%
  summarise(mean_age = mean(age_days),
            n = n())

df_nopsych %>%
  dplyr::filter(study == "NDAR_IBIS") %>%
  group_by(dx) %>%
  summarise(n=n(),
            min_age = min(age_days),
            max_age = max(age_days))
```


subjects missing diagnoses are not consistently infants

### Scan QC filtering
```{r}
study_mad <- df_nopsych %>%
  drop_na(SurfaceHoles) %>%
  group_by(study) %>%
  summarize(study_euler_mad = mad(SurfaceHoles),
            study_euler_median = median(SurfaceHoles)) %>%
  mutate(euler_threshold = study_euler_median+(4*study_euler_mad))

left_join(df_nopsych, study_mad) %>%
  mutate(scan_pass = ifelse(SurfaceHoles <= euler_threshold, "yes", "no")) %>%
  group_by(study) %>%
  summarise(fail_percent = (sum(scan_pass == "no") / n())*100) %>%
  na.omit() %>%
  summarise(max_fail = max(fail_percent),
            min_fail=min(fail_percent),
            mean_fail=mean(fail_percent))

df_qc_filt <- left_join(df_nopsych, study_mad) %>%
  dplyr::filter(SurfaceHoles <= euler_threshold | QC.SS < 0.65 | is.na(QC_Metric)) #using SS recommended threshold

nrow(df_nopsych) - nrow(df_qc_filt)
```


### Final Filtering

filter to just controls, single scan per subject, drop sites with fewer than 5 ppl

```{r}
unique(df_qc_filt$dx)[grep("cn|control|none", unique(df_nopsych$dx), ignore.case=TRUE)]

df_clean <- df_qc_filt %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(age_days <= (101*365.25)) %>%
  dplyr::filter(dx == "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup() %>%
  group_by(study_site) %>%
  filter(n_distinct(INDEX.ID) >=5) %>% #remove sites with < 5 ppl
  ungroup()

#look at completeness for infant/toddler data
df_clean %>%
  dplyr::filter(age_days <= (365.25*3 + 280)) %>%
  skim()
#some phenotypes around 50%


df_clean %>%
  dplyr::filter(age_days <= (365.25 + 280)) %>%
  skim()
#some as low as 25%

nrow(df_clean)

```

determine equal knot placements 
```{r}
range(df_clean$logAge_days)

knot_lists <- c()
for (degree in seq(2, 24)){
  list_name <- paste0("df", degree)
  knot_lists[[list_name]] <- dlnm::equalknots(df_clean$logAge_days,
                                             fun="ns",
                                             df= degree)
}

#also get quantiles
quant_knots <- quantile(df_clean$logAge_days, names=FALSE)
dec_knots <- quantile(df_clean$logAge_days, probs=seq(0,1, 0.1), names=FALSE)

quant_knot_lists <- c()
dec_knot_lists <-c()
for (name in names(knot_lists)){
  quant_knot_lists[[name]] <- c(knot_lists[[name]], quant_knots) %>% sort()
  dec_knot_lists[[name]] <- c(knot_lists[[name]], dec_knots) %>% sort()
}

saveRDS(knot_lists, "code/knot_lists.RDS")
saveRDS(quant_knot_lists, "code/quant_knot_lists.RDS")
saveRDS(dec_knot_lists, "code/dec_knot_lists.RDS")
```

some phenotypes seem to only have younger data - looking into this
```{r}
df_clean %>%
  ggplot() +
  geom_point(aes(x=age_days, y=SUBC.Right.Lentiform.Nucleus))

approx_max <- max(df_clean$age_days) - 5*365.25
approx_min <- min(df_clean$age_days) + 5*365.25

get_sample_range <- function(pheno, df){
  df %>%
    as.data.table %>%
    na.omit(cols=pheno) %>%
    summarise(max_age = max(age_days),
              min_age = min (age_days))
}

list_ranges <- lapply(pheno_list, get_sample_range, df=df_clean)
names(list_ranges) <- pheno_list


df_ranges <- bind_rows(list_ranges, .id="pheno")

restricted <-df_ranges %>%
  dplyr::filter(max_age < approx_max | min_age > approx_min)
```

remove subcortical vols restricted to infants
```{r}
df_clean2 <- df_clean %>%
  dplyr::select(!unique(restricted$pheno))

#also updating subcort lists
subcort_list <- subcort_list[!(subcort_list %in% unique(restricted$pheno))]
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")
pheno_list <- pheno_list[!(pheno_list %in% unique(restricted$pheno))]

fwrite(df_clean2, "./data/SBC_LBCC_CN_cleaned.csv")
```

###Split for CV Modeling

Splitting data into 50-50 train and testing, stratified by site and sex. Using this [reference](https://stackoverflow.com/questions/74573270/stratified-data-splitting-in-r).

```{r}
table(df_clean2$study_site, df_clean2$sexMale)
```

going to drop small sites (<10 subjects per sex) for now, but alternatives could be grouping small sitest together? (go back and include sites with <5 total subjs)

```{r}
valid_sites <- df_clean2 %>%
  mutate(sex=as.factor(sexMale)) %>%
  group_by(study_site, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(study_site) %>%
  summarise(
    valid = all(c("0", "1") %in% sex) && all(n[match(c("0", "1"), sex, nomatch = 0)] >= 10),
    .groups = "drop"
  ) %>%
  filter(valid) %>%
  pull(study_site)

df_clean2_sites <- df_clean2 %>%
  filter(study_site %in% valid_sites)

nrow(df_clean2) - nrow(df_clean2_sites)

ggplot(df_clean2) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))
ggplot(df_clean2_sites) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))

#table(df_filt$study_site, df_filt$sexMale)
```

Still have almost 100k subjects, proceeding to divide into A and B samples

```{r}
set.seed(5678)
#add sex factor var
df_clean2_sites <- df_clean2_sites %>% mutate(sex=as.factor(sexMale))

split_samples <- splitstackshape::stratified(df_clean2_sites, group=c("study_site", "sex"), size=0.5, bothSets = TRUE)

samp_A <- split_samples$SAMP1
samp_B <- split_samples$SAMP2

ggplot(samp_A) +
  geom_histogram(aes(x=age_days, fill=sex))
ggplot(samp_B) +
  geom_histogram(aes(x=age_days, fill=sex))

samp_A$sample <- "A"
samp_B$sample <- "B"
full_df <- rbind(samp_A, samp_B) %>%
  unite(sample_sex, c(sex, sample))
kruskal.test(age_days ~ sample_sex, full_df)
ggplot(full_df)+
  geom_violin(aes(x=sample_sex, y=age_days))
```
pretty good distributions of ages

save
```{r}
fwrite(split_samples$SAMP1, "./data/cv_sample_A.csv")
fwrite(split_samples$SAMP2, "./data/cv_sample_B.csv")
```

###Subdivide by Feature Category

Write out separate csvs for each pheno category. These will be further subdivided into separate CSVs with maximally complete data in order to ComBat features (`subjobs_datasplit.sh` and `split_data_for_combat.R`), so for each feature (target feature to ComBat) keep as many other features as possible (as priors) without including any NAs.

```{r}
#save out QC separately since these would only be used for sensitivity analyses
df_covars <- df_clean2 %>%
  dplyr::select(all_of(covar_list))
fwrite(df_covars, file="./data/SBC_LBCC_CN_covars.csv")

covar_subset <- covar_list[!covar_list %in% c("QC_Metric", "run", "eTIV")]

#global vols
df_global <- df_clean2 %>%
  dplyr::select(all_of(c(covar_subset, glob_vol_list))) %>%
  filter_at(glob_vol_list, any_vars(complete.cases(.)))
fwrite(df_global, "./data/csvs_by_pheno/global_vols.csv")

#regional gmv
df_gmv <- df_clean2 %>%
  dplyr::select(all_of(c(covar_subset, gmv_list)))%>%
  filter_at(gmv_list, any_vars(complete.cases(.)))
fwrite(df_gmv, "./data/csvs_by_pheno/cortical_vols.csv")

#regional sa
df_sa <- df_clean2 %>%
  dplyr::select(all_of(c(covar_subset, sa_list)))%>%
  filter_at(sa_list, any_vars(complete.cases(.)))
fwrite(df_sa, "./data/csvs_by_pheno/cortical_surf.csv")

#regional ct
df_ct <- df_clean2 %>%
  dplyr::select(all_of(c(covar_subset, ct_list)))%>%
  filter_at(ct_list, any_vars(complete.cases(.)))
fwrite(df_ct, "./data/csvs_by_pheno/cortical_thickness.csv")

#subcortical vols
df_sub <- df_clean2 %>%
  dplyr::select(any_of(c(covar_subset, subcort_list)))%>%
  filter_at(vars(any_of(subcort_list)), any_vars(complete.cases(.)))
fwrite(df_sub, "./data/csvs_by_pheno/subcortical_vols.csv")
```

# Visualize Data

### Euler QC
looking at associations between Euler and demographics (in control sample and held-out psychosis sample)

```{r}
skim(df_clean2$QC_Metric)

#see studies w/o QC
no_eu <- df_clean2 %>%
  filter(is.na(QC_Metric)) %>%
  group_by(study) %>%
  summarise(n=n()) %>%
  arrange(desc(n))

euler_df <- df_clean2 %>%
  filter(fs_version != "synthseg-robust") %>%
  mutate(sexMale = as.factor(sexMale))

skim(euler_df$QC_Metric)

euler_df %>%
  ggplot(aes(x=sexMale, y=QC_Metric)) +
  geom_boxplot() +
  stat_compare_means()

euler_df %>%
  ggplot(aes(x=age_days, y=QC_Metric)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")
```

Try filtering more?
```{r}
names(euler_df)

euler_df_strict <- euler_df %>%
  mutate(euler_threshold = study_euler_median + (3.5*study_euler_mad)) %>%
  dplyr::filter(SurfaceHoles <= euler_threshold)

euler_df_strict %>%
  ggplot(aes(x=age_days, y=QC_Metric)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")

nrow(euler_df) - nrow(euler_df_strict)
```
loose a lot of subjects, doesn't seem to affect relationship

```{r}
euler_df_strict %>%
  ggplot(aes(x=age_days, y=eTIV)) +
  geom_point() +
  geom_hline(yintercept=4553071)
```

Look at psychosis sample
```{r}
psych_df_eu <- psychosis_df %>%
  filter(fs_version != "synthseg-robust") %>%
  mutate(sexMale = as.factor(sexMale))

skim(psych_df_eu$QC_Metric)
```

```{r}
unique(psych_df_eu$dx)
#collapsing all non-CN diagnoses for now
psych_df_eu <- psych_df_eu %>%
  mutate(dx_sum = case_when(dx=="CN" ~ "CN",
                            is.na(dx) ~ NA,
                            TRUE ~ "PS"))
psych_df_eu %>%
  ggplot(aes(x=sexMale, y=QC_Metric, fill=dx_sum)) +
  geom_boxplot() +
  stat_compare_means()

psych_df_eu %>%
  ggplot(aes(x=age_days, y=QC_Metric, fill=dx_sum, color=dx_sum)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")
```

### Global Phenotypes
```{r}
ggplot(df_global) + 
  geom_point(aes(x=age_days, y=GMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=WMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=sGMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=CSF,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))
```

### Cortical volumes
```{r}
ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.lingual,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.parsorbitalis,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.fusiform,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

#compare to original data

df_test_dx <- df %>%
  mutate(dx = case_when(is.na(dx) ~ NA,
                        dx == "CN" ~ "CN",
                        TRUE ~ "notCN"))
ggplot(df_test_dx) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal, color=dx), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))
```

###Cortical Thickness
```{r}
ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

#compare to full dataset
ggplot(df_clean) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.parahippocampal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.caudalanteriorcingulate, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.transversetemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Cortical SA
```{r}
ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.pericalcarine, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.lateralorbitofrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.entorhinal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
df_sa %>%
  dplyr::filter(rh.SA.rostralmiddlefrontal < 3000 & age_days > 2500) %>%
  group_by(fs_version, study) %>%
  summarise(n = n())

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=lh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)
```

values seem low, re-check post-ComBat

###Subcortical Volumes
```{r}
ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Hippocampus, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Accumbens.area, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Amygdala, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Right.Thalamus.Proper, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###ICV scaling

```{r}
df_clean2 %>%
  filter(eTIV < 3e6) %>%
  ggplot(aes(x=eTIV, y=SUBC.Left.Amygdala)) +
  geom_point(alpha=0.6, color="gray") +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x, bs = "cs", k=5)) +
  theme_bw()
```

###Visualize Batch Effects

using simple PCA
```{r}
df_phenos <- df_clean2 %>% dplyr::select(all_of(c("study_site", "study", pheno_list))) %>% na.omit()

pc_phenos <- prcomp(df_phenos[,-(1:2)], center=TRUE, scale=TRUE)

as.data.frame(pc_phenos$x) %>% 
  mutate(site = df_phenos$study_site) %>%
  ggplot(aes(x=PC1, y=PC2), alpha=0.4) +
  geom_point(aes(color=site))+
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none")

as.data.frame(pc_phenos$x) %>% 
  mutate(site = df_phenos$study) %>%
  ggplot(aes(x=PC1, y=PC2), alpha=0.4) +
  geom_point(aes(color=site))+
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none")
```
