#Curation

Script for cleaning and preparing LBCC data. 


```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel, ggpubr, stringr, pwr)
#devtools::install_github("BGDlab/gamlssTools", force = TRUE) 
library(gamlssTools)
```

load data - prelim version of LBCC 3.0, waiting on Lena to look into missing sex info
```{r}
df <- fread("/mnt/isilon/bgdlab_processing/LBCC_releases/v3_Penn/v3_0_0_penn.csv", na.strings = c("", NA))
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$site))
sum(is.na(df$dx))
sum(is.na(df$session))
sum(is.na(df$study_session))

hist(df$MAD.Euler)
```

write out lists of phenotypes by category
```{r}
#cerebellum
cerebel <- names(df)[grep("Cerebell|cerebell|CERE", names(df))]
  df %>%
  select(all_of(cerebel)) %>%
  skim()

#rename Cerebellum.all to CBV
df <- df %>%
  rename(CBV = Cerebellum.all)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF", "CBV", "mean.CT", "total.SA")
write_lines(glob_vol_list, "./pheno_lists/global_vols.txt")
saveRDS(glob_vol_list, "./pheno_lists/global_vols.rds")
```


```{r, eval=FALSE}
ss_subcort_list <- c("SUBC.Left.Thalamus.Proper", "SUBC.Right.Thalamus.Proper", 
                     "SUBC.Left.Caudate", "SUBC.Right.Caudate",
                     "SUBC.Left.Putamen", "SUBC.Right.Putamen",
                     "SUBC.Left.Pallidum", "SUBC.Right.Pallidum",
                     "SUBC.Left.Hippocampus", "SUBC.Right.Hippocampus",
                     "SUBC.Left.Amygdala", "SUBC.Right.Amygdala",
                     "SUBC.Left.Accumbens.area", "SUBC.Right.Accumbens.area")

ss_ventricles_list <- c("SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle",
                        "SUBC.Left.Inf.Lat.Vent", "SUBC.Right.Inf.Lat.Vent",
                        "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

#try to fill in missing global volumes with other cols
df2 <- df %>%
  mutate(GMV = case_when(!is.na(GMV) ~ GMV,
                         !is.na(CortexVol) ~ CortexVol,
                         !is.na(rhCortexVol) & !is.na(lhCortexVol) ~ rhCortexVol + lhCortexVol,
                         TRUE ~ NA),
         sGMV = case_when(!is.na(sGMV) ~ sGMV,
            !is.na(SubCortGM) ~ SubCortGM,
            TRUE ~  rowSums(across(all_of(ss_subcort_list)), na.rm=FALSE)),
         WMV = case_when(!is.na(WMV) ~ WMV,
            !is.na(CerebralWhiteMatterVol) ~ CerebralWhiteMatterVol,
            #!is.na(lhCerebralWhiteMatterVol) & !is.na(rhCerebralWhiteMatterVol) ~  lhCerebralWhiteMatterVol + rhCerebralWhiteMatterVol,
            #!is.na(right.cerebral.white.matter) & !is.na(left.cerebral.white.matter) ~ right.cerebral.white.matter + left.cerebral.white.matter
            ),
         CSF = case_when(!is.na(CSF) ~ CSF,
           !is.na(BrainSegVol) & !is.na(BrainSegVolNotVent) ~ 
                           BrainSegVol - BrainSegVolNotVent,
                          TRUE ~  rowSums(across(all_of(ss_ventricles_list)), na.rm=FALSE)),
         CSF.test = case_when(!is.na(BrainSegVol) & !is.na(BrainSegVolNotVent) ~ 
                           BrainSegVol - BrainSegVolNotVent,
                          TRUE ~  rowSums(across(all_of(ss_ventricles_list)), na.rm=FALSE)),
         eTIV = case_when(!is.na(eTIV) ~ eTIV,
                          TRUE ~ EstimatedTotalIntraCranialVol)
         )
all.equal(df2, df)

ggplot(df2) +
  geom_point(aes(x=age_days, y=WMV), alpha=0.4)

ggplot(df) +
  geom_point(aes(x=age_days, y=WMV), alpha=0.4) #added a lot of bad data, not going to do this!

ggplot(df2) +
  geom_point(aes(x=CSF, y=CSF.test, color=fs_version), alpha=0.4)
```

```{r}
#cortical gmv
gmv_list <- names(df)[grep("GM\\.", names(df))]
stopifnot(length(gmv_list) == 68)
write_lines(gmv_list, "./pheno_lists/cortical_vols.txt")
saveRDS(gmv_list, "./pheno_lists/cortical_vols.rds")

#subcortical gmv - updated to include wm and csv vols as well!
subcort <- names(df)[grep("SUBC\\.", names(df))]
subcort_list <- subcort[-grep("choroid|vessel|hypointensities|CSF|Chiasm", subcort)] #drop WM, cerebellum, etc
subcort_list <- subcort_list[-grep("Lentiform|Inf.Lat.Vent", subcort_list)] #drop segments only in infants
subcort_list <- subcort_list[subcort_list != "SUBC.CC"]
subcort_list <- subcort_list[subcort_list != "SUBC.5th.Ventricle"] #too hard to segment
subcort_list <- subcort_list[subcort_list != "SUBC.Left.Cerebellum"]
subcort_list <- subcort_list[subcort_list != "SUBC.Right.Cerebellum"]#only in a couple studies, use SUBC.Left.Cerebellum.Cortex
length(subcort_list)
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")

df %>%
  select(all_of(subcort_list)) %>%
  skim()

#cortical thickness
ct_list <- names(df)[grep("CT\\.", names(df))]
stopifnot(length(ct_list) == 68)
write_lines(ct_list, "./pheno_lists/cortical_thickness.txt")
saveRDS(ct_list, "./pheno_lists/cortical_thickness.rds")

#cortical SA
sa_list <- names(df)[grep("SA\\.", names(df))]
stopifnot(length(sa_list) == 68)
write_lines(sa_list, "./pheno_lists/cortical_surf.txt")
saveRDS(sa_list, "./pheno_lists/cortical_surf.rds")
```

## HBCD
Adding HBCD release

```{r}
hbcd <- fread("/mnt/isilon/bgdlab_hbcd/v2/derivatives/HBCD_idps.csv")
hbcd_rename <- hbcd %>%
  rename_with(~ gsub("-", ".", .)) %>%
  rename_with(~ gsub("SubcGM", "SUBC", .))

setdiff(names(hbcd_rename), names(df))

#Add Missing cols
hbcd_rename$study <- "HBCD"
hbcd_rename$dx <- "CN"
hbcd_rename$fs_version <- "FSInfant"
hbcd_rename$fs_version_GM <- "FSInfant"
hbcd_rename$fs_version_SA  <- "FSInfant"
hbcd_rename$fs_version_CT <- "FSInfant"
```

get demographics

```{r}
hbcd_demo <- fread("/mnt/isilon/bgdlab_hbcd/v2/rawdata/phenotype/sed_basic_demographics.tsv")

#get age at scan
hbcd_demo_scan <- fread("/mnt/isilon/bgdlab_hbcd/v2/derivatives/hbcd_scan_info.csv") %>%
  select(subject, session, site, age, gestational_age, QC, filename) %>%
  filter(str_detect(filename, regex("anat/sub-.*_ses-V0.?_run-.?_T.?w\\.nii\\.gz"))) %>%   #keep just basic T1 or T2 anat scans
  mutate(age_days_birth = 365.25*age) 

table(hbcd_demo_scan$QC) #all scans usable
n_unique(hbcd_demo_scan$subject)

#make names compatible for merge
hbcd_demo2 <- hbcd_demo %>%
  rename("participant" = "participant_id",
         "session" = "session_id")

hbcd_demo_scan2 <- hbcd_demo_scan %>%
  rename("participant" = "subject") %>%
  select(!filename) %>%
  distinct()

table(hbcd_demo_scan2$session)
n_unique(hbcd_demo_scan2$participant)

#merge into main demographics
hbcd_demo_merge <- inner_join(hbcd_demo2, hbcd_demo_scan2)

unique(df$sex)

#sex levels from sed_basic_demographics.json

hbcd_demo_reduced <- hbcd_demo_merge %>%
  rename_with(~ gsub("sed_basic_demographics_", "", .)) %>%
  mutate(session = as.integer(gsub("ses-V0", "", session)),
         sex = case_when(sex==0 ~ "female",
                         sex==1 ~ "male",
                         sex==2 ~ NA),
         gestational_age_delivery = gestational_age_delivery*7) %>%
  mutate(age_days = gestational_age_delivery + age_days_birth,
         site = as.character(site)) %>%
  select(participant, session, sex, site, age_days) %>%
  #drop session 1, which has no scans
  filter(session !=1)

names(hbcd_demo_reduced)
```

merge demographics
```{r}
nrow(hbcd_rename)
nrow(hbcd_demo_reduced)

hbcd_final <- left_join(hbcd_rename, hbcd_demo_reduced)
nrow(hbcd_final) == n_unique(hbcd_final$participant)
sum(is.na(hbcd_final))

length(setdiff(hbcd_rename$participant, hbcd_demo_reduced$participant))
```

merge in
```{r}
df_with_hbcd <- full_join(df, hbcd_final)
stopifnot(nrow(df_with_hbcd) == nrow(df) + nrow(hbcd_final))
```

## SLIP 2025-03

csv created with `build_your_cohort_dev_mg.Rmd`

```{r}
slip <- fread("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/data/slip_202503_release_updated.csv") #updated b/c mismatched scans
names(slip)
skim(slip)

# unlist(unique(names(slip)[names(slip) %in% names(df_with_hbcd)]))
# setdiff(names(slip), names(df_with_hbcd))
# setdiff(names(df_with_hbcd), names(slip))
```

```{r}
slip_rename <- slip %>%
  rename_with(~ str_replace(.x, "^GrayVol\\_lh_", "lh.GM.")) %>%
  rename_with(~ str_replace(.x, "^GrayVol\\_rh_", "rh.GM.")) %>%
  rename_with(~ str_replace(.x, "^SurfArea\\_lh_", "lh.SA.")) %>%
  rename_with(~ str_replace(.x, "^SurfArea\\_rh_", "rh.SA.")) %>%
  rename_with(~ str_replace(.x, "^ThickAvg\\_lh_", "lh.CT.")) %>%
  rename_with(~ str_replace(.x, "^ThickAvg\\_rh_", "rh.CT.")) %>%
  rename(age_days = adjusted_age_in_days,
         participant = participant_id,
         session = session_id) %>%
  filter(!is.na(device_serial_number)) %>%
  mutate(site = gsub(" ", "", paste0("CHOP", device_serial_number)),
         QC.RCA = do.call(pmin, c(across(starts_with("QC")), na.rm = TRUE)),
         euler = euler_rh + euler_lh) %>%
  mutate(SurfaceHoles = (2-euler)/2)

slip_rename %>%
  summarise(n_pass = sum(QC.RCA > 0.65),
            pct_pass = sum(QC.RCA > 0.65)/n())

slip_rename %>%
  filter(age_days <= 280 + (365.25/2)) %>%
  summarise(n_pass = sum(QC.RCA > 0.65),
            pct_pass = (sum(QC.RCA > 0.65)/n())*100,
            n_total=n())

# setdiff(names(df_with_hbcd), names(slip_rename))
# setdiff(names(slip_rename), names(df_with_hbcd))
```

very few subjects pass

rename cols as needed:

[1] "SUBC.Left.Lateral.Ventricle"        "SUBC.Left.Cerebellum.White.Matter"  "SUBC.Left.Cerebellum.Cortex"       
 [4] "SUBC.Left.Thalamus.Proper"          "SUBC.Left.Caudate"                  "SUBC.Left.Putamen"                 
 [7] "SUBC.Left.Pallidum"                 "SUBC.3rd.Ventricle"                 "SUBC.4th.Ventricle"                
[10] "SUBC.Brain.Stem"                    "SUBC.Left.Hippocampus"              "SUBC.Left.Amygdala"                
[13] "SUBC.Left.Accumbens.area"           "SUBC.Left.VentralDC"                "SUBC.Right.Cerebellum.White.Matter"
[16] "SUBC.Right.Cerebellum.Cortex"       "SUBC.Right.Thalamus.Proper"         "SUBC.Right.Caudate"                
[19] "SUBC.Right.Putamen"                 "SUBC.Right.Pallidum"                "SUBC.Right.Hippocampus"            
[22] "SUBC.Right.Amygdala"                "SUBC.Right.Accumbens.area"          "SUBC.Right.VentralDC"              
[25] "SUBC.Right.Lateral.Ventricle"       "SUBC.CC_Posterior"                  "SUBC.CC_Mid_Posterior"             
[28] "SUBC.CC_Central"                    "SUBC.CC_Mid_Anterior"               "SUBC.CC_Anterior"                  

```{r}
slip_rename2 <- slip_rename %>%
  mutate(run = as.numeric(str_extract(scan_id, "(?<=run-)\\d+"))) %>%
rename_with(
    ~ .x %>%
      str_replace_all("_|\\s+", " ") %>%     # Replace _ or space with single space
      str_to_title() %>%                     # Capitalize each word
      str_replace_all(" ", ".") %>%          # Replace spaces with .
      paste0("SUBC.", .),                    # Add prefix
    .cols = matches("^left_|^right_")
  ) %>%
  rename(CBV = cerebellum,
         SUBC.Left.Thalamus.Proper = SUBC.Left.Thalamus,
         SUBC.Right.Thalamus.Proper = SUBC.Right.Thalamus,
         SUBC.3rd.Ventricle = "3rd_ventricle",
         SUBC.4th.Ventricle = "4th_ventricle",
         SUBC.Brain.Stem = "brain_stem",
         SUBC.Left.Accumbens.area = SUBC.Left.Accumbens.Area,
         SUBC.Right.Accumbens.area = SUBC.Right.Accumbens.Area,
         SUBC.Right.VentralDC = SUBC.Right.Ventral.Dc,
         SUBC.Left.VentralDC = SUBC.Left.Ventral.Dc,
         SUBC.Left.Inf.Lat.Vent = SUBC.Left.Inferior.Lateral.Ventricle,
         SUBC.Right.Inf.Lat.Vent = SUBC.Right.Inferior.Lateral.Ventricle,
         slip_session = session) 
```

```{r}
ss_ventricles_list <- c("SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle",
                        "SUBC.Left.Inf.Lat.Vent", "SUBC.Right.Inf.Lat.Vent",
                        "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

vent_test <- slip_rename2 %>%
  mutate(diff = case_when((!is.na(BrainSegVol) & !is.na(BrainSegVolNotVent)) ~
                            BrainSegVol - BrainSegVolNotVent,
                          TRUE ~ NA),
         vent_sum = rowSums(across(all_of(ss_ventricles_list)), na.rm=FALSE))

ggplot(vent_test) +
  geom_point(aes(x=vent_sum, y=ventricles))
ggplot(vent_test) +
  geom_point(aes(x=diff, y=ventricles))
ggplot(vent_test) +
  geom_point(aes(x=csf, y=ventricles))

range(vent_test$ventricles)
```

Trying Ventricles for CSF for SLIP based on suggestions from Aaron

```{r}
slip_glob <- slip_rename2 %>%
  mutate(GMV = case_when(!is.na(CortexVol) ~ CortexVol,
                         !is.na(rhCortexVol) & !is.na(lhCortexVol) ~ rhCortexVol + lhCortexVol,
                         TRUE ~ NA),
         sGMV = case_when(!is.na(SubCortGrayVol) ~ SubCortGrayVol,
            TRUE ~ NA),
            #TRUE ~  rowSums(across(all_of(ss_subcort_list)), na.rm=FALSE)),
         WMV = case_when(!is.na(CerebralWhiteMatterVol) ~ CerebralWhiteMatterVol,
            !is.na(lhCerebralWhiteMatterVol) & !is.na(rhCerebralWhiteMatterVol) ~  lhCerebralWhiteMatterVol + rhCerebralWhiteMatterVol,
            TRUE ~ NA
            #!is.na(right.cerebral.white.matter) & !is.na(left.cerebral.white.matter) ~ right.cerebral.white.matter + left.cerebral.white.matter
            ),
         CSF = ventricles,
         eTIV = case_when(!is.na(eTIV) ~ eTIV,
                          TRUE ~ total_intracranial))


#LABEL FS VERSION
slip_glob$fs_version <- "synthseg-robust"
slip_glob$fs_version_GM <- "synthseg-robust"
slip_glob$fs_version_SA  <- "synthseg-robust"  
slip_glob$fs_version_CT <- "synthseg-robust"
slip_glob$study <- "SLIP"
slip_glob$dx <- "CN"
```

```{r}
slip_glob %>%
  ggplot() +
  geom_point(aes(x=age_days, y=CSF))
```

check that all cols included

```{r}
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)

setdiff(pheno_list, names(slip_glob))

# setdiff(names(df_with_hbcd), names(slip_glob))
# setdiff(names(slip_glob), names(df_with_hbcd))
```

Looks pretty good, going to remove old SLIP data and add new

```{r}
drp_slip <- df_with_hbcd %>%
  dplyr::filter(study != "SLIP")

df_with_slip <- full_join(drp_slip, slip_glob)
stopifnot(nrow(df_with_slip) == sum(nrow(drp_slip), nrow(slip_glob)))
```

```{r}
#check global volumes
df_slip_check <- df_with_slip %>%
  mutate(is.slip = ifelse(study=="SLIP", T, F))

df_slip_check %>%
  ggplot() +
  geom_point(aes(x=age_days, y=CSF, color=is.slip), alpha=0.5)

df_slip_check %>%
  ggplot() +
  geom_point(aes(x=age_days, y=sGMV, color=is.slip), alpha=0.5)

df_slip_check %>%
  ggplot() +
  geom_point(aes(x=age_days, y=WMV, color=is.slip), alpha=0.5)

df_slip_check %>%
  ggplot() +
  geom_point(aes(x=age_days, y=GMV, color=is.slip), alpha=0.5)
```

looks good

## Data Cleaning

remove intersex subj, dummy-code sex and sex*age, add totals for CT and SA

```{r}
unique(df_with_slip$sex)
sum(is.na(df_with_slip$sex))

df_filt1 <- df_with_slip %>%
  filter(study != "SYS-Adolescents")  #weird bad sGMV values, removing
```

```{r}
#get starting counts
df_filt1 %>%
  mutate(is_slip = ifelse(study=="SLIP", T, F),
         INDEX.ID = paste(study, participant, sex, sep="|")) %>%
  group_by(is_slip) %>%
  summarise(n = n(),
            n_subj = n_unique(INDEX.ID),
            n_study = n_unique(study))

nrow(df_filt1)
n_unique(df_filt1$study)


df_filt <- df_filt1 %>%
  dplyr::filter(sex != "intersex" & sex != "Unknown") %>%
  mutate(sexMale = ifelse(sex=="male" | sex == "Male", 1, 0),
         sexMale_x_age = ifelse(sex=="male"| sex == "Male", age_days, 0),
         logAge_days = log(age_days, base=10)) %>%
  mutate(logAge_days = ifelse(is.infinite(logAge_days),0.001, logAge_days)) %>% #make sure no inf values
  mutate(sexMale_x_logAge = sexMale * logAge_days, #add log-scaled ages
         mean.CT = rowMeans(across(all_of(ct_list)), na.rm=FALSE),
         total.SA = rowSums(across(all_of(sa_list)), na.rm=FALSE),
         study = as.factor(study),
         study_site = paste0("study-", study, "_", "site-", site),
         INDEX.ID = paste(study, participant, sex, sep="|")) %>%
  mutate(study_site = as.factor(stringr::str_replace(study_site, "site-site-", "site-"))) #remove duplicates

nrow(df_filt) -  nrow(df_filt1)
#skim(df_filt)
```

keep only columns we need (phenotypes to model, covariates)

```{r}
covar_list <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age",
                "fs_version", "fs_version_SA", "fs_version_CT", "fs_version_GM",
                "age_days", "logAge_days", "sexMale_x_logAge", "site", "country", "study_site", 
                "QC.SS", "QC.RCA", "SurfaceHoles", 
                "eTIV", "TBV", "mean.CT", "total.SA")

df_select <- df_filt %>%
  mutate(TBV = WMV + sGMV + GMV + CBV) %>% #carriess forward NAs
  dplyr::select(all_of(c(pheno_list, covar_list)))

#skim(df_select)
```

### Scan QC filtering
```{r}
qc_df <- df_select %>%
  mutate(QC.SS = ifelse(is.na(QC.SS), FALSE, TRUE),
         QC.RCA = ifelse(is.na(QC.RCA), FALSE, TRUE),
         SurfaceHoles = ifelse(is.na(SurfaceHoles), FALSE, TRUE))

table(qc_df$fs_version, qc_df$QC.SS)
table(qc_df$fs_version, qc_df$QC.RCA)

#euler filt
study_mad <- df_select %>%
  filter(is.na(QC.SS) & is.na(QC.RCA)) %>% #only use euler for non-ss
  drop_na(SurfaceHoles) %>%
  group_by(study) %>%
  summarize(study_euler_mad = mad(SurfaceHoles),
            study_euler_median = median(SurfaceHoles)) %>%
  mutate(euler_threshold = study_euler_median+(4*study_euler_mad))

n_unique(study_mad$study)

left_join(df_select, study_mad) %>%
  mutate(scan_pass = ifelse(SurfaceHoles <= euler_threshold, "yes", "no")) %>%
  filter(!is.na(scan_pass)) %>%
  summarise(n_fail = sum(scan_pass=="no"))

left_join(df_select, study_mad) %>%
  mutate(scan_pass = ifelse(SurfaceHoles <= euler_threshold, "yes", "no")) %>%
  filter(!is.na(scan_pass)) %>%
  group_by(study) %>%
  summarise(fail_percent = (sum(scan_pass == "no") / n())*100) %>%
  na.omit() %>%
  summarise(max_fail = max(fail_percent),
            min_fail=min(fail_percent),
            mean_fail=mean(fail_percent))

left_join(df_select, study_mad) %>%
  filter(!is.na(study_euler_median)) %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=SurfaceHoles)) +
  geom_point(aes(x=study, y=(study_euler_median+4*study_euler_mad)), color="red") +
  theme(axis.text.x = element_text(angle=90, hjust=0.5))

ss <- df_select %>%
  filter(QC.SS <= 0.65 | QC.RCA <= 0.65)

df_qc_filt <- left_join(df_select, study_mad) %>%
  mutate(has_qc = ifelse(is.na(SurfaceHoles) & is.na(QC.SS) & is.na(QC.RCA),
                        FALSE, TRUE)) %>%
  dplyr::filter(SurfaceHoles <= euler_threshold | QC.SS > 0.65 | QC.RCA > 0.65 | has_qc == FALSE) #using SS recommended threshold

nrow(df_select) - nrow(df_qc_filt)
```

```{r}
both_qc <- df_select %>%
  filter(!is.na(QC.SS) & !is.na(QC.RCA))
nrow(both_qc)

unique(both_qc$study)
unique(both_qc$fs_version)

both_qc %>% 
  filter(QC.SS > 0.65 & QC.RCA <= 0.65) %>%
  nrow()

both_qc %>% 
  filter(QC.SS <= 0.65 & QC.RCA > 0.65) %>%
  nrow()

both_qc %>%
  pivot_longer(cols=c(QC.SS, QC.RCA), values_to = "QC", names_to = "type") %>%
  ggplot() +
  geom_histogram(aes(x=QC, fill=type), position="identity", alpha=.6) +
  geom_vline(xintercept=0.65)
```


look at IDPs=0
```{r}
df_zeros <- df_qc_filt %>%
  filter(if_any(any_of(pheno_list), ~ . == 0))

df_zeros %>%
  select(all_of(pheno_list)) %>%
  map_int(~ sum(.x == 0, na.rm=TRUE))

ggplot(df_zeros) +
  geom_histogram(aes(x=age_days, fill=study)) +
  theme(legend.position="none")
```

Going to remove all subjects with 0s

```{r}
df_nozero <- df_qc_filt %>%
  filter(if_all(pheno_list, ~ is.na(.) | . != 0))
stopifnot((nrow(df_qc_filt) - nrow(df_nozero)) == nrow(df_zeros))
```


Look into scans w/o standard QC metrics to see if there are other measures to use

```{r}
no_qc <- df_nozero %>%
  filter(has_qc == FALSE)
(nrow(no_qc)/nrow(df_nozero))*100 # % subjects w/o QC scores

unique(no_qc$study)

#studies with QC missing from at least 50% subjects
df_nozero %>%
  group_by(study) %>%
  summarise(pct_noqc = sum(has_qc==FALSE)/n()) %>%
  filter(pct_noqc > .5)
```

```{r}
ggplot(df_nozero) +
  geom_histogram(aes(x=age_days, fill=has_qc))
```


### Final Filtering

filter to just controls, single scan per subject, drop sites with fewer than 5 ppl

```{r}
unique(df_nozero$dx)[grep("cn|control|none", unique(df_nozero$dx), ignore.case=TRUE)]

df_clean1 <- df_nozero %>% 
  dplyr::filter(age_days <= (100*365.25+280))
nrow(df_nozero) - nrow(df_clean1)

df_clean2 <- df_clean1 %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(dx == "CN")
nrow(df_clean1) - nrow(df_clean2)
n_unique(df_clean2$INDEX.ID)

#fwrite(df_clean2, "./data/euler_tmp.csv")

df_clean <- df_clean2 %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup() %>%
  group_by(study_site) %>%
  filter(n_distinct(INDEX.ID) >=5) %>% #remove sites with < 5 ppl
  ungroup() %>%
#remove SA from dHCP due to segmentation issues
  mutate(across(
    contains("SA."),
    ~ if_else(study == "dHCP-fetal", NA_real_, .x)
  ))
```


basic demographics

```{r}
nrow(df_clean)
table(df_clean$sexMale)
table(df_clean$sexMale)[[1]] -> n_female
n_female/nrow(df_clean) * 100

min(df_clean$age_days)
(max(df_clean$age_days) - 280)/365.25 #yrs from birth
```


visualize IDPs for consistency 

```{r, eval=FALSE, message=FALSE}
for (pheno in pheno_list){
  plt <- ggplot(df_clean) +
    geom_point(aes(x=logAge_days, y=.data[[pheno]], color=study)) +
    theme(legend.position="none") +
    ggtitle(pheno)
  
  print(plt)
  
  ggsave(paste0("./figs/", pheno, "logAge.png"), plt)
}
```


```{r}
df_clean %>%
  filter(CSF > 2e5) %>%
  summarise(study = unique(study))

df_clean %>%
  filter(logAge_days > 3 & CBV < 50000) %>%
  summarise(study=unique(study))

df_clean %>%
  mutate(is.study = ifelse(study=="OpenNeuro_Wang", TRUE, FALSE)) %>%
  ggplot() +
  geom_point(aes(x=logAge_days, y=CBV, color=is.study))
```

write out full dataframe
```{r}
#fwrite(df_clean, "./data/v3_CN_cleaned.csv")
```

###Split for CV Modeling

Splitting data into 50-50 train and testing, stratified by site and sex. Using this [reference](https://stackoverflow.com/questions/74573270/stratified-data-splitting-in-r).

```{r, eval=FALSE}
table(df_clean$study_site, df_clean$sexMale)
```

going to drop small sites (<10 subjects per sex) for now, but alternatives could be grouping small sites together? 

```{r}
valid_sites <- df_clean %>%
  mutate(sex=as.factor(sexMale)) %>%
  group_by(study_site, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(study_site) %>%
  summarise(
    valid = all(c("0", "1") %in% sex) && all(n[match(c("0", "1"), sex, nomatch = 0)] >= 10),
    .groups = "drop"
  ) %>%
  filter(valid) %>%
  pull(study_site)

df_clean_sites <- df_clean %>%
  filter(study_site %in% valid_sites)

nrow(df_clean) - nrow(df_clean_sites)

ggplot(df_clean) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))
ggplot(df_clean_sites) +
  geom_histogram(aes(x=age_days, fill=as.factor(sexMale)))

#table(df_filt$study_site, df_filt$sexMale)
```

Still have almost 100k subjects, proceeding to divide into A and B samples

```{r}
set.seed(5678)
#add sex factor var
df_clean_sites <- df_clean_sites %>% mutate(sex=as.factor(sexMale))

split_samples <- splitstackshape::stratified(df_clean_sites, group=c("study_site", "sex"), size=0.5, bothSets = TRUE)

samp_A <- split_samples$SAMP1
samp_B <- split_samples$SAMP2

ggplot(samp_A) +
  geom_histogram(aes(x=age_days, fill=sex))
ggplot(samp_B) +
  geom_histogram(aes(x=age_days, fill=sex))

samp_A$sample <- "A"
samp_B$sample <- "B"
full_df <- rbind(samp_A, samp_B)

samp_plt <- ggplot(full_df)+
  geom_violin(aes(x=sample, y=age_days, fill=as.factor(sexMale))) +
  theme_linedraw() +
  labs(fill = "Male")
ggsave("figs/cv_samples.png", samp_plt)
```
pretty good distributions of ages and Euler within sex, no significant differences between samples 

save
```{r}
#fwrite(split_samples$SAMP1, "./data/cv_sample_A.csv")
#fwrite(split_samples$SAMP2, "./data/cv_sample_B.csv")
```

also saving sim_df() versions for plotting consistently on across phenos
```{r}
rm_list <- c(gmv_list, subcort_list, ct_list, sa_list)

sampleA <- fread("./data/cv_sample_A.csv") %>%
  select(!rm_list)
sampleB <- fread("./data/cv_sample_B.csv") %>%
  select(!rm_list)

sampleA_sim <- sim_data(sampleA, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")
sampleB_sim <- sim_data(sampleB, "logAge_days", factor_var="sexMale", special_term = "sexMale_x_logAge = sexMale * logAge_days")

#saveRDS(sampleA_sim, "./data/cv_sample_A_sim.rds")
#saveRDS(sampleB_sim, "./data/cv_sample_B_sim.rds")
```

###Subdivide by Feature Category

Write out separate csvs for each pheno category. These will be further subdivided into separate CSVs with maximally complete data in order to ComBat features (`subjobs_datasplit.sh` and `split_data_for_combat.R`), so for each feature (target feature to ComBat) keep as many other features as possible (as priors) without including any NAs.

```{r}
#save out QC separately since these would only be used for sensitivity analyses
df_covars <- df_clean %>%
  dplyr::select(all_of(covar_list))
#fwrite(df_covars, file="./data/v3_CN_covars.csv")

covar_subset <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age", 
                  "age_days", "logAge_days", "sexMale_x_logAge", "study_site")

#global vols
df_global <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, glob_vol_list, 
                         "fs_version", "fs_version_SA", "fs_version_CT", "fs_version_GM", 
                         "eTIV", "TBV"))) %>%
  filter_at(glob_vol_list, any_vars(complete.cases(.)))
#fwrite(df_global, "./data/csvs_by_pheno/global_vols.csv")

#regional gmv
df_gmv <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, gmv_list, "fs_version_GM", "eTIV")))%>%
  filter_at(gmv_list, any_vars(complete.cases(.)))
#fwrite(df_gmv, "./data/csvs_by_pheno/cortical_vols.csv")

#regional sa
df_sa <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, sa_list, "fs_version_SA", "total.SA")))%>%
  filter_at(sa_list, any_vars(complete.cases(.)))
#fwrite(df_sa, "./data/csvs_by_pheno/cortical_surf.csv")

#regional ct
df_ct <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, ct_list, "fs_version_CT", "mean.CT")))%>%
  filter_at(ct_list, any_vars(complete.cases(.)))
#fwrite(df_ct, "./data/csvs_by_pheno/cortical_thickness.csv")

#subcortical vols
df_sub <- df_clean %>%
  dplyr::select(any_of(c(covar_subset, subcort_list, "fs_version_GM", "eTIV")))%>%
  filter_at(vars(any_of(subcort_list)), any_vars(complete.cases(.)))
#fwrite(df_sub, "./data/csvs_by_pheno/subcortical_vols.csv")
```

## Patient Sample

save patient sample
```{r}
df_pt <- df_clean1 %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(dx != "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

nrow(df_pt)
table(df_pt$dx)

#check most common dx
df_pt %>%
  count(dx) %>%
  arrange(desc(n)) %>%
  slice(1:40)
```

```{r}
#recode dx
#edit to remove family/risk categories
exclude_family_risk <- function(x) {
  x[!grepl("\\b(risk|family|relative|sibling)\\b", x, ignore.case = TRUE)]
}
mdd_list <- c(grep("Major Depressive Disorder|Dysthymia|depression", unique(df_pt$dx), ignore.case=TRUE, value=TRUE), "Depression", "MDD") %>% 
  unique() %>% 
  exclude_family_risk()

alz_list <- c(grep("alz", unique(df_pt$dx), ignore.case=TRUE, value=TRUE), "AD") %>% 
  unique() %>%
  exclude_family_risk()

gad_list <- c(grep("generalized anxiety|GAD", unique(df_pt$dx), ignore.case=TRUE, value=TRUE), "GAD") %>% 
  unique() %>%
  exclude_family_risk()

asd_list <- c(grep("autism|ASD", unique(df_pt$dx), ignore.case=TRUE, value=TRUE), "ASD") %>% 
  unique() %>%
  exclude_family_risk() 
#editing to remove subthreshold, 22q
asd_list <- asd_list[!grepl("22q|Not full", asd_list, ignore.case = TRUE)]

scz_list <- c(grep("schizo|SCZ", unique(df_pt$dx), ignore.case=TRUE, value=TRUE), "SCZ", "Psychosis") %>% 
  unique() %>%
  exclude_family_risk()
#remove schizotypy
scz_list <- scz_list[!grepl("SCZTypy", scz_list, ignore.case = TRUE)]

adhd_list <- c("ADHD", "Attention-Deficit/Hyperactivity Disorder")

mci_list <- c(grep("mci", unique(df_pt$dx), ignore.case=TRUE, value=TRUE), "MCI", "LMCI", "EMCI") %>% 
  unique() %>%
  exclude_family_risk()

df_pt_recode <- df_pt %>%
  mutate(dx_recode = case_when(
    dx %in% scz_list ~ "SCZ",
    dx %in% alz_list ~ "ALZ",
    dx %in% asd_list ~ "ASD",
    dx %in% mci_list ~ "MCI",
    dx %in% mdd_list ~ "MDD",
    dx %in% gad_list ~ "GAD",
    dx %in% adhd_list ~ "ADHD",
    TRUE ~ NA
  )) %>%
  filter(study_site %in% valid_sites) #sites to match controls

table(df_pt_recode$dx_recode)
sum(!is.na(df_pt_recode$dx_recode))/nrow(df_pt)

df_pt_recode %>%
  filter(is.na(dx_recode)) %>%
  count(dx) %>%
  arrange(desc(n)) %>%
  slice(1:40)

df_pt_recode %>%
  filter(study=="UKB") %>%
  group_by(dx_recode) %>%
  summarise(n=n())

#drop NAs
df_pt_recode2 <- df_pt_recode %>%
  tidyr::drop_na(dx_recode) %>% 
  #removing MCI, too small
  filter(dx_recode != "MCI") %>%
  #remove UKB ASD and SCZ - very few pts vs many controls -> skews sample
  filter(!(study=="UKB" & (dx_recode=="ASD" | dx_recode=="SCZ")))
  
#fwrite(df_pt_recode2, "./data/v3_pts_cleaned.csv")
```

demographics
```{r}
ggplot(df_pt_recode2) +
  geom_histogram(aes(x=age_days, fill=dx_recode), position = "identity", alpha=.4) +
  gamlssTools::format_x_axis("lifespan_fetal", df_pt_recode2$age_days)

df_pt_recode2 %>%
  mutate(age_yrs = (age_days-280)/365.25) %>%
  group_by(dx_recode) %>%
  summarise(n = n(),
            n_male = sum(sexMale),
            pct_male = (sum(sexMale)/n())*100,
            min_age = min(age_yrs),
            max_age = max(age_yrs),
            mean_age = mean(age_yrs)
            )
#table(df_pt_recode2$dx_recode, df_pt_recode2$study_site)
```

site match controls

```{r}
#samp_A <- fread("./data/cv_sample_A.csv")
#samp_B <- fread("./data/cv_sample_B.csv")
cn_df_list <- list(samp_A, samp_B)
names(cn_df_list) <- c("A", "B")
dx_list <- unique(df_pt_recode2$dx_recode)

for (dx_val in dx_list){
  print(dx_val)
  pts <- df_pt_recode2 %>%
    filter(dx_recode == dx_val)
  
  sites <- unique(pts$study_site)
  print(length(sites))
  
  for (n in names(cn_df_list)){
  cn_df <- cn_df_list[[n]]
  
  new_df <- cn_df %>%
    filter(study_site %in% sites) %>%
    mutate(dx_recode = paste(dx, dx_val, sep="_"))
    
  full_samp <- rbind(new_df, pts, fill=TRUE)
  
  print(full_samp %>%
          mutate(age_yrs = (age_days-280)/365.25) %>%
  group_by(dx_recode) %>%
  summarise(n = n(),
            n_male = sum(sexMale),
            pct_male = (sum(sexMale)/n())*100,
            min_age = min(age_yrs),
            max_age = max(age_yrs),
            mean_age = mean(age_yrs)
            ))
  
  print(ggplot(full_samp) +
          geom_histogram(aes(x=age_days, fill=dx_recode), position="identity", alpha=.5))
  
  fname <- paste0("./data/case_control_dfs/cv_sample_", n, "_", dx_val, ".csv")
  fwrite(full_samp, fname)
  
  }
}

```


```{r}
set.seed(5678)
#add sex factor var
df_pt_recode2 <- df_pt_recode2 %>% mutate(sex=as.factor(sexMale))

split_samples <- splitstackshape::stratified(df_pt_recode2, group=c("study_site", "sex"), size=0.5, bothSets = TRUE)

samp_A <- split_samples$SAMP1 %>% droplevels()
samp_B <- split_samples$SAMP2 %>% droplevels()

ggplot(samp_A) +
  geom_histogram(aes(x=age_days, fill=sex))
ggplot(samp_B) +
  geom_histogram(aes(x=age_days, fill=sex))

samp_A$sample <- "A"
samp_B$sample <- "B"

full_df <- rbind(samp_A, samp_B)

#check diffs in study site between pts and controls
cv_A <- fread("./data/cv_sample_A.csv")
cv_B <- fread("./data/cv_sample_B.csv")
setdiff(unique(samp_A$study_site), unique(cv_A$study_site))
setdiff(unique(samp_B$study_site), unique(cv_B$study_site))

ggplot(full_df)+
  geom_violin(aes(x=sample, y=age_days, fill=as.factor(sexMale))) +
  theme_linedraw() +
  labs(fill = "Male")
```

```{r}
# fwrite(split_samples$SAMP1, "./data/pt_sample_A.csv")
# fwrite(split_samples$SAMP2, "./data/pt_sample_B.csv")
```


Power analyses for sex differences within diagnoses
```{r}
n_comp <- length(unique(full_df$dx_recode)) * length(global_list)

full_df %>%
  group_by(sample, dx_recode) %>%
  summarise(n_female = sum(sexMale==0),
            n_male = sum(sexMale==1)) %>%
  rowwise() %>%
  mutate(d = pwr.t2n.test(n1=n_female, n2=n_male, sig.level=(0.05/n_comp), power=.8)$d) %>%
  ungroup() %>%
  arrange(d)
```

only powered for pretty large effects


# Visualize Data

```{r, eval=FALSE}
df_clean <- fread("./data/v3_CN_cleaned.csv", stringsAsFactors=T)
glob_vol_list <- readRDS("./pheno_lists/global_vols.rds")
gmv_list <- readRDS("./pheno_lists/cortical_vols.rds")
subcort_list <- readRDS("./pheno_lists/subcortical_vols.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)
```


```{r}
library(gamlssTools)
for (pheno in glob_vol_list){
  df <- df_clean %>%
  dplyr::select(all_of(c(pheno, "fs_version_GM", "logAge_days", "sexMale_x_logAge", "sexMale", "study_site"))) %>%
    na.omit() %>%
    trunc_coverage("logAge_days")
  
    pheno_sym <- sym(pheno)
  
  df <- df %>%
    mutate(!!pheno_sym := log_scale(!!pheno_sym))
    
  
  model <- gamlss_lambda(pheno,
                          fs_ver="fs_version_GM",
                          fs_moment="all",
                          fam="BCCG",
                          nu_form="sexMale + pb(logAge_days, method='GAIC', k=log(nrow(df)), control = pb.control(order = 3))")
  
  model$call$data <- "df"
model$call$family <- "BCCG"
  
  fan_plot <- make_centile_fan(gamlssModel=model, df=df,
                             x_var="logAge_days",
                             color_var="sexMale",
                             get_peaks=FALSE, 
                             desiredCentiles=c(0.05, 0.25, 0.5, 0.75, 0.95),
                             remove_point_effect = "study_site",
                             y_scale=unscale)
  print(fan_plot)
}

```

check ranges

```{r}
get_min <- function(p, df){
  min(df[[p]], na.rm=T)
}
pheno_mins <- lapply(pheno_list, get_min, df_clean)
```

look at data in ages <= 3 years

```{r}
df_clean_inf <- df_clean %>%
  select(c(pheno_list, sexMale, age_days, country, study, logAge_days)) %>%
  dplyr::filter(age_days <= (365.25*3)+280) %>%
  mutate(across(all_of(pheno_list), ~ ifelse(!is.na(.), TRUE, FALSE))) %>%
  pivot_longer(pheno_list, names_to = "pheno", values_to = "has_data") %>%
  dplyr::filter(has_data == TRUE) %>%
  mutate(pheno_cat = case_when(pheno %in% glob_vol_list ~ "global",
                               pheno %in% gmv_list ~ "cortical_vol",
                               pheno %in% subcort_list ~ "subcortical",
                               pheno %in% ct_list ~ "thickness",
                               pheno %in% sa_list ~ "surface_area"))

ggplot(df_clean_inf) +
  geom_histogram(aes(x=logAge_days, fill=pheno), position = "identity", alpha=0.4) +
  theme(legend.position="none") +
  geom_vline(xintercept=log(280, base=10)) + 
  # geom_vline(xintercept=(28*7)) + #3rd trimester starts
  # geom_vline(xintercept=(13*7)) + #2rd trimester starts
  facet_wrap(~pheno_cat)

ggplot(df_clean_inf) +
  geom_histogram(aes(x=logAge_days, fill=pheno_cat, color=pheno_cat), position = "identity", alpha=0.4) +
  geom_vline(xintercept=log(280, base=10))
```

for whatever reason there's very little data ~ 2 months post-birth



### Euler QC
looking at associations between Euler and demographics (in control sample and held-out psychosis sample)

```{r}
euler_df <- df_clean %>%
  filter(!is.na(SurfaceHoles)) %>%
  mutate(sexMale = as.factor(sexMale))

euler_df %>%
  ggplot(aes(x=sexMale, y=SurfaceHoles)) +
  geom_boxplot() +
  stat_compare_means()

euler_df %>%
  ggplot(aes(x=age_days, y=SurfaceHoles)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")
```

Try filtering more?
```{r}
names(euler_df)

euler_df_strict <- euler_df %>%
  mutate(euler_threshold = study_euler_median + (3.5*study_euler_mad)) %>%
  dplyr::filter(SurfaceHoles <= euler_threshold)

euler_df_strict %>%
  ggplot(aes(x=age_days, y=SurfaceHoles)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")

nrow(euler_df) - nrow(euler_df_strict)
```
loose a lot of subjects, doesn't seem to affect relationship

```{r}
euler_df_strict %>%
  ggplot(aes(x=age_days, y=eTIV)) +
  geom_point() +
  geom_hline(yintercept=4553071)
```

Look at psychosis sample
```{r}
psych_df_eu <- psychosis_df %>%
   filter(!is.na(SurfaceHoles)) %>%
  mutate(sexMale = as.factor(sexMale))
```

```{r}
unique(psych_df_eu$dx)
#collapsing all non-CN diagnoses for now
psych_df_eu <- psych_df_eu %>%
  mutate(dx_sum = case_when(dx=="CN" ~ "CN",
                            is.na(dx) ~ NA,
                            TRUE ~ "PS"))
psych_df_eu %>%
  ggplot(aes(x=sexMale, y=SurfaceHoles, fill=dx_sum)) +
  geom_boxplot() +
  stat_compare_means()

psych_df_eu %>%
  ggplot(aes(x=age_days, y=SurfaceHoles, fill=dx_sum, color=dx_sum)) +
  geom_point(alpha=0.4) +
  geom_smooth(method = "lm")
```

### Global Phenotypes
```{r}
df_gmv <- df_global %>%
  dplyr::select(age_days, sexMale, GMV) %>%
  na.omit()

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=GMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  scale_y_continuous(limits = c(0, NA))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=WMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  scale_y_continuous(limits = c(0, NA))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=sGMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  scale_y_continuous(limits = c(0, NA))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=CSF,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  scale_y_continuous(limits = c(0, NA))
```
```{r}
ggplot(df_global, aes(x=log(age_days), y=eTIV)) + 
  geom_point(aes(color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  # geom_smooth(method=lm, color="#264762") +
  # geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x)) +
  scale_y_continuous(limits = c(0, NA))

ggplot(df_global, aes(x=age_days, y=eTIV)) + 
  geom_point(aes(color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x)) +
  scale_y_continuous(limits = c(0, NA))

df_global <- df_global %>%
  mutate(TBV = WMV + sGMV + GMV)

ggplot(df_global, aes(x=log(age_days), y=TBV)) + 
  geom_point(aes(color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280)) +
  # geom_smooth(method=lm, color="#264762") +
  # geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x)) +
  scale_y_continuous(limits = c(0, NA))
```

```{r}
mad_etiv <- mad(df_clean$eTIV, na.rm = T)
med_etiv <- median(df_clean$eTIV, na.rm = T)

ggplot(df_clean, aes(y=eTIV)) + 
  geom_boxplot() +
  geom_hline(yintercept = med_etiv + 8*mad_etiv) +
  scale_y_continuous(limits = c(0, NA))

mad_tbv <- mad(df_clean$TBV, na.rm = T)
med_tbv <- median(df_clean$TBV, na.rm = T)

ggplot(df_clean, aes(y=TBV)) + 
  geom_boxplot() +
  geom_hline(yintercept = med_tbv + 8*mad_tbv) +
  scale_y_continuous(limits = c(0, NA))

mad_ct <- mad(df_clean$mean.CT, na.rm = T)
med_ct <- median(df_clean$mean.CT, na.rm = T)

ggplot(df_clean, aes(y=mean.CT)) + 
  geom_boxplot() +
  geom_hline(yintercept = med_ct + 8*mad_ct) +
  scale_y_continuous(limits = c(0, NA))

mad_sa <- mad(df_clean$total.SA, na.rm = T)
med_sa <- median(df_clean$total.SA, na.rm = T)

ggplot(df_clean, aes(y=total.SA)) + 
  geom_boxplot() +
  geom_hline(yintercept = med_sa + 8*mad_sa) +
  scale_y_continuous(limits = c(0, NA))
```

### Cortical volumes
```{r}
ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.lingual,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.parsorbitalis,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.fusiform,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

#compare to original data

ggplot(df) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal, color=sex), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))
```

###Cortical Thickness
```{r}
ggplot(df_ct) + 
  geom_point(aes(x=log(age_days), y=mean.CT, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))+
  scale_y_continuous(limits = c(0, NA))

#compare to full dataset
ggplot(df_clean) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.parahippocampal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.caudalanteriorcingulate, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.transversetemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Cortical SA
```{r}
ggplot(df_sa) + 
  geom_point(aes(x=log(age_days), y=total.SA, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))+
  scale_y_continuous(limits = c(0, max(df_sa$total.SA, na.rm=TRUE)))

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.pericalcarine, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.lateralorbitofrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.entorhinal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
df_sa %>%
  dplyr::filter(rh.SA.rostralmiddlefrontal < 3000 & age_days > 2500) %>%
  group_by(fs_version_SA, study) %>%
  summarise(n = n())

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=lh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)
```

values seem low, re-check post-ComBat

###Subcortical Volumes
```{r}
ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Hippocampus, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Accumbens.area, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Amygdala, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Right.Thalamus.Proper, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###eTIV scaling

```{r}
df_clean %>%
  filter(eTIV < 3e6) %>%
  ggplot(aes(x=eTIV, y=SUBC.Left.Amygdala)) +
  geom_point(alpha=0.6, color="gray") +
  geom_smooth(method=lm, color="#264762") +
  geom_smooth(color="#3e5410", method = "gam", formula = y ~ s(x, bs = "cs", k=5)) +
  theme_bw()
```

###Visualize Batch Effects

using simple PCA
```{r}
df_phenos <- df_clean %>% dplyr::select(all_of(c("study_site", "study", ct_list, gmv_list, sa_list, subcort_list))) %>% na.omit()

pc_phenos <- prcomp(df_phenos[,-(1:2)], center=TRUE, scale=TRUE)

as.data.frame(pc_phenos$x) %>% 
  mutate(site = df_phenos$study_site) %>%
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(aes(color=site), alpha=0.4)+
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none")

as.data.frame(pc_phenos$x) %>% 
  mutate(site = df_phenos$study) %>%
  ggplot(aes(x=PC1, y=PC2)) +
  geom_point(aes(color=site), alpha=0.4)+
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme(legend.position="none")
```
