#EDA

Script for cleaning and preparing data. Using dataset compiled by Lena, tentatively to be LBCC release v 5.0

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats)
```

load data
```{r}
df <- fread("./data/240820-L.R.LBCC.csv")
names(df)
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$dx))
sum(is.na(df$session))
```

rename/calculate global volumes

```{r}
#cortical GMV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'TotalGM'] <- 'GMV'

#sGMV
names(df)[grepl("subc", names(df), ignore.case=T)]
names(df)[names(df) == 'SubCortGM'] <- 'sGMV'

#WMV
names(df)[grepl("white", names(df), ignore.case=T)]
names(df)[names(df) == 'CerebralWhiteMatterVol'] <- 'WMV'

#CSF (ventricles)
names(df)[grepl("brainseg", names(df), ignore.case=T)]
df <- df %>%
  mutate(CSF = BrainSegVol - BrainSegVolNotVent)
```


set aside selected psychosis studies
```{r}
psych_dx_list <- c(unique(df$dx)[grepl("scz|sz|psych|schizo", unique(df$dx), ignore.case=T)], "CN")

unique(df$study)
df %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(study, dx) %>%
  summarise(count = n())

scz_study_list <- c("ALSPAC", "AMP-SCZ", "BHRCS", "COBRE_FS6_part1", "Dublin_FS6", "HCP-EarlyPsychosis", "MCI", "Maastricht_FS6",
                    "NAPLS3", "NDAR-CCDP", "NDAR-PsychiatricCerebralBlood", "NDAR-PsychosisOMS", "NDAR-SCZAging", "NDAR-SCZEmot",
                    "NDAR-SCZGlu", "NDAR-SocialBrainSchizoAutism", "NDAR_BSNIP", "NDAR_PARDIP", "NDAR_SCZIowa", "OpenNeuro_LA5c",
                    "SRPBS1600", "TOPSY")

psychosis_df <- df %>%
  dplyr::filter(study %in% scz_study_list) %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

ggplot(psychosis_df) +
  geom_histogram(aes(x=(age_days/365.25), fill=dx))

psychosis_df %>%
  group_by(dx) %>%
  summarise(count=n())
```

write out csv
```{r}
fwrite(psychosis_df, file="./data/240820-L.R.LBCC_scz_studies.csv")
```

FOR MORE SCZ STUDIES TO BE ADDED, SEE NDAR PROPOSAL GOOGLE DOC 

remove psychosis studies from main dataset so there's no data leakage
```{r}
df_nopsych <- df %>%
  dplyr::filter(!(study %in% scz_study_list))
```


look into subjects missing dx
```{r}
no_dx <- df_nopsych %>% dplyr::filter(is.na(dx))
hist((no_dx$age_days)/365.25)
```

subjects missing diagnoses are not consistently infants, going to remove

filter to just controls, single scan per subject

```{r}
unique(df_nopsych$dx)[grep("cn|control|none", unique(df_nopsych$dx), ignore.case=TRUE)]

df_filt <- df_nopsych %>%
  dplyr::filter(sex != "intersex") %>% #removing 5 intersex scans
  tidyr::drop_na(dx) %>%
  dplyr::filter(age_days <= (101*365.25)) %>%
  dplyr::filter(dx == "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()
```

```{r}
ggplot(df_filt) +
  geom_histogram(aes(x=(age_days/365.25), fill=sex), position="identity", alpha=.6)
```

write out data csv

```{r}
fwrite(df_filt, file="./data/SBC_LBCC_controls.csv")
```

write out lists of phenotypes by category
```{r}
names(df_filt)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF")
write_lines(glob_vol_list, "global_vols.txt")

#cortical gmv
gmv_list <- names(df_filt)[grep("GM\\.", names(df_filt))]
length(gmv_list) == 68
write_lines(gmv_list, "cortical_vols.txt")

#subcortical gmv
subcort <- names(df_filt)[grep("SUBC\\.Right|SUBC\\.Left", names(df_filt))]
subcort_list <- subcort[-grep("Cerebellum|WM|Vent|choroid|vessel", subcort)] #drop WM, cerebellum, etc
length(subcort_list)
write_lines(subcort_list, "subcortical_vols.txt")

#cortical thickness
ct_list <- names(df_filt)[grep("CT\\.", names(df_filt))]
length(ct_list) == 68
write_lines(ct_list, "cortical_thickness.txt")

#cortical SA
sa_list <- names(df_filt)[grep("SA\\.", names(df_filt))]
length(sa_list) == 68
write_lines(sa_list, "cortical_surf.txt")
```
