#EDA

Script for cleaning and preparing data. Using dataset compiled by Lena, tentatively to be LBCC release v 5.0

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr)
```

load data
```{r}
df <- fread("./data/240820-L.R.LBCC.csv")
names(df)
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$dx))
sum(is.na(df$session))
```

```{r}
#look at missing country info - should be findable
df %>%
  dplyr::filter(is.na(country)) %>%
  group_by(study) %>%
  summarise(n = n())

#adding back country data based on OG nature paper

df <- df %>%
  dplyr::mutate(country = case_when(is.na(country) & (study == "SYS-Adolescents" | study == "SYS-Adults") ~ "Canada",
                                    is.na(country) & study == "TOPSY" ~ "Canada",
                                    is.na(country) & study == "dHCP-fetal" ~ "UK",
                                    TRUE ~ country))
```

##Calculate Global Vols

first from freesurfer-derived outputs (rename)

```{r}
#cortical GMV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'TotalGM'] <- 'GMV'

#sGMV
names(df)[grepl("subc", names(df), ignore.case=T)]
names(df)[names(df) == 'SubCortGM'] <- 'sGMV'

#WMV
names(df)[grepl("white", names(df), ignore.case=T)]
names(df)[names(df) == 'CerebralWhiteMatterVol'] <- 'WMV'

#CSF (ventricles)
names(df)[grepl("brainseg", names(df), ignore.case=T)]
df <- df %>%
  mutate(CSF = BrainSegVol - BrainSegVolNotVent)
```

then from synthseg scans
```{r}
#add ID for merging scans
df <- df %>% mutate(merge_id = row_number())

df_synthseg <- df %>%
  dplyr::filter(fs_version == "synthseg-robust")

#confirm all missing
stopifnot(sum(is.na(df_synthseg$GMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$sGMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$WMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$CSF)) == nrow(df_synthseg))

#find appropriate cols to sum
df_synthseg %>%
  dplyr::select(all_of(c(contains("white"), contains("cort"), contains("hemi"), contains("ven")))) %>%
  skim()

df_synthseg %>%
  dplyr::select(contains("SUBC")) %>%
  skim()

ss_subcort_list <- c("SUBC.Left.Thalamus.Proper", "SUBC.Right.Thalamus.Proper", 
                     "SUBC.Left.Caudate", "SUBC.Right.Caudate",
                     "SUBC.Left.Putamen", "SUBC.Right.Putamen",
                     "SUBC.Left.Pallidum", "SUBC.Right.Pallidum",
                     "SUBC.Left.Hippocampus", "SUBC.Right.Hippocampus",
                     "SUBC.Left.Amygdala", "SUBC.Right.Amygdala",
                     "SUBC.Left.Accumbens.area", "SUBC.Right.Accumbens.area")

#calculate
df_glob <- df %>%
  mutate(GMV = ifelse(fs_version == "synthseg-robust", 
                      rowSums(across(c(rhCortexVol, lhCortexVol)), na.rm=FALSE),
                      GMV),
         WMV = ifelse(fs_version == "synthseg-robust",
                      rowSums(across(c(rhCerebralWhiteMatterVol, lhCerebralWhiteMatterVol)), na.rm=FALSE),
                      WMV),
         CSF = ifelse(fs_version == "synthseg-robust", Ventricles.all, CSF),
         sGMV = ifelse(fs_version == "synthseg-robust",
                       rowSums(across(all_of(ss_subcort_list)), na.rm=FALSE),
                       sGMV))
stopifnot(nrow(df_glob) == nrow(df))
```

plot for comparisons
```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

sum(is.na(df$GMV)) - sum(is.na(df_glob$GMV))
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280) +
  facet_wrap(~fs_version)
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

CSF has s ome outliers but overall looks good, going to check with Lena but proceed for now

check for any more big groups of missing global volumes
```{r}
df_glob %>%
  dplyr::filter(is.na(GMV&sGMV&WMV&CSF)) %>%
  group_by(fs_version) %>%
  summarise(n = n())
```

## Data Cleaning
remove intersex subj, dummy-code sex and sex*age, combine Euler and SS QC metrics (because of missingness)
also fix age calculations for `OpenNeuro-PaingenPlacebo` study - study reported years, accidentally calculated as months

```{r}
df_filt <- df_glob %>%
  dplyr::filter(sex != "intersex") %>%
  mutate(age_days = ifelse(study != "OpenNeuro-PaingenPlacebo", age_days,
                           ((age_days-280)*(12/365.25))*365.25+280)) %>% #convert back from days to 'months', then re-convert years to days
  mutate(sexMale = ifelse(sex=="male", 1, 0),
         sexMale_x_age = ifelse(sex=="male", age_days, 0),
         QC_Metric = ifelse(fs_version == "synthseg-robust", QC.SS, SurfaceHoles),
         logAge_days = log(age_days, base=10)) %>% #only keep QC.SS for scans processed with synthseg]
  mutate(sexMale_x_logAge = sexMale * logAge_days) #add log-scaled ages
```

write out lists of phenotypes by category
```{r}
#names(df_filt)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF")
write_lines(glob_vol_list, "./pheno_lists/global_vols.txt")
saveRDS(glob_vol_list, "./pheno_lists/global_vols.rds")

#cortical gmv
gmv_list <- names(df_filt)[grep("GM\\.", names(df))]
stopifnot(length(gmv_list) == 68)
write_lines(gmv_list, "./pheno_lists/cortical_vols.txt")
saveRDS(gmv_list, "./pheno_lists/cortical_vols.rds")

#subcortical gmv
subcort <- names(df_filt)[grep("SUBC\\.Right|SUBC\\.Left", names(df))]
subcort_list <- subcort[-grep("Cerebellum|WM|Vent|choroid|vessel", subcort)] #drop WM, cerebellum, etc
length(subcort_list)
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")

#cortical thickness
ct_list <- names(df_filt)[grep("CT\\.", names(df))]
stopifnot(length(ct_list) == 68)
write_lines(ct_list, "./pheno_lists/cortical_thickness.txt")
saveRDS(ct_list, "./pheno_lists/cortical_thickness.rds")

#cortical SA
sa_list <- names(df_filt)[grep("SA\\.", names(df))]
stopifnot(length(sa_list) == 68)
write_lines(sa_list, "./pheno_lists/cortical_surf.txt")
saveRDS(sa_list, "./pheno_lists/cortical_surf.rds")
```

keep only columns we need (phenotypes to model, covariates)

```{r}
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)
covar_list <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age",
                "age_days", "logAge_days", "sexMale_x_logAge", "site", "country", "study_site", "QC_Metric", "fs_version",
                "EstimatedTotalIntraCranialVol")

df_select <- df_filt %>%
  dplyr::select(all_of(c(pheno_list, covar_list))) %>%
  dplyr::select(where(~mean(is.na(.)) < 0.5)) #drop cols with <50% completion

skim(df_select)
```


### Remove Psychosis Studies
set aside selected psychosis studies
```{r}
psych_dx_list <- c(unique(df$dx)[grepl("scz|sz|psych|schizo|cos", unique(df$dx), ignore.case=T)], "CN")

unique(df_select$study)
df %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(study, dx) %>%
  summarise(count = n())

scz_study_list <- c("ALSPAC", "AMP-SCZ", "BHRCS", "COBRE_FS6_part1", "Dublin_FS6", "HCP-EarlyPsychosis", "MCI", "Maastricht_FS6",
                    "NAPLS3", "NDAR-CCDP", "NDAR-PsychiatricCerebralBlood", "NDAR-PsychosisOMS", "NDAR-SCZAging", "NDAR-SCZEmot",
                    "NDAR-SCZGlu", "NDAR-SocialBrainSchizoAutism", "NDAR_BSNIP", "NDAR_PARDIP", "NDAR_SCZIowa", "OpenNeuro_LA5c",
                    "SRPBS1600", "TOPSY", "NDAR_NIH-IRP") #NDAR_NIH-IRP = controls for Armin's COS dataset

psychosis_df <- df_select %>%
  dplyr::filter(study %in% scz_study_list) %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

ggplot(psychosis_df) +
  geom_histogram(aes(x=(age_days/365.25), fill=dx))

psychosis_df %>%
  group_by(dx) %>%
  summarise(count=n())
```

write out csv - *not filtering/removing duplicate scans because planning to add more data!*
```{r}
fwrite(psychosis_df, file="./data/240820-L.R.LBCC_scz_studies.csv")
```

FOR MORE SCZ STUDIES TO BE ADDED, SEE NDAR PROPOSAL GOOGLE DOC 

remove psychosis studies from main dataset so there's no data leakage
```{r}
df_nopsych <- df_select %>%
  dplyr::filter(!(study %in% scz_study_list))
```


look into subjects missing dx
```{r}
no_dx <- df_nopsych %>% dplyr::filter(is.na(dx))
ggplot(no_dx) +
  geom_histogram(aes(x=age_days)) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

no_dx %>%
  group_by(study) %>%
  summarise(mean_age = mean(age_days),
            n = n())

df_nopsych %>%
  dplyr::filter(study == "NDAR_IBIS") %>%
  group_by(dx) %>%
  summarise(n=n(),
            min_age = min(age_days),
            max_age = max(age_days))
```


subjects missing diagnoses are not consistently infants

filter to just controls, single scan per subject, drop studies with fewer than 5 ppl

```{r}
unique(df_nopsych$dx)[grep("cn|control|none", unique(df_nopsych$dx), ignore.case=TRUE)]

df_clean <- df_nopsych %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(age_days <= (101*365.25)) %>%
  dplyr::filter(dx == "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup() %>%
  group_by(study) %>%
  filter(n_distinct(INDEX.ID) >=5) %>%
  ungroup()

#look at completeness for infant/toddler data
df_clean %>%
  dplyr::filter(age_days <= (365.25*3 + 280)) %>%
  skim()
#some phenotypes around 50%


df_clean %>%
  dplyr::filter(age_days <= (365.25 + 280)) %>%
  skim()
#some as low as 25%
```

write out separate csvs for each pheno category (because combatLS and gamlss can't handle NAs and not all subjects have all phenos)

```{r}
#save out QC separately since these would only be used for sensitivity analyses and i dont want to loose lots of ppts because they're missing
df_covars <- df_clean %>%
  dplyr::select(all_of(covar_list))
fwrite(df_covars, file="./data/SBC_LBCC_CN_covars.csv")

covar_subset <- covar_list[!covar_list %in% c("QC_Metric", "run")]

#global vols
df_global <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, glob_vol_list))) %>%
  na.omit()
nrow(df_global)/nrow(df_clean)
fwrite(df_global, file="./data/SBC_LBCC_CN_globvol_tocombat.csv")

#regional gmv
df_gmv <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, gmv_list))) %>%
  na.omit()
nrow(df_gmv)/nrow(df_clean)
fwrite(df_gmv, file="./data/SBC_LBCC_CN_gmv_tocombat.csv")

#regional sa
df_sa <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, sa_list))) %>%
  na.omit()
nrow(df_sa)/nrow(df_clean)
fwrite(df_sa, file="./data/SBC_LBCC_CN_sa_tocombat.csv")

#regional ct
df_ct <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, ct_list))) %>%
  na.omit()
nrow(df_ct)/nrow(df_clean)
fwrite(df_ct, file="./data/SBC_LBCC_CN_ct_tocombat.csv")

#subcortical vols
df_sub <- df_clean %>%
  dplyr::select(any_of(c(covar_subset, subcort_list))) %>%
  na.omit()
nrow(df_sub)/nrow(df_clean)
fwrite(df_sub, file="./data/SBC_LBCC_CN_sub_tocombat.csv")
```
Also writing out full, unique dataset in case i want to run combat separately by pheno to try to preserve infants
(many of whom are missing 1+ phenotypes)
```{r}
fwrite(df_clean, file="./data/SBC_LBCC_allunique_CN.csv")
```


Visualize remaining sample
```{r}
df_list <- list(df_clean, df_global, df_gmv, df_ct, df_sub, df_sa)

ggplot(df) + geom_histogram(aes(x=age_days))

for(dataset in df_list){
  plt <- ggplot(dataset) + geom_histogram(aes(x=age_days))
  print(plt)
}
```

## Visualize Data

### Global Phenotypes
```{r}
ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=GMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=WMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=sGMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=CSF,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))
```

### Cortical volumes
```{r}
ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.lingual,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.parsorbitalis,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.fusiform,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

#compare to original data

df_test_dx <- df %>%
  mutate(dx = case_when(is.na(dx) ~ NA,
                        dx == "CN" ~ "CN",
                        TRUE ~ "notCN"))
ggplot(df_test_dx) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal, color=dx), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))
```

###Cortical Thickness
```{r}
ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

#compare to full dataset
ggplot(df_clean) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.parahippocampal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.caudalanteriorcingulate, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.transversetemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Cortical SA
```{r}
ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.pericalcarine, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.lateralorbitofrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.entorhinal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Subcortical Volumes
```{r}
ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Hippocampus, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Accumbens.area, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Amygdala, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Right.Thalamus.Proper, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

## some test models

```{r}
library(gamlss)
devtools::install_github("BGDlab/gamlssTools")
library(gamlssTools)
```

standard sex-mod model with pb()
```{r}
mod1 <- gamlss(GMV ~ pb(age_days, method="GAIC", k=log(length(df_global$GMV))) + sexMale + pb(sexMale_x_age, method="GAIC", k=log(length(df_global$GMV))) + study, family = GG, data=df_global)
make_centile_fan(mod1, df_global, "age_days", "sexMale")
```
sex-mod model with pb() on log-scaled age
```{r}
mod2 <- gamlss(GMV ~ pb(logAge_days, method="GAIC", k=log(length(df_global$GMV))) + sexMale + pb(sexMale_x_logAge, method="GAIC", k=log(length(df_global$GMV))) + study, family = GG, data=df_global)
make_centile_fan(mod2, df_global, "logAge_days", "sexMale")
```
looks better!

```{r}
irismod1 <-gamlss(Sepal.Length ~ ns(Sepal.Width) + Species, data=iris)
make_centile_fan(irismod1, iris, "Sepal.Width", "Species")

irismod2 <-gamlss(Sepal.Length ~ ns(Sepal.Width, knots=(c(mean(iris$Sepal.Width)-0.5, mean(iris$Sepal.Width), mean(iris$Sepal.Width)+0.5))) + Species, data=iris)
make_centile_fan(irismod2, iris, "Sepal.Width", "Species")

irismod3 <-gamlss(Sepal.Length ~ ns(Sepal.Width, df=4) + Species, data=iris)
make_centile_fan(irismod3, iris, "Sepal.Width", "Species")
```

sex-mod model with ns() on log-scaled age and biologically-informed spline placement
```{r}
ages_yrs<-c(1, 3, 7, 16, 25, 45, 70)
ages_logdays <- lapply(ages_yrs, function(x){log(x*365.25, base=10)}) %>% unlist

mod3 <- gamlss(GMV ~ ns(logAge_days, knots=c(3.039711, 3.407688, 3.766710, 3.960530, 4.215803, 4.407688)) + sexMale + ns(sexMale_x_logAge, df=10) + study, family = GG, data=df_global)
make_centile_fan(mod3, df_global, "logAge_days", "sexMale")
```

sex-mod model with ns() on log-scaled age
```{r}
mod4 <- gamlss(GMV ~ ns(logAge_days, df=30) + sexMale + ns(sexMale_x_logAge, df=10) + study, family = GG, data=df_global)
make_centile_fan(mod4, df_global, "logAge_days", "sexMale")
```

sex-strat model with ns() on log-scaled age and biologically-informed spline placement
```{r}
mod5 <- gamlss(GMV ~ ns(logAge_days, knots=c(2.562590, 3.039711, 3.407688, 3.766710, 3.960530, 4.215803, 4.407688)) + sexMale + study, family = GG, data=df_global)
make_centile_fan(mod5, df_global, "logAge_days", "sexMale")
```

sex-mod model with ns() on log-scaled age
```{r}
mod6 <- gamlss(GMV ~ ns(logAge_days, df=30) + sexMale + study, family = GG, data=df_global)
make_centile_fan(mod6, df_global, "logAge_days", "sexMale")
```