#Curation

Script for cleaning and preparing data. Using dataset compiled by Lena, tentatively to be LBCC release v 5.0

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel)
```

load data
```{r}
df <- fread("./data/240820-L.R.LBCC.csv")
names(df)
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$dx))
sum(is.na(df$session))
```

```{r}
#look at missing country info - should be findable
df %>%
  dplyr::filter(is.na(country)) %>%
  group_by(study) %>%
  summarise(n = n())

#adding back country data based on OG nature paper

df <- df %>%
  dplyr::mutate(country = case_when(is.na(country) & (study == "SYS-Adolescents" | study == "SYS-Adults") ~ "Canada",
                                    is.na(country) & study == "TOPSY" ~ "Canada",
                                    is.na(country) & study == "dHCP-fetal" ~ "UK",
                                    TRUE ~ country))
```

##Calculate Global Vols

first from freesurfer-derived outputs (rename)

```{r}
#cortical GMV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'TotalGM'] <- 'GMV'

#sGMV
names(df)[grepl("subc", names(df), ignore.case=T)]
names(df)[names(df) == 'SubCortGM'] <- 'sGMV'

#WMV
names(df)[grepl("white", names(df), ignore.case=T)]
names(df)[names(df) == 'CerebralWhiteMatterVol'] <- 'WMV'

#CSF (ventricles)
names(df)[grepl("brainseg", names(df), ignore.case=T)]
df <- df %>%
  mutate(CSF = BrainSegVol - BrainSegVolNotVent)

#eTIV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'EstimatedTotalIntraCranialVol'] <- 'eTIV'
```

then from synthseg scans
```{r}
#add ID for merging scans
df <- df %>% mutate(merge_id = row_number())

df_synthseg <- df %>%
  dplyr::filter(fs_version == "synthseg-robust")

#confirm all missing
stopifnot(sum(is.na(df_synthseg$GMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$sGMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$WMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$CSF)) == nrow(df_synthseg))

#find appropriate cols to sum
df_synthseg %>%
  dplyr::select(all_of(c(contains("white"), contains("cort"), contains("hemi"), contains("ven")))) %>%
  skim()

df_synthseg %>%
  dplyr::select(contains("SUBC")) %>%
  skim()

ss_subcort_list <- c("SUBC.Left.Thalamus.Proper", "SUBC.Right.Thalamus.Proper", 
                     "SUBC.Left.Caudate", "SUBC.Right.Caudate",
                     "SUBC.Left.Putamen", "SUBC.Right.Putamen",
                     "SUBC.Left.Pallidum", "SUBC.Right.Pallidum",
                     "SUBC.Left.Hippocampus", "SUBC.Right.Hippocampus",
                     "SUBC.Left.Amygdala", "SUBC.Right.Amygdala",
                     "SUBC.Left.Accumbens.area", "SUBC.Right.Accumbens.area")

#calculate
df_glob <- df %>%
  mutate(GMV = ifelse(fs_version == "synthseg-robust", 
                      rowSums(across(c(rhCortexVol, lhCortexVol)), na.rm=FALSE),
                      GMV),
         WMV = ifelse(fs_version == "synthseg-robust",
                      rowSums(across(c(rhCerebralWhiteMatterVol, lhCerebralWhiteMatterVol)), na.rm=FALSE),
                      WMV),
         CSF = ifelse(fs_version == "synthseg-robust", Ventricles.all, CSF),
         sGMV = ifelse(fs_version == "synthseg-robust",
                       rowSums(across(all_of(ss_subcort_list)), na.rm=FALSE),
                       sGMV),
         eTIV = ifelse(fs_version == "synthseg-robust" & is.na(eTIV),
                       total.intracranial, eTIV)
         )
stopifnot(nrow(df_glob) == nrow(df))
```

plot for comparisons
```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

sum(is.na(df$GMV)) - sum(is.na(df_glob$GMV))
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280) +
  facet_wrap(~fs_version)
```

```{r}
df %>%
  dplyr::filter(sGMV > (1*10^5)) %>%
  group_by(fs_version, study) %>%
  summarise(n =n())
```

SYS-Adolescents looks way off on sGMV, removing this phenotype

```{r}
df_glob_nosys <- df_glob %>%
  mutate(sGMV = ifelse(study == "SYS-Adolescents", NA, sGMV))
```


```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

CSF has some outliers but overall looks good, going to check with Lena but proceed for now

check for any more big groups of missing global volumes
```{r}
df_glob %>%
  dplyr::filter(is.na(GMV&sGMV&WMV&CSF)) %>%
  group_by(fs_version) %>%
  summarise(n = n())
```

## Data Cleaning
remove intersex subj, dummy-code sex and sex*age, combine Euler and SS QC metrics (because of missingness)
also fix age calculations for `OpenNeuro-PaingenPlacebo` study - study reported years, accidentally calculated as months

```{r}
df_filt <- df_glob_nosys %>%
  dplyr::filter(sex != "intersex") %>%
  mutate(age_days = ifelse(study != "OpenNeuro-PaingenPlacebo", age_days,
                           ((age_days-280)*(12/365.25))*365.25+280)) %>% #convert back from days to 'months', then re-convert years to days
  mutate(sexMale = ifelse(sex=="male", 1, 0),
         sexMale_x_age = ifelse(sex=="male", age_days, 0),
         QC_Metric = ifelse(fs_version == "synthseg-robust", QC.SS, SurfaceHoles),
         logAge_days = log(age_days, base=10)) %>% #only keep QC.SS for scans processed with synthseg]
  mutate(sexMale_x_logAge = sexMale * logAge_days) #add log-scaled ages
```

write out lists of phenotypes by category
```{r}
#names(df_filt)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF","eTIV")
write_lines(glob_vol_list, "./pheno_lists/global_vols.txt")
saveRDS(glob_vol_list, "./pheno_lists/global_vols.rds")

#cortical gmv
gmv_list <- names(df_filt)[grep("GM\\.", names(df))]
stopifnot(length(gmv_list) == 68)
write_lines(gmv_list, "./pheno_lists/cortical_vols.txt")
saveRDS(gmv_list, "./pheno_lists/cortical_vols.rds")

#subcortical gmv
subcort <- names(df_filt)[grep("SUBC\\.Right|SUBC\\.Left", names(df))]
subcort_list <- subcort[-grep("Cerebellum|WM|Vent|choroid|vessel", subcort)] #drop WM, cerebellum, etc
length(subcort_list)
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")

#cortical thickness
ct_list <- names(df_filt)[grep("CT\\.", names(df))]
stopifnot(length(ct_list) == 68)
write_lines(ct_list, "./pheno_lists/cortical_thickness.txt")
saveRDS(ct_list, "./pheno_lists/cortical_thickness.rds")

#cortical SA
sa_list <- names(df_filt)[grep("SA\\.", names(df))]
stopifnot(length(sa_list) == 68)
write_lines(sa_list, "./pheno_lists/cortical_surf.txt")
saveRDS(sa_list, "./pheno_lists/cortical_surf.rds")
```

keep only columns we need (phenotypes to model, covariates)

```{r}
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)
covar_list <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age",
                "age_days", "logAge_days", "sexMale_x_logAge", "site", "country", "study_site", "QC_Metric", "fs_version", "eTIV")

df_select <- df_filt %>%
  dplyr::select(all_of(c(pheno_list, covar_list)))

skim(df_select)
```


### Remove Psychosis Studies
set aside selected psychosis studies
```{r}
psych_dx_list <- c(unique(df$dx)[grepl("scz|sz|psych|schizo|cos", unique(df$dx), ignore.case=T)], "CN")

unique(df_select$study)
df %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(study, dx) %>%
  summarise(count = n())

scz_study_list <- c("ALSPAC", "AMP-SCZ", "BHRCS", "COBRE_FS6_part1", "Dublin_FS6", "HCP-EarlyPsychosis", "MCI", "Maastricht_FS6",
                    "NAPLS3", "NDAR-CCDP", "NDAR-PsychiatricCerebralBlood", "NDAR-PsychosisOMS", "NDAR-SCZAging", "NDAR-SCZEmot",
                    "NDAR-SCZGlu", "NDAR-SocialBrainSchizoAutism", "NDAR_BSNIP", "NDAR_PARDIP", "NDAR_SCZIowa", "OpenNeuro_LA5c",
                    "SRPBS1600", "TOPSY", "NDAR_NIH-IRP") #NDAR_NIH-IRP = controls for Armin's COS dataset

psychosis_df <- df_select %>%
  dplyr::filter(study %in% scz_study_list) %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

ggplot(psychosis_df) +
  geom_histogram(aes(x=(age_days/365.25), fill=dx))

psychosis_df %>%
  group_by(dx) %>%
  summarise(count=n())
```

write out csv - *not filtering/removing duplicate scans because planning to add more data!*
```{r}
fwrite(psychosis_df, file="./data/240820-L.R.LBCC_scz_studies.csv")
```

FOR MORE SCZ STUDIES TO BE ADDED, SEE NDAR PROPOSAL GOOGLE DOC 

remove psychosis studies from main dataset so there's no data leakage
```{r}
df_nopsych <- df_select %>%
  dplyr::filter(!(study %in% scz_study_list))
```


look into subjects missing dx
```{r}
no_dx <- df_nopsych %>% dplyr::filter(is.na(dx))
ggplot(no_dx) +
  geom_histogram(aes(x=age_days)) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

no_dx %>%
  group_by(study) %>%
  summarise(mean_age = mean(age_days),
            n = n())

df_nopsych %>%
  dplyr::filter(study == "NDAR_IBIS") %>%
  group_by(dx) %>%
  summarise(n=n(),
            min_age = min(age_days),
            max_age = max(age_days))
```


subjects missing diagnoses are not consistently infants

filter to just controls, single scan per subject, drop studies with fewer than 5 ppl

```{r}
unique(df_nopsych$dx)[grep("cn|control|none", unique(df_nopsych$dx), ignore.case=TRUE)]

df_clean <- df_nopsych %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(age_days <= (101*365.25)) %>%
  dplyr::filter(dx == "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup() %>%
  group_by(study) %>%
  filter(n_distinct(INDEX.ID) >=5) %>%
  ungroup()

#look at completeness for infant/toddler data
df_clean %>%
  dplyr::filter(age_days <= (365.25*3 + 280)) %>%
  skim()
#some phenotypes around 50%


df_clean %>%
  dplyr::filter(age_days <= (365.25 + 280)) %>%
  skim()
#some as low as 25%
```

determine equal knot placements 
```{r}
range(df_clean$logAge_days)

knot_lists <- c()
for (degree in seq(2, 24)){
  list_name <- paste0("df", degree)
  knot_lists[[list_name]] <- dlnm::equalknots(df_clean$logAge_days,
                                             fun="ns",
                                             df= degree)
}

saveRDS(knot_lists, "./code/knot_lists.RDS")
```

###Subdivide by Missingness

Making separate CSVs with maximally complete data in order to ComBat features. For each feature (target feature to ComBat) keep as many other features as possible (as priors) without including any NAs
write out separate csvs for each pheno category, further subdivide using `subjobs_datasplit.sh` and `split_data_for_combat.R`

```{r}
#save out QC separately since these would only be used for sensitivity analyses
df_covars <- df_clean %>%
  dplyr::select(all_of(covar_list))
fwrite(df_covars, file="./data/SBC_LBCC_CN_covars.csv")

covar_subset <- covar_list[!covar_list %in% c("QC_Metric", "run", "eTIV")]

#global vols
df_global <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, glob_vol_list))) %>%
  filter_at(glob_vol_list, any_vars(complete.cases(.)))
fwrite(df_global, "./data/csvs_by_pheno/global_vols.csv")

#regional gmv
df_gmv <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, gmv_list)))%>%
  filter_at(gmv_list, any_vars(complete.cases(.)))
fwrite(df_gmv, "./data/csvs_by_pheno/cortical_vols.csv")

#regional sa
df_sa <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, sa_list)))%>%
  filter_at(sa_list, any_vars(complete.cases(.)))
fwrite(df_sa, "./data/csvs_by_pheno/cortical_surf.csv")

#regional ct
df_ct <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, ct_list)))%>%
  filter_at(ct_list, any_vars(complete.cases(.)))
fwrite(df_ct, "./data/csvs_by_pheno/cortical_thickness.csv")

#subcortical vols
df_sub <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, subcort_list)))%>%
  filter_at(subcort_list, any_vars(complete.cases(.)))
fwrite(df_sub, "./data/csvs_by_pheno/subcortical_vols.csv")
```

## Visualize Data

### Global Phenotypes
```{r}
ggplot(df_global) + 
  geom_point(aes(x=age_days, y=GMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=WMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=sGMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=CSF,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))
```

### Cortical volumes
```{r}
ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.lingual,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.parsorbitalis,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.fusiform,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

#compare to original data

df_test_dx <- df %>%
  mutate(dx = case_when(is.na(dx) ~ NA,
                        dx == "CN" ~ "CN",
                        TRUE ~ "notCN"))
ggplot(df_test_dx) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal, color=dx), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))
```

###Cortical Thickness
```{r}
ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

#compare to full dataset
ggplot(df_clean) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.parahippocampal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.caudalanteriorcingulate, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.transversetemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Cortical SA
```{r}
ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.pericalcarine, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.lateralorbitofrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.entorhinal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
df_sa %>%
  dplyr::filter(rh.SA.rostralmiddlefrontal < 3000 & age_days > 2500) %>%
  group_by(fs_version, study) %>%
  summarise(n = n())

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)

df_sa %>%
  mutate(check = ifelse(study=="OpenNeuro_Wang", 1, 0)) %>%
  ggplot() +
  geom_point(aes(x=age_days, y=lh.SA.rostralmiddlefrontal, color=as.factor(check)), alpha=.4)
```

values seem low, re-check post-ComBat

###Subcortical Volumes
```{r}
ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Hippocampus, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Accumbens.area, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Amygdala, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Right.Thalamus.Proper, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

## some test models

```{r}
library(gamlss)
devtools::install_github("BGDlab/gamlssTools")
library(gamlssTools)

df_global <- fread("./data/csvs_by_pheno/global_vols.csv")
```

```{r}
irismod1 <-gamlss(Sepal.Length ~ ns(Sepal.Width) + Species, data=iris)
make_centile_fan(irismod1, iris, "Sepal.Width", "Species", label_centiles = FALSE)

irismod2 <-gamlss(Sepal.Length ~ ns(Sepal.Width, knots=(c(mean(iris$Sepal.Width)-0.5, mean(iris$Sepal.Width), mean(iris$Sepal.Width)+0.5))) + Species, data=iris)
make_centile_fan(irismod2, iris, "Sepal.Width", "Species", label_centiles = FALSE)

irismod3 <-gamlss(Sepal.Length ~ ns(Sepal.Width, df=4) + Species, data=iris)
make_centile_fan(irismod3, iris, "Sepal.Width", "Species", label_centiles = FALSE)
```

sex-mod model with ns() on log-scaled age and biologically-informed spline placement
```{r}
ages_yrs<-c(1, 3, 7, 16, 25, 45, 70)
ages_logdays <- lapply(ages_yrs, function(x){log(x*365.25, base=10)}) %>% unlist

mod3 <- gamlss(GMV ~ ns(logAge_days, knots=c(3.039711, 3.407688, 3.766710, 3.960530, 4.215803, 4.407688)) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod3, na.omit(df_global), "logAge_days", "sexMale", label_centiles = FALSE, show_points = FALSE)
```

non log-scaled
```{r}
knots_days <- lapply(ages_yrs, function(x){x*365.25+280}) %>% unlist

mod3_nolog <- gamlss(GMV ~ ns(age_days, knots=c(645.25, 1375.75, 2836.75, 6124.00, 9411.25, 16716.25, 25847.50)) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod3_nolog, na.omit(df_global), "age_days", "sexMale", label_centiles = FALSE)
```
looks decent, trying to add more smooths
```{r}
ages_yrs_up <- c(ages_yrs, 0, 5) %>% sort()

knots_days2 <- lapply(ages_yrs_up, function(x){x*365.25+280}) %>% unlist

mod4_nolog <- gamlss(GMV ~ ns(age_days, knots=c(280.00, 645.25, 1375.75, 2106.25, 2836.75, 6124.00, 9411.25, 16716.25, 25847.50)) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod4_nolog, na.omit(df_global), "age_days", "sexMale", label_centiles = FALSE)
```

still kinda bumpy in early years, try with log-scale

```{r}
ages_logdays2 <- lapply(ages_yrs_up, function(x){log(x*365.25, base=10)}) %>% unlist

mod4 <- gamlss(GMV ~ ns(logAge_days, knots=c(0, 2.562590, 3.039711, 3.407688, 3.766710, 3.960530, 4.215803, 4.407688)) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod4, na.omit(df_global), "logAge_days", "sexMale", label_centiles = FALSE)

```
similar overall

ns() on log-scaled age
```{r}
mod5 <- gamlss(GMV ~ ns(logAge_days, df=30) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod5, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F, show_points = F)
```

ns() on un-scaled age
```{r}
mod5_nolog <- gamlss(GMV ~ ns(age_days, df=30) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod5_nolog, na.omit(df_global), "age_days", "sexMale", label_centiles = F, show_points = F)
```

def too wiggly, try fewer dfs
```{r}
mod6 <- gamlss(GMV ~ ns(logAge_days, df=10) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod6, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)
```

ns() on un-scaled age
```{r}
mod6_nolog <- gamlss(GMV ~ ns(age_days, df=10) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(mod6_nolog, na.omit(df_global), "age_days", "sexMale", label_centiles = F)
```
not good. think that best option will be log-scaled date with a low df

```{r}
GMV_mod_pb <- gamlss(GMV ~ pb(logAge_days) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_pb, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df3 <- gamlss(GMV ~ ns(logAge_days, df=3) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df3, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df5 <- gamlss(GMV ~ ns(logAge_days, df=5) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df5, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df6 <- gamlss(GMV ~ ns(logAge_days, df=6) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df6, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df7 <- gamlss(GMV ~ ns(logAge_days, df=7) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df7, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df8 <- gamlss(GMV ~ ns(logAge_days, df=8) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df8, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df9 <- gamlss(GMV ~ ns(logAge_days, df=9) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df9, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

GMV_mod_df10 <- gamlss(GMV ~ ns(logAge_days, df=10) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(GMV_mod_df10, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)
```

```{r}
WMV_mod_pb <- gamlss(WMV ~ pb(logAge_days) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_pb, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df3 <- gamlss(WMV ~ ns(logAge_days, df=3) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df3, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df5 <- gamlss(WMV ~ ns(logAge_days, df=5) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df5, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df6 <- gamlss(WMV ~ ns(logAge_days, df=6) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df6, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df7 <- gamlss(WMV ~ ns(logAge_days, df=7) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df7, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df8 <- gamlss(WMV ~ ns(logAge_days, df=8) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df8, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df9 <- gamlss(WMV ~ ns(logAge_days, df=9) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df9, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_df10 <- gamlss(WMV ~ ns(logAge_days, df=10) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_df10, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)
```
penalized splines
```{r}
WMV_mod_pb <- gamlss(WMV ~ pb(logAge_days) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_pb, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

#play with lambda
WMV_mod_pb5 <- gamlss(WMV ~ pb(logAge_days, lambda=5) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_pb5, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

WMV_mod_pb10 <- gamlss(WMV ~ pb(logAge_days, lambda=10) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_pb10, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)

n <- nrow(na.omit(df_global))
WMV_mod_pbgaic <- gamlss(WMV ~ pb(logAge_days, method="GAIC", k=log(n)) + sexMale + study, family = GG, data=na.omit(df_global))
make_centile_fan(WMV_mod_pbgaic, na.omit(df_global), "logAge_days", "sexMale", label_centiles = F)
```


playing around with spacing splines based on epochs in [Kang et al, 2013](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3566780/#SD1).

```{r}
month_to_days <- function(x){
  d <- (x*365.25)/12 + 280
  return(d)
}

yr_to_days <- function(x){
  d <- x*365.25 + 280
  return(d)
}

#convert age windows to days
kang_windows <- c(4*7, 8*7, 10*7, 13*7, 16*7, 19*7, 24*7, 0, 
                  month_to_days(6),
                  month_to_days(12),
                  yr_to_days(6),
                  yr_to_days(12),
                  yr_to_days(20),
                  yr_to_days(40),
                  yr_to_days(60))
```