#Curation

Script for cleaning and preparing data. Using dataset compiled by Lena, tentatively to be LBCC release v 5.0

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel)
```

load data
```{r}
df <- fread("./data/240820-L.R.LBCC.csv")
names(df)
```

```{r}
table(df$sex)
table(df$session)
table(df$run)
length(unique(df$src_subject_id))
names(df)[grepl("index", names(df), ignore.case=T)]

length(unique(df$INDEX.ID))
sum(is.na(df$INDEX.ID))


sum(is.na(df$participant))
sum(is.na(df$study))
sum(is.na(df$dx))
sum(is.na(df$session))
```

```{r}
#look at missing country info - should be findable
df %>%
  dplyr::filter(is.na(country)) %>%
  group_by(study) %>%
  summarise(n = n())

#adding back country data based on OG nature paper

df <- df %>%
  dplyr::mutate(country = case_when(is.na(country) & (study == "SYS-Adolescents" | study == "SYS-Adults") ~ "Canada",
                                    is.na(country) & study == "TOPSY" ~ "Canada",
                                    is.na(country) & study == "dHCP-fetal" ~ "UK",
                                    TRUE ~ country))
```

##Calculate Global Vols

first from freesurfer-derived outputs (rename)

```{r}
#cortical GMV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'TotalGM'] <- 'GMV'

#sGMV
names(df)[grepl("subc", names(df), ignore.case=T)]
names(df)[names(df) == 'SubCortGM'] <- 'sGMV'

#WMV
names(df)[grepl("white", names(df), ignore.case=T)]
names(df)[names(df) == 'CerebralWhiteMatterVol'] <- 'WMV'

#CSF (ventricles)
names(df)[grepl("brainseg", names(df), ignore.case=T)]
df <- df %>%
  mutate(CSF = BrainSegVol - BrainSegVolNotVent)

#eTIV
names(df)[grepl("total", names(df), ignore.case=T)]
names(df)[names(df) == 'EstimatedTotalIntraCranialVol'] <- 'eTIV'
```

then from synthseg scans
```{r}
#add ID for merging scans
df <- df %>% mutate(merge_id = row_number())

df_synthseg <- df %>%
  dplyr::filter(fs_version == "synthseg-robust")

#confirm all missing
stopifnot(sum(is.na(df_synthseg$GMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$sGMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$WMV)) == nrow(df_synthseg))
stopifnot(sum(is.na(df_synthseg$CSF)) == nrow(df_synthseg))

#find appropriate cols to sum
df_synthseg %>%
  dplyr::select(all_of(c(contains("white"), contains("cort"), contains("hemi"), contains("ven")))) %>%
  skim()

df_synthseg %>%
  dplyr::select(contains("SUBC")) %>%
  skim()

ss_subcort_list <- c("SUBC.Left.Thalamus.Proper", "SUBC.Right.Thalamus.Proper", 
                     "SUBC.Left.Caudate", "SUBC.Right.Caudate",
                     "SUBC.Left.Putamen", "SUBC.Right.Putamen",
                     "SUBC.Left.Pallidum", "SUBC.Right.Pallidum",
                     "SUBC.Left.Hippocampus", "SUBC.Right.Hippocampus",
                     "SUBC.Left.Amygdala", "SUBC.Right.Amygdala",
                     "SUBC.Left.Accumbens.area", "SUBC.Right.Accumbens.area")

#calculate
df_glob <- df %>%
  mutate(GMV = ifelse(fs_version == "synthseg-robust", 
                      rowSums(across(c(rhCortexVol, lhCortexVol)), na.rm=FALSE),
                      GMV),
         WMV = ifelse(fs_version == "synthseg-robust",
                      rowSums(across(c(rhCerebralWhiteMatterVol, lhCerebralWhiteMatterVol)), na.rm=FALSE),
                      WMV),
         CSF = ifelse(fs_version == "synthseg-robust", Ventricles.all, CSF),
         sGMV = ifelse(fs_version == "synthseg-robust",
                       rowSums(across(all_of(ss_subcort_list)), na.rm=FALSE),
                       sGMV),
         eTIV = ifelse(fs_version == "synthseg-robust" & is.na(eTIV),
                       total.intracranial, eTIV)
         )
stopifnot(nrow(df_glob) == nrow(df))
```

plot for comparisons
```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=GMV, color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

sum(is.na(df$GMV)) - sum(is.na(df_glob$GMV))
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=sGMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280) +
  facet_wrap(~fs_version)
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=WMV,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

```{r}
ggplot(df) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_glob) + 
  geom_point(aes(x=age_days, y=CSF,color=fs_version), alpha=.4) +
  geom_vline(xintercept = 280)
```

CSF has s ome outliers but overall looks good, going to check with Lena but proceed for now

check for any more big groups of missing global volumes
```{r}
df_glob %>%
  dplyr::filter(is.na(GMV&sGMV&WMV&CSF)) %>%
  group_by(fs_version) %>%
  summarise(n = n())
```

## Data Cleaning
remove intersex subj, dummy-code sex and sex*age, combine Euler and SS QC metrics (because of missingness)
also fix age calculations for `OpenNeuro-PaingenPlacebo` study - study reported years, accidentally calculated as months

```{r}
df_filt <- df_glob %>%
  dplyr::filter(sex != "intersex") %>%
  mutate(age_days = ifelse(study != "OpenNeuro-PaingenPlacebo", age_days,
                           ((age_days-280)*(12/365.25))*365.25+280)) %>% #convert back from days to 'months', then re-convert years to days
  mutate(sexMale = ifelse(sex=="male", 1, 0),
         sexMale_x_age = ifelse(sex=="male", age_days, 0),
         QC_Metric = ifelse(fs_version == "synthseg-robust", QC.SS, SurfaceHoles),
         logAge_days = log(age_days, base=10)) %>% #only keep QC.SS for scans processed with synthseg]
  mutate(sexMale_x_logAge = sexMale * logAge_days) #add log-scaled ages
```

write out lists of phenotypes by category
```{r}
#names(df_filt)

#global volumes
glob_vol_list <- c("sGMV", "GMV", "WMV", "CSF","eTIV")
write_lines(glob_vol_list, "./pheno_lists/global_vols.txt")
saveRDS(glob_vol_list, "./pheno_lists/global_vols.rds")

#cortical gmv
gmv_list <- names(df_filt)[grep("GM\\.", names(df))]
stopifnot(length(gmv_list) == 68)
write_lines(gmv_list, "./pheno_lists/cortical_vols.txt")
saveRDS(gmv_list, "./pheno_lists/cortical_vols.rds")

#subcortical gmv
subcort <- names(df_filt)[grep("SUBC\\.Right|SUBC\\.Left", names(df))]
subcort_list <- subcort[-grep("Cerebellum|WM|Vent|choroid|vessel", subcort)] #drop WM, cerebellum, etc
length(subcort_list)
write_lines(subcort_list, "./pheno_lists/subcortical_vols.txt")
saveRDS(subcort_list, "./pheno_lists/subcortical_vols.rds")

#cortical thickness
ct_list <- names(df_filt)[grep("CT\\.", names(df))]
stopifnot(length(ct_list) == 68)
write_lines(ct_list, "./pheno_lists/cortical_thickness.txt")
saveRDS(ct_list, "./pheno_lists/cortical_thickness.rds")

#cortical SA
sa_list <- names(df_filt)[grep("SA\\.", names(df))]
stopifnot(length(sa_list) == 68)
write_lines(sa_list, "./pheno_lists/cortical_surf.txt")
saveRDS(sa_list, "./pheno_lists/cortical_surf.rds")
```

keep only columns we need (phenotypes to model, covariates)

```{r}
pheno_list <- c(glob_vol_list, gmv_list, subcort_list, ct_list, sa_list)
covar_list <- c("INDEX.ID", "session", "run", "dx", "study", "sexMale", "sexMale_x_age",
                "age_days", "logAge_days", "sexMale_x_logAge", "site", "country", "study_site", "QC_Metric", "fs_version", "eTIV")

df_select <- df_filt %>%
  dplyr::select(all_of(c(pheno_list, covar_list)))

skim(df_select)
```


### Remove Psychosis Studies
set aside selected psychosis studies
```{r}
psych_dx_list <- c(unique(df$dx)[grepl("scz|sz|psych|schizo|cos", unique(df$dx), ignore.case=T)], "CN")

unique(df_select$study)
df %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(study, dx) %>%
  summarise(count = n())

scz_study_list <- c("ALSPAC", "AMP-SCZ", "BHRCS", "COBRE_FS6_part1", "Dublin_FS6", "HCP-EarlyPsychosis", "MCI", "Maastricht_FS6",
                    "NAPLS3", "NDAR-CCDP", "NDAR-PsychiatricCerebralBlood", "NDAR-PsychosisOMS", "NDAR-SCZAging", "NDAR-SCZEmot",
                    "NDAR-SCZGlu", "NDAR-SocialBrainSchizoAutism", "NDAR_BSNIP", "NDAR_PARDIP", "NDAR_SCZIowa", "OpenNeuro_LA5c",
                    "SRPBS1600", "TOPSY", "NDAR_NIH-IRP") #NDAR_NIH-IRP = controls for Armin's COS dataset

psychosis_df <- df_select %>%
  dplyr::filter(study %in% scz_study_list) %>%
  dplyr::filter(dx %in% psych_dx_list) %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup()

ggplot(psychosis_df) +
  geom_histogram(aes(x=(age_days/365.25), fill=dx))

psychosis_df %>%
  group_by(dx) %>%
  summarise(count=n())
```

write out csv - *not filtering/removing duplicate scans because planning to add more data!*
```{r}
fwrite(psychosis_df, file="./data/240820-L.R.LBCC_scz_studies.csv")
```

FOR MORE SCZ STUDIES TO BE ADDED, SEE NDAR PROPOSAL GOOGLE DOC 

remove psychosis studies from main dataset so there's no data leakage
```{r}
df_nopsych <- df_select %>%
  dplyr::filter(!(study %in% scz_study_list))
```


look into subjects missing dx
```{r}
no_dx <- df_nopsych %>% dplyr::filter(is.na(dx))
ggplot(no_dx) +
  geom_histogram(aes(x=age_days)) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

no_dx %>%
  group_by(study) %>%
  summarise(mean_age = mean(age_days),
            n = n())

df_nopsych %>%
  dplyr::filter(study == "NDAR_IBIS") %>%
  group_by(dx) %>%
  summarise(n=n(),
            min_age = min(age_days),
            max_age = max(age_days))
```


subjects missing diagnoses are not consistently infants

filter to just controls, single scan per subject, drop studies with fewer than 5 ppl

```{r}
unique(df_nopsych$dx)[grep("cn|control|none", unique(df_nopsych$dx), ignore.case=TRUE)]

df_clean <- df_nopsych %>%
  tidyr::drop_na(dx) %>%
  dplyr::filter(age_days <= (101*365.25)) %>%
  dplyr::filter(dx == "CN") %>%
  group_by(INDEX.ID) %>% #remove duplicate participants
  slice_sample() %>%
  ungroup() %>%
  group_by(study) %>%
  filter(n_distinct(INDEX.ID) >=5) %>%
  ungroup()

#look at completeness for infant/toddler data
df_clean %>%
  dplyr::filter(age_days <= (365.25*3 + 280)) %>%
  skim()
#some phenotypes around 50%


df_clean %>%
  dplyr::filter(age_days <= (365.25 + 280)) %>%
  skim()
#some as low as 25%
```

###Subdivide by Missingness

Making separate CSVs with maximally complete data in order to ComBat features. For each feature (target feature to ComBat) keep as many other features as possible (as priors) without including any NAs
write out separate csvs for each pheno category

```{r}
#save out QC separately since these would only be used for sensitivity analyses
df_covars <- df_clean %>%
  dplyr::select(all_of(covar_list))
fwrite(df_covars, file="./data/SBC_LBCC_CN_covars.csv")

covar_subset <- covar_list[!covar_list %in% c("QC_Metric", "run", "eTIV")]

#global vols
df_global <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, glob_vol_list))) %>%
  filter_at(glob_vol_list, any_vars(complete.cases(.)))
fwrite(df_global, "./data/csvs_by_pheno/global_vols.csv")

#regional gmv
df_gmv <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, gmv_list)))%>%
  filter_at(gmv_list, any_vars(complete.cases(.)))
fwrite(df_gmv, "./data/csvs_by_pheno/cortical_vols.csv")

#regional sa
df_sa <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, sa_list)))%>%
  filter_at(sa_list, any_vars(complete.cases(.)))
fwrite(df_sa, "./data/csvs_by_pheno/cortical_surf.csv")

#regional ct
df_ct <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, ct_list)))%>%
  filter_at(ct_list, any_vars(complete.cases(.)))
fwrite(df_ct, "./data/csvs_by_pheno/cortical_thickness.csv")

#subcortical vols
df_sub <- df_clean %>%
  dplyr::select(all_of(c(covar_subset, subcort_list)))%>%
  filter_at(subcort_list, any_vars(complete.cases(.)))
fwrite(df_sub, "./data/csvs_by_pheno/subcortical_vols.csv")
```

start with global vols
```{r}
gg_miss_upset(df_global)

gg_miss_var(df_global,
            facet = fs_version)
```
Taking too much memory, writing `split_data_for_combat.R` to handle subdividing csvs 

Functions
```{r}
#function to split out complete dataframes
get_complete_df <- function(var, df){
  #print(var)
  new_df <- df %>%
    dplyr::filter(!is.na(!!sym(var))) %>% #can't handle var name as string, fix this!
    select_if(~ !any(is.na(.)))
  return(new_df)
}

#re-combine dfs for any phenotypes that are identical - code with help from ChatGPT
merge_identical_dfs <- function(df_list) {
  df_names <- names(df_list)
  processed <- rep(FALSE, length(df_list)) # Track which dataframes have been processed
  new_list <- list() # Initialize an empty list to store merged dataframes
  
  for (i in seq_along(df_list)) {
    if (!processed[i]) {
      identical_dfs <- c(df_names[i]) # Start with the current dataframe's name
      
      # Compare the current dataframe with others
      for (j in (i + 1):length(df_list)) {
        if (!processed[j] && identical(df_list[[i]], df_list[[j]])) {
          identical_dfs <- c(identical_dfs, df_names[j])
          processed[j] <- TRUE # Mark the duplicate dataframe as processed
        }
      }
      
      # If there are any identical dataframes, rename the first one
      if (length(identical_dfs) > 1) {
        new_name <- paste(identical_dfs, collapse = "_") # Concatenate the names
        new_list[[new_name]] <- df_list[[i]] # Add the merged dataframe with the new name
      } else {
        # If no identicals were found, keep the original dataframe
        new_list[[df_names[i]]] <- df_list[[i]]
      }
      
      processed[i] <- TRUE # Mark the current dataframe as processed
    }
  }
  
  return(new_list)
}

rename_cols <- function(df_list, col_list, ncores=5){
   mclapply(names(df_list), function(df_name) {
    # Construct the file path using the dataframe's name
    combat_list <- strsplit(gsub("_", " ", df_name), " ")[[1]]
   
    #find any phenotypes that are NOT the combat targets
    not_target <- setdiff(col_list, combat_list)
    #print(not_target)
    
    #see which ones are in the dataframe
    not_target <- intersect(not_target, names(df_list[[df_name]]))
    #print(not_target)
    
    if (length(not_target)>0){
    df_list[[df_name]] <- df_list[[df_name]] %>%
      rename_with(~ paste(., "X", sep = "_"), any_of(not_target))
    }
    return(df_list[[df_name]])
  }, mc.cores=ncores)
}

#write out csvs - writtenwith help from ChatGPT
save_csv_list <- function(df_list, path = ".") {
  lapply(seq(length(df_list)), function(idx) {
    # Construct the file path using the dataframe's name
    file_path <- paste0(path, idx, ".csv")
    
    # Use fwrite to save the dataframe
    fwrite(df_list[[idx]], file_path)
  })
}
```

```{r}
df_global_list <- lapply(c(glob_vol_list, "eTIV"), get_complete_df, df_global)
names(df_global_list) <- c(glob_vol_list, "eTIV")
lapply(df_global_list, ncol)
df_global_final <- rename_cols(df_global_list, c(glob_vol_list, "eTIV"), 20)

#save
save_csv_list(df_global_list, "./data/to_combat/global_")
rm(df_global_list)
```

```{r}
df_sub_list <- lapply(subcort_list, get_complete_df, df_sub)
names(df_sub_list) <- subcort_list

df_sub_list_clean <- merge_identical_dfs(df_sub_list)
lapply(df_sub_list_clean, ncol)

df_sub_list_final <- rename_cols(df_sub_list_clean, subcort_list, 20)

#save
save_csv_list(df_sub_list_final, "./data/to_combat/sub_")
rm(df_sub_list_final, df_sub_list, df_sub_list_clean)
```

```{r}
df_gmv_list <- lapply(gmv_list, get_complete_df, df_gmv)
names(df_gmv_list) <- gmv_list

df_gmv_list_clean <- merge_identical_dfs(df_gmv_list)
lapply(df_gmv_list_clean, ncol)

df_gmv_list_final <- rename_cols(df_gmv_list_clean, df_gmv_list, 20)

#save
save_csv_list(df_gmv_list_final, "./data/to_combat/gmv_")
rm(df_gmv_list_final, df_gmv_list)
```

```{r}
df_ct_list <- lapply(ct_list, get_complete_df, df_ct)
names(df_ct_list) <- ct_list

df_ct_list_clean <- merge_identical_dfs(df_ct_list)
```

```{r}
df_sa_list <- lapply(sa_list, get_complete_df, df_sa)
names(df_sa_list) <- sa_list

df_sa_list_clean <- merge_identical_dfs(df_sa_list)
```

## Visualize Data

### Global Phenotypes
```{r}
ggplot(df_global) + 
  geom_point(aes(x=age_days, y=GMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=WMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=sGMV,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))

ggplot(df_global) + 
  geom_point(aes(x=log(age_days), y=CSF,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = log(280))
```

### Cortical volumes
```{r}
ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.lingual,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.parsorbitalis,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=rh.GM.fusiform,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_gmv) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))

#compare to original data

df_test_dx <- df %>%
  mutate(dx = case_when(is.na(dx) ~ NA,
                        dx == "CN" ~ "CN",
                        TRUE ~ "notCN"))
ggplot(df_test_dx) + 
  geom_point(aes(x=age_days, y=lh.GM.inferiortemporal, color=dx), alpha=.4) +
  geom_vline(xintercept = 280) +
  geom_vline(xintercept = (365.25*3))
```

###Cortical Thickness
```{r}
ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

#compare to full dataset
ggplot(df_clean) + 
  geom_point(aes(x=age_days, y=rh.CT.superiorparietal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.parahippocampal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=rh.CT.caudalanteriorcingulate, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_ct) + 
  geom_point(aes(x=age_days, y=lh.CT.transversetemporal,color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Cortical SA
```{r}
ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.pericalcarine, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.lateralorbitofrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=lh.SA.entorhinal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sa) + 
  geom_point(aes(x=age_days, y=rh.SA.rostralmiddlefrontal, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

###Subcortical Volumes
```{r}
ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Hippocampus, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Accumbens.area, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Left.Amygdala, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)

ggplot(df_sub) + 
  geom_point(aes(x=age_days, y=SUBC.Right.Thalamus.Proper, color=as.factor(sexMale)), alpha=.4) +
  geom_vline(xintercept = 280)
```

## some test models

```{r}
library(gamlss)
devtools::install_github("BGDlab/gamlssTools")
library(gamlssTools)
```

standard sex-mod model with pb()
```{r}
mod1 <- gamlss(GMV ~ pb(age_days, method="GAIC", k=log(length(df_global$GMV))) + sexMale + pb(sexMale_x_age, method="GAIC", k=log(length(df_global$GMV))) + study, family = GG, data=df_global)
make_centile_fan(mod1, df_global, "age_days", "sexMale")
```
sex-mod model with pb() on log-scaled age
```{r}
mod2 <- gamlss(GMV ~ pb(logAge_days, method="GAIC", k=log(length(df_global$GMV))) + sexMale + pb(sexMale_x_logAge, method="GAIC", k=log(length(df_global$GMV))) + study, family = GG, data=df_global)
make_centile_fan(mod2, df_global, "logAge_days", "sexMale")
```
looks better!

```{r}
irismod1 <-gamlss(Sepal.Length ~ ns(Sepal.Width) + Species, data=iris)
make_centile_fan(irismod1, iris, "Sepal.Width", "Species")

irismod2 <-gamlss(Sepal.Length ~ ns(Sepal.Width, knots=(c(mean(iris$Sepal.Width)-0.5, mean(iris$Sepal.Width), mean(iris$Sepal.Width)+0.5))) + Species, data=iris)
make_centile_fan(irismod2, iris, "Sepal.Width", "Species")

irismod3 <-gamlss(Sepal.Length ~ ns(Sepal.Width, df=4) + Species, data=iris)
make_centile_fan(irismod3, iris, "Sepal.Width", "Species")
```

sex-mod model with ns() on log-scaled age and biologically-informed spline placement
```{r}
ages_yrs<-c(1, 3, 7, 16, 25, 45, 70)
ages_logdays <- lapply(ages_yrs, function(x){log(x*365.25, base=10)}) %>% unlist

mod3 <- gamlss(GMV ~ ns(logAge_days, knots=c(3.039711, 3.407688, 3.766710, 3.960530, 4.215803, 4.407688)) + sexMale + ns(sexMale_x_logAge, df=10) + study, family = GG, data=df_global)
make_centile_fan(mod3, df_global, "logAge_days", "sexMale")
```

sex-mod model with ns() on log-scaled age
```{r}
mod4 <- gamlss(GMV ~ ns(logAge_days, df=30) + sexMale + ns(sexMale_x_logAge, df=10) + study, family = GG, data=df_global)
make_centile_fan(mod4, df_global, "logAge_days", "sexMale")
```

sex-strat model with ns() on log-scaled age and biologically-informed spline placement
```{r}
mod5 <- gamlss(GMV ~ ns(logAge_days, knots=c(2.562590, 3.039711, 3.407688, 3.766710, 3.960530, 4.215803, 4.407688)) + sexMale + study, family = GG, data=df_global)
make_centile_fan(mod5, df_global, "logAge_days", "sexMale")
```

sex-mod model with ns() on log-scaled age
```{r}
mod6 <- gamlss(GMV ~ ns(logAge_days, df=30) + sexMale + study, family = GG, data=df_global)
make_centile_fan(mod6, df_global, "logAge_days", "sexMale")
```

```{r}
mod7 <- gamlss(GMV ~ ns(logAge_days, 5) + sexMale + study, family = GG, data=df_global)
make_centile_fan(mod7, df_global, "logAge_days", "sexMale")
```