---
title: "Prelim_HBCD_eda"
output: html_document
date: "2025-03-14"
---

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

Incorporating preliminary HBCD data (pre-release) into LBCC v3  for growth charts

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel, ggpubr)
```

#tabular imaging df

```{r}
imaging_df <- fread("/mnt/isilon/bgdlab_processing/Data/HBCD/HBCD_tabular_downloads/imaging/bibset_segmentation/data.csv", na.strings = c("", NA))
#names(imaging_df)
names(imaging_df)[grep("age", names(imaging_df))]
skim(imaging_df$img_mriqc_T2w_candidate_age)

imaging_df %>%
  select(all_of(names(imaging_df)[grep("age", names(imaging_df))])) %>%
  skim()

imaging_df_filt <- imaging_df %>%
  rename(age = img_mriqc_T2w_candidate_age)
skim(imaging_df_filt)

colnames(imaging_df_filt) <- gsub("img_bibsnet_space-T2w_desc-aseg_volumes_", "", colnames(imaging_df_filt))
```

```{r}
imaging_df_filt %>%
  select(starts_with("img_mriqc_T2w")) %>%
  dplyr::select(where(is.numeric)) %>%
  select(-contains("summary")) %>%
  pivot_longer(cols = everything(),
               names_to = "IQM",
               values_to = "value") %>%
  mutate(IQM = gsub("img_mriqc_T2w_", "", IQM)) %>%
  ggplot() +
  geom_histogram(aes(x = value, fill = IQM), position = "identity", bins = 30) +
  theme_minimal() +
  facet_wrap(~IQM, scales="free") +
  theme(legend.position = "none") +
  ggtitle("T2w Scan IQMs")

```


```{r}
ggplot(imaging_df) +
  geom_point(aes(x=img_mriqc_T2w_candidate_age, y= img_mriqc_T1w_candidate_age))

imaging_df %>%
  rename(img_bibsnet_space_T2w_desc_aseg_volumes_candidate_age = "img_bibsnet_space-T2w_desc-aseg_volumes_candidate_age") %>%
ggplot() +
  geom_point(aes(x=img_mriqc_T2w_candidate_age, y=img_bibsnet_space_T2w_desc_aseg_volumes_candidate_age))
```


#M-CRIB Imaging Outputs
need to compile across subject-specific dirs

```{r}
mcrib_df <- fread("/mnt/isilon/bgdlab_processing/Data/HBCD/HBCD_justMCRIB_21March2025/HBCD_beta_MCRIB_freesurfer_aparcaseg_volumes.csv")
skim(mcrib_df)

mcrib_df <- mcrib_df %>%
  rename("id" = "subject") %>%
  filter_all(any_vars(!is.na(.)))
```


compare dfs
```{r}
n_unique(mcrib_df$id)
n_unique(imaging_df_filt$id)

intersect(names(imaging_df_filt), names(mcrib_df))

skim(mcrib_df)
skim(imaging_df_filt)
```

test merge
```{r}
test_df <- full_join(imaging_df_filt, mcrib_df)
names(test_df)

test_df %>% filter(is.na(img_mriqc_T1w_candidate_age))

# ggplot(test_df) +
#   geom_point(aes(x=age, y=CSF), alpha=.4)
ggplot(test_df) +
  geom_point(aes(x=age, y=`Right-Amygdala`), alpha=.4)
ggplot(test_df) +
  geom_point(aes(x=age, y=`Right-Lateral-Ventricle`), alpha=.4)
```

looks pretty reasonable

#Demographics

```{r}
demo_df <- fread("/mnt/isilon/bgdlab_processing/Data/HBCD/HBCD_tabular_downloads/nonimaging/demographics/DATA_CSV14096344619509598011.csv", na.strings = c("", NA))
#names(demo_df)
unique(demo_df$par_visit_data_par_visit_data_cohort)
unique(demo_df$par_visit_data_par_visit_data_project)

demo_sub <- demo_df %>%
  dplyr::select(id, visit_label, cohort, site, sed_basic_demographics_sed_basic_demographics_sex) %>%
  rename(sex = sed_basic_demographics_sed_basic_demographics_sex) %>%
  mutate(sex = ifelse(sex=="Unknown", NA, sex)) %>%
  na.omit()
```

Merge!
```{r}
n_unique(demo_sub$id)
n_unique(test_df$id)

hbcd <- inner_join(demo_sub, test_df) %>%
  select_if(~!all(is.na(.))) %>% #drop col if all is empty
  drop_na(age, sex) %>%
  mutate(age_days = 365.25*age + 280) #confirmed age years

skim(hbcd)

n_unique(hbcd$id)

length(intersect(demo_sub$id, mcrib_df$id))
```

Add minimal QC filter (T2w CJV) based on slack convo

```{r}
cjv_mad = mad(hbcd$img_mriqc_T2w_cjv, na.rm = T)
cjv_med = median(hbcd$img_mriqc_T2w_cjv, na.rm=T)
thresh <- cjv_med + 4*cjv_mad

ggplot(hbcd) +
  geom_histogram(aes(x=img_mriqc_T1w_cjv)) +
  geom_vline(xintercept = cjv_med, color="blue") +
  geom_vline(xintercept = thresh, color="red")

qi2_mad = mad(hbcd$img_mriqc_T2w_qi_2, na.rm = T)
qi2_med = median(hbcd$img_mriqc_T2w_qi_2, na.rm=T)
thresh <- qi2_med + 4*qi2_mad

ggplot(hbcd) +
  geom_histogram(aes(x=img_mriqc_T2w_qi_2)) +
  geom_vline(xintercept = qi2_med, color="blue") +
  geom_vline(xintercept = thresh, color="red")
```

using 4 MAD threshold for CJV excludes no one. going to err on the side of being more inclusive, might circle back to remove scans based on lower threshold and/or QI2 as necessary

````{r}
fwrite(hbcd, "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/data/beta_hbcd.csv")
```