---
title: "Puberty EDA"
output: html_document
date: "2024-11-01"
---

Collating puberty data for Aim 2 sample.

##Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="/mnt/isilon/bgdlab_processing")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel, ggpubr, fuzzyjoin, cowplot, ggpubr)
```


set paths
```{r}
my_path <- "./Margaret/sex_mod_braincharts"
data_path <- "./Data/LBCC"
```

##LBCC Demos
just get basic demographics from puberty studies included in current LBCC

```{r}
lbcc_df <- fread(paste0(my_path, "/data/240820-L.R.LBCC.csv"))
unique(lbcc_df$study) %>% sort()

lbcc.puberty_studies <- c("ABCD", "PNC", "HCP-D", "HBN", "SYS-Adolescents", "PACCT")

lbcc_puberty <- lbcc_df %>%
  dplyr::filter(study %in% lbcc.puberty_studies) %>%
  dplyr::filter(sex != "intersex") %>%
  dplyr::select(c(INDEX.ID, age_days, sex, study, dx)) %>%
  mutate(age_yrs = (age_days-280)/365.25) %>%
  dplyr::filter(age_yrs <= 25 & age_yrs >= 5)

#trying to rank by youngest age in study
lbcc_puberty_rank <- lbcc_puberty%>%
  group_by(INDEX.ID) %>%
  arrange(age_days) %>%
  slice_head() %>% 
  ungroup() %>%
  mutate(age_rank = rank(age_days, ties.method="random")) %>%
  select(INDEX.ID, age_rank)

lbcc_puberty_rank2 <- left_join(lbcc_puberty, lbcc_puberty_rank, by="INDEX.ID")

lbcc_puberty_rank2 %>%
  ggplot(aes(x=age_yrs, y=age_rank, color=sex, group=INDEX.ID), alpha=0.5)+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(lbcc_puberty_rank$INDEX.ID) == n_unique(lbcc_puberty_rank2$age_rank)
```


##Get Raw Data
Rough, still need to go back later and filter to just CN

###Stanford NDAR2135

```{r}
path_2135 <- paste0(data_path, "/NDAR2135/data_dump")
image_df <- fread(paste0(path_2135, "/image03.txt"))
skim(image_df)
names(image_df)

table(image_df$scan_type)

#drop subheader row
image_df2 = image_df[-1,]

image_df_filt <- image_df2 %>%
  dplyr::select(subjectkey, src_subject_id, interview_date, interview_age, sex) %>%
  mutate(age_yrs = as.numeric(interview_age)/12,
         scan_date = as.Date(interview_date, "%m/%d/%y")) %>%
  rename(scan_age = interview_age) %>%
  select(!interview_date)

image_df_filt %>%
  filter(str_detect(subjectkey, "^N")) %>%
  ggplot(aes(x=age_yrs, y=subjectkey, color=sex, group=subjectkey))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(image_df_filt$subjectkey)
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_2135, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
pdf_df_filt <- pds_df2 %>% 
  select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join (https://stackoverflow.com/questions/58718287/fuzzyjoin-with-dates-in-r) with 30-day margin

```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2135 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date) %>%
  unique()

n_unique(NDAR2135$subjectkey)
```



###TAG NDAR2315

no scan info on NDAR, just using pds timepoints
```{r}
path_2315 <- paste0(data_path, "/NDAR2315/data_dump")
pds_df <- fread(paste0(path_2315, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
TAG <- pds_df2 %>% select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

###Nebraska NDAR2714

```{r}
path_2714 <- paste0(data_path, "/NDAR2714/data_dump")
image_df <- fread(paste0(path_2714, "/image03.txt"))
skim(image_df)
names(image_df)

table(image_df$scan_type)

#drop subheader row
image_df2 = image_df[-1,]

image_df_filt <- image_df2 %>%
  dplyr::select(subjectkey, src_subject_id, interview_date, interview_age, sex) %>%
  mutate(age_yrs = as.numeric(interview_age)/12,
         scan_date = as.Date(interview_date, "%m/%d/%y")) %>%
  rename(scan_age = interview_age) %>%
  select(!interview_date)

image_df_filt %>%
  filter(str_detect(subjectkey, "^N")) %>%
  ggplot(aes(x=age_yrs, y=subjectkey, color=sex, group=subjectkey))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(image_df_filt$subjectkey)
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_2714, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
pdf_df_filt <- pds_df2 %>% 
  select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join
```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2714 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date) %>%
  unique()

n_unique(NDAR2714$subjectkey)
```


###RDOC-Twin NDAR2105

```{r}
path_2105 <- paste0(data_path, "/NDAR-RDOCTwin/data_dump")
image_df <- fread(paste0(path_2105, "/image03.txt"))
skim(image_df)
names(image_df)

table(image_df$scan_type)

#drop subheader row
image_df2 = image_df[-1,]

image_df_filt <- image_df2 %>%
  dplyr::select(subjectkey, src_subject_id, interview_date, interview_age, sex) %>%
  mutate(age_yrs = as.numeric(interview_age)/12,
         scan_date = as.Date(interview_date, "%m/%d/%y")) %>%
  rename(scan_age = interview_age) %>%
  select(!interview_date)

image_df_filt %>%
  filter(str_detect(subjectkey, "^N")) %>%
  ggplot(aes(x=age_yrs, y=subjectkey, color=sex, group=subjectkey))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(image_df_filt$subjectkey)
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_2105, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
pdf_df_filt <- pds_df2 %>% 
  select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join
```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2105 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date) %>%
  unique()

n_unique(NDAR2105$subjectkey)
```

###ABCD
```{r}
path_abcd <- "./Data/ABCD/release5/core"
image_df <- fread(paste0(path_abcd, "/imaging/mri_y_adm_info.csv"))
skim(image_df)

image_df2 <- image_df %>%
  mutate(scan_date = as.Date(as.character(mri_info_studydate), "%Y%m%d"))
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_abcd, "/physical-health/ph_y_pds.csv"))
skim(pds_df)

#get ages
demo_df <- fread(paste0(path_abcd, "/abcd-general/abcd_y_lt.csv"))
skim(demo_df)
demo_df2 <- demo_df %>%
  mutate(interview_date = as.Date(interview_date, "%d/%m/%y"))

#merge
pds_demo <- left_join(pds_df, demo_df2)
```

inner join - too big to use fuzzy join
```{r}
abcd <- inner_join(pds_demo, image_df2,
                             by = c("src_subject_id", "eventname")) %>%
  mutate(pds_date_diff = scan_date - interview_date,
         sex=case_when(pds_sex_y == 1 ~ "male",
                       pds_sex_y == 2 ~ "female"))

n_unique(abcd$src_subject_id)

abcd %>%
  ggplot(aes(x=interview_age, y=src_subject_id, color=sex, group=src_subject_id))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")
```

keeping all for now, but unclear what date PDS was actually assessed and how close it was to scan date

##NDAR Combine

```{r}
ndar_df <- bind_rows(list(NDAR2135, TAG, NDAR2714, NDAR2105), .id="study") %>%
  mutate(age_yrs = as.numeric((interview_age))/12)

skim(ndar_df)

#drop puberty data for now
ndar_demos <- ndar_df %>%
  select(subjectkey, interview_date, sex, age_yrs, study)
```

##All Demos
add new NDAR demos into LBCC data

```{r}
full_demo <- full_join(lbcc_puberty, ndar_df, by=c("INDEX.ID"="subjectkey", "age_yrs" = "age_yrs", "sex"="sex", "study" = "study"))%>%
  dplyr::filter(age_yrs <= 25 & age_yrs >= 5) %>%
  dplyr::filter(!(age_yrs > 17 & study=="ABCD")) %>%#drop 1 old ABCD ppt based on release 5 matching
  mutate(sex = case_when(sex == "M" | sex =="male" ~ "Male",
                         sex == "F" | sex =="female" ~ "Female",
                         TRUE ~ NA),
         is.pnc = ifelse(study=="PNC", TRUE, FALSE))

demo_nopnc <- full_demo %>%
  dplyr::filter(is.pnc == FALSE)

#trying to rank by youngest age in study
full_demo_rank <- demo_nopnc%>%
  group_by(INDEX.ID) %>%
  arrange(age_yrs) %>%
  slice_head() %>% 
  ungroup() %>%
  group_by(sex) %>%
  mutate(age_rank = rank(age_yrs, ties.method="random")) %>%
  ungroup() %>%
  select(INDEX.ID, age_rank)
  

full_demo_rank2 <- left_join(demo_nopnc, full_demo_rank, by="INDEX.ID")

demo_plt <- full_demo_rank2 %>%
  ggplot(aes(x=age_yrs, y=age_rank, color=sex, group=INDEX.ID))+
  geom_vline(xintercept=17, color='gray', linetype="dotted") +
  geom_line(alpha=0.4, linewidth=.3) +
  geom_point(size=1) +
  xlim(5,25)+
  theme(legend.position="none") +
  theme_cowplot() +
  labs(x=element_blank(), y="N") +
  facet_wrap(~sex) +
  theme(legend.position="none")
```

```{r}
pnc <- full_demo %>%
  filter(is.pnc ==TRUE)

pnc.plt <- pnc %>%
  ggplot() +
  geom_histogram(aes(x=age_yrs, fill=sex, color=sex), alpha=.7) +
  facet_wrap(~sex) +
  theme_cowplot() +
  xlim(5,25) +
  labs(x= "Age (years)", y="N") +
  theme(strip.background = element_blank(),
  strip.text.x = element_blank(),
  legend.position="none")
```

```{r}
pubert_plt <- ggarrange(demo_plt, pnc.plt,
          ncol=1, nrow=2,
          align = "v",
          heights=c(4,1))
ggsave("./Margaret/sex_mod_braincharts/puberty_sample.png", pubert_plt)
```
