---
title: "Puberty EDA"
output: html_document
date: "2024-11-01"
---

Collating puberty data for Aim 2 sample.

##Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir="/mnt/isilon/bgdlab_processing")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, naniar, rlang, parallel, ggrepel, ggpubr, fuzzyjoin)
```


set paths
```{r}
my_path <- "./Margaret/sex_mod_braincharts"
data_path <- "./Data/LBCC"
```

##Get Raw Demographics

###Stanford NDAR2135

```{r}
path_2135 <- paste0(data_path, "/NDAR2135/data_dump")
image_df <- fread(paste0(path_2135, "/image03.txt"))
skim(image_df)
names(image_df)

table(image_df$scan_type)

#drop subheader row
image_df2 = image_df[-1,]

image_df_filt <- image_df2 %>%
  dplyr::select(subjectkey, src_subject_id, interview_date, interview_age, sex) %>%
  mutate(age_yrs = as.numeric(interview_age)/12,
         scan_date = as.Date(interview_date, "%m/%d/%y")) %>%
  rename(scan_age = interview_age) %>%
  select(!interview_date)

image_df_filt %>%
  filter(str_detect(subjectkey, "^N")) %>%
  ggplot(aes(x=age_yrs, y=subjectkey, color=sex, group=subjectkey))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(image_df_filt$subjectkey)
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_2135, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
pdf_df_filt <- pds_df2 %>% 
  select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join (https://stackoverflow.com/questions/58718287/fuzzyjoin-with-dates-in-r) with 30-day margin

```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2135 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date)

n_unique(NDAR2135$subjectkey)
```



###TAG NDAR2315

no scan info on NDAR, just using pds timepoints
```{r}
path_2315 <- paste0(data_path, "/NDAR2315/data_dump")
pds_df <- fread(paste0(path_2315, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
TAG <- pds_df2 %>% select(where(~mean(is.na(.)) < .5))
```

###Nebraska NDAR2714

```{r}
path_2714 <- paste0(data_path, "/NDAR2714/data_dump")
image_df <- fread(paste0(path_2714, "/image03.txt"))
skim(image_df)
names(image_df)

table(image_df$scan_type)

#drop subheader row
image_df2 = image_df[-1,]

image_df_filt <- image_df2 %>%
  dplyr::select(subjectkey, src_subject_id, interview_date, interview_age, sex) %>%
  mutate(age_yrs = as.numeric(interview_age)/12,
         scan_date = as.Date(interview_date, "%m/%d/%y")) %>%
  rename(scan_age = interview_age) %>%
  select(!interview_date)

image_df_filt %>%
  filter(str_detect(subjectkey, "^N")) %>%
  ggplot(aes(x=age_yrs, y=subjectkey, color=sex, group=subjectkey))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(image_df_filt$subjectkey)
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_2714, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
pdf_df_filt <- pds_df2 %>% 
  select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join
```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2714 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date)

n_unique(NDAR2714$subjectkey)
```


###RDOC-Twin NDAR2105

```{r}
path_2105 <- paste0(data_path, "/NDAR-RDOCTwin/data_dump")
image_df <- fread(paste0(path_2105, "/image03.txt"))
skim(image_df)
names(image_df)

table(image_df$scan_type)

#drop subheader row
image_df2 = image_df[-1,]

image_df_filt <- image_df2 %>%
  dplyr::select(subjectkey, src_subject_id, interview_date, interview_age, sex) %>%
  mutate(age_yrs = as.numeric(interview_age)/12,
         scan_date = as.Date(interview_date, "%m/%d/%y")) %>%
  rename(scan_age = interview_age) %>%
  select(!interview_date)

image_df_filt %>%
  filter(str_detect(subjectkey, "^N")) %>%
  ggplot(aes(x=age_yrs, y=subjectkey, color=sex, group=subjectkey))+
  geom_point() +
  geom_line() +
  theme(legend.position="none")

n_unique(image_df_filt$subjectkey)
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_2105, "/pds01.txt")) %>% mutate_all(~na_if(., ''))
#drop subheader row
pds_df2 = pds_df[-1,]
skim(pds_df2)

#drop unused cols
pdf_df_filt <- pds_df2 %>% 
  select(where(~mean(is.na(.)) < .5)) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join
```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2105 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date)

n_unique(NDAR2105$subjectkey)
```

###ABCD
```{r}
path_abcd <- "./Data/ABCD/release5/core"
image_df <- fread(paste0(path_abcd, "/imaging/mri_y_adm_info.csv"))
skim(image_df)

image_df2 <- image_df %>%
  mutate(scan_date = as.Date(as.character(mri_info_studydate), "%Y%m%d"))
```
now adding corresponding puberty info
```{r}
pds_df <- fread(paste0(path_abcd, "/physical-health/ph_y_pds.csv"))
skim(pds_df)

#drop unused cols
pdf_df_filt <- pds_df %>% 
  select(!contains("cat")) %>%
  mutate(interview_date = as.Date(interview_date, "%m/%d/%y"))
```

fuzzy inner join
```{r}
image_df_mrg <- image_df_filt %>%
  mutate(scan_date_plus = scan_date + 30,
         scan_date_minus = scan_date- 30)

NDAR2105 <- fuzzy_inner_join(pdf_df_filt, image_df_mrg,
                             by = c("subjectkey" = "subjectkey",
                                    "sex" = "sex",
                                    'interview_date' = "scan_date_minus",
                                    'interview_date' = "scan_date_plus"),
                                 match_fun = list(`==`, `==`, `>`, `<`)) %>%
  dplyr::select(!c(scan_date_plus, scan_date_minus, subjectkey.y, src_subject_id.y, sex.y)) %>%
  rename(subjectkey = subjectkey.x,
         src_subject_id = src_subject_id.x,
         sex = sex.x) %>%
  mutate(pds_date_diff = scan_date - interview_date)

n_unique(NDAR2105$subjectkey)
```
