---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, sf, cowplot, subcortexVisualizationR, rstatix, freesurferformats, mgcv)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
#remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)

#source spin test
source("https://raw.githubusercontent.com/frantisekvasa/rotate_parcellation/refs/heads/master/R/perm.sphere.p.R")
source("https://raw.githubusercontent.com/frantisekvasa/rotate_parcellation/refs/heads/master/R/rotate.parcellation.R")
fs_path <- "/mnt/isilon/bgdlab_processing/code/freesurfer-6.0/subjects/fsaverage/"
```

```{r}
source("code/plot_cv_brain.R")
```

#Defining integral funs

```{r}
#fun for derivative
get_deriv <- function(x, age) {
  c(NA, diff(x) / diff(age))
}

#riemann_stieltjes
rs_integral <- function(x, y, g = function(x) log10(x)) {
  stopifnot(
    length(x) == length(y),
    length(x) >= 2
  )

  # sort just in case
  ord <- order(x)
  x <- x[ord]
  y <- y[ord]

  gx <- g(x)

  # monotonicity check (important!)
  if (any(diff(gx) < 0)) {
    stop("g(x) must be non-decreasing for a Riemann–Stieltjes integral")
  }

  sum(y[-length(y)] * diff(gx))
}

#euclidean distance fun from https://www.geeksforgeeks.org/r-language/how-to-calculate-euclidean-distance-in-r/
euclidean_dist <- function(vect1, vect2) sqrt(sum((vect1 - vect2)^2))
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

length(output_fread)

LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  dplyr::select(!file)
```

```{r, eval=FALSE}
LR.df %>%
  group_by(cv_sample, total, weighted) %>%
  summarise(n_phenos = n_unique(pheno))

A_totalFALSE <- LR.df %>%
  filter(cv_sample == "A" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalFALSE)

A_totalTRUE <- LR.df %>%
  filter(cv_sample == "A" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalTRUE)

B_totalFALSE <- LR.df %>%
  filter(cv_sample == "B" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalFALSE)

B_totalTRUE <- LR.df %>%
  filter(cv_sample == "B" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalTRUE)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
    pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)))

table(LR.df2$effect, LR.df2$total)
```



```{r}
LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$label), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE)

LR.df_uncor %>%
  group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T),
            mean_fsq = mean(sig_fsq, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_uncor_novent <- LR.df_uncor %>%
  filter(!(pheno %in% vent_list))
range(LR.df_uncor_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_uncor_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)

ggsave("figs/supplement/resi_uncorrected.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

```{r}
#test plotting just split A
LR.df_uncor_novent_B <- LR.df_uncor_novent %>%
  filter(cv_sample=="B")
B_plt_uncor <- plot_cv_brain(df = LR.df_uncor_novent_B, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = FALSE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)
saveRDS(B_plt_uncor, "figs/ggobjs/agebysex_B.rds")
```
###Weighted
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==FALSE) %>%
  group_by(cv_sample, weighted) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total == FALSE) %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(weighted) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  filter(total==FALSE & weighted==TRUE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL) #keeping same scale from unweighted models

ggsave("figs/supplement/resi_uncorrected_weight.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE)

table(LR.df_tot$sig_fdr)

LR.df_tot %>%
group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_tot_novent <- LR.df_tot %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_novent$resi)

full_plt_tot <- plot_cv_brain(df = LR.df_tot_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)

ggsave("figs/supplement/resi_tbv.png", full_plt_tot, dpi=600, height=8.5, width=7.1, bg = "white")
```

###Weighted
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age") %>%
  group_by(cv_sample, weighted) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age") %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(weighted) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```
more regions are significant in weighted analyses

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted==TRUE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL) #keeping same scale from unweighted models

ggsave("figs/supplement/resi_tbv_weight.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```
Note phenos with age*sex effects, per LR tests

```{r}
#consistent age*sex effects in both CV splits
consistent_agesex <- LR.df_tot %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  filter(sig_fdr_A == TRUE & sig_fdr_B == TRUE) %>%
  pull(pheno)

#just within each sample
sig_agesex <- LR.df_tot %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  filter(sig_fdr==TRUE) %>%
  pull(samp_pheno)
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE)

table(LR.df_tot_sex$sig_fdr, LR.df_tot_sex$cv_sample)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)

LR.df_tot_sex %>%
  filter(sig_fdr == TRUE) %>%
  group_by(cv_sample) %>%
  summarise(min_resi = min(sig_resi),
            max_resi = max(sig_resi),
            mean_resi = mean(sig_resi))
```

```{r}
#removing ventricles for visualization:
LR.df_tot_sex_novent <- LR.df_tot_sex %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_sex_novent$resi)

full_plt_tot_sex <- plot_cv_brain(df = LR.df_tot_sex_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)

ggsave("figs/supplement/resi_tbv_allsex.png", full_plt_tot_sex, dpi=600, height=8.5, width=7.1, bg = "white")
```

plot just in B
```{r}
#removing ventricles for visualization:
LR.df_tot_sex_noventB <- LR.df_tot_sex_novent %>%
  filter(cv_sample=="A")
full_plt_tot_sex_B <- plot_cv_brain(df = LR.df_tot_sex_noventB, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = FALSE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)
  
saveRDS(full_plt_tot_sex_B, "figs/ggobjs/full_plt_tot_sex_B.rds")
```

###Weighted
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all") %>%
  group_by(cv_sample, weighted) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_resi = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all") %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(weighted) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```
RESIs and n significant regions are basically identical across weighted and unweighted models 

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted==TRUE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)
  
ggsave("figs/supplement/resi_tbv_allsex_weight.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

##Save Table
save out csv with all significance tests

```{r}
LR_table <- LR.df_plot %>%
  select(total, effect, weighted, pheno_cat, pheno, cv_sample, chi, df, p_val, fsq, resi, p.val_fdr, sig_fdr) %>%
  rename(total_size_covar = total,
         effect_tested = effect,
         surfhole_weighted = weighted,
         idp_category = pheno_cat,
         idp = pheno,
         p.val = p_val,
         significant_fdr = sig_fdr) %>%
  mutate(effect_tested = ifelse(effect_tested=="sex_age", "age-varying sex component", "any sex effect"))

fwrite(LR_table, "figs/supplement/significance_test_results.csv")
```

##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

sexdiffs.df_all <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         pheno = stringr::str_remove(basename(file), "_(uniform_)?sexdiffs\\.csv"),
         uniform = ifelse(grepl("uniform", file), TRUE, FALSE)) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    age_days = un_log(logAge_days)) %>%
  dplyr::select(!file) %>%
  group_by(total, weighted, cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(#chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age,
         deriv_logcv_M_div_F.unlog = deriv_logcv_M_div_F*deriv_age)

sexdiffs.df <- sexdiffs.df_all %>% filter(uniform==FALSE)
sexdiffs.df.u <- sexdiffs.df_all %>% filter(uniform==TRUE)


rm(output_fread_sd) #save space
rm(sexdiffs.df_all)
```


```{r}
sexdiffs_tbv <- sexdiffs.df %>%
  filter(weighted==FALSE & total==TRUE)

sexdiffs <- sexdiffs.df %>%
  filter(weighted==FALSE & total==FALSE)
```

Plot

```{r}
for (df in list(sexdiffs_tbv, sexdiffs)){
  t <- as.character(unique(df$total))
  print(t)
  
  #median centile
  cent_df <- df %>%
    dplyr::select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  cent_plt <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_cent.png")
  ggsave(n, cent_plt)
  print(cent_plt)
  
  cent_plt2 <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Scaled 50th Centile Sex-Difference")
  
  print(cent_plt2)
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_cent_scaled.png")
  ggsave(n, cent_plt2)

  #sigma
  sig_df <- df %>%
    dplyr::select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  sig_plt <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="log CV Ratio")
  print(sig_plt)
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_cv.png")
  ggsave(n, sig_plt)
  
  #derivatives
  deriv_df <- df %>%
    dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  deriv_plt <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="50th Centile Difference Derivative")
  print(deriv_plt)
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_deriv.png")
  ggsave(n, deriv_plt)
  
  deriv_plt2 <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="Scaled 50th Centile Difference Derivative")
  print(deriv_plt2)
  
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_deriv_scaled.png")
  ggsave(n, deriv_plt2)
}
```

interactive plots
```{r, eval=FALSE}
cent_plt <- sexdiffs_tbv %>%
  dplyr::select(centile_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    labs(y="Scaled 50th Centile Sex-Difference")
ggplotly(cent_plt)

deriv_plt <- sexdiffs_tbv %>%
  dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    labs(y="Scaled 50th Centile Difference Derivative")
ggplotly(deriv_plt)
```

###Plots

Average across splits to make summary plots for global phenos

```{r}
sum_cols <- setdiff(names(sexdiffs.df.u), c("logAge_days", "cv_sample", "total", "weighted", "pheno", "uniform", "pheno_cat", "age_days"))

glob_mean <- sexdiffs.df.u %>%
  filter(pheno %in% global_list & total==TRUE) %>%
  group_by(pheno, age_days) %>%
  #filtering outliers
  filter(age_days >= 370) %>%
  summarise(across(all_of(sum_cols), mean)) %>%
  #rescale
  group_by(pheno) %>%
  mutate(centile_M_minus_F_z = centile_M_minus_F_z / max(abs(centile_M_minus_F_z)) * 100) %>%
  ungroup()

glob_mean_sums <- glob_mean %>%
  summarize(#age at peak sex diff
            max_sex_diff = extr_val(centile_M_minus_F_z, which_max_gc(abs(centile_M_minus_F_z))),
            age_max_sex_diff = extr_val(age_days, which_max_gc(abs(centile_M_minus_F_z))),
            #age at min sex diff
            min_sex_diff = extr_val(centile_M_minus_F_z, which_min_gc(abs(centile_M_minus_F_z))),
            age_min_sex_diff = extr_val(age_days, which_min_gc(abs(centile_M_minus_F_z))),
            #age peak change in sex diff
            max_deriv = extr_val(centile_M_minus_F_z, which_max_gc(abs(deriv_M_minus_F_z))),
            age_max_deriv = extr_val(age_days, which_max_gc(abs(deriv_M_minus_F_z))),
            #age at male peak
            max_M = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Male_z)),
            age_max_M = extr_val(age_days, which_max_gc(median_centile_Male_z)),
            #age at female peak
            max_F = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Female_z)),
            age_max_F = extr_val(age_days, which_max_gc(median_centile_Female_z)),
            #age at male peak growth
            max_deriv_M = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Male_z)),
            age_max_deriv_M = extr_val(age_days, which_max_gc(deriv_median_centile_Male_z)),
            #age at female peak growth
            max_deriv_F = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Female_z)),
            age_max_deriv_F = extr_val(age_days, which_max_gc(deriv_median_centile_Female_z)),
            .by = pheno
            )

#add age at peak rate divergence!!

#pivot and rejoin
values_long <- glob_mean_sums %>%
  select(!starts_with("age_")) %>%
  pivot_longer(cols = -c(pheno), names_to = "metric", values_to = "value")

ages_long <- glob_mean_sums %>%
  select(pheno, starts_with("age_")) %>%
  pivot_longer(cols = starts_with("age_"), names_to = "metric", values_to = "age") %>%
  mutate(metric = gsub("^age_", "", metric))

glob_mean_sums_long <- left_join(values_long, ages_long, by = c("pheno", "metric"))


#reformat x to keep birth label
reformat_x <- function(x_values, simplified=FALSE){
  tickLabels <- c()
  tickMarks <- c()
  if(simplified==TRUE){
    for (year in c(0, 25, 50, 75, 100)) {
     tickMarks <- append(tickMarks, year * 365.25 + 280)
    }
    tickLabels <- c("Birth", "25", "50", "75", "100")
  } else {
    for (year in seq(0, 100, by = 10)) {
      tickMarks <- append(tickMarks, year * 365.25 + 280)
    }
    tickLabels <- c("Birth", "10", "20", "30", "40", "50", "60", "70", "80", "90", "100")
  }
  
  unit_lab <- "(years)"
  list(scale_x_continuous(breaks = tickMarks, labels = tickLabels), labs(x = paste("Age at Scan", unit_lab)))
}

cent_plt_pct_mean <- glob_mean %>%
  ggplot() +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
  geom_point(
    data = glob_mean_sums_long,
    aes(x = age, y = value, color = pheno, shape = metric),
    size = 3
  ) +
  scale_shape_manual(
    values = c(
      "max_sex_diff" = 19,
      "min_sex_diff" = 15,
      "max_deriv"    = 18,
      "max_M"        = 2,
      "max_F"        = 6,
      "max_deriv_M"  = 3,
      "max_deriv_F"  = 4
    ),
    labels = c(
      "max_sex_diff" = "Max Sex Bias",
      "min_sex_diff" = "Min Sex Bias",
      "max_deriv"    = "Max Change Sex Bias",
      "max_M"        = "Max Male",
      "max_F"        = "Max Female",
      "max_deriv_M"  = "Max Change Male",
      "max_deriv_F"  = "Max Change Female"
    ),
    name = "Metric"
  ) +
  scale_color_discrete(
    labels = c(
      "total.SA" = "Total SA",
      "mean.CT"  = "Mean CT",
      "CBV"      = "Cerebellum",
      "CSF"      = "Ventricles",
      "GMV"      = "Cortical GM",
      "sGMV"     = "Subcortical GM",
      "WMV"      = "WM"
    ),
    name = "Phenotype"
  ) +
  theme_linedraw() +
  reformat_x(glob_mean$age_days) +
  labs(y="% Maximum Sex-Difference")
saveRDS(cent_plt_pct_mean, "figs/ggobjs/glob_trajectories.rds")
ggsave("figs/supplement/glob_trajectories.png", cent_plt_pct_mean, dpi=600, width=11, height=5)
```

```{r}
biggest_agesex_resi <- LR.df_tot %>%
  mintyr:::top_perc(perc = 0.1,                # Select top 10%
         trait = c("resi"),  # Column to analyze
         by = "cv_sample",
         keep_data = TRUE)

biggest_agesex_resi <- biggest_agesex_resi$resi_0.1$data %>% ungroup() %>% select(!pheno_cat)
biggest_agesex_resi$top10 <- TRUE

biggest_sex_resi <- LR.df_tot_sex %>%
  mintyr:::top_perc(perc = 0.1,                # Select top 10%
         trait = c("resi"),  # Column to analyze
         by = "cv_sample",
         keep_data = TRUE)

biggest_sex_resi <- biggest_sex_resi$resi_0.1$data %>% ungroup() %>% select(!pheno_cat)
biggest_sex_resi$top10 <- TRUE

#highlight trajectories of phenos with large effects
top_sex_plt <- full_join(sexdiffs_tbv, biggest_sex_resi) %>%
  dplyr::select(centile_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, top10) %>%
  mutate(top10 = ifelse(is.na(top10), "no", "yes")) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(top10 == "yes", calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    labs(title="Male-Female Trajectory Differences in Phenos\nwith Strongest Sex Effects (Top 10%)", 
         y="Scaled 50th Centile Sex-Difference")

top_agesex_plt <- full_join(sexdiffs_tbv, biggest_agesex_resi) %>%
  dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, top10) %>%
  mutate(top10 = ifelse(is.na(top10), "no", "yes")) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
  gghighlight(top10 == "yes", calculate_per_facet = TRUE) +
    labs(title="Male-Female Rate-of-Growth Differences in Phenos\nwith Strongest Age*Sex Effects (Top 10%)", 
         y="Scaled 50th Centile Difference Derivative")

ggsave("figs/misc/top_sexdiffs_centiles.png", top_sex_plt)
ggsave("figs/misc/top_agesexdiffs_derivs.png", top_agesex_plt)
```

###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#calculate integral of deriv of diff^2 (all in z-scores)
sexdiffs_tbv_sq <- sexdiffs_tbv %>%
  mutate(sq_centile_diff_z = (centile_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_deriv_z = get_deriv(sq_centile_diff_z, logAge_days)) %>%
  mutate(sq_deriv_z.unlog = sq_deriv_z*deriv_age) %>% #chain rule to convert derivative
  ungroup()

#try getting integral
convergence_int <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv_z),
            sq_int_z.unlog = rs_integral(logAge_days, sq_deriv_z)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging")) #agree regardless of whether log-age or non-log-age is used
```

finding points at which pt derivative crosses 0 increasing vs decreasing

```{r}
cent_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv_z.unlog > 0 ~ "diverging",
    sq_deriv_z.unlog == 0 ~ "constant",
    sq_deriv_z.unlog < 0 ~ "converging",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- cent_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv_z, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

cent_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", cent_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))

#just look at phenos with consistent age*sex effects across both cvs
cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)
```


####Sigma

*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of lnCVR squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
sigma_df <- sexdiffs_tbv %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_cv_diff = (logcv_M_div_F)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_cv_deriv = get_deriv(sq_cv_diff, logAge_days)) %>%
  ungroup() %>%
  mutate(sq_cv_deriv.unlog = sq_cv_deriv*deriv_age)

sigma_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_cv_deriv, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sigma_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Variability")

#try getting integral
sigma_convergence_df <- sigma_df %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_cv_deriv),
            sq_int_z.unlog = rs_integral(logAge_days, sq_cv_deriv)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging"))
```

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)
Calculating across age, not log-age, for interpretability

```{r}
dominance_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = rs_integral(logAge_days, centile_M_minus_F_z),
            sigma_int.unlog =  rs_integral(logAge_days, logcv_M_div_F),
            age_range = max(age_days) - min(age_days)) %>%
  # rescaling integrals to 100 years
  mutate(centile_int_norm.unlog = (centile_int.unlog/age_range)*365.25*100,
         sigma_int_norm.unlog = (sigma_int.unlog/age_range)*365.25*100) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

dominance_df %>%
  group_by(cv_sample) %>%
  summarise(pct_male=sum(dominance=="Male")/n()*100,
            pct_female=sum(dominance=="Female")/n()*100)

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  dplyr::count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5)) +
  coord_flip()
```

restrict to phenos with significant sex effects

```{r}
dominance_plt_df <- full_join(LR.df_tot_sex, dominance_df %>% dplyr::select(!pheno_cat))
dominance_plt_df %>%
  mutate(dominance = ifelse(sig_fdr==TRUE, dominance, "NA")) %>%
  group_by(cv_sample) %>%
  summarise(F_dom = sum(dominance=="Female")/n(),
            M_dom = sum(dominance=="Male")/n())
```

plot dominance on brain

```{r}
#removing ventricles for visualization:
dominance_plt_df_novent <- dominance_plt_df %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_novent$centile_int_norm.unlog)

full_dom_plt <- plot_cv_brain(df=dominance_plt_df_novent, 
                              fill_var=centile_int_norm.unlog, 
                              fill_name = "Net Sex\nDifference\nover\n100 yrs", 
                              fill_color="ggthemes::Orange-Blue Diverging", 
                              fill_limits= c(-0.07, 0.07), 
                              height_list=c(1.03, 4.5, .93), 
                              direction=1)

ggsave("figs/supplement/dominance_tbvTRUE.png", full_dom_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```

How much of the lifespan are phenos male vs female dominant?

```{r}
#define function
prop_above <- function(x, y, threshold=0){
  crosses <- find_threshold_crosses(y, x, threshold, return="x", return_endpoints = FALSE)
  crosses <- c(crosses, min(x), max(x)) %>% 
    sort()
  #overall age modeled
  total_range <- max(x) - min(x)
  
  ranges <- diff(crosses)
  
  #if y > threshold at min(x) -> sum odd number ranges
  odd_indices_logical <- (1:length(ranges)) %% 2 == 1
  
  if (y[1] > threshold) {
    items_at_odd_indices <- ranges[odd_indices_logical]
    time_above <- sum(items_at_odd_indices)
  } else {
    items_at_even_indices <- ranges[!odd_indices_logical]
    time_above <- sum(items_at_even_indices)
  }
  
  prop_above <- time_above/total_range
  
  return(prop_above)
}

#test function
test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="sGMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="WMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="GMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="B" & pheno=="GMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="B" & pheno=="lh.SA.pericalcarine")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))
```

```{r}
prop_above_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, centile_M_minus_F_z)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male < .05 ~ "F",
    prop_male > .95 ~ "M",
    TRUE ~ "mixed"),
    strict_class = case_when(
    prop_male == 0 ~ "F",
    prop_male == 1 ~ "M",
    TRUE ~ "mixed")
  )

#add back sex-effect significance, remove ventricles for visualization
prop_above_df_novent <- full_join(LR.df_tot_sex, prop_above_df %>% dplyr::select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_plt <- plot_cv_brain(df=prop_above_df_novent, 
                                fill_var=prop_male, 
                                fill_name = "Proportion\nMale-\nDominant",
                                fill_color = "ggthemes::Orange-Blue Diverging", 
                                fill_limits=c(0, 1), 
                                height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_tbvTRUE.png", full_above_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```
####Sigma

```{r}
dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  dplyr::count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()


dominance_plt_df %>%
  #restrict to significant sex effects
  mutate(sigma_dominance = ifelse(sig_fdr==TRUE, sigma_dominance, "NA")) %>%
  group_by(cv_sample) %>%
  summarise(pct_male=sum(sigma_dominance=="Male")/n()*100,
            pct_female=sum(sigma_dominance=="Female")/n()*100)
```


```{r}
range(dominance_plt_df_novent$sigma_int_norm.unlog)

full_dom_plt_sigma <- plot_cv_brain(df=dominance_plt_df_novent, 
                                    fill_var=sigma_int_norm.unlog, 
                                    fill_name = "Net Sex\nDifference\nover\n100 yrs", 
                                    fill_color = "ggthemes::Orange-Blue Diverging", 
                                    fill_limits =  c(-.03, .03),
                                    height_list= c(1.03, 4.5, .93))

ggsave("figs/supplement/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, dpi=600, height=8.5, width=7.5, bg = "white")
```

```{r}
prop_above_sigma_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, logcv_M_div_F)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male < .05 ~ "F",
    prop_male > .95 ~ "M",
    TRUE ~ "mixed"),
    strict_class = case_when(
    prop_male == 0 ~ "F",
    prop_male == 1 ~ "M",
    TRUE ~ "mixed")
  )

table(prop_above_sigma_df$cv_sample, prop_above_sigma_df$class)

prop_above_sigma_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="F"),
            n_male = sum(class=="M"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_sigma_df_novent <- full_join(LR.df_tot_sex, prop_above_sigma_df %>% dplyr::select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_sig_plt <- plot_cv_brain(df=prop_above_sigma_df_novent, 
                                    fill_var=prop_male, 
                                    fill_name="Proportion\nMale-\nDominant", 
                                    fill_color="ggthemes::Orange-Blue Diverging", 
                                    fill_limits= c(0, 1), 
                                    height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_sigma_tbvTRUE.png", full_above_sig_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```

Prep for fig 3

```{r}
dominance_plt_df_B <- dominance_plt_df_novent %>%
  filter(cv_sample=="B")

#centiles
range(dominance_plt_df_B$centile_int_norm.unlog)
full_dom_pltB <- plot_cv_brain(df=dominance_plt_df_B, 
                               fill_var=centile_int_norm.unlog, 
                               fill_name="Normalized\nSex\nDifference", 
                               fill_color="ggthemes::Orange-Blue Diverging", 
                               fill_limits=c(-0.07, 0.07),
                               height_list=c(1.03, 4.5, .93), 
                               return_plt = FALSE)
saveRDS(full_dom_pltB, "figs/ggobjs/cent_dom_B.rds")

#sigma
range(dominance_plt_df_B$sigma_int_norm.unlog)
full_dom_plt_sigmaB <- plot_cv_brain(df=dominance_plt_df_B, 
                                     fill_var=sigma_int_norm.unlog, 
                                     fill_name="Normalized\nSigma Sex\nDifference", 
                                     fill_color="ggthemes::Orange-Blue Diverging", 
                                     fill_limits=c(-0.03, 0.03), 
                                     height_list=c(1.03, 4.5, .93), 
                                     return_plt = FALSE)
saveRDS(full_dom_plt_sigmaB, "figs/ggobjs/sig_dom_B.rds")
```

####Spin Test

Check whether sex-dominance in trajectories is associated with sex-dominance in variance

setup based on code from Val and claude AI

```{r}
#read in files
sphere.lh <- read.fs.surface(file.path(fs_path, "/surf/lh.sphere"))
sphere.rh <- read.fs.surface(file.path(fs_path, "/surf/rh.sphere"))
annot.lh <- freesurfer:::read_annotation(file.path(fs_path, "/label/lh.aparc.annot"))
annot.rh <- freesurfer:::read_annotation(file.path(fs_path, "/label/rh.aparc.annot"))

# Compute centroid (except for medial wall)
rm <- which(annot.lh$colortable$label %in% c("unknown", "corpuscallosum"))
#left hemisphere
lhcoords <- t(sapply(c(1:length(annot.lh$colortable$label))[-rm], function(x) 
  colMeans(sphere.lh$vertices[which(annot.lh$label == annot.lh$colortable$code[x]), ])
))
rownames(lhcoords) <- annot.lh$colortable$label[-rm]
#right hemisphere
rhcoords <- t(sapply(c(1:length(annot.rh$colortable$label))[-rm], function(x) 
  colMeans(sphere.rh$vertices[which(annot.rh$label == annot.rh$colortable$code[x]), ])
))
rownames(rhcoords) <- annot.rh$colortable$label[-rm]
```

spin permutations
```{r}
perm_file <- "spin_perms_DK68_n10000.rds"
 
if (file.exists(perm_file)) {
  message("Loading existing permutations from ", perm_file)
  perm.id <- readRDS(perm_file)
} else {
  message("Generating 10,000 spin permutations (this may take a few minutes)...")
  perm.id <- rotate.parcellation(lhcoords, rhcoords, nrot = 10000, method = "hungarian")
  saveRDS(perm.id, perm_file)
  message("Permutations saved to ", perm_file)
}
```

prep dominance dfs

```{r}
spin_prep <- dominance_df %>%
  filter(pheno %in% c(sa_list, ct_list, vol_list)) %>%
  select(pheno, cv_sample, pheno_cat, centile_int_norm.unlog, sigma_int_norm.unlog) %>%
  mutate(hemi = stringr::str_extract(pheno, "(lh|rh)"),
         pheno = str_extract(pheno, "(?<=\\.)[^.]+$"))

#check pheno names match
stopifnot(all(unique(spin_prep$pheno[spin_prep$hemi == "lh"]) %in% rownames(lhcoords)))
stopifnot(all(unique(spin_prep$pheno[spin_prep$hemi == "rh"]) %in% rownames(rhcoords)))
#check region counts = 34
region_counts <- spin_prep %>%
  group_by(cv_sample, pheno_cat, hemi) %>%
  summarise(n = n(), .groups = "drop") %>%
  filter(n != 34)
```

run spin test
```{r}
spin_prep %>%
  group_by(cv_sample, pheno_cat) %>%
  group_modify(~ {
    # Order LH and RH rows to match coordinate matrices
    lh_data <- .x %>%
      filter(hemi == "lh") %>%
      arrange(match(pheno, rownames(lhcoords)))

    rh_data <- .x %>%
      filter(hemi == "rh") %>%
      arrange(match(pheno, rownames(rhcoords)))

    # Subset coordinate matrices to only regions present in this group
    lhcoords_sub <- lhcoords[rownames(lhcoords) %in% lh_data$pheno, , drop = FALSE]
    rhcoords_sub <- rhcoords[rownames(rhcoords) %in% rh_data$pheno, , drop = FALSE]

    # Re-order data to match subsetted coordinate matrices
    lh_data <- lh_data %>% arrange(match(pheno, rownames(lhcoords_sub)))
    rh_data <- rh_data %>% arrange(match(pheno, rownames(rhcoords_sub)))

    # Generate group-specific permutations if regions are missing
    n_regions <- nrow(lh_data) + nrow(rh_data)
    if (n_regions < 68) {
      message("Group ", unique(.x$cv_sample[1]), " / ", unique(.x$pheno_cat[1]),
              ": only ", n_regions, " regions — generating group-specific permutations")
      perm.id_group <- rotate.parcellation(lhcoords_sub, rhcoords_sub,
                                           nrot = 10000, method = "hungarian")
    } else {
      perm.id_group <- perm.id
    }

    # Build map vectors
    map_centile <- c(lh_data$centile_int_norm.unlog, rh_data$centile_int_norm.unlog)
    map_sigma   <- c(lh_data$sigma_int_norm.unlog,   rh_data$sigma_int_norm.unlog)

    obs_r  <- cor(map_centile, map_sigma, method = "spearman")
    p_spin <- perm.sphere.p(map_centile, map_sigma, perm.id_group, corr.type = "spearman")

    tibble(obs_r = obs_r, p_spin = p_spin, n_regions = n_regions)
  }) %>%
  ungroup()

```

no significant association between mean and variance dominance

###Grouping
grouping by M vs F, converging vs diverging

```{r}
ggplot(prop_above_df) +
  geom_histogram(aes(x=prop_male)) +
  geom_vline(xintercept = .05) +
  geom_vline(xintercept = .95)

table(prop_above_df$class, prop_above_df$cv_sample) # > 95% lifespan in one sex
table(prop_above_df$strict_class, prop_above_df$cv_sample) #100% lifespan in one sex

#add age*sex significance info
prop_above_df2 <- full_join(prop_above_df, LR.df_tot)

#add info about convergence vs divergence
prop_above_df3 <- full_join(prop_above_df2, convergence_int) %>%
  mutate(class = paste(class, overall_trajectory, sep="_"))

prop_above_df3 %>%
  filter(sig_fdr==TRUE) %>%
  group_by(cv_sample, class) %>%
  summarise(n = n())
```

what's going on with the mixed phenotypes?

```{r}
#add back trajectory
group_cd <- full_join(sexdiffs_tbv, prop_above_df3)

mixed_plt <- group_cd %>%
  filter(sig_fdr==TRUE & (class=="mixed_converging" | class=="mixed_diverging")) %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y")
ggplotly(mixed_plt)

#try simple grouping by sign at start and end (Lena's method)
group_cd_mix <- group_cd %>%
  filter(class=="mixed_converging" | class=="mixed_diverging") %>%
  select(logAge_days, cv_sample, pheno, centile_M_minus_F_z) %>%
  arrange(logAge_days) %>%
  group_by(cv_sample, pheno) %>%
  slice(c(1,n())) %>%
  mutate(sex = ifelse(centile_M_minus_F_z > 0 , "M", "F")) %>%
  summarise(transition = paste(first(sex), last(sex), sep = "_to_"), .groups = "drop")
table(group_cd_mix$transition, group_cd_mix$cv_sample)

mixed_plt2 <- full_join(group_cd, group_cd_mix) %>%
  filter(class=="mixed_converging" | class=="mixed_diverging" & sig_fdr==TRUE) %>%
  #filter(pheno != "rh.CT.rostralanteriorcingulate") %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(transition~cv_sample, scales="free_y") +
  theme(legend.position="none")

ggplotly(mixed_plt2)
```

```{r}
group_traj_final <- full_join(group_cd, group_cd_mix) %>%
  mutate(cent_group=case_when(
    class=="mixed_converging" | class=="mixed_diverging" ~ transition,
    TRUE ~ class
  )) %>%
  #regroup U trajectories
  rename(cent_group_full = cent_group) %>%
  mutate(cent_group = case_when(
    cent_group_full == "F_to_F" | cent_group_full == "M_to_M" ~ "other",
    TRUE ~ cent_group_full
  )) %>%
  mutate(parent_group = case_when(
    sig_fdr==FALSE ~ "Stable",
    cent_group == "F_diverging" | cent_group == "M_diverging" ~ "Diverging",
    cent_group == "F_converging" | cent_group == "M_converging" ~ "Converging",
    cent_group == "M_to_F" | cent_group == "F_to_M" ~ "Flip",
    TRUE ~ str_to_title(cent_group)
  ))


group_final <- group_traj_final %>%
  select(cv_sample, total, weighted, pheno, pheno_cat, class, prop_male, strict_class, resi, sig_fdr, overall_trajectory, transition, cent_group, label, parent_group) %>%
  unique() %>%
  #filter(!(pheno %in% vent_list)) %>%
  mutate(cent_group = ifelse(sig_fdr==FALSE, "NS", cent_group)) 
```

```{r}
#GET FINAL COUNTS
#group
table(group_final$cent_group, group_final$cv_sample)

#parent groups
group_final %>%
  group_by(cv_sample, parent_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)
```

plot formatting prep

```{r}
group_values <- c("F_diverging"="#9E3D22FF", "F_converging"="#C06636FF","F_to_M" ="#E8B960FF", 
             "M_to_F" = "#528AA9", "M_converging"="#426f89","M_diverging"="#154063","other" = "gray40",
             "NS" = "gray75")

group_color_scale <- scale_colour_manual(
  values = group_values,
  breaks = c("F_diverging", "F_converging", "F_to_M", 
             "M_to_F", "M_converging", "M_diverging", 
             "other", "NS"),
  labels = c("Female Diverging", "Female Converging", "Female to Male Bias", 
             "Male to Female Bias", "Male Converging", "Male Diverging", 
             "Other", "Stable"),
  name = "Group",
  na.value="gray90"
)

group_fill_scale <- scale_fill_manual(
  values = group_values,
  breaks = c("F_diverging", "F_converging", "F_to_M", 
             "M_to_F", "M_converging", "M_diverging", 
             "other", "NS"),
  labels = c("Female Diverging", "Female Converging", "Female to Male Bias", 
             "Male to Female Bias", "Male Converging", "Male Diverging", 
             "Other", "Stable"),
  name = "Group",
  na.value="gray90"
)

#alternate version with different spacing for labels
group_fill_scale2 <- scale_fill_manual(
  values = group_values,
  breaks = c("F_diverging", "F_converging", "F_to_M", 
             "M_to_F", "M_converging", "M_diverging", 
             "other", "NS"),
  labels = c("Female\nDiverging", "Female\nConverging", "Female to\nMale Bias", 
             "Male to\nFemale Bias", "Male\nConverging", "Male\nDiverging", 
             "Other", "Stable"),
  name = "Group",
  na.value="gray90"
)

bottom_group_legend <- get_legend(
  ggplot(group_final) +
    geom_bar(aes(x=cv_sample, fill=cent_group)) +
    group_fill_scale2 +
    theme(legend.position = "bottom",
          legend.text = element_text(margin = margin(r = .5, unit = "cm"))) +
    guides(fill = guide_legend(nrow = 1)),
  legend="bottom"
  )
```

```{r}
group_bar_plt_total <- group_final %>%
  group_by(cv_sample) %>% 
  mutate(cent_group = as.character(cent_group)) %>%
  mutate(cent_group = ifelse(sig_fdr ==TRUE, cent_group, NA)) %>%
  mutate(cent_group = factor(cent_group, levels = names(group_values))) %>%  
  dplyr::count(cent_group, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x="TOTAL", y=perc, fill=cent_group)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
  geom_text(aes(label = n), 
            position = position_fill(vjust = 0.5),
            size = 1.5) +
  scale_y_continuous(labels=scales::percent) +
  labs(y="Percentage") +
  theme_linedraw() +
  theme(legend.position = "none",
        axis.title.x= element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  group_fill_scale

group_bar_plt <- group_final %>%
  #filter(sig_fdr==TRUE & pheno_cat != "Global Vol") %>%
  ggplot() +
  geom_bar(aes(x=parent_group, fill=cent_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  group_fill_scale +
  guides(x = guide_axis(angle = 45)) +
  labs(y="Count") +
  theme_linedraw() +
  theme(axis.title.x = element_blank(),
        legend.position = "none")

group_bar_plt_all <- plot_grid(
    group_bar_plt_total, 
    group_bar_plt,
    ncol = 2,
    rel_widths = c(1,4),
    align = "hv",
    axis = "lr",
    labels=c("A", "B"))

#ggsave("figs/supplement/group_bar_plt.png", group_bar_plt_all, dpi=600, height=9, width=10, bg = "white")
```


```{r}
#just plot B
barplt_final <- group_final %>%
  filter(sig_fdr==TRUE & cv_sample=="B") %>%
  mutate(parent_group = case_when(
    cent_group == "F_diverging" | cent_group == "M_diverging" ~ "Diverging",
    cent_group == "F_converging" | cent_group == "M_converging" ~ "Converging",
    cent_group == "M_to_F" | cent_group == "F_to_M" ~ "Flip",
    TRUE ~ str_to_title(cent_group)
  )) %>%
  dplyr::count(pheno_cat, parent_group, cent_group) %>%
  group_by(pheno_cat) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  ggplot() +
  geom_bar(aes(x=parent_group, y=pct, fill=cent_group), stat="identity") +
  facet_wrap(~pheno_cat, nrow=1) +
  scale_y_continuous(labels=scales::percent) +
  group_fill_scale +
  guides(x = guide_axis(angle = 45)) +
  labs(y="Percentage") +
  theme_linedraw() +
  theme(axis.title.x = element_blank(), legend.position = "bottom")
saveRDS(barplt_final, "figs/ggobjs/cent_groups_barplt_B.rds")
```

brain plots

```{r}
group_final_novent <- group_final %>%
  filter(!(pheno %in% vent_list))

group_brain_plt <- plot_cv_brain(group_final_novent, cent_group, "Group", fill_limits=levels(group_final$cent_group), height_list=c(1.03, 4.5, .93), return_plt = FALSE, fill_values=group_values)

group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale

group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale

group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale

#recreate legends
legend <- get_legend(
  ggplot(group_final) +
    geom_bar(aes(x=cv_sample, fill=cent_group)) + group_fill_scale
  )

plts <- cowplot::plot_grid(
    group_brain_plt[[1]],
    group_brain_plt[[2]],
    group_brain_plt[[3]],
    ncol = 1,
    nrow = 3,
    rel_heights = c(1.03, 4.5, .93),
    align = "v",
    axis = "lr",
    greedy = TRUE
  ) 
  
full_plt_obj <- plot_grid(plts, legend, ncol=2, rel_widths = c(3, 1))

#ggsave("figs/supplement/group_brain_plt.png", full_plt_obj, dpi=600, height=8.5, width=8, bg = "white")
```

plot trajectories by group

```{r}
group_traj_final %>%
  arrange(pheno, age_days) %>%
  ggplot(aes(x=age_days, y=centile_M_minus_F_z, group=pheno, color=cent_group)) +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  geom_line(aes(alpha=sig_fdr)) +
  #gghighlight(sig_fdr==TRUE, calculate_per_facet = TRUE) +
  #facet_grid(cent_group ~ cv_sample, scales="free_y") +
  ggh4x:::facet_nested(cent_group ~ pheno_cat + cv_sample, scales="free_y") +
  group_color_scale +
  scale_alpha_manual(values=c("TRUE"=0.7, "FALSE"=0.1))+
  theme_bw() +
  format_x_axis("lifespan_fetal", group_traj_final$age_days)+
  theme(legend.position="none")

group_traj_full <- group_traj_final %>%
  filter(sig_fdr==TRUE & cent_group != "other") %>%
  arrange(pheno, age_days) %>%
  ggplot(aes(x=age_days, y=centile_M_minus_F_z, color=cent_group)) +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  geom_line(aes(group=pheno), alpha=.1) +
  facet_grid(parent_group ~ cv_sample, scales="free_y") +
  geom_smooth(method = "gam") +
  group_color_scale +
  reformat_x(group_traj_final$age_days) +
  theme_linedraw() +
  #theme(legend.position="none") +
  labs(y="Scaled Median Sex Difference")
#ggsave("figs/supplement/group_traj_plt.png", group_traj_full, dpi=600, height=8, width=10, bg = "white")
```

```{r}
#save combined group plots for supplement
sub_plt <- plot_grid(NULL, plts,
                      group_traj_full,
                      nrow=1, rel_widths =c(.1,1,2),
                      labels=c("C", "", "D"))

group_plot_supplement <- plot_grid(group_bar_plt_all,
                          sub_plt,
                          bottom_group_legend,
                          nrow=3,
                          ncol=1,
                          rel_heights = c(1, 1.1, .1))
ggsave("figs/supplement/groups_all_plt.png", group_plot_supplement, dpi=600, height=16, width=16, bg = "white")
```

also just plot B
```{r}
##BRAIN PLOT
group_finalB <- group_final %>% filter(cv_sample=="B")
group_brain_plt <- plot_cv_brain(group_finalB, cent_group, "Group", fill_limits=levels(group_finalB$cent_group), height_list=c(1.03, 4.5, .93), return_plt = FALSE, fill_values=group_values)

group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale

group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale

group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale

#recreate legends
group_brain_plt[[4]] <- bottom_group_legend
saveRDS(group_brain_plt, "figs/ggobjs/cent_groups_brain_B.rds")

##TRAJECTORIES
group_traj_plt_B <- group_traj_final %>%
  filter(sig_fdr==TRUE & cv_sample=="B" & cent_group != "other") %>%
  arrange(pheno, age_days) %>%
  ggplot(aes(x=age_days, y=centile_M_minus_F_z, color=cent_group)) +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  #geom_line(aes(group=pheno), alpha=.2) +
  facet_wrap(~parent_group, nrow=3) +
  geom_smooth(method = "gam") +
  group_color_scale +
  reformat_x(group_traj_final$age_days) +
  theme_linedraw() +
  theme(legend.position="none") +
  labs(y="Scaled Median Sex Difference")
saveRDS(group_traj_plt_B, "figs/ggobjs/cent_groups_traj_B.rds")
```

####Sigma

```{r}
ggplot(prop_above_sigma_df) +
  geom_histogram(aes(x=prop_male)) +
  geom_vline(xintercept = .05) +
  geom_vline(xintercept = .95)

table(prop_above_sigma_df$class, prop_above_sigma_df$cv_sample)
table(prop_above_sigma_df$strict_class, prop_above_sigma_df$cv_sample)

#look at age*sex significance
prop_above_sigma_df2 <- full_join(prop_above_sigma_df, LR.df_tot)

#see convergence vs divergence
prop_above_sigma_df3 <- full_join(prop_above_sigma_df2, sigma_convergence_df) %>%
  mutate(class = paste(class, overall_trajectory, sep="_"))

prop_above_sigma_df3 %>%
  filter(sig_fdr==TRUE) %>%
  group_by(cv_sample, class) %>%
  summarise(n = n())
```

what's going on with the mixed phenotypes?

```{r}
#add back in trajectory data
group_sig_cd <- full_join(sexdiffs_tbv, prop_above_sigma_df3)

group_sig_cd %>%
  filter(sig_fdr==TRUE & (class=="mixed_converging" | class=="mixed_diverging")) %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  theme(legend.position = "none")

#should work to just do m to f and f to m
group_sig_cd_mix <- group_sig_cd %>%
  filter(class=="mixed_converging" | class=="mixed_diverging") %>%
  select(logAge_days, cv_sample, pheno, logcv_M_div_F) %>%
  arrange(logAge_days) %>%
  group_by(cv_sample, pheno) %>%
  slice(c(1,n())) %>%
  mutate(sex = ifelse(logcv_M_div_F > 0 , "M", "F")) %>%
  summarise(transition = paste(first(sex), last(sex), sep = "_to_"), .groups = "drop")
table(group_sig_cd_mix$transition, group_sig_cd_mix$cv_sample)

mixed_sig_plt2 <- full_join(group_sig_cd, group_sig_cd_mix) %>%
  filter(class=="mixed_converging" | class=="mixed_diverging" & sig_fdr==TRUE) %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(transition~cv_sample, scales="free_y") +
  theme(legend.position="none")

ggplotly(mixed_sig_plt2)
```

sample A rh.CT.lateraloccipital is the only feature thats wiggly, doesn't just transition M-F or vice versa

```{r}
group_sig_traj_final <- full_join(group_sig_cd, group_sig_cd_mix) %>%
  mutate(sig_group=case_when(
    class=="mixed_converging" | class=="mixed_diverging" ~ transition,
    TRUE ~ class
  )) %>%
  mutate(sig_group = case_when(
    sig_group == "F_to_F" | sig_group == "M_to_M" ~ "other",
    TRUE ~ sig_group
  )) %>%
  mutate(parent_group = case_when(
    sig_fdr == FALSE ~ "Stable",
    sig_group == "F_diverging" | sig_group == "M_diverging" ~ "Diverging",
    sig_group == "F_converging" | sig_group == "M_converging" ~ "Converging",
    sig_group == "M_to_F" | sig_group == "F_to_M" ~ "Flip",
    TRUE ~ str_to_title(sig_group)
  ))

group_sig_final <- group_sig_traj_final %>%
  select(cv_sample, total, weighted, pheno, pheno_cat, class, prop_male, strict_class, resi, sig_fdr, overall_trajectory, transition, sig_group, label, parent_group) %>%
  unique() %>%
  #filter(!(pheno %in% vent_list))  %>%
  mutate(sig_group = ifelse(sig_fdr==FALSE, "NS", sig_group))
```

```{r}
#GET FINAL COUNTS
#group
group_sig_final %>%
  group_by(cv_sample, sig_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)

#parent groups
group_sig_final %>%
  group_by(cv_sample, parent_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)

#group - just looking at significant age*sex
group_sig_final %>%
  #filter(sig_group != "NS") %>%
  group_by(cv_sample, sig_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)%>%
  arrange(desc(pct), by_group=T)
```

bar plot

```{r}
sig_group_bar_plt_total <- group_sig_final %>%
  group_by(cv_sample) %>% 
  mutate(sig_group = as.character(sig_group)) %>%
  mutate(sig_group = ifelse(sig_fdr ==TRUE, sig_group, NA)) %>%
  mutate(sig_group = factor(sig_group, levels = names(group_values))) %>%  
  dplyr::count(sig_group, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x="TOTAL", y=perc, fill=sig_group)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
  geom_text(aes(label = n), 
            position = position_fill(vjust = 0.5),
            size = 1.5) +
  scale_y_continuous(labels=scales::percent) +
  labs(y="Percentage") +
  theme_linedraw() +
  theme(legend.position = "none",
        axis.title.x= element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  group_fill_scale

sig_group_bar_plt <- group_sig_final %>%
  #filter(sig_fdr==TRUE & pheno_cat != "Global Vol") %>%
  ggplot() +
  geom_bar(aes(x=parent_group, fill=sig_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  group_fill_scale +
  guides(x = guide_axis(angle = 45)) +
  labs(y="Count") +
  theme_linedraw() +
  theme(axis.title.x = element_blank(),
        legend.position = "none")

sig_group_bar_plt_all <- plot_grid(
    sig_group_bar_plt_total, 
    sig_group_bar_plt,
    ncol = 2,
    rel_widths = c(1,4),
    align = "hv",
    axis = "lr",
    labels=c("A", "B"))


#ggsave("figs/supplement/sig_group_bar_plt.png", sig_group_bar_plt_all, dpi=600, height=9, width=10, bg = "white")
```

Plot
```{r}
sig_group_brain_plt <- plot_cv_brain(group_sig_final, sig_group, "Group", fill_limits=levels(group_sig_final$sig_group), height_list=c(1.03, 4.5, .93), return_plt = FALSE, fill_values=group_values)

sig_group_brain_plt[[1]] <- sig_group_brain_plt[[1]] + group_fill_scale

sig_group_brain_plt[[2]] <- sig_group_brain_plt[[2]] + group_fill_scale

sig_group_brain_plt[[3]] <- sig_group_brain_plt[[3]] + group_fill_scale

#recreate legends
legend <- get_legend(
  ggplot(group_sig_final) +
    geom_bar(aes(x=cv_sample, fill=sig_group)) + group_fill_scale
  )

plts <- cowplot::plot_grid(
    sig_group_brain_plt[[1]],
    sig_group_brain_plt[[2]],
    sig_group_brain_plt[[3]],
    ncol = 1,
    nrow = 3,
    rel_heights = c(1.03, 4.5, .93),
    align = "v",
    axis = "lr",
    greedy = TRUE
  ) 
  
full_plt_obj <- plot_grid(plts, legend, ncol=2, rel_widths = c(3, 1))

#ggsave("figs/supplement/sigma_group_brain_plt.png", full_plt_obj, dpi=600, height=8.5, width=8, bg = "white")
```


plot trajectories by group

```{r}
group_sig_traj_full <- group_sig_traj_final %>%
  filter(sig_fdr==TRUE & sig_group != "other") %>%
  arrange(pheno, age_days) %>%
  ggplot(aes(x=age_days, y=logcv_M_div_F, color=sig_group)) +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  geom_line(aes(group=pheno), alpha=.1) +
  facet_grid(parent_group ~ cv_sample, scales="free_y") +
  geom_smooth(method = "gam") +
  group_color_scale +
  reformat_x(group_sig_traj_final$age_days) +
  theme_linedraw() +
  theme(legend.position="none") +
  labs(y="Scaled Median Sex Difference")
```

```{r}
#save combined group plots for supplement
sub_plt <- plot_grid(NULL, plts,
                      group_sig_traj_full,
                      nrow=1, rel_widths =c(.1,1,2),
                      labels=c("C", "", "D"))

group_sigma_plot_supplement <- plot_grid(sig_group_bar_plt_all,
                          sub_plt,
                          bottom_group_legend,
                          nrow=3,
                          ncol=1,
                          rel_heights = c(1, 1.1, .1))
ggsave("figs/supplement/groups_sigma_all_plt.png", group_sigma_plot_supplement, dpi=600, height=16, width=16, bg = "white")
```

also just plot B
```{r}
##BRAIN PLOT
group_sig_finalB <- group_sig_final %>% filter(cv_sample=="B")
group_brain_plt <- plot_cv_brain(group_sig_finalB, sig_group, "Group", fill_limits=levels(group_sig_finalB$sig_group), height_list=c(1.03, 4.5, .93), return_plt = FALSE, fill_values=group_values)

group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale

group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale

group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale

#recreate legends
group_brain_plt[[4]] <- get_legend(
  ggplot(group_sig_finalB) +
    geom_bar(aes(x=cv_sample, fill=sig_group)) + group_fill_scale
  )


saveRDS(group_brain_plt, "figs/ggobjs/sig_groups_brain_B.rds")

##TRAJECTORIES
group_sig_traj_plt_B <- group_sig_traj_final %>%
  filter(sig_fdr==TRUE & cv_sample=="B" & sig_group != "other") %>%
  arrange(pheno, age_days) %>%
  ggplot(aes(x=age_days, y=logcv_M_div_F, color=sig_group)) +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  geom_line(aes(group=pheno), alpha=.2) +
  facet_wrap(~parent_group, nrow=3, scales="free_y") +
  geom_smooth(method = "gam", linewidth=1) +
  group_color_scale +
  reformat_x(group_traj_final$age_days) +
  theme_linedraw() +
  theme(legend.position="none") +
  labs(y="Variance Ratio")
saveRDS(group_sig_traj_plt_B, "figs/ggobjs/sig_groups_traj_B.rds")
```


###Data-Driven Clustering

```{r}
clust_fun <- function(df, clust_var, k){
  medoids<-NULL
  #filter df
  df <- df %>%
    select(pheno, logAge_days, clust_var, pheno_cat, age_days)
  
  #pivot wide - rows = pheno levels, cols = time points
  ts_wide <- df %>%
    arrange(pheno, age_days) %>%
    select(pheno, age_days, clust_var) %>%
    pivot_wider(
      names_from  = age_days,
      values_from = clust_var
    ) %>%
    filter(!if_all(.cols = !pheno, ~is.na(.))) %>%
    select(where(~ !any(is.na(.))))
  
  # Store pheno labels, then convert to matrix
  pheno_labels <- ts_wide$pheno
  ts_matrix <- ts_wide %>%
    select(-pheno) %>%
    as.matrix()

  rownames(ts_matrix) <- pheno_labels
  
  #drop any missing vals
  ts_matrix_complete <- ts_matrix[, colSums(is.na(ts_matrix)) == 0]
  ts_matrix_complete <- ts_matrix_complete[rowSums(is.na(ts_matrix_complete)) == 0, ]
  
  #euclidean distance
  dist_matrix <- dist(ts_matrix_complete, method = "euclidean")
  
  #Cluster
  set.seed(42)
  
  # Elbow plot using average silhouette width (more natural for PAM)
    elbow_plt <- factoextra:::fviz_nbclust(ts_matrix_complete, cluster:::pam, method = "silhouette")
    print(elbow_plt)
  
    pm <- cluster:::pam(dist_matrix, k = k, diss = TRUE)
    clusters <- setNames(pm$clustering, rownames(ts_matrix_complete))
    
    medoids <- pm$medoids

  # Add cluster assignments back to original data
  cluster_df <- tibble(
    pheno   = names(clusters),
    cluster = clusters
  )
  
  data_with_clusters <- df %>%
    left_join(cluster_df, by = "pheno")

  #plot time series by cluster
  plt_traj <- data_with_clusters %>%
    mutate(cluster = factor(cluster)) %>%
    ggplot(aes(x = age_days, y = !!sym(clust_var),
               group = pheno, colour = cluster)) +
    geom_line(alpha = 0.3) +
    facet_wrap(~cluster) +
    labs(
      title = "Pheno Time Series by Cluster",
      colour = "Cluster"
    ) +
    reformat_x(data_with_clusters$age_days, simplified=TRUE) +
    theme_minimal()
  print(plt_traj)

#plot medoids
  med_plt <- data_with_clusters %>%
    filter(pheno %in% medoids) %>%
    mutate(cluster = factor(cluster)) %>%
    ggplot(aes(x = age_days, y = !!sym(clust_var),
                group = pheno, colour = cluster)) +
    geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
    geom_line(alpha = 0.7) +
    facet_wrap(~pheno, nrow=2) +
    labs(title = "Cluster Medoid"
    ) +
  reformat_x(data_with_clusters$age_days, simplified=TRUE) +
      theme_minimal()
    print(med_plt)

#plot list
plt_list <- c(elbow_plt, plt_traj, med_plt)
return(plt_list)
}
```

prep dfs
```{r}
LR.df_tot.u <- full_join(LR.df_tot, sexdiffs.df.u) %>%
  filter(sig_fdr==TRUE & total==TRUE & weighted==FALSE)

LR.df_tot.uA <- LR.df_tot.u %>%
  filter(cv_sample=="A")

LR.df_tot.uB <- LR.df_tot.u %>%
  filter(cv_sample=="B")
```

cluster and plot - wrapping for simplicity
```{r}
for (df in list(LR.df_tot.uA, LR.df_tot.uB)){
  cv_sample <- unique(df[["cv_sample"]])
  
  #Cluster
  cent_clust_plt <- clust_fun(df, "centile_M_minus_F_z", 2)
  
  #also save just skree plt
  rds_name <- paste0("figs/ggobjs/cent_elbow_", cv_sample, ".rds")
  saveRDS(cent_clust_plt[[1]], rds_name)

  #update plots
  clust_legend <- get_legend(cent_clust_plt[[2]])
  cent_clust_plt[[2]] <- cent_clust_plt[[2]] +
  labs(y="Scaled Median Sex Difference") +
  theme(legend.position="none")
  
  cent_clust_plt[[3]] <- cent_clust_plt[[3]] + 
    labs(y="Scaled Median Sex Difference") + 
    theme(legend.position = "none")

  #also save just medoid plt
  rds_name <- paste0("figs/ggobjs/cent_kmedoid_", cv_sample, ".rds")
  saveRDS(cent_clust_plt[[3]], rds_name)
  
  cent_clust_wrap <- plot_grid(
    cent_clust_plt[[1]],
    cent_clust_plt[[2]],
    cent_clust_plt[[3]],
    clust_legend,
    labels = c("A", "B", "C", "")
  )
  
  file_name <- paste0("figs/supplement/cent_kmedoid_", cv_sample, ".png")
  ggsave(file_name, cent_clust_wrap, dpi=600, height=9, width=16, bg = "white")
}
```

####Sigma

```{r}
for (df in list(LR.df_tot.uA, LR.df_tot.uB)){
  cv_sample <- unique(df[["cv_sample"]])
  
  #Cluster
  sigma_clust_plt <- clust_fun(df, "logcv_M_div_F", 2)
  
  #also save just skree plt
  rds_name <- paste0("figs/ggobjs/sigma_elbow_", cv_sample, ".rds")
  saveRDS(sigma_clust_plt[[1]], rds_name)

  #update plots
  clust_legend <- get_legend(sigma_clust_plt[[2]])
  sigma_clust_plt[[2]] <- sigma_clust_plt[[2]] +
  labs(y="Variance Ratio") +
  theme(legend.position="none")
  
  sigma_clust_plt[[3]] <- sigma_clust_plt[[3]] +
    labs(y="Variance Ratio") +
    theme(legend.position="none")
  #also save just medoid plt
  rds_name <- paste0("figs/ggobjs/sigma_kmedoid_", cv_sample, ".rds")
  saveRDS(sigma_clust_plt[[3]], rds_name)
  
  sigma_clust_wrap <- plot_grid(
    sigma_clust_plt[[1]],
    sigma_clust_plt[[2]],
    sigma_clust_plt[[3]],
    clust_legend,
    labels = c("A", "B", "C", "")
  )
  
  file_name <- paste0("figs/supplement/sigma_kmedoid_", cv_sample, ".png")
  ggsave(file_name, sigma_clust_wrap, dpi=600, height=9, width=16, bg = "white")
}
```
