---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue")
library(ggsegTissue)
remotes::install_github("ggseg/ggsegExtra")
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")


LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  select(!file)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
         pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
)

table(LR.df2$effect, LR.df2$total)
```

Prep atlases for plotting

```{r}
#clean up atlases to remove non-modeled phenos
#write out new dk atlas w/o medial wall
dk_clean <- dk
dk_clean$data <- dk_clean$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")

aseg_clean <- aseg
aseg_clean$data <- aseg_clean$data %>%
  dplyr::filter(!is.na(label))

aseg_all_clean <- aseg_all
aseg_all_clean$data <- aseg_all_clean$data %>%
  dplyr::filter(label != "Brain-Stem")
```

```{r}
# sub_list
# unique(aseg$data$label)

LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(tissue_class = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    pheno == "SUBC.Brain.Stem" ~ "Brainstem",
    TRUE ~ NA_character_
  )) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$tissue_class), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas

check scales for plotting

```{r}
LR.df_plot %>%
  group_by(effect) %>%
  summarise(max_resi = max(resi))
```


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE)

table(LR.df_uncor$sig_fdr)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```


```{r}
cort_plt_uncor <- LR.df_uncor %>%
  filter(pheno_cat == "Regional Vol" | pheno_cat == "Regional SA" | pheno_cat == "Regional CT") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_clean, 
             position = position_brain(hemi ~ side),
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Greens 3", name = "RESI", limits = c(0, 0.18), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample) 

sub_plt_uncor <- LR.df_uncor %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_clean, 
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Greens 3", name = "RESI", limits = c(0, 0.18), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample)

LR.df_plot_glob_uncor <- inner_join(LR.df_uncor, aseg_all_clean$data, by="tissue_class")
glob_plt_uncor <- LR.df_plot_glob_uncor %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_all_clean, 
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Greens 3", name = "RESI", limits = c(0, 0.18), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample)
```

```{r}
full_plt_uncor <- glob_plt_uncor /plot_spacer()/ cort_plt_uncor / plot_spacer()/ sub_plt_uncor +
  plot_layout(guides = 'collect', heights = unit(c(3, -1.2, 6, -1.2, 3), c('in', 'in', 'in')))

ggsave("figs/resi_uncorrected.png", full_plt_uncor, width=7, height=10)
```

#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE)

table(LR.df_tot$sig_fdr)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
cort_plt_tot <- LR.df_tot %>%
  filter(pheno_cat == "Regional Vol" | pheno_cat == "Regional SA" | pheno_cat == "Regional CT") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_clean, 
             position = position_brain(hemi ~ side),
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Greens 3", name = "RESI", limits = c(0, 0.18), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample) 

sub_plt_tot <- LR.df_tot %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_clean, 
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Greens 3", name = "RESI", limits = c(0, 0.18), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample)

LR.df_plot_glob_tot <- inner_join(LR.df_tot, aseg_all_clean$data, by="tissue_class")
glob_plt_tot <- LR.df_plot_glob_tot %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_all_clean, 
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Greens 3", name = "RESI", limits = c(0, 0.18), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample)
```

```{r}
full_plt_tot <- glob_plt_tot /plot_spacer()/ cort_plt_tot / plot_spacer()/ sub_plt_tot +
  plot_layout(guides = 'collect', heights = unit(c(3, -1.2, 6, -1.2, 3), c('in', 'in', 'in')))
full_plt_tot
ggsave("figs/resi_tbv.png", full_plt_tot, width=7, height=10)
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE)

table(LR.df_tot_sex$sig_fdr)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)
```

```{r}
cort_plt_tot_sex <- LR.df_tot_sex %>%
  filter(pheno_cat == "Regional Vol" | pheno_cat == "Regional SA" | pheno_cat == "Regional CT") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_clean, 
             position = position_brain(hemi ~ side),
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Purple-Blue", name = "RESI", limits = c(0, 0.3), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample) 

sub_plt_tot_sex <- LR.df_tot_sex %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_clean, 
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Purple-Blue", name = "RESI", limits = c(0, 0.3), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample)

LR.df_plot_glob_tot_sex <- inner_join(LR.df_tot_sex, aseg_all_clean$data, by="tissue_class")
glob_plt_tot_sex <- LR.df_plot_glob_tot_sex %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = aseg_all_clean, 
            aes(fill=sig_resi)) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c("grDevices::Purple-Blue", name = "RESI", limits = c(0, 0.3), direction = -1) +
  facet_grid(pheno_cat ~ cv_sample)
```

```{r}
full_plt_tot_sex <- glob_plt_tot_sex /plot_spacer()/ cort_plt_tot_sex / plot_spacer()/ sub_plt_tot_sex +
  plot_layout(guides = 'collect', heights = unit(c(3, -1.2, 6, -1.2, 3), c('in', 'in', 'in')))
full_plt_tot_sex
ggsave("figs/resi_tbv_allsex.png", full_plt_tot_sex, width=7, height=10)
```

##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

sexdiffs.df <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         pheno = stringr::str_remove(basename(file), "_sexdiffs\\.csv")) %>%
  select(!file)

rm(sex_diffs_paths) #save space
```
scale diffs within each pheno cat
```{r}
#visualizing wm separately for trajectories
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

sexdiffs_df_sc <- sexdiffs.df %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  group_by(weighted, total, pheno_cat) %>%
  mutate(centile_M_minus_F_sc = scale(centile_M_minus_F, center=FALSE),
         sigma_M_minus_F_sc = scale(sigma_M_minus_F, center=FALSE),
         deriv_M_minus_F_sc = scale(deriv_M_minus_F, center=FALSE)) %>%
  ungroup()
```

Plot

```{r}
sexdiffs_tbv <- sexdiffs_df_sc %>%
  filter(weighted==FALSE & total==TRUE)

sexdiffs <- sexdiffs_df_sc %>%
  filter(weighted==FALSE & total==FALSE)

for (df in list(sexdiffs_tbv, sexdiffs)){
  t <- as.character(unique(df$total))
  print(t)
  
  #median centile
  cent_df <- df %>%
    select(centile_M_minus_F_sc, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  cent_plt <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_sc, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile")
  
  n <- paste0("figs/sexdiffs_tbv",t,"_cent.png")
  ggsave(n, cent_plt)
  #ggplotly(cent_plt)

  #sigma
  sig_df <- df %>%
    select(sigma_M_minus_F_sc, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  sig_plt <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sigma_M_minus_F_sc, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="Variance (Sigma)")
  
  n <- paste0("figs/sexdiffs_tbv",t,"_sig.png")
  ggsave(n, sig_plt)
  #ggplotly(sig_plt)

  #derivatives
  deriv_df <- df %>%
    select(deriv_M_minus_F_sc, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  deriv_plt <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_sc, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="50th Centile Derivative")
  
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv.png")
  ggsave(n, deriv_plt)
  #ggplotly(deriv_plt)
}
```

###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*
plot absolute value centile sex-diffs -> magnitude sex-diffs

```{r}
#regions w sig sex effects in both samples
LR.tot_sex_wide <- LR.df_tot_sex %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label, tissue_class), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  mutate(rep = case_when(
    sig_fdr_A == TRUE & sig_fdr_B == TRUE ~ "consistent_sig",
    sig_fdr_A == FALSE & sig_fdr_B == FALSE ~ "consistent_ns",
    TRUE ~ "mixed"
  ))

table(LR.tot_sex_wide$rep)

LR.tot_sex_wide %>%
  filter(rep == "consistent_sig") %>%
  ggplot() +
  geom_histogram(aes(x=resi_A))
LR.tot_sex_wide %>%
  filter(rep == "consistent_sig") %>%
  ggplot() +
  geom_histogram(aes(x=resi_B))

consistent_sex <- LR.tot_sex_wide %>%
  filter(rep == "consistent_sig") %>%
  pull(pheno)

#consistent age*sex effects
LR.tot_wide <- LR.df_tot %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label, tissue_class), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  mutate(rep = case_when(
    sig_fdr_A == TRUE & sig_fdr_B == TRUE ~ "consistent_sig",
    sig_fdr_A == FALSE & sig_fdr_B == FALSE ~ "consistent_ns",
    TRUE ~ "mixed"
  ))

consistent_agesex <- LR.tot_wide %>%
  filter(rep == "consistent_sig") %>%
  pull(pheno)
```

```{r}
cent_df <- sexdiffs_df_sc %>%
  filter(weighted==FALSE & total==TRUE) %>%
  select(centile_M_minus_F_sc, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F) %>%
  na.omit() %>%
  arrange(pheno, logAge_days)

cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=abs(centile_M_minus_F_sc), group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(pheno %in% consistent_sex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Magnitude Sex Diff in 50th Centile")
```

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
tol <- 1e-5 #setting arbitrarily, need to update when zscored

consistent_agesex <- LR.tot_wide %>%
  filter(rep == "consistent_sig") %>%
  pull(pheno)

#quick dirty code to get abs diff
cent_df2 <- cent_df %>%
  mutate(sq_centile_diff = (centile_M_minus_F)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    sq_deriv = c(
      NA,
      diff(sq_centile_diff) / diff(logAge_days)
    )
  ) %>%
  ungroup()

cent_df2 %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff")
  

#try getting integral
diff_int <- cent_df2 %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int = trapz(logAge_days, sq_deriv)
            )
  
diff_int_wide <- diff_int %>%
  pivot_wider(id_cols=c(pheno, pheno_cat), names_from=cv_sample, values_from=sq_int) %>%
  mutate(trajectory = case_when(
    A > tol & B > tol ~ "diverging",
    A < -tol & B < -tol ~ "converging",
    abs(A) <= tol & abs(B) <= tol ~ "constant",
    is.na(A) | is.na(B) ~ NA,
    TRUE ~ "inconsistent"
  ))

diff_int_wide %>%
  filter(pheno %in% consistent_agesex) %>%
  ggplot() +
  geom_bar(aes(x=pheno_cat, fill=trajectory))
```


try finding points at which derivative crosses 0 increasing vs decreasing

```{r}
cent_df3 <- cent_df2 %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv > tol ~ "diverging",
    abs(sq_deriv) <= tol ~ "constant",
    sq_deriv < -tol ~ "converging",
  ))

#order by age first increasing
pheno_order_mean <- cent_df3 %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv, x = logAge_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

#highlighting phenos with significant age*sex effects in both samples
cent_df3 %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=logAge_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("log_lifespan_fetal", cent_df3$logAge_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0)) +
  gghighlight(pheno %in% consistent_agesex, use_direct_label = FALSE, unhighlighted_params = list(colour = NULL, alpha = 0.3), calculate_per_facet = TRUE)


cent_df3 %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=logAge_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("log_lifespan_fetal", cent_df3$logAge_days)

cent_df3 %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=logAge_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("log_lifespan_fetal", cent_df3$logAge_days)
```

recolor centile diff trajectories by whether males and females are converging or diverging overall
using examples from https://albert-rapp.de/posts/ggplot2-tips/07_four_ways_colors_more_efficiently/07_four_ways_colors_more_efficiently.html

```{r}
consistent_trajectory <- diff_int_wide %>% 
  filter(trajectory != "inconsistent") %>%
  pull(pheno)

consistent_trajectory_sig_sex <- intersect(consistent_trajectory, consistent_sex)

cent_df_trajectory_sum <- full_join(cent_df, diff_int)

cent_df_trajectory_sum %>%
   mutate(trajectory = case_when(
     abs(sq_int) <= tol ~ "constant",
    sq_int > tol ~ "diverging",
    sq_int < -tol  ~ "converging"
  )) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_sc, group=pheno, color=trajectory)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Difference") +
  gghighlight(pheno %in% consistent_trajectory_sig_sex, use_direct_label = FALSE, unhighlighted_params = list(colour = NULL, alpha = 0.3), calculate_per_facet = TRUE)
```

Get max male and female biases and plot on centile differences

```{r}
peak_diffs_sum <- cent_df %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(centile_M_minus_F, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(centile_M_minus_F)),
            female_peak = min_gc(centile_M_minus_F, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(centile_M_minus_F))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")


peak_diffs_sum %>%
  ggplot() +
  geom_density(aes(x=peak_age, fill=sex), position="identity", alpha=.4) +
  format_x_axis("log_lifespan_fetal", peak_diffs_sum$peak_age)

cent_df %>%
    filter(pheno %in% consistent_agesex) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% consistent_agesex), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
```

differences tend to peak around beginning and end of life - likely d/t smaller sample sizes at extreme ages

*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*
```{r}
sigma_df <- sexdiffs_df_sc %>%
  filter(weighted==FALSE & total==TRUE) %>%
  select(sigma_M_minus_F_sc, pheno, logAge_days, pheno_cat, cv_sample, sigma_M_minus_F) %>%
  na.omit() %>%
  arrange(pheno, logAge_days)

sigma_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=abs(sigma_M_minus_F_sc), group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Magnitude Sex Diff in Sigma")
```

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#quick dirty code to get abs diff
sigma_df2 <- sigma_df %>%
  mutate(sq_sigma_diff = (sigma_M_minus_F)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    sq_sigma_deriv = c(
      NA,
      diff(sq_sigma_diff) / diff(logAge_days)
    )
  ) %>%
  ungroup()

sigma_df2 %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_sigma_deriv, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sigma_df2$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Variability")
  

#try getting integral
diff_sigma_int <- sigma_df2 %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_sigma_int = trapz(logAge_days, sq_sigma_deriv)
            )
  
diff_sigma_int_wide <- diff_sigma_int %>%
  pivot_wider(id_cols=c(pheno, pheno_cat), names_from=cv_sample, values_from=sq_sigma_int) %>%
  mutate(trajectory = case_when(
    A > tol & B > tol ~ "diverging",
    A < -tol & B < -tol ~ "converging",
    abs(A) <= tol & abs(B) <= tol ~ "constant",
    is.na(A) | is.na(B) ~ NA,
    TRUE ~ "inconsistent"
  ))

diff_sigma_int_wide %>%
  filter(pheno %in% consistent_agesex) %>%
  ggplot() +
  geom_bar(aes(x=pheno_cat, fill=trajectory))
```

```{r}
consistent_sigma_trajectory <- diff_sigma_int_wide %>% 
  filter(trajectory != "inconsistent") %>%
  pull(pheno)

consistent_sigma_trajectory_sig_sex <- intersect(consistent_sigma_trajectory, consistent_sex)

sigma_df_trajectory_sum <- full_join(sigma_df, diff_sigma_int)

#displaying phenos with significant sex effects whose sigma trajectories are overall similar
sigma_df_trajectory_sum %>%
   mutate(trajectory = case_when(
     abs(sq_sigma_int) <= tol ~ "constant",
    sq_sigma_int > tol ~ "diverging",
    sq_sigma_int < -tol  ~ "converging"
  )) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sigma_M_minus_F_sc, group=pheno, color=trajectory)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", sigma_df$logAge_days) +
    labs(y="Sigma Difference") +
  gghighlight(pheno %in% consistent_sigma_trajectory_sig_sex, use_direct_label = FALSE, unhighlighted_params = list(colour = NULL, alpha = 0.3), calculate_per_facet = TRUE)
```


###Cross-sample Correlations
Get across-sample correlations
```{r}
cor_df <- data.frame()
for (pheno_val in pheno_list){
  df <- sexdiffs_tbv %>%
    filter(pheno==pheno_val) %>%
    arrange(logAge_days)
  
  dfA <- df %>%
    filter(cv_sample=="A")
  
  dfB <- df %>%
    filter(cv_sample=="B")
  
  if (nrow(dfA) == 0 | nrow(dfB) == 0) next
  
  cent_cor <- cor(dfA$centile_M_minus_F, dfB$centile_M_minus_F, use="complete.obs")
  sigma_cor <- cor(dfA$sigma_M_minus_F, dfB$sigma_M_minus_F, use="complete.obs")
  deriv_cor <- cor(dfA$deriv_M_minus_F, dfB$deriv_M_minus_F, use="complete.obs")
  
  result <- data.frame(
    "pheno" = pheno_val,
    "centile_cor" = cent_cor,
    "sigma_cor" = sigma_cor,
    "deriv_cor" = deriv_cor
  )
  cor_df <- bind_rows(cor_df, result)
}

ggplot(cor_df) +
  geom_histogram(aes(x=centile_cor)) +
  geom_vline(xintercept = .9, color="red")

ggplot(cor_df) +
  geom_histogram(aes(x=sigma_cor)) +
  geom_vline(xintercept = .9, color="red")

ggplot(cor_df) +
  geom_histogram(aes(x=deriv_cor)) +
  geom_vline(xintercept = .9, color="red")

cor_df %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  pivot_longer(cols=c(ends_with("cor")), names_to="type", values_to="cor") %>%
  group_by(type, pheno_cat) %>%
  summarise(mean_cor = mean(cor),
            n_pass = sum(cor>.9),
            pct_pass = (sum(cor>.9)/n())*100,
            n_neg = sum(cor<0),
            pct_neg = (sum(cor<0)/n())*100
            )
```

filtering to 1+ or 2+ yrs old doesn't change much

```{r, eval=FALSE, message=FALSE}
for (type in c("centile", "sigma", "deriv")){
  cor_col <- paste0(type, "_cor")
  val_col <- paste0(type, "_M_minus_F")

  for (pheno_val in pheno_list){
    #pull correlations
    cor_val <- cor_df %>% 
      filter(pheno==pheno_val) %>%
      pull(!!sym(cor_col))
    
    plt <- sexdiffs_tbv %>%
      filter(pheno==pheno_val) %>%
      select(!!sym(val_col), pheno, logAge_days, cv_sample) %>%
      na.omit() %>%
      arrange(pheno, logAge_days) %>%
      ggplot() +
      geom_line(aes(x=logAge_days, y=!!sym(val_col), group=cv_sample, color=cv_sample)) +
      theme(legend.position="none") +
      ggtitle(paste(pheno_val, "r =", cor_val))
    
    print(plt)
    
    ggsave(paste0("./figs/", pheno_val, "_", type, "_sexdiff_plt.png"), plt)
  }
}
```

Look at just regions with significant sex effects overall

```{r}
LR.tot_sex_wide <- LR.df_tot_sex %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label, tissue_class), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  mutate(rep = case_when(
    sig_fdr_A == TRUE & sig_fdr_B == TRUE ~ "consistent_sig",
    sig_fdr_A == FALSE & sig_fdr_B == FALSE ~ "consistent_ns",
    TRUE ~ "mixed"
  ))

table(LR.tot_sex_wide$rep)

LR.tot_sex_wide %>%
  filter(rep == "consistent_sig") %>%
  ggplot() +
  geom_histogram(aes(x=resi_A))
LR.tot_sex_wide %>%
  filter(rep == "consistent_sig") %>%
  ggplot() +
  geom_histogram(aes(x=resi_B))
```

```{r}
cor_df %>%
  filter(pheno %in% consistent_sex) %>%
  pivot_longer(cols=c(ends_with("cor")), names_to="type", values_to="cor") %>%
  group_by(type) %>%
  summarise(mean_cor = mean(cor),
            n_pass = sum(cor>.9),
            pct_pass = (sum(cor>.9)/n())*100,
            n_neg = sum(cor<0),
            pct_neg = (sum(cor<0)/n())*100
            )
```


Check same with age*sex effects

```{r}
LR.tot_wide <- LR.df_tot %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label, tissue_class), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  mutate(rep = case_when(
    sig_fdr_A == TRUE & sig_fdr_B == TRUE ~ "consistent_sig",
    sig_fdr_A == FALSE & sig_fdr_B == FALSE ~ "consistent_ns",
    TRUE ~ "mixed"
  ))

table(LR.tot_wide$rep)

LR.tot_wide %>%
  filter(rep == "consistent_sig") %>%
  ggplot() +
  geom_histogram(aes(x=resi_A))
LR.tot_wide %>%
  filter(rep == "consistent_sig") %>%
  ggplot() +
  geom_histogram(aes(x=resi_B))
```

```{r}
cor_df %>%
  filter(pheno %in% consistent_agesex) %>%
  pivot_longer(cols=c(ends_with("cor")), names_to="type", values_to="cor") %>%
  group_by(type) %>%
  summarise(mean_cor = mean(cor),
            n_pass = sum(cor>.9),
            pct_pass = (sum(cor>.9)/n())*100,
            n_neg = sum(cor<0),
            pct_neg = (sum(cor<0)/n())*100
            )
```
Higher but not a big difference tbh

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)

```{r}
dominance_df <- sexdiffs_df_sc %>%
  filter(total==TRUE & weighted==FALSE) %>%
  select(!contains("deriv")) %>%
  na.omit() %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int = trapz(logAge_days, centile_M_minus_F),
            sigma_int = trapz(logAge_days, sigma_M_minus_F)) %>%
  mutate(dominance = case_when(
    abs(centile_int) <= tol ~ "Balanced",
    centile_int < -tol ~ "Female",
    centile_int > tol ~ "Male"
  ),
  sigma_dominance = case_when(
    abs(sigma_int) <= tol ~ "Balanced",
    sigma_int < -tol ~ "Female",
    sigma_int > tol ~ "Male"
  ))

dominance_df %>%
  filter(pheno %in% consistent_sex) %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()

dominance_df %>%
  filter(pheno %in% consistent_sex) %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()
```

plot dominance on brain

```{r}
dominance_df2 <- dominance_df %>% select(!pheno_cat)
LR.df_tot_dominance <- left_join(LR.df_tot, dominance_df2)

reg_dom_plt <- LR.df_tot_dominance %>%
  filter(pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")) %>% 
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    aes(fill = dominance),
    atlas = dk_clean,
    alpha=0.3,
    position = position_brain(hemi ~ side)
  ) +
  geom_brain(
    data = LR.df_tot_dominance %>% 
      filter(pheno %in% consistent_sex),
    aes(fill = dominance),
    atlas = dk_clean,
    position = position_brain(hemi ~ side)
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  facet_grid(rows = vars(pheno_cat), cols = vars(cv_sample), drop = TRUE) 


sub_dom_plt <- LR.df_tot_dominance %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    aes(fill = dominance),
    atlas = aseg_clean,
    alpha=0.3
  ) +
  geom_brain(
    data = LR.df_tot_dominance %>% 
      filter(pheno %in% consistent_sex & pheno_cat == "Subcortical Vol"),
    aes(fill = dominance),
    atlas = aseg_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  facet_grid(pheno_cat ~ cv_sample)

LR.df_tot_dominance_glob <- inner_join(LR.df_tot_dominance, aseg_all_clean$data, 
                                       by="tissue_class", relationship = "many-to-many")
glob_dom_plt <- LR.df_tot_dominance_glob %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    aes(fill = dominance),
    atlas = aseg_all_clean,
    alpha=0.3
  ) +
  geom_brain(
    data = LR.df_tot_dominance_glob %>% 
      filter(pheno %in% consistent_sex & pheno_cat == "Global Vol"),
    aes(fill = dominance),
    atlas = aseg_all_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  facet_grid(pheno_cat ~ cv_sample)
```

```{r}
full_dom_plt <- glob_dom_plt /plot_spacer()/ reg_dom_plt / plot_spacer()/ sub_dom_plt +
  plot_layout(guides = 'collect', heights = unit(c(3.5, -1.6, 6, -1.6, 3.5), c('in', 'in', 'in', 'in', 'in')))
full_dom_plt
ggsave("figs/dominance_tbvTRUE.png", full_dom_plt, width=7, height=8.5)
```

plot sigma dominance brain

```{r}
reg_dom_plt_sigma <- LR.df_tot_dominance %>%
  filter(pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")) %>% 
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    aes(fill = sigma_dominance),
    atlas = dk_clean,
    alpha=0.3,
    position = position_brain(hemi ~ side)
  ) +
  geom_brain(
    data = LR.df_tot_dominance %>% 
      filter(pheno %in% consistent_sex & pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")),
    aes(fill = sigma_dominance),
    atlas = dk_clean,
    position = position_brain(hemi ~ side)
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  facet_grid(rows = vars(pheno_cat), cols = vars(cv_sample), drop = TRUE)  +
  scale_fill_manual(values = c("Balanced" = "#F8B150FF", 
                                "Female" = "#00BA38", 
                                "Male" = "#619CFF"))


sub_dom_plt_sigma <- LR.df_tot_dominance %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    aes(fill = sigma_dominance),
    atlas = aseg_clean,
    alpha=0.3
  ) +
  geom_brain(
    data = LR.df_tot_dominance %>% 
      filter(pheno %in% consistent_sex & pheno_cat == "Subcortical Vol"),
    aes(fill = sigma_dominance),
    atlas = aseg_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "none") +
  facet_grid(pheno_cat ~ cv_sample) +
  scale_fill_manual(values = c("Balanced" = "#F8B150FF", 
                                "Female" = "#00BA38", 
                                "Male" = "#619CFF"))

glob_dom_plt_sigma <- LR.df_tot_dominance_glob %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    aes(fill = sigma_dominance),
    atlas = aseg_all_clean,
    alpha=0.3
  ) +
  geom_brain(
    data = LR.df_tot_dominance_glob %>% 
      filter(pheno %in% consistent_sex),
    aes(fill = sigma_dominance),
    atlas = aseg_all_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "none") +
  facet_grid(pheno_cat ~ cv_sample) +
  scale_fill_manual(values = c("Balanced" = "#F8B150FF", 
                                "Female" = "#00BA38", 
                                "Male" = "#619CFF"))
```

```{r}
full_dom_plt_sigma <- glob_dom_plt_sigma /plot_spacer()/ reg_dom_plt_sigma / plot_spacer()/ sub_dom_plt_sigma +
  plot_layout(guides = 'collect', heights = unit(c(3.5, -2, 6, -2, 3.5), c('in', 'in', 'in', 'in', 'in')))
full_dom_plt_sigma
ggsave("figs/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, width=7, height=8)
```


```{r}
dominance_df2 <- cent_df2 %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(dominance = case_when(
    centile_M_minus_F > tol ~ "Male",
    abs(centile_M_minus_F) <= tol ~ "Balanced",
    centile_M_minus_F < -tol ~ "Female",
  ))

#order by age first increasing
pheno_order_mean <- dominance_df2 %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = centile_M_minus_F, x = logAge_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

#highlighting phenos with significant sex effects in both samples
dominance_df2 %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=logAge_days, y=pheno, color=dominance)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("log_lifespan_fetal", dominance_df2$logAge_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0)) +
  gghighlight(pheno %in% consistent_sex, use_direct_label = FALSE, unhighlighted_params = list(colour = NULL, alpha = 0.3), calculate_per_facet = TRUE)
```

###Shape Differences
*At what points in the lifespan are males and females growing differently?*
integral of f''(M) - f''(F)

```{r}
#quick dirty code to get second deriv
second_deriv_df <- sexdiffs_df_sc %>%
  filter(total==TRUE & weighted==FALSE) %>%
  select(!contains("centile") & !contains("sigma")) %>%
  na.omit() %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    deriv2_Male = c(
      NA,
      diff(deriv_Male) / diff(logAge_days)
    ),
    deriv2_Female = c(
      NA,
      diff(deriv_Female) / diff(logAge_days)
    )) %>%
  mutate(deriv2_M_minus_F = deriv2_Male - deriv2_Female)

second_deriv_df_sum <- second_deriv_df %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  na.omit() %>%
  summarise(int2 = trapz(logAge_days, deriv2_M_minus_F)
            )


#order by integral
pheno_order_mean <- second_deriv_df_sum %>%
  group_by(pheno) %>%
  summarise(ord = mean(int2), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

second_deriv_df_sum %>%
  filter(pheno %in% consistent_agesex) %>%
  ggplot() +
  geom_col(aes(x=fct_reorder(pheno, int2), y=int2, fill=cv_sample), position="dodge") +
  theme(axis.text.x = element_blank()) +
  facet_wrap(~pheno_cat, scales="free")
```

int2 > 0 -> males grow faster than females
int2 < 0 -> females grow faster
int2 = 0 -> same rate

```{r}
second_deriv_df_wide <- second_deriv_df_sum %>%
  pivot_wider(id_cols=c(pheno, pheno_cat), names_from=cv_sample, values_from=int2) %>%
  mutate(dominance = case_when(
    A > tol & B > tol ~ "Male",
    A < -tol & B < -tol ~ "Female",
    abs(A) <= tol & abs(B) <= tol ~ "Balanced",
    is.na(A) | is.na(B) ~ NA,
    TRUE ~ "inconsistent"
  ))

second_deriv_df_wide %>%
  filter(pheno %in% consistent_agesex) %>%
  ggplot() +
  geom_bar(aes(x=pheno_cat, fill=dominance))
```

```{r}
consistent_deriv_trajectory <- second_deriv_df_wide %>% 
  filter(dominance != "inconsistent") %>%
  pull(pheno)

consistent_deriv_trajectory_sig_sex <- intersect(consistent_deriv_trajectory, consistent_agesex)

deriv_df <- sexdiffs_df_sc %>%
  filter(total==TRUE & weighted==FALSE) %>%
  select(deriv_M_minus_F_sc, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F) %>%
  na.omit() %>%
  arrange(pheno, logAge_days)

deriv_df2 <- full_join(deriv_df, second_deriv_df_sum) %>%
  mutate(dominance = case_when(
    int2 > tol ~ "Male",
    int2 < -tol ~ "Female",
    abs(int2) <= tol ~ "Balanced",
    TRUE ~ NA
  ))


#displaying phenos with significant age*sex effects whose deriv trajectories are overall similar
deriv_df2 %>%
  #mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=logAge_days, y=deriv_M_minus_F, group=pheno, color=dominance)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("log_lifespan_fetal", deriv_df2$logAge_days) +
  gghighlight(pheno %in% consistent_deriv_trajectory_sig_sex, use_direct_label = FALSE, unhighlighted_params = list(colour = NULL, alpha = 0.3), calculate_per_facet = TRUE)
```