---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, ggcorrplot, sf)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

#ggseg Prep

```{r}
### dk for cortical ###
#write out new dk atlas w/o medial wall
dk_clean <- dk
dk_clean$data <- dk_clean$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")

#shrink for outlines
dk_region_gap <- 0.3        # Space between regions
dk_border_inset <- 0.3      # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
dk_fill <- dk_clean
dk_fill$data <- dk_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -dk_region_gap) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
dk_border <- dk_clean
dk_border$data <- dk_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(dk_region_gap + dk_border_inset)) %>%
  as_tibble() %>%
  st_as_sf()

### aseg for subcortical ###
#shrink for outlines
aseg_region_gap <- 0        # Space between regions
aseg_border_inset <- 1e-5     # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
aseg_fill <- aseg
aseg_fill$data <- aseg$data %>%
  st_as_sf() %>%
  st_buffer(dist = 0, 
            joinStyle = "MITRE",
            nQuadSegs=500) %>%
  st_buffer(0) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
aseg_border <- aseg
aseg_border$data <- aseg$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(aseg_region_gap + aseg_border_inset), nQuadSegs=500) %>%
  as_tibble() %>%
  st_as_sf()

### aseg for global ###
aseg_all_clean <- aseg_all
aseg_all_clean$data <- aseg_all_clean$data %>%
  dplyr::filter(label != "Brain-Stem")

#shrink for outlines
glob_region_gap <- 1e-10        # Space between regions
glob_border_inset <- 1e-10      # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
glob_fill <- aseg_all_clean
glob_fill$data <- aseg_all_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -glob_region_gap) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
glob_border <- aseg_all_clean
glob_border$data <- aseg_all_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(glob_region_gap + glob_border_inset)) %>%
  as_tibble() %>%
  st_as_sf()
```

trying to write funs for easier plotting
```{r}
plot_cv_brain <- function(df, fill_var, fill_name, fill_color, fill_limits, height_list){

#cortex
cortex_plt <- df %>%
  filter(pheno_cat == "Regional Vol" | pheno_cat == "Regional SA" | pheno_cat == "Regional CT") %>%
  group_by(cv_sample, pheno_cat) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = dk_fill,
             position = position_brain(hemi ~ side),
            aes(fill={{fill_var}}),#
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = dk_border,
                    position = position_brain(hemi ~ side),
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)

subcort_plt <- df %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  group_by(cv_sample, pheno_cat) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = aseg_fill,
            aes(fill={{fill_var}}),
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = aseg_border,
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)


#no buffer for now
glob_plt <- df %>%
  select(-any_of(c("geometry", "tissue_class"))) %>%
  group_by(cv_sample, pheno_cat) %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = aseg_all_clean,
            aes(fill={{fill_var}}), #
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = aseg_all_clean,
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)

#testing new spacing
# glob_plt <- glob_plt +
#   theme(plot.margin = margin(0, 0, 5, 0))

cortex_plt <- cortex_plt +
  theme(plot.margin = margin(5, 0, 5, 0))

full_plt_obj <- glob_plt / cortex_plt / subcort_plt +
  plot_layout(guides = "collect", heights = height_list)

# full_plt_obj <- glob_plt /plot_spacer()/ cortex_plt / plot_spacer()/ subcort_plt +
#   plot_layout(guides = 'collect', heights = unit(height_list, c('in', 'in', 'in')))

return(full_plt_obj)
}
```

#Defining integral funs

```{r}
#fun for derivative
get_deriv <- function(x, age) {
  c(NA, diff(x) / diff(age))
}

#riemann_stieltjes
rs_integral <- function(x, y, g = function(x) log10(x)) {
  stopifnot(
    length(x) == length(y),
    length(x) >= 2
  )

  # sort just in case
  ord <- order(x)
  x <- x[ord]
  y <- y[ord]

  gx <- g(x)

  # monotonicity check (important!)
  if (any(diff(gx) < 0)) {
    stop("g(x) must be non-decreasing for a Riemann–Stieltjes integral")
  }

  sum(y[-length(y)] * diff(gx))
}

#euclidean distance fun from https://www.geeksforgeeks.org/r-language/how-to-calculate-euclidean-distance-in-r/
euclidean_dist <- function(vect1, vect2) sqrt(sum((vect1 - vect2)^2))
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")


LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  select(!file)
```

```{r, eval=FALSE}
LR.df %>%
  group_by(cv_sample, total, weighted) %>%
  summarise(n_phenos = n_unique(pheno))

A_totalFALSE <- LR.df %>%
  filter(cv_sample == "A" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalFALSE)

A_totalTRUE <- LR.df %>%
  filter(cv_sample == "A" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalTRUE)

B_totalFALSE <- LR.df %>%
  filter(cv_sample == "B" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalFALSE)

B_totalTRUE <- LR.df %>%
  filter(cv_sample == "B" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalTRUE)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
         pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
)

table(LR.df2$effect, LR.df2$total)
```



```{r}
# sub_list
# unique(aseg$data$label)

LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$tissue_class), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas

check scales for plotting

```{r}
LR.df_plot %>%
  group_by(effect) %>%
  summarise(max_resi = max(resi))
```


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE)

LR.df_uncor %>%
  group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n()) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_uncor_novent <- LR.df_uncor %>%
  filter(!(pheno %in% vent_list))
range(LR.df_uncor_novent$resi)
full_plt_uncor <- plot_cv_brain(LR.df_uncor_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.1), c(1, 5, 1.1))

ggsave("figs/resi_uncorrected.png", full_plt_uncor, width=7, height=8.5, dpi=600)
```

#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE)

table(LR.df_tot$sig_fdr)

LR.df_tot %>%
group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n()) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_tot_novent <- LR.df_tot %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_novent$resi)
full_plt_tot <- plot_cv_brain(LR.df_tot_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.1), c(1, 5, 1.1))

ggsave("figs/resi_tbv.png", full_plt_tot, width=7, height=8.5, dpi=600)
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE)

table(LR.df_tot_sex$sig_fdr, LR.df_tot_sex$cv_sample)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)

LR.df_tot_sex %>%
  filter(sig_fdr == TRUE) %>%
  group_by(cv_sample) %>%
  summarise(min_resi = min(sig_resi),
            max_resi = max(sig_resi),
            mean_resi = mean(sig_resi))
```

```{r}
#removing ventricles for visualization:
LR.df_tot_sex_novent <- LR.df_tot_sex %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_sex_novent$resi)
full_plt_tot_sex <- plot_cv_brain(LR.df_tot_sex_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.2), c(1, 5, 1.1))

ggsave("figs/resi_tbv_allsex.png", full_plt_tot_sex, width=7, height=8.5, dpi=600)
```


##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

sexdiffs.df_all <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         pheno = stringr::str_remove(basename(file), "_(uniform_)?sexdiffs\\.csv")) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    age_days = un_log(logAge_days))

#split into regular df and uniform df for correlations
sexdiffs.df_uniform <- sexdiffs.df_all[grepl("uniform", sexdiffs.df_all$file), ]
sexdiffs.df <- sexdiffs.df_all[!grepl("uniform", sexdiffs.df_all$file), ]

rm(output_fread_sd) #save space
rm(sexdiffs.df_all)
```

Fix scaling, derivatives

```{r}
sexdiffs.df_sc <- sexdiffs.df %>%
  group_by(total, weighted, cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    #tmp fix for sigma as CV, getting ln CV ratio
    logcv_M_div_F = log(sigma_Male/sigma_Female),
    #chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age)
```

Plot

```{r}
sexdiffs_tbv <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==TRUE)

sexdiffs <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==FALSE)

for (df in list(sexdiffs_tbv, sexdiffs)){
  t <- as.character(unique(df$total))
  print(t)
  
  #median centile
  cent_df <- df %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  cent_plt <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
  n <- paste0("figs/sexdiffs_tbv",t,"_cent.png")
  ggsave(n, cent_plt)
  print(cent_plt)
  
  cent_plt2 <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Scaled 50th Centile Sex-Difference")
  
  print(cent_plt2)
  n <- paste0("figs/sexdiffs_tbv",t,"_cent_scaled.png")
  ggsave(n, cent_plt2)
  #ggplotly(cent_plt)

  #sigma
  sig_df <- df %>%
    select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  sig_plt <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="log CV Ratio")
  print(sig_plt)
  n <- paste0("figs/sexdiffs_tbv",t,"_cv.png")
  ggsave(n, sig_plt)
  
  #derivatives
  deriv_df <- df %>%
    select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  deriv_plt <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="50th Centile Difference Derivative")
  print(deriv_plt)
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv.png")
  ggsave(n, deriv_plt)
  
  deriv_plt2 <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="Scaled 50th Centile Difference Derivative")
  print(deriv_plt2)
  
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv_scaled.png")
  ggsave(n, deriv_plt2)
  #ggplotly(deriv_plt)
}
```

compare derivatives with chain rule
```{r}
sexdiffs_tbv %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("log_lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="50th Centile Difference Derivative")

sexdiffs_tbv %>%
  #filter(pheno_cat == "Subcortical Vol" & age_days > (365.25*40)) %>%
    ggplot() +
    geom_line(aes(x=age_days, y=deriv_M_minus_F_z.unlog, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("lifespan_fetal", sexdiffs_tbv$age_days) +
    labs(y="50th Centile Difference Derivative")
```

###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

Bootstrap!

```{r}
#calculate integral of deriv of diff^2 (all in z-scores)
sexdiffs_tbv_sq <- sexdiffs_tbv %>%
  mutate(sq_centile_diff_z = (centile_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_deriv_z = get_deriv(sq_centile_diff_z, logAge_days)) %>%
  mutate(sq_deriv_z.unlog = sq_deriv_z*deriv_age) %>% #chain rule to convert derivative
  ungroup()


#CHECK THAT THESE PLOTS MAKE SENSE!!
sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv_z, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv_sq$logAge_days) +
    labs(y="Change in Magnitude Sex Diff")

sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=age_days, y=sq_deriv_z.unlog, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("lifespan_fetal", sexdiffs_tbv_sq$age_days) +
    labs(y="Change in Magnitude Sex Diff")

#try getting integral
diff_int <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv_z),
            sq_int_z.unlog = rs_integral(logAge_days, sq_deriv_z))
```


try finding points at which pt derivative crosses 0 increasing vs decreasing

```{r}
cent_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv_z.unlog > 0 ~ "diverging",
    sq_deriv_z.unlog == 0 ~ "constant",
    sq_deriv_z.unlog < 0 ~ "converging",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- cent_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv_z, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

cent_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", cent_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)
```
Get max male and female biases and plot on centile differences

```{r}
peak_diffs_sum <- cent_df %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(centile_M_minus_F_z, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(centile_M_minus_F)),
            female_peak = min_gc(centile_M_minus_F_z, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(centile_M_minus_F))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")


peak_diffs_sum %>%
  ggplot() +
  geom_histogram(aes(x=peak_age, fill=sex), position="identity", alpha=.4) +
  format_x_axis("log_lifespan_fetal", peak_diffs_sum$peak_age) +
  facet_grid(pheno_cat~cv_sample, scales="free_y")

cent_df %>%
    filter(pheno %in% consistent_agesex) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% consistent_agesex), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
```

differences tend to peak around beginning and end of life - likely d/t smaller sample sizes at extreme ages

plot convergence/divergence on brain

```{r}
#removing ventricles for visualization:
diff_int_novent <- full_join(LR.df_tot_sex, diff_int) %>%
  filter(!(pheno %in% vent_list))
range(diff_int_novent$sq_int_z.unlog)

full_converge_plt <- plot_cv_brain(diff_int_novent, sq_int_z.unlog, "Change in \nMagnitude \nSex Diff", "ggthemes::Orange-Blue Diverging", c(-.21, .21), c(2, 5, 2))

ggsave("figs/convergence_tbvTRUE.png", full_converge_plt, width=7, height=8, dpi=600)
```

####Sigma

*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of lnCVR squared: (not sure this totally makes sense tbh)
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#quick dirty code to get abs diff
sigma_df <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==TRUE) %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_cv_diff = (logcv_M_div_F)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_cv_deriv = get_deriv(sq_cv_diff, logAge_days)) %>%
  ungroup() %>%
  mutate(sq_cv_deriv.unlog = sq_cv_deriv*deriv_age)

sigma_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_cv_deriv, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sigma_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Variability")
```

###Cross-sample Consistency
Get across-sample euclidean distance
```{r}
dist_df <- data.frame()
for (pheno_val in pheno_list){
  df <- sexdiffs_tbv %>%
    filter(pheno==pheno_val) %>%
    arrange(logAge_days)
  
  dfA <- df %>%
    filter(cv_sample=="A")
  
  dfB <- df %>%
    filter(cv_sample=="B")
  
  if (nrow(dfA) == 0 | nrow(dfB) == 0) next
  
  cent_dist <- euclidean_dist(dfA$centile_M_minus_F_z, dfB$centile_M_minus_F_z)
  sigma_dist <- euclidean_dist(dfA$logcv_M_div_F, dfB$logcv_M_div_F)
  deriv_dist <- euclidean_dist(na.omit(dfA$deriv_M_minus_F_z), na.omit(dfB$deriv_M_minus_F_z))
  
  #rm nas for integral
  dfA <- dfA %>%
    select(logAge_days, deriv_M_minus_F_z) %>%
    na.omit()
  
  dfB <- dfB %>%
    select(logAge_days, deriv_M_minus_F_z) %>%
    na.omit()
  
  deriv_j <- trapz(dfA$logAge_days, abs(dfA$deriv_M_minus_F_z - dfB$deriv_M_minus_F_z))
  
  result <- data.frame(
    "pheno" = pheno_val,
    "centile_dist" = cent_dist,
    "sigma_dist" = sigma_dist,
    "deriv_dist" = deriv_dist,
    "deriv_trapz" = deriv_j
  )
  dist_df <- bind_rows(dist_df, result)
}

dist_df <- dist_df %>%
  mutate(centile_sim = 1/(1+centile_dist),
         sigma_sim = 1/(1+sigma_dist),
         deriv_sim = 1/(1+deriv_dist)
         )

thresholds <- dist_df %>%
  summarize(centile_dist_mad = mad(centile_dist),
            centile_dist_median = median(centile_dist),
            sigma_dist_mad = mad(sigma_dist),
            sigma_dist_median = median(sigma_dist),
            deriv_dist_mad = mad(deriv_dist),
            deriv_dist_median = median(deriv_dist)) %>%
  mutate(centile_dist_threshold = centile_dist_median+(1*centile_dist_mad),
         sigma_dist_threshold = sigma_dist_median+(1*sigma_dist_mad),
         deriv_dist_threshold = deriv_dist_median+(1*deriv_dist_mad))

ggplot(dist_df) +
  geom_histogram(aes(x=centile_sim), alpha=.5, fill="purple") +
  geom_histogram(aes(x=sigma_sim), alpha=.5, fill="green") +
  geom_histogram(aes(x=deriv_sim), alpha=.5, fill="orange")

dist_df %>%
  summarise(n_pass_cent = sum(centile_dist <= thresholds$centile_dist_threshold),
            n_pass_sigma = sum(sigma_dist <= thresholds$sigma_dist_threshold),
            n_pass_deriv = sum(deriv_dist <= thresholds$deriv_dist_threshold),
            n = n()) %>%
  mutate(pct_pass_cent = (n_pass_cent/n)*100,
         pct_pass_sigma = (n_pass_sigma/n)*100,
         pct_pass_deriv = (n_pass_deriv/n)*100)
  
```


```{r}
sexdiffs_tbv2 <- full_join(sexdiffs_tbv, dist_df)

cent_dist_plt <- sexdiffs_tbv2 %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z, centile_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    gghighlight(centile_dist <= thresholds$centile_dist_threshold, calculate_per_facet = TRUE) +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

cent_dist_plt

cent_dist_plt_sig <- sexdiffs_tbv2 %>%
    select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample, sigma_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(sigma_dist <= thresholds$sigma_dist_threshold, calculate_per_facet = TRUE) +
    theme(legend.position = "none") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Sex-Difference in Variability (lnCVR)")

cent_dist_plt_sig

cent_dist_plt_d <- sexdiffs_tbv2 %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F_z, deriv_dist) %>%
  #filter(deriv_dist <= thresholds$deriv_dist_threshold) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(deriv_dist <= thresholds$deriv_dist_threshold, calculate_per_facet = TRUE) +
    theme(legend.position = "none") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Sex-Difference in Growth")

ggplotly(cent_dist_plt_d)

cent_dist_plt_d2 <- sexdiffs_tbv2 %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F_z, deriv_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=deriv_dist)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

#ggplotly(cent_dist_plt_d2)
```

compare consistent trajectories to regions with consistent sex/ consistent age*sex effects

```{r}
dist_df2 <- dist_df %>%
  mutate(consistent_sex = ifelse(pheno %in% consistent_sex, TRUE, FALSE),
         consistent_agesex = ifelse(pheno %in% consistent_agesex, TRUE, FALSE))
  
ggplot(dist_df2) +
  geom_jitter(aes(x=1, y=deriv_dist, color=consistent_agesex), alpha=.4) +
  theme_bw()

ggplot(dist_df) +
  geom_jitter(aes(x=1, y=cent_dist), alpha=.4, color="red") +
  #gghighlight(pheno %in% consistent_sex) +
  theme_bw()
```

re-assessing peak male-female differences only in phenos with consistent derivatives

```{r}
consistent_deriv <- dist_df %>% 
  filter(deriv_dist <= thresholds$deriv_dist_threshold) %>%
  pull(pheno)

keep_list <- intersect(consistent_deriv, consistent_agesex)

peak_diffs_sum %>%
  filter(pheno %in% keep_list) %>%
  mutate(peak_age_yrs = (un_log(peak_age)-280)/365.25) %>%
  ggplot() +
  geom_histogram(aes(x=peak_age_yrs, fill=sex), position="identity", alpha=.4) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  geom_vline(xintercept = 50, color='gray') +
  geom_vline(xintercept = 9, color='gray') +
  geom_vline(xintercept = 18, color='gray') +
  geom_vline(xintercept = 0, color='gray')

peak_plt <- sexdiffs_tbv %>%
    filter(pheno %in% keep_list) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% keep_list), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

ggplotly(peak_plt)
```

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)
Calculating across age, not log-age, for interpretability

```{r}
dominance_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = rs_integral(logAge_days, centile_M_minus_F_z),
            sigma_int.unlog =  rs_integral(logAge_days, logcv_M_div_F),
            age_range = max(age_days) - min(age_days)) %>%
  # rescaling integrals to 100 years
  mutate(centile_int_norm.unlog = (centile_int.unlog/age_range)*365.25*100,
         sigma_int_norm.unlog = (sigma_int.unlog/age_range)*365.25*100) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5)) +
  coord_flip()

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()
```


```{r}
dominance_plt_df <- full_join(LR.df_tot_sex, dominance_df %>% select(!pheno_cat))
dominance_plt_df %>%
  mutate(dominance = ifelse(sig_fdr==TRUE, dominance, "NA")) %>%
  group_by(cv_sample) %>%
  summarise(F_dom = sum(dominance=="Female")/n(),
            M_dom = sum(dominance=="Male")/n())
```

plot dominance on brain

```{r}
#removing ventricles for visualization:
dominance_plt_df_novent <- dominance_plt_df %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_novent$centile_int_norm.unlog)

full_dom_plt <- plot_cv_brain(dominance_plt_df_novent, centile_int_norm.unlog, "Normalized Sex\nDifference", "ggthemes::Orange-Blue Diverging", c(-0.07, 0.07), c(2, 5, 2))

ggsave("figs/dominance_tbvTRUE.png", full_dom_plt, width=7, height=8, dpi=600)
```

How much of the lifespan are phenos male vs female dominant?

```{r}
#define function
prop_above <- function(x, y, threshold=0){
  crosses <- find_threshold_crosses(y, x, threshold, return="x", return_endpoints = FALSE)
  crosses <- c(crosses, min(x), max(x)) %>% 
    sort()
  #overall age modeled
  total_range <- max(x) - min(x)
  
  ranges <- diff(crosses)
  
  #if y > threshold at min(x) -> sum odd number ranges
  odd_indices_logical <- (1:length(ranges)) %% 2 == 1
  
  if (y[1] > threshold) {
    items_at_odd_indices <- ranges[odd_indices_logical]
    time_above <- sum(items_at_odd_indices)
  } else {
    items_at_even_indices <- ranges[!odd_indices_logical]
    time_above <- sum(items_at_even_indices)
  }
  
  prop_above <- time_above/total_range
  
  return(prop_above)
}

# #test function
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="sGMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
# 
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="WMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
```

```{r}
prop_above_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, centile_M_minus_F_z)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_df$cv_sample, prop_above_df$class)

prop_above_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_df_novent <- full_join(LR.df_tot_sex, prop_above_df %>% select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_plt <- plot_cv_brain(prop_above_df_novent, prop_male, "Proportion Male-Dominant", "ggthemes::Orange-Blue Diverging", c(0, 1), c(2, 5, 2))

ggsave("figs/prop_male_tbvTRUE.png", full_above_plt, width=7, height=8, dpi=600)
```


####Sigma

```{r}
range(dominance_plt_df_novent$sigma_int_norm.unlog)

full_dom_plt_sigma <- plot_cv_brain(dominance_plt_df_novent, sigma_int_norm.unlog, "Normalized Sigma\nSex Difference", "ggthemes::Orange-Blue Diverging", c(-.03, .03), c(2, 5, 2))

ggsave("figs/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, width=7, height=8, dpi=600)
```

```{r}
prop_above_sigma_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, logcv_M_div_F)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_sigma_df$cv_sample, prop_above_sigma_df$class)

prop_above_sigma_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_sigma_df_novent <- full_join(LR.df_tot_sex, prop_above_sigma_df %>% select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_sig_plt <- plot_cv_brain(prop_above_sigma_df_novent, prop_male, "Proportion Male-Dominant\nSigma", "ggthemes::Orange-Blue Diverging", c(0, 1), c(2, 5, 2))

ggsave("figs/prop_male_sigma_tbvTRUE.png", full_above_sig_plt, width=7, height=8, dpi=600)
```



###Shape Differences
*At what points in the lifespan are males and females growing differently?*
difference of first derivatives!
can also get magnitude differences

```{r}
deriv_df_sum <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(int2_z = trapz(logAge_days, deriv_M_minus_F_z),
            int2_z.unlog = rs_integral(logAge_days, deriv_M_minus_F_z))
```

int > 0 -> males grow faster than females
int < 0 -> females grow faster
int = 0 -> same rate

```{r}
deriv_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(faster_rate = case_when(
    deriv_M_minus_F_z.unlog > 0 ~ "Males",
    deriv_M_minus_F_z.unlog == 0 ~ "Same",
    deriv_M_minus_F_z.unlog < 0 ~ "Females",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- deriv_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = deriv_M_minus_F_z.unlog, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

deriv_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=faster_rate)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", deriv_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))
```

do times with greatest differences in growth rate line up with puberty, menopause, etc? 

```{r}
peak_deriv_diffs <- sexdiffs_tbv %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(deriv_M_minus_F_z, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(deriv_M_minus_F_z)),
            female_peak = min_gc(deriv_M_minus_F_z, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(deriv_M_minus_F_z))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")

peak_deriv_hist <- peak_deriv_diffs %>%
  filter(pheno %in% keep_list) %>%
  mutate(peak_age_yrs = (un_log(peak_age)-280)/365.25) %>%
  ggplot() +
  geom_histogram(aes(x=peak_age_yrs, fill=sex), position="identity", alpha=.4) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  geom_vline(xintercept = 50, color='gray') +
  geom_vline(xintercept = 9, color='gray') +
  geom_vline(xintercept = 18, color='gray') +
  geom_vline(xintercept = 0, color='gray')

peak_deriv_hist

ggplotly(peak_deriv_hist)

yr_to_logAge_days <- function(x){
  log((x*365.25 + 280), base=10)
}

peak_plt_deriv <- sexdiffs_tbv %>%
    filter(pheno %in% keep_list) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_deriv_diffs, pheno %in% keep_list), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference") +
    geom_vline(xintercept = yr_to_logAge_days(50), color='gray') +
    geom_vline(xintercept = yr_to_logAge_days(9), color='gray') +
    geom_vline(xintercept = yr_to_logAge_days(18), color='gray') +
    geom_vline(xintercept = yr_to_logAge_days(0), color='gray')

peak_plt_deriv
```

###fPCA

getting and plotting principle component of centile differences, sigma differences, and derivative differences
just using phenos with consistent results across cv splits

prep split A for now

```{r}
sexdiffs.df_uniform.sc <- sexdiffs.df_uniform %>%
  filter(total=="TRUE" & weighted==FALSE) %>%
  group_by(cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    #tmp fix for sigma as CV, getting ln CV ratio
    logcv_M_div_F = log(sigma_Male/sigma_Female),
    #chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age)
```

define function (written with help from claude)

```{r}
library(fda)

plot_fpc1 <- function(data,
                      time_var = "logAge_days",
                      value_var = "centile_M_minus_F_z",
                      cv_sample_value = "A",
                      keep_phenos = NULL,
                      nbasis = 20,
                      nharm = 10,
                      x_axis_formatter = "log_lifespan_fetal",
                      plot_title = NULL,
                      x_label = NULL,
                      y_label = "Value") {
  
  # Filter data if keep_phenos is provided
  if (!is.null(keep_phenos)) {
    data <- data %>% filter(pheno %in% keep_phenos)
  }
  
  # Filter by cv_sample if provided
  if (!is.null(cv_sample_value)) {
    data <- data %>% filter(cv_sample == cv_sample_value)
  }
  
  # Create wide format
  wide_data <- data %>%
    pivot_wider(id_cols = time_var, 
                names_from = pheno, 
                values_from = value_var) %>%
    na.omit()
  
  # Extract time points
  time_points <- wide_data[[time_var]]
  
  # Create matrix (transpose so rows = phenos, cols = time points)
  matrix_data <- wide_data %>%
    arrange(!!sym(time_var)) %>%
    dplyr::select(-(time_var)) %>%
    t()
  
  # Create B-spline basis
  basis <- create.bspline.basis(rangeval = range(time_points, na.rm = TRUE), 
                                 nbasis = nbasis)
  
  # Convert to functional data object
  brain_fd <- Data2fd(argvals = time_points, 
                      y = t(matrix_data), 
                      basisobj = basis)
  
  # Perform functional PCA
  fpca_result <- pca.fd(fdobj = brain_fd, 
                        nharm = nharm)
  
  # Extract PC1 variance explained
  pc1_var <- fpca_result$varprop[1] * 100
  
  # Create fine grid for smooth plotting
  time_grid <- seq(min(time_points), max(time_points), length.out = 200)
  
  # Evaluate PC1 at grid points
  pc1_values <- as.numeric(eval.fd(time_grid, fpca_result$harmonics[1]))
  
  # Create plot data
  plot_data <- data.frame(
    time = time_grid,
    pc1 = pc1_values
  )
  
  # Create base plot
  p <- ggplot(plot_data, aes(x = time, y = pc1)) +
    geom_line(linewidth = 1.2, color = "darkblue") +
    labs(title = paste(plot_title, "PC 1"),
        subtitle = sprintf("%.1f%% variance", pc1_var),
         x = if(!is.null(x_label)) x_label else time_var,
         y = y_label) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
    format_x_axis(x_axis_formatter, plot_data$time) 
  
  # Same for PC2
  pc2_var <- fpca_result$varprop[2] * 100
  pc2_values <- as.numeric(eval.fd(time_grid, fpca_result$harmonics[2]))
  
  plot_data2 <- data.frame(
    time = time_grid,
    pc2 = pc2_values
  )
  
  # Create base plot
  p2 <- ggplot(plot_data2, aes(x = time, y = pc2)) +
    geom_line(linewidth = 1.2, color = "darkblue") +
    labs(title = paste(plot_title, "PC 2"),
        subtitle = sprintf("%.1f%% variance", pc2_var),
         x = if(!is.null(x_label)) x_label else time_var,
         y = y_label) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
    format_x_axis(x_axis_formatter, plot_data2$time) 
  
  # Return list with plot and fpca results
  return(list(
    plot = p,
    plot2 = p2,
    fpca_result = fpca_result,
    pc1_variance = pc1_var,
    time_grid = time_grid,
    pc1_values = pc1_values,
    cv_sample = cv_sample_value,
    phenos = keep_phenos,
    var = value_var,
    time_var = time_var
  ))
}

```

centiles

```{r}
consistent_cent <- dist_df %>% 
  filter(centile_dist <= thresholds$centile_dist_threshold) %>%
  pull(pheno)

#phenos that def have sex effects & have consistent centile diff trajectories
keep_list_cent <- intersect(consistent_cent, consistent_sex)
length(keep_list_cent)

centile_A_fpc <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "centile_M_minus_F_z",
  cv_sample_value = "A",
  keep_phenos = keep_list_cent,
  plot_title = "Centile Split A",
  y_label = "Centile Difference PC1")

centile_B_fpc <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "centile_M_minus_F_z",
  cv_sample_value = "B",
  keep_phenos = keep_list_cent,
  plot_title = "Centile Split B",
  y_label = "Centile Difference PC1")


centile_A_fpc$plot
centile_A_fpc$plot2
centile_B_fpc$plot
centile_B_fpc$plot2
```

sigma

```{r}
consistent_sigma <- dist_df %>% 
  filter(sigma_dist <= thresholds$sigma_dist_threshold) %>%
  pull(pheno)

#phenos that def have sex effects & have consistent centile diff trajectories
keep_list_sigma <- intersect(consistent_sigma, consistent_sex)
length(keep_list_sigma)

sigma_A_fpc <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "logcv_M_div_F",
  cv_sample_value = "A",
  keep_phenos = keep_list_sigma,
  plot_title = "Variance Diff Split A",
  y_label = "log CV Ratio")

sigma_B_fpc <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "logcv_M_div_F",
  cv_sample_value = "B",
  keep_phenos = keep_list_sigma,
  plot_title = "Variance Split B",
  y_label = "log CV Ratio")


sigma_A_fpc$plot
sigma_A_fpc$plot2

sigma_B_fpc$plot
sigma_B_fpc$plot2
```


Derivatives
```{r}
consistent_deriv <- dist_df %>% 
  filter(deriv_dist <= thresholds$deriv_dist_threshold) %>%
  pull(pheno)

#all consistent sex-differences
keep_list_deriv1 <- intersect(consistent_deriv, consistent_sex)
length(keep_list_deriv1)

deriv_A_fpc <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "deriv_M_minus_F_z",
  cv_sample_value = "A",
  keep_phenos = keep_list_deriv1,
  plot_title = "Sex-Difference Deriv Split A",
  y_label = "Derivative")

deriv_B_fpc <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "deriv_M_minus_F_z",
  cv_sample_value = "B",
  keep_phenos = keep_list_deriv1,
  plot_title = "Sex-Difference Deriv Split B",
  y_label = "Derivative")

deriv_A_fpc$plot
deriv_A_fpc$plot2
deriv_B_fpc$plot
deriv_B_fpc$plot2
```


derivatives just in phenos with significant age*sex effects

```{r}
#consistent age*sex effects
keep_list_deriv2 <- intersect(consistent_deriv, consistent_agesex)
length(keep_list_deriv2)

deriv_A_fpc2 <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "deriv_M_minus_F_z",
  cv_sample_value = "A",
  keep_phenos = keep_list_deriv2,
  plot_title = "Sex-Difference Deriv Split A",
  y_label = "Derivative")

deriv_B_fpc2 <- plot_fpc1(
  data = sexdiffs.df_uniform.sc,
  time_var = "logAge_days",
  value_var = "deriv_M_minus_F_z",
  cv_sample_value = "B",
  keep_phenos = keep_list_deriv2,
  plot_title = "Sex-Difference Deriv Split B",
  y_label = "Derivative")

deriv_A_fpc2$plot
deriv_A_fpc2$plot2
deriv_B_fpc2$plot
deriv_B_fpc2$plot2
```

####Clustering

Clustering same phenos as above on fPCA

```{r}
cluster_fpca_trajectories <- function(fpca_output,
                                      data,
                                      n_pcs = NULL,
                                      variance_threshold = 0.80,
                                      k = 5,
                                      method = "kmeans",
                                      nstart = 25,
                                      plot_clusters = TRUE,
                                      x_axis_formatter = "log_lifespan_fetal",
                                      x_label = NULL) {
  
  # Extract components from fpca_output
  fpca_result <- fpca_output$fpca_result
  keep_phenos <- fpca_output$phenos
  cv_sample_value <- fpca_output$cv_sample
  value_var <- fpca_output$var
  time_var <- fpca_output$time_var
  
  # Get fPCA scores
  fpca_scores <- fpca_result$scores
  rownames(fpca_scores) <- keep_phenos
  
  # Determine number of PCs to use
  variance_explained <- fpca_result$varprop
  cumulative_variance <- cumsum(variance_explained)
  
  if (is.null(n_pcs)) {
    # Use variance threshold
    n_pcs <- which(cumulative_variance >= variance_threshold)[1]
    if (is.na(n_pcs)) n_pcs <- length(cumulative_variance)
  }
  
  cat(sprintf("Using %d PCs (explaining %.1f%% of variance)\n", 
              n_pcs, cumulative_variance[n_pcs] * 100))
  
  # Select PC scores to use
  scores_subset <- fpca_scores[, 1:n_pcs, drop = FALSE]
  
  # Perform clustering
  set.seed(123)
  if (method == "kmeans") {
    cluster_result <- kmeans(scores_subset, centers = k, nstart = nstart)
    clusters <- cluster_result$cluster
  } else if (method == "hierarchical") {
    dist_matrix <- dist(scores_subset)
    hc <- hclust(dist_matrix, method = "ward.D2")
    clusters <- cutree(hc, k = k)
  } else {
    stop("Method must be 'kmeans' or 'hierarchical'")
  }
  
  # Print cluster sizes
  cat("\nCluster Sizes:\n")
  print(table(clusters))
  
  # Create results dataframe
  results_df <- data.frame(
    pheno = keep_phenos,
    cluster = clusters,
    stringsAsFactors = FALSE
  )
  
  # Add PC scores
  for (i in 1:n_pcs) {
    results_df[[paste0("PC", i)]] <- fpca_scores[, i]
  }
  
  # Sort by cluster
  results_df <- results_df %>% arrange(cluster, pheno)
  
  # Print phenotypes by cluster
  cat("\nPhenotypes by Cluster:\n")
  for (i in 1:k) {
    cat(sprintf("\nCluster %d (%d phenotypes):\n", i, sum(clusters == i)))
    print(keep_phenos[clusters == i])
  }
  
  # Generate plots if requested
  plots <- list()
  
  if (plot_clusters) {
    # Filter and prepare data for plotting
    plot_data <- data %>%
      filter(pheno %in% keep_phenos) %>%
      filter(cv_sample == cv_sample_value) %>%
      left_join(results_df %>% dplyr::select(pheno, cluster), by = "pheno") %>%
      mutate(cluster = factor(cluster))
    
    # Plot 1: Trajectories colored by cluster
    p_trajectories <- ggplot(plot_data, 
                             aes(x = !!sym(time_var), 
                                 y = !!sym(value_var), 
                                 group = pheno, 
                                 color = cluster)) +
      geom_line(alpha = 0.5, linewidth = 0.7) +
      facet_wrap(~cluster, ncol = 2) +
      labs(title = "Trajectories by Cluster",
           x = if(!is.null(x_label)) x_label else time_var,
           y = value_var) +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
            legend.position = "none") +
      format_x_axis(x_axis_formatter, plot_data[[time_var]])
    
    plots$trajectories_faceted <- p_trajectories
    
    # Plot 2: All trajectories colored by cluster (no facets)
    p_trajectories_all <- ggplot(plot_data, 
                                  aes(x = !!sym(time_var), 
                                      y = !!sym(value_var), 
                                      group = pheno, 
                                      color = cluster)) +
      geom_line(alpha = 0.5, linewidth = 0.7) +
      labs(title = "All Trajectories Colored by Cluster",
           x = if(!is.null(x_label)) x_label else time_var,
           y = value_var,
           color = "Cluster") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
      format_x_axis(x_axis_formatter, plot_data[[time_var]])
    
    plots$trajectories_all <- p_trajectories_all
    
    # Plot 3: PC1 vs PC2 scatter colored by cluster
    pc_plot_data <- data.frame(
      PC1 = fpca_scores[, 1],
      PC2 = fpca_scores[, 2],
      cluster = factor(clusters),
      pheno = keep_phenos
    )
    
    p_pc_scatter <- ggplot(pc_plot_data, aes(x = PC1, y = PC2, color = cluster, label = pheno)) +
      geom_point(size = 3, alpha = 0.7) +
      geom_text(size = 2.5, hjust = 0, vjust = 0, nudge_x = 0.01, nudge_y = 0.01, 
                check_overlap = TRUE, show.legend = FALSE) +  # Added text labels
      labs(title = "Clusters in PC Space",
       x = sprintf("PC1 (%.1f%% variance)", variance_explained[1] * 100),
       y = sprintf("PC2 (%.1f%% variance)", variance_explained[2] * 100),
       color = "Cluster") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
    
    plots$pc_scatter <- p_pc_scatter
    
    # Plot 4: Elbow plot (if kmeans)
    if (method == "kmeans") {
      max_k <- min(10, nrow(scores_subset) - 1)
      wss <- sapply(1:max_k, function(k) {
        kmeans(scores_subset, centers = k, nstart = 25)$tot.withinss
      })
      
      elbow_data <- data.frame(k = 1:max_k, wss = wss)
      
      p_elbow <- ggplot(elbow_data, aes(x = k, y = wss)) +
        geom_line(color = "blue", linewidth = 1) +
        geom_point(color = "blue", size = 3) +
        geom_vline(xintercept = k, color = "red", linetype = "dashed") +
        labs(title = "Elbow Plot for Optimal k",
             x = "Number of Clusters",
             y = "Total Within-Cluster Sum of Squares") +
        theme_minimal() +
        scale_x_continuous(breaks = 1:max_k)
      
      plots$elbow <- p_elbow
    }
    
    # Plot 5: Dendrogram (if hierarchical)
    if (method == "hierarchical") {
      dist_matrix <- dist(scores_subset)
      hc <- hclust(dist_matrix, method = "ward.D2")
      plots$dendrogram <- hc  # Store hclust object for plotting separately
    }
  }
  
  # Return comprehensive results
  return(list(
    clusters = clusters,
    results_df = results_df,
    cluster_result = if(method == "kmeans") cluster_result else NULL,
    n_pcs_used = n_pcs,
    variance_explained = cumulative_variance[n_pcs],
    fpca_scores = fpca_scores,
    plots = plots,
    method = method,
    k = k
  ))
}
```


centiles

```{r}
centile_clusters <- cluster_fpca_trajectories(fpca_output = centile_A_fpc,
                                      data = sexdiffs.df_uniform.sc,
                                      n_pcs = 2,
                                      variance_threshold = 0.80,
                                      k = 3,
                                      method = "kmeans",
                                      nstart = 25,
                                      plot_clusters = TRUE,
                                      x_axis_formatter = "log_lifespan_fetal",
                                      x_label = NULL)

print(centile_clusters$plots$trajectories_faceted)
print(centile_clusters$plots$trajectories_all)
print(centile_clusters$plots$pc_scatter)
print(centile_clusters$plots$elbow)
```

```{r}
sigma_clusters <- cluster_fpca_trajectories(fpca_output = sigma_A_fpc,
                                      data = sexdiffs.df_uniform.sc,
                                      n_pcs = 2,
                                      variance_threshold = 0.80,
                                      k = 3,
                                      method = "kmeans",
                                      nstart = 25,
                                      plot_clusters = TRUE,
                                      x_axis_formatter = "log_lifespan_fetal",
                                      x_label = NULL)

print(sigma_clusters$plots$trajectories_faceted)
print(sigma_clusters$plots$trajectories_all)
print(sigma_clusters$plots$pc_scatter)
#ggplotly(sigma_clusters$plots$pc_scatter)
print(sigma_clusters$plots$elbow)
```

```{r}
deriv_clusters <- cluster_fpca_trajectories(fpca_output = deriv_A_fpc,
                                      data = sexdiffs.df_uniform.sc,
                                      n_pcs = 2,
                                      variance_threshold = 0.80,
                                      k = 3,
                                      method = "kmeans",
                                      nstart = 25,
                                      plot_clusters = TRUE,
                                      x_axis_formatter = "log_lifespan_fetal",
                                      x_label = NULL)

print(deriv_clusters$plots$trajectories_faceted)
print(deriv_clusters$plots$trajectories_all)
print(deriv_clusters$plots$pc_scatter)
print(deriv_clusters$plots$elbow)
```