---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, ggcorrplot, sf)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

#ggseg Prep

```{r}
### dk for cortical ###
#write out new dk atlas w/o medial wall
dk_clean <- dk
dk_clean$data <- dk_clean$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")

#shrink for outlines
dk_region_gap <- 0.3        # Space between regions
dk_border_inset <- 0.3      # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
dk_fill <- dk_clean
dk_fill$data <- dk_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -dk_region_gap) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
dk_border <- dk_clean
dk_border$data <- dk_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(dk_region_gap + dk_border_inset)) %>%
  as_tibble() %>%
  st_as_sf()

### aseg for subcortical ###
#shrink for outlines
aseg_region_gap <- 0        # Space between regions
aseg_border_inset <- 1e-5     # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
aseg_fill <- aseg
aseg_fill$data <- aseg$data %>%
  st_as_sf() %>%
  st_buffer(dist = 0, 
            joinStyle = "MITRE",
            nQuadSegs=500) %>%
  st_buffer(0) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
aseg_border <- aseg
aseg_border$data <- aseg$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(aseg_region_gap + aseg_border_inset), nQuadSegs=500) %>%
  as_tibble() %>%
  st_as_sf()

### aseg for global ###
aseg_all_clean <- aseg_all
aseg_all_clean$data <- aseg_all_clean$data %>%
  dplyr::filter(label != "Brain-Stem")

#shrink for outlines
glob_region_gap <- 1e-10        # Space between regions
glob_border_inset <- 1e-10      # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
glob_fill <- aseg_all_clean
glob_fill$data <- aseg_all_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -glob_region_gap) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
glob_border <- aseg_all_clean
glob_border$data <- aseg_all_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(glob_region_gap + glob_border_inset)) %>%
  as_tibble() %>%
  st_as_sf()
```

trying to write funs for easier plotting
```{r}
plot_cv_brain <- function(df, fill_var, fill_name, fill_color, fill_limits, height_list){

#cortex
cortex_plt <- df %>%
  filter(pheno_cat == "Regional Vol" | pheno_cat == "Regional SA" | pheno_cat == "Regional CT") %>%
  group_by(cv_sample, pheno_cat) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = dk_fill,
             position = position_brain(hemi ~ side),
            aes(fill={{fill_var}}),#
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = dk_border,
                    position = position_brain(hemi ~ side),
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)

subcort_plt <- df %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  group_by(cv_sample, pheno_cat) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = aseg_fill,
            aes(fill={{fill_var}}),
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = aseg_border,
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)


#no buffer for now
glob_plt <- df %>%
  select(-any_of(c("geometry", "tissue_class"))) %>%
  group_by(cv_sample, pheno_cat) %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = aseg_all_clean,
            aes(fill={{fill_var}}), #
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = aseg_all_clean,
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)

#testing new spacing
# glob_plt <- glob_plt +
#   theme(plot.margin = margin(0, 0, 5, 0))

cortex_plt <- cortex_plt +
  theme(plot.margin = margin(5, 0, 5, 0))

full_plt_obj <- glob_plt / cortex_plt / subcort_plt +
  plot_layout(guides = "collect", heights = height_list)

# full_plt_obj <- glob_plt /plot_spacer()/ cortex_plt / plot_spacer()/ subcort_plt +
#   plot_layout(guides = 'collect', heights = unit(height_list, c('in', 'in', 'in')))

return(full_plt_obj)
}
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")


LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  select(!file)
```
```{r, eval=FALSE}
LR.df %>%
  group_by(cv_sample, total, weighted) %>%
  summarise(n_phenos = n_unique(pheno))

A_totalFALSE <- LR.df %>%
  filter(cv_sample == "A" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalFALSE)

A_totalTRUE <- LR.df %>%
  filter(cv_sample == "A" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalTRUE)

B_totalFALSE <- LR.df %>%
  filter(cv_sample == "B" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalFALSE)

B_totalTRUE <- LR.df %>%
  filter(cv_sample == "B" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalTRUE)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
         pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
)

table(LR.df2$effect, LR.df2$total)
```



```{r}
# sub_list
# unique(aseg$data$label)

LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$tissue_class), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas

check scales for plotting

```{r}
LR.df_plot %>%
  group_by(effect) %>%
  summarise(max_resi = max(resi))
```


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE)

table(LR.df_uncor$sig_fdr)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_uncor_novent <- LR.df_uncor %>%
  filter(!(pheno %in% vent_list))
range(LR.df_uncor_novent$resi)
full_plt_uncor <- plot_cv_brain(LR.df_uncor_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.1), c(1, 5, 1.1))

ggsave("figs/resi_uncorrected.png", full_plt_uncor, width=7, height=8.5, dpi=600)
```

#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE)

table(LR.df_tot$sig_fdr)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_tot_novent <- LR.df_tot %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_novent$resi)
full_plt_tot <- plot_cv_brain(LR.df_tot_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.1), c(1, 5, 1.1))

ggsave("figs/resi_tbv.png", full_plt_tot, width=7, height=8.5, dpi=600)
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE)

table(LR.df_tot_sex$sig_fdr, LR.df_tot_sex$cv_sample)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)

LR.df_tot_sex %>%
  filter(sig_fdr == TRUE) %>%
  group_by(cv_sample) %>%
  summarise(min_resi = min(sig_resi),
            max_resi = max(sig_resi),
            mean_resi = mean(sig_resi))
```

```{r}
#removing ventricles for visualization:
LR.df_tot_sex_novent <- LR.df_tot_sex %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_sex_novent$resi)
full_plt_tot_sex <- plot_cv_brain(LR.df_tot_sex_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.2), c(1, 5, 1.1))

ggsave("figs/resi_tbv_allsex.png", full_plt_tot_sex, width=7, height=8.5, dpi=600)
```


##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

sexdiffs.df_all <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         pheno = stringr::str_remove(basename(file), "_(uniform_)?sexdiffs\\.csv")) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    age_days = un_log(logAge_days))

#split into regular df and uniform df for correlations
sexdiffs.df_uniform <- sexdiffs.df_all[grepl("uniform", sexdiffs.df_all$file), ]
sexdiffs.df <- sexdiffs.df_all[!grepl("uniform", sexdiffs.df_all$file), ]

rm(output_fread_sd) #save space
rm(sexdiffs.df_all)
```

Fix scaling, derivatives

```{r}
sexdiffs.df_sc <- sexdiffs.df %>%
  group_by(total, weighted, cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    sigma_M_minus_F_z = sigma_Male_z - sigma_Female_z,
    centile_M_minus_F_z = median_centile_Male_z-median_centile_Female_z) %>%
  mutate(deriv_M_minus_F_z = c(
      NA,
      diff(centile_M_minus_F_z) / diff(logAge_days)),
      deriv_Male_z = c(
      NA,
      diff(median_centile_Male_z) / diff(logAge_days)),
      deriv_Female_z = c(
      NA,
      diff(median_centile_Female_z) / diff(logAge_days)),
      deriv_age = c(
      NA,
      diff(logAge_days) / diff(age_days)) #chain rule per taki's suggestion
      ) %>%
  ungroup() %>%
  #chain rule per taki's suggestion
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age)

sexdiffs.df_sc %>%
  group_by(total, weighted, cv_sample, pheno) %>%
  summarise(min_sig_diff=min(sigma_M_minus_F_z, na.rm=T),
            max_sig_diff=max(sigma_M_minus_F_z, na.rm=T))
```

Plot

```{r}
sexdiffs_tbv <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total=="TRUE")

sexdiffs <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==FALSE)

for (df in list(sexdiffs_tbv, sexdiffs)){
  t <- as.character(unique(df$total))
  print(t)
  
  #median centile
  cent_df <- df %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  cent_plt <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
  n <- paste0("figs/sexdiffs_tbv",t,"_cent.png")
  ggsave(n, cent_plt)
  print(cent_plt)
  
  cent_plt2 <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Scaled 50th Centile Sex-Difference")
  
  print(cent_plt2)
  n <- paste0("figs/sexdiffs_tbv",t,"_cent_scaled.png")
  ggsave(n, cent_plt2)
  #ggplotly(cent_plt)

  #sigma
  sig_df <- df %>%
    select(sigma_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, sigma_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  sig_plt <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sigma_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="Variance (Sigma) Sex-Difference")
  print(sig_plt)
  n <- paste0("figs/sexdiffs_tbv",t,"_sig.png")
  ggsave(n, sig_plt)
  
  sig_plt2 <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sigma_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="Scaled Variance (Sigma) Sex-Difference")
  print(sig_plt2)
  n <- paste0("figs/sexdiffs_tbv",t,"_sig_scaled.png")
  ggsave(n, sig_plt2)
  #ggplotly(sig_plt)
  
  #derivatives
  deriv_df <- df %>%
    select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  deriv_plt <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="50th Centile Difference Derivative")
  print(deriv_plt)
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv.png")
  ggsave(n, deriv_plt)
  
  deriv_plt2 <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="Scaled 50th Centile Difference Derivative")
  print(deriv_plt2)
  
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv_scaled.png")
  ggsave(n, deriv_plt2)
  #ggplotly(deriv_plt)
}
```

compare derivatives with chain rule
```{r}
sexdiffs_tbv %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="50th Centile Difference Derivative")

sexdiffs_tbv %>%
  #filter(pheno_cat == "Subcortical Vol" & age_days > (365.25*40)) %>%
    ggplot() +
    geom_line(aes(x=age_days, y=deriv_M_minus_F_z.unlog, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("lifespan_fetal", sexdiffs_tbv$age_days) +
    labs(y="50th Centile Difference Derivative")
```


Label phenos by whether sex effects are significant per LR test
```{r}
#regions w sig sex effects in both samples
LR.tot_sex_wide <- LR.df_tot_sex %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  mutate(rep = case_when(
    sig_fdr_A == TRUE & sig_fdr_B == TRUE ~ "consistent_sig",
    sig_fdr_A == FALSE & sig_fdr_B == FALSE ~ "consistent_ns",
    is.na(sig_fdr_A) | is.na(sig_fdr_B) ~ NA,
    TRUE ~ "mixed"
  ))

table(LR.tot_sex_wide$rep)

consistent_sex <- LR.tot_sex_wide %>%
  filter(rep == "consistent_sig") %>%
  pull(pheno)

#to temporarily account for missingness while remaining test models run
consistent_or_missing.sex <- LR.tot_sex_wide %>%
  filter(rep == "consistent_sig" | is.na(rep)) %>%
  pull(pheno)

#consistent age*sex effects
LR.tot_wide <- LR.df_tot %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  mutate(rep = case_when(
    sig_fdr_A == TRUE & sig_fdr_B == TRUE ~ "consistent_sig",
    sig_fdr_A == FALSE & sig_fdr_B == FALSE ~ "consistent_ns",
    is.na(sig_fdr_A) | is.na(sig_fdr_B) ~ NA,
    TRUE ~ "mixed"
  ))
table(LR.tot_wide$rep)

consistent_agesex <- LR.tot_wide %>%
  filter(rep == "consistent_sig") %>%
  pull(pheno)
```


###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#calculate integral of deriv of diff^2 (all in z-scores)
sexdiffs_tbv_sq <- sexdiffs_tbv %>%
  mutate(sq_centile_diff_z = (centile_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    sq_deriv_z = c(
      NA,
      diff(sq_centile_diff_z) / diff(logAge_days)
    )) %>%
  mutate(sq_deriv_z.unlog = sq_deriv_z*deriv_age) %>%
  ungroup()


sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv_z, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv_sq$logAge_days) +
    labs(y="Change in Magnitude Sex Diff")

sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=age_days, y=sq_deriv_z.unlog, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("lifespan_fetal", sexdiffs_tbv_sq$age_days) +
    labs(y="Change in Magnitude Sex Diff")

#try getting integral
diff_int <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv_z),
            sq_int_z.unlog = trapz(age_days, sq_deriv_z.unlog)
            )
```

How small of a derivative is effectively zero?

```{r}
#no sex*age effects -> derivative of sex diff is ~ 0
test_age.sex_tol <- full_join(LR.df_tot[, names(LR.df_tot) != "pheno_cat"], sexdiffs_tbv_sq)

test_age.sex_tol %>%
  group_by(sig_fdr) %>%
  summarise(max_der_diff = max(abs(sq_deriv_z), na.rm=T),
            mean_der_diff = mean(abs(sq_deriv_z), na.rm=T),
            min_der_diff = min(abs(sq_deriv_z), na.rm=T))

test_age.sex_tol %>%
  filter(sig_fdr == FALSE) %>%
  ggplot() +
  geom_histogram(aes(x=abs(sq_deriv_z))) +
  geom_vline(xintercept = 0.004)

sq_der_tol<-0.004
```


try finding points at which pt derivative crosses 0 increasing vs decreasing

```{r}
tol <- 1e-10
cent_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv_z >= tol ~ "diverging",
    abs(sq_deriv_z) < tol ~ "constant",
    sq_deriv_z <= -tol ~ "converging",
  )) %>% 
  ungroup()

cent_df %>%
  #filter(pheno %in% consistent_ns.agesex) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv_z, group=pheno, color=trajectory)) +
    facet_grid(~pheno_cat, scales="free_y")

#order by age first increasing
pheno_order_mean <- cent_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv_z, x = logAge_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

cent_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=logAge_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=logAge_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("log_lifespan_fetal", cent_df$logAge_days)

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=logAge_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("log_lifespan_fetal", cent_df$logAge_days)
```
Get max male and female biases and plot on centile differences

```{r}
peak_diffs_sum <- cent_df %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(centile_M_minus_F_z, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(centile_M_minus_F)),
            female_peak = min_gc(centile_M_minus_F_z, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(centile_M_minus_F))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")


peak_diffs_sum %>%
  ggplot() +
  geom_histogram(aes(x=peak_age, fill=sex), position="identity", alpha=.4) +
  format_x_axis("log_lifespan_fetal", peak_diffs_sum$peak_age) +
  facet_grid(pheno_cat~cv_sample, scales="free_y")

cent_df %>%
    #filter(pheno %in% consistent_agesex) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% consistent_agesex), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
```

differences tend to peak around beginning and end of life - likely d/t smaller sample sizes at extreme ages

*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#quick dirty code to get abs diff
sigma_df <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==TRUE) %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_sigma_diff_z = (sigma_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    sq_sigma_deriv_z = c(
      NA,
      diff(sq_sigma_diff_z) / diff(logAge_days))) %>%
  ungroup() %>%
  mutate(sq_sigma_deriv_z.unlog = sq_sigma_deriv_z*deriv_age)

sigma_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_sigma_deriv_z, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sigma_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Variability")
```

###Cross-sample Correlations
Get across-sample correlations
```{r}
cor_df <- data.frame()
for (pheno_val in pheno_list){
  df <- sexdiffs_tbv %>%
    filter(pheno==pheno_val) %>%
    arrange(logAge_days)
  
  dfA <- df %>%
    filter(cv_sample=="A")
  
  dfB <- df %>%
    filter(cv_sample=="B")
  
  if (nrow(dfA) == 0 | nrow(dfB) == 0) next
  
  cent_cor <- cor(dfA$centile_M_minus_F_z, dfB$centile_M_minus_F_z, use="complete.obs")
  sigma_cor <- cor(dfA$sigma_M_minus_F_z, dfB$sigma_M_minus_F_z, use="complete.obs")
  deriv_cor <- cor(dfA$deriv_M_minus_F_z, dfB$deriv_M_minus_F_z, use="complete.obs")
  
  result <- data.frame(
    "pheno" = pheno_val,
    "centile_cor" = cent_cor,
    "sigma_cor" = sigma_cor,
    "deriv_cor" = deriv_cor
  )
  cor_df <- bind_rows(cor_df, result)
}

ggplot(cor_df) +
  geom_histogram(aes(x=centile_cor)) +
  geom_vline(xintercept = .9, color="red")

ggplot(cor_df) +
  geom_histogram(aes(x=sigma_cor)) +
  geom_vline(xintercept = .9, color="red")

ggplot(cor_df) +
  geom_histogram(aes(x=deriv_cor)) +
  geom_vline(xintercept = .9, color="red")

cor_df %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  pivot_longer(cols=c(ends_with("cor")), names_to="type", values_to="cor") %>%
  group_by(type, pheno_cat) %>%
  summarise(mean_cor = mean(cor),
            n_pass = sum(cor>.9),
            pct_pass = (sum(cor>.9)/n())*100,
            n_neg = sum(cor<0),
            pct_neg = (sum(cor<0)/n())*100
            )
```

filtering to 1+ or 2+ yrs old doesn't change much

```{r, eval=FALSE, message=FALSE}
for (type in c("centile", "sigma", "deriv")){
  cor_col <- paste0(type, "_cor")
  val_col <- paste0(type, "_M_minus_F_z")

  for (pheno_val in pheno_list){
    #pull correlations
    cor_val <- cor_df %>% 
      filter(pheno==pheno_val) %>%
      pull(!!sym(cor_col))
    
    plt <- sexdiffs_tbv %>%
      filter(pheno==pheno_val) %>%
      select(!!sym(val_col), pheno, logAge_days, cv_sample) %>%
      na.omit() %>%
      arrange(pheno, logAge_days) %>%
      ggplot() +
      geom_line(aes(x=logAge_days, y=!!sym(val_col), group=cv_sample, color=cv_sample)) +
      theme(legend.position="none") +
      ggtitle(paste(pheno_val, "r =", cor_val))
    
    #print(plt)
    
    ggsave(paste0("./figs/", pheno_val, "_", type, "_sexdiff_cor_plt.png"), plt)
  }
}
```


```{r}
cor_df %>%
  filter(pheno %in% consistent_sex) %>%
  pivot_longer(cols=c(ends_with("cor")), names_to="type", values_to="cor") %>%
  group_by(type) %>%
  summarise(mean_cor = mean(cor),
            n_pass = sum(cor>.9),
            pct_pass = (sum(cor>.9)/n())*100,
            n_neg = sum(cor<0),
            pct_neg = (sum(cor<0)/n())*100
            )
```

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)
Calculating across age, not log-age, for interpretability (pending Taki's confirmation on slack)

```{r}
dominance_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = trapz(age_days, centile_M_minus_F_z),
            sigma_int.unlog = trapz(age_days, sigma_M_minus_F_z),
            log_age_range = max(logAge_days) - min(logAge_days),
            age_range = max(age_days) - min(age_days)) %>%
  # normalizing integrals by age range covered
  mutate(centile_int_norm.unlog = centile_int.unlog/age_range,
         sigma_int_norm.unlog = sigma_int.unlog/age_range) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

dominance_df %>%
  filter(dominance != dominance_norm) %>%
  nrow()
dominance_df %>%
  filter(sigma_dominance != sigma_dominance_norm) %>%
  nrow()

dominance_df %>%
  filter(pheno %in% consistent_sex) %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()

dominance_df %>%
  filter(pheno %in% consistent_sex) %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()
```


```{r}
dominance_plt_df <- full_join(LR.df_tot_sex, dominance_df %>% select(!pheno_cat))
dominance_plt_df %>%
  mutate(dominance = ifelse(sig_fdr==TRUE, dominance, "NA")) %>%
  group_by(cv_sample) %>%
  summarise(F_dom = sum(dominance=="Female")/n(),
            M_dom = sum(dominance=="Male")/n())
```

plot dominance on brain

```{r}
#removing ventricles for visualization:
dominance_plt_df_novent <- dominance_plt_df %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_novent$centile_int_norm.unlog)

full_dom_plt <- plot_cv_brain(dominance_plt_df_novent, centile_int_norm.unlog, "Normalized Sex\nDifference", "ggthemes::Orange-Blue Diverging", c(-.3, .3), c(2, 5, 2))

ggsave("figs/dominance_tbvTRUE.png", full_dom_plt, width=, height=8.5, dpi=600)
```

How much of the lifespan are phenos male vs female dominant?

```{r}
#define function
prop_above <- function(x, y, threshold=0){
  crosses <- find_threshold_crosses(y, x, threshold, return="x", return_endpoints = FALSE)
  crosses <- c(crosses, min(x), max(x)) %>% 
    sort()
  #overall age modeled
  total_range <- max(x) - min(x)
  
  ranges <- diff(crosses)
  
  #if y > threshold at min(x) -> sum odd number ranges
  odd_indices_logical <- (1:length(ranges)) %% 2 == 1
  
  if (y[1] > threshold) {
    items_at_odd_indices <- ranges[odd_indices_logical]
    time_above <- sum(items_at_odd_indices)
  } else {
    items_at_even_indices <- ranges[!odd_indices_logical]
    time_above <- sum(items_at_even_indices)
  }
  
  prop_above <- time_above/total_range
  
  return(prop_above)
}

#test function
test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="sGMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="WMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))
```

```{r}
prop_above_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, centile_M_minus_F_z)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_df$cv_sample, prop_above_df$class)

prop_above_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_df_novent <- full_join(LR.df_tot_sex, prop_above_df %>% select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_plt <- plot_cv_brain(prop_above_df_novent, prop_male, "Proportion Male-Dominant", "ggthemes::Orange-Blue Diverging", c(0, 1), c(3.5, -2, 6, -2, 3.5))

ggsave("figs/prop_male_tbvTRUE.png", full_above_plt, width=7, height=8, dpi=600)
```

```{r}
prop_above_plt <- left_join(LR.df_plot, prop_above_df %>% select(!pheno_cat))

reg_above_plt <- prop_above_plt %>%
  filter(pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")) %>% 
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    fill="gray",
    atlas = dk_clean,
    position = position_brain(hemi ~ side)
  ) +
  geom_brain(
    data = prop_above_plt %>% 
      filter(pheno %in% consistent_or_missing.sex & pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")),
    aes(fill = prop_male),
    atlas = dk_clean,
    position = position_brain(hemi ~ side)
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  facet_grid(rows = vars(pheno_cat), cols = vars(cv_sample), drop = TRUE)


sub_above_plt <- prop_above_plt %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    fill="gray",
    atlas = aseg_clean,
  ) +
  geom_brain(
    data = prop_above_plt %>% 
      filter(pheno %in% consistent_or_missing.sex & pheno_cat == "Subcortical Vol"),
    aes(fill = prop_male),
    atlas = aseg_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "none") +
  facet_grid(pheno_cat ~ cv_sample)

prop_above_plt_glob <- inner_join(prop_above_plt, aseg_all_clean$data, by="tissue_class")
glob_above_plt <- prop_above_plt_glob %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    fill="gray",
    atlas = aseg_all_clean
  ) +
  geom_brain(
    data = prop_above_plt_glob %>% 
      filter(pheno %in% consistent_or_missing.sex & pheno_cat == "Global Vol"),
    aes(fill = prop_male),
    atlas = aseg_all_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "none") +
  facet_grid(pheno_cat ~ cv_sample)
```

```{r}
full_above_plt <- glob_above_plt /plot_spacer()/ reg_above_plt / plot_spacer()/ sub_above_plt +
  plot_layout(guides = 'collect', heights = unit(c(3.5, -1.7, 6, -1.5, 3.5), c('in', 'in', 'in', 'in', 'in')))
full_above_plt
ggsave("figs/prop_male_tbvTRUE.png", full_above_plt, width=7, height=8)
```


####Sigma

recalc for ages 3+
```{r}
no_inf_dominance_df <- sexdiffs_tbv %>%
  filter(age_days >= (365.25*3)+280) %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = trapz(age_days, centile_M_minus_F_z),
            sigma_int.unlog = trapz(age_days, sigma_M_minus_F_z),
            log_age_range = max(logAge_days) - min(logAge_days),
            age_range = max(age_days) - min(age_days)) %>%
  # normalizing integrals by age range covered
  mutate(centile_int_norm.unlog = centile_int.unlog/age_range,
         sigma_int_norm.unlog = sigma_int.unlog/age_range) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

ggplot(no_inf_dominance_df) +
  geom_histogram(aes(x=sigma_int.unlog, fill=pheno_cat), position="identity", alpha=.5)
ggplot(no_inf_dominance_df) +
  geom_histogram(aes(x=sigma_int_norm.unlog, fill=pheno_cat), position="identity", alpha=.5)
```

CT still having really big sex effects

```{r}
range(LR.df_tot_dominance$sigma_int_norm)

reg_dom_plt_sigma <- LR.df_tot_dominance %>%
  filter(pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")) %>% 
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    fill="gray",
    atlas = dk_clean,
    position = position_brain(hemi ~ side)
  ) +
  geom_brain(
    data = LR.df_tot_dominance %>% 
      filter(pheno %in% consistent_sex & pheno_cat %in% c("Regional Vol", "Regional SA", "Regional CT")),
    aes(fill = sigma_int_norm),
    atlas = dk_clean,
    position = position_brain(hemi ~ side)
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  facet_grid(rows = vars(pheno_cat), cols = vars(cv_sample), drop = TRUE)  +
  scale_fill_gradient2(low = "#CD71B2FF", mid = "white", high = "#00ACAFFF", midpoint = 0, limits=c(-.02,.02))


sub_dom_plt_sigma <- LR.df_tot_dominance %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    fill="gray",
    atlas = aseg_clean,
  ) +
  geom_brain(
    data = LR.df_tot_dominance %>% 
      filter(pheno %in% consistent_sex & pheno_cat == "Subcortical Vol"),
    aes(fill = sigma_int_norm),
    atlas = aseg_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "none") +
  facet_grid(pheno_cat ~ cv_sample) +
  scale_fill_gradient2(low = "#CD71B2FF", mid = "white", high = "#00ACAFFF", midpoint = 0, limits=c(-.02,.02))

glob_dom_plt_sigma <- LR.df_tot_dominance_glob %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  geom_brain(
    fill="gray",
    atlas = aseg_all_clean
  ) +
  geom_brain(
    data = LR.df_tot_dominance_glob %>% 
      filter(pheno %in% consistent_sex),
    aes(fill = sigma_int_norm),
    atlas = aseg_all_clean
  ) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "none") +
  facet_grid(pheno_cat ~ cv_sample) +
  scale_fill_gradient2(low = "#CD71B2FF", mid = "white", high = "#00ACAFFF", midpoint = 0, limits=c(-.02,.02))
```

```{r}
full_dom_plt_sigma <- glob_dom_plt_sigma /plot_spacer()/ reg_dom_plt_sigma / plot_spacer()/ sub_dom_plt_sigma +
  plot_layout(guides = 'collect', heights = unit(c(3.5, -1.7, 6, -1.5, 3.5), c('in', 'in', 'in', 'in', 'in')))
full_dom_plt_sigma
ggsave("figs/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, width=7, height=8)
```


###Shape Differences
*At what points in the lifespan are males and females growing differently?*
integral of f''(M) - f''(F)

```{r}
#quick dirty code to get second deriv
second_deriv_df <- sexdiffs_tbv %>%
  filter(total==TRUE & weighted==FALSE) %>%
  select(!contains("centile") & !contains("sigma")) %>%
  na.omit() %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    deriv2_Male_z = c(
      NA,
      diff(deriv_Male_z) / diff(logAge_days)
    ),
    deriv2_Female_z = c(
      NA,
      diff(deriv_Female_z) / diff(logAge_days)
    ),
    deriv2_age = c(
      NA,
      diff(deriv_age) / diff(age_days)
    )) %>%
  mutate(deriv2_M_minus_F_z = deriv2_Male_z - deriv2_Female_z) %>%
         #chain rule with di bruno generalization - CHECK THIS
  mutate(deriv2_M_minus_F_z.unlog = deriv2_M_minus_F_z*(deriv_age)^2 + deriv_M_minus_F_z*deriv2_age)

second_deriv_df_sum <- second_deriv_df %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  na.omit() %>%
  summarise(int2_z = trapz(logAge_days, deriv2_M_minus_F_z),
            int2_z.unlog = trapz(age_days, deriv2_M_minus_F_z.unlog))

# ggplot(second_deriv_df_sum)+
#   geom_point(aes(x=int2, y=int2.unlog))
```

int2 > 0 -> males grow faster than females
int2 < 0 -> females grow faster
int2 = 0 -> same rate

```{r}
second_deriv_df %>%
  #mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=logAge_days, y=deriv2_M_minus_F_z, group=pheno, color=pheno)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  theme(legend.position = "none") +
  format_x_axis("log_lifespan_fetal", second_deriv_df$logAge_days)

second_deriv_df %>%
  #mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=deriv2_M_minus_F_z.unlog, group=pheno, color=pheno)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  theme(legend.position = "none") +
  format_x_axis("lifespan_fetal", second_deriv_df$age_days)
```

###Correlation

```{r}
df_cor_cent <- sexdiffs.df_uniform %>%
  filter(total=="TRUE" & weighted==FALSE) %>%
  select(logAge_days, centile_M_minus_F_z, pheno) %>%
  pivot_wider(id_cols=logAge_days, names_from=pheno, values_from=centile_M_minus_F_z)

c <- cor(df_cor)

```