---
title: "Review Centile Tests"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, RESI)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue")
library(ggsegTissue)
#remotes::install_github("ggseg/ggsegExtra")
```

#Read & Prep CSVs

centiles

*need to figure out how to just get control centiles ONCE*

```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "cent_csvs/*_PT_*cent.csv"))[1:100] #troubleshooting with subset

# Read all matching CSVs
pt_cent_df_list <- lapply(paths, function(p) {
  fread(p) %>%
    filter(dx != "CN") %>%
    mutate(
      cv_sample = stringr::str_extract(p, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
      total     = stringr::str_extract(p, "(?<=total)(TRUE|FALSE)"),
      weighted  = stringr::str_detect(p, "weight"),
      control = if_else(str_detect(dx_recode, "^CN_"), "CN", "PT"),
      dx_tested  = str_remove(dx_recode, "^CN_")
    ) %>%
    select(!c(sexMale_x_logAge, sexMale, study_site, INDEX.ID))
})

length(pt_cent_df_list)
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")
pt_cent_df <- rbindlist(pt_cent_df_list, fill=T)

rm(pt_cent_df_list)
```

```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "cent_csvs/*_CN_*cent.csv"))[1:100]

# Read all matching CSVs
cn_cent_df_list <- lapply(paths, function(p) {
  fread(p) %>%
    mutate(
      cv_sample = stringr::str_extract(p, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
      total     = stringr::str_extract(p, "(?<=total)(TRUE|FALSE)"),
      weighted  = stringr::str_detect(p, "weight"),
      control = if_else(str_detect(dx_recode, "^CN_"), "CN", "PT"),
      dx_tested  = str_remove(dx_recode, "^CN_")
    ) %>%
    select(!c(sexMale_x_logAge, sexMale, study_site, INDEX.ID, dx_recode))
})


length(cn_cent_df_list)
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")
cn_cent_df <- cn_cent_df_list %>%
  reduce(full_join) 

rm(cn_cent_df_list)
```

Case-Control results
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "cent_csvs/*cent_pt*_casecontrol_test.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")


cent_test.df <- bind_rows(output_fread, .id="file") %>%
  mutate(pheno = sapply(basename(file), function(f) {
           matched <- pheno_list[startsWith(f, pheno_list)]
           if (length(matched) > 0) matched[which.max(nchar(matched))] else NA_character_
         }),
         cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  select(!file) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo ??apply fdr correction within each model??
  mutate(
    dx_p.val_fdr = p.adjust(
      case.control_p.val,
      method = "fdr"
    ),
    case.control_d = -case.control_d) %>% #flipping sign so that neg effect means pts < CN
  ungroup() %>%
  mutate(dx_sig_fdr = ifelse(dx_p.val_fdr <=0.05, TRUE, FALSE))
```

lm test results
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "cent_csvs/*cent_pt*_lm.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")


cent_lm.df <- rbindlist(output_fread, .id="file") %>%
  mutate(pheno = stringr::str_extract(basename(file), "^[^_]+"),
          cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight"),
         dx = stringr::str_extract(basename(file), "(?<=pt)[^_]+")) %>%
  select(!file) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p.value,
      method = "fdr"
    )) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE))
```

## Non TBV-Corrected

```{r}
#full model
cent_test.df_uncor <- cent_test.df %>%
  filter(total == "FALSE" & weighted == "FALSE")

cent_test.df_uncor %>%
  ggplot() +
  geom_bar(aes(x=dx, fill=dx_sig_fdr)) +
  facet_grid(model~cv_sample)

cent_test.df_uncor %>%
  ggplot() +
  geom_bar(aes(x=pheno, fill=cv_sample)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~dx) +
  gghighlight(dx_sig_fdr == TRUE, calculate_per_facet = TRUE, unhighlighted_params = list(fill = NULL, alpha = 0.3))

uncor_d_plt <- cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=pheno, y= case.control_d, fill=model), position="dodge") +
  theme(axis.text.x = element_blank()) +
  facet_grid(dx~cv_sample, scales="free_y") 
ggplotly(uncor_d_plt)

cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE & pheno %in% global_list) %>%
  ggplot() +
  geom_col(aes(x=dx, y= case.control_d, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~pheno)
```

```{r}
cent_plt_df <- cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE) %>%
  arrange(pheno_cat, pheno) %>%
  mutate(pheno_nested = interaction(pheno, pheno_cat, sep = "__"))

cent_plt_uncor <- ggplot(cent_plt_df) +
  geom_col(aes(x = pheno_nested, y = case.control_d, fill = model), position = "dodge") +
  guides(x = legendry::guide_axis_nested(
  key = key_range_auto(sep = "__"),
  levels_text = list(
    element_blank(),        # inner (pheno) - hide
    element_text(size = 8)  # outer (pheno_cat) - show
  )
)) +
  theme_linedraw() +
  theme(
    axis.ticks.x    = element_blank(),
    legend.position = "right",
    axis.title.x = element_blank()
  ) +
  scale_fill_discrete(name = "Sex-Effect", labels = c("Age-Varying", "Intercept Only")) +
  facet_grid(dx~cv_sample, scales = "free_y") +
  labs(y = "Cohen's D")
ggsave("figs/misc/cent_test_col_uncorrected.png", cent_plt_uncor, dpi = 600)
```

```{r, eval=FALSE}
dx_levels <- unique(cent_test.df_uncor$dx)

for (dx_name in dx_levels) {
  
  # Filter and rename
  dx.df_uncor <- cent_test.df_uncor %>%
    filter(dx == dx_name) %>%
    dplyr::rename(sig_fdr = dx_sig_fdr)
  
  # Compute symmetric range for color scale
  d_range <- range(dx.df_uncor$case.control_d, na.rm = TRUE)
  sym_range <- c(-max(abs(d_range)), max(abs(d_range)))
  
  # Split by model
  dx.df_full <- dx.df_uncor %>% filter(model == "full")
  dx.df_null <- dx.df_uncor %>% filter(model == "null")
  
  # Plot
  plt_full <- plot_cv_brain(dx.df_full, case.control_d, "Cohen's D", "ggthemes::Gold-Purple Diverging", sym_range, c(1.03, 4.5, .93))
  plt_null <- plot_cv_brain(dx.df_null, case.control_d, "Cohen's D", "ggthemes::Gold-Purple Diverging", sym_range, c(1.03, 4.5, .93))
  
  # Save
  dx_lower <- tolower(dx_name)
  ggsave(paste0("figs/misc/", dx_lower, "_full_uncorrected.png"),  plt_full, dpi = 600, height = 8.5, width = 7.1, bg = "white")
  ggsave(paste0("figs/misc/", dx_lower, "_null_uncorrected.png"),  plt_null, dpi = 600, height = 8.5, width = 7.1, bg = "white")
  
  message("Saved plots for: ", dx_name)
}
```

**need to add significance markers to boxplots!

```{r}
cent_df_uncor <- cent_df %>%
  filter(total == "FALSE" & weighted == "FALSE")

cent_df_full_long <- cent_df_uncor %>%
  select(logAge_days, control, dx_tested, sex, ends_with("std_score_full")) %>%
  pivot_longer(cols = ends_with("std_score_full"),
    names_to = "pheno",
    names_pattern = "^(.*)_std_score_full$",
    values_to = "std_score_full") %>%
  na.omit(std_score_full)

ggplot(cent_df_full_long) +
  geom_boxplot(aes(x=pheno, y=std_score_full, fill=control)) +
  facet_wrap(~dx_tested)

#quick and dirty filtering for significance
cent_test.df_uncor2 <- cent_test.df_uncor %>%
  mutate(dx_pheno = paste(dx,pheno, sep="_"))

sig_dx_pheno <- cent_test.df_uncor2 %>%
  filter(dx_sig_fdr==TRUE) %>%
  pull(dx_pheno)

cent_df_full_long %>%
   mutate(dx_pheno = paste(dx_tested,pheno, sep="_")) %>%
  filter(dx_pheno %in% sig_dx_pheno) %>%
ggplot() +
  geom_boxplot(aes(x=dx_tested, y=std_score_full, fill=control)) +
  facet_wrap(~pheno)
```

patient sex differences

```{r}
cent_test.df_uncor %>%
  ggplot() +
  geom_bar(aes(x=dx, fill=pt.sex_sig_fdr)) +
  facet_grid(model~cv_sample)

cent_test.df_uncor %>%
  filter(pt.sex_sig_fdr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=pheno, y= pt.sex_d, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~dx)
```

properly accounting for sex -> fewer differences between male and female patients

```{r}
cent_lm.df %>%
  filter(str_detect(term, ":")) %>%
  filter(total==FALSE & sig_fdr==TRUE) %>%
  ggplot() +
  geom_col(aes(x=pheno, y=estimate, fill=cv_sample), position="dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~dx)
```

```{r}
cent_df_full_long %>%
  filter(control== "PT") %>%
  ggplot() +
  geom_boxplot(aes(x=pheno, y=std_score_full, fill=sex)) +
  facet_wrap(~dx_tested)

sample_sum <- cent_df_full_long %>%
  filter(control== "PT") %>%
  group_by(dx_tested, pheno, sex) %>%
  summarise(n=n())

ggplot(sample_sum) +
  geom_col(aes(x=pheno, y=n, fill=sex), position="dodge") +
  facet_wrap(~dx_tested)
```

## TBV-Corrected

```{r}
#full model
cent_test.df_tbv <- cent_test.df %>%
  filter(total == "TRUE" & weighted == "FALSE")

cent_test.df_tbv %>%
  ggplot() +
  geom_bar(aes(x=dx, fill=dx_sig_fdr)) +
  facet_grid(model~cv_sample)

cent_test.df_tbv %>%
  ggplot() +
  geom_bar(aes(x=pheno, fill=cv_sample)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~dx) +
  gghighlight(dx_sig_fdr == TRUE, calculate_per_facet = TRUE, unhighlighted_params = list(fill = NULL, alpha = 0.3))

cent_test.df_tbv %>%
  filter(dx_sig_fdr == TRUE & model != "null") %>%
  ggplot() +
  geom_col(aes(x=pheno, y= case.control_d, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_blank()) +
  facet_grid(dx~model, scales="free_y") 
```

```{r}
cent_df_tbv <- cent_df %>%
  filter(total == "TRUE" & weighted == "FALSE")

cent_df_tbv_full_long <- cent_df_tbv %>%
  select(logAge_days, control, dx_tested, sex, ends_with("std_score_full")) %>%
  pivot_longer(cols = ends_with("std_score_full"),
    names_to = "pheno",
    names_pattern = "^(.*)_std_score_full$",
    values_to = "std_score_full") %>%
  na.omit(std_score_full)

ggplot(cent_df_tbv_full_long) +
  geom_boxplot(aes(x=pheno, y=std_score_full, fill=control)) +
  facet_wrap(~dx_tested)
```

patient sex differences

```{r}
cent_test.df_tbv %>%
  ggplot() +
  geom_bar(aes(x=dx, fill=pt.sex_sig_fdr)) +
  facet_grid(model~cv_sample)

cent_test.df_tbv %>%
  filter(pt.sex_sig_fdr == TRUE & model != "null") %>%
  ggplot() +
  geom_col(aes(x=pheno, y= pt.sex_d, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~dx)

cent_test.df_tbv %>%
  filter(model != "null") %>%
  ggplot() +
  geom_col(aes(x=pheno, y= pt.sex_d, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~dx) +
  gghighlight(pt.sex_sig_fdr == TRUE, calculate_per_facet = TRUE, unhighlighted_params = list(fill = NULL, alpha = 0.3))

```

properly accounting for sex -> fewer differences between male and female patients
```{r}
cent_df_tbv_full_long %>%
  filter(control== "PT") %>%
  ggplot() +
  geom_boxplot(aes(x=pheno, y=std_score_full, fill=sex)) +
  facet_wrap(~dx_tested)
```

```{r}
sample_sum_tbv <- cent_df_tbv_full_long %>%
  filter(control== "PT") %>%
  group_by(dx_tested, pheno, sex) %>%
  summarise(n=n())

ggplot(sample_sum_tbv) +
  geom_col(aes(x=pheno, y=n, fill=sex), position="dodge") +
  facet_wrap(~dx_tested)
```

*power analyses
