---
title: "Review Centile Tests"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, ggpubr, cowplot, EnvStats, binom)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
#just looking at non-TBV-corrected results for now
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/dx_tests"

dx_levels <- c("SCZ", "ALZ", "ASD", "MDD", "GAD", "ADHD")

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue")
library(ggsegTissue)
#remotes::install_github("ggseg/ggsegExtra")
source("code/plot_cv_brain.R")
```

#Read & Prep CSVs

just visualize global centiles for now

```{r}
#CASES
paths <- Sys.glob(file.path(test_path, "*mean_scores.csv")) %>% unlist()

#CONTROLS
glob_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*global*totalFALSE*/"
# Use Sys.glob to expand wildcard paths
cn_paths <- Sys.glob(file.path(glob_path, "cent_csvs/*_*_*cent.csv")) %>%
  grep("CN", ., value = TRUE)
# Read all matching CSVs
cent_df_list <- lapply(c(paths, cn_paths), function(f){
  fread(f) %>%
    select(any_of(c("dx_recode", "dx_tested", "sex", "sexMale")), 
           ends_with("std_score_full")) %>%
    mutate(control = ifelse(grepl("CN", dx_recode), "CN", "PT"),
           dx_tested  = sub("CN_", "", dx_recode)) %>%
    pivot_longer(cols = ends_with("std_score_full"),
    names_to = "pheno",
    names_pattern = "^(.*)_std_score_full$",
    values_to = "std_score_full") %>%
  filter(pheno %in% global_list) %>% #just keep global pheno's z scores from full model for now
  na.omit(std_score_full)
})

length(cent_df_list)
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")
cent_df <- rbindlist(cent_df_list, fill=TRUE)
rm(cent_df_list)
```


Case-Control results
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "*total*_casecontrol_test.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths


cent_test.df_full <- bind_rows(output_fread, .id="file") %>%
  mutate(dx_tested = stringr::str_extract(basename(file), "[^_]+"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         pi = n_cn/(n_cn+n_pt),
         #flipping sign so that neg effect means pts < CN
         d = -d) %>%
  select(!c(contains("centile"), "file")) %>% #just looking at z score stats
  separate(variable, into = c("pheno", "model"), sep = "_std_score_") %>%
  #update pheno naming for plotting
  mutate(pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(model, total) %>% #apply FDR correction within each model/total
  mutate(
    dx_p.val_fdr = p.adjust(
      p_value,
      method = "fdr"
    )) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(
    dx_sig_fdr = ifelse(dx_p.val_fdr <= 0.05, TRUE, FALSE),
    case.control_resi = d2S(d = d, pi = pi)
  ) %>%
  #get back sign from d - pending approval from simon
  mutate(case.control_resi = ifelse(d < 0, -case.control_resi, case.control_resi)) %>%
  ungroup()
```

##Case-Control Diffs

```{r}
#just non-tbv-corrected
cent_test.df_uncor <- cent_test.df_full %>%
  filter(total==FALSE)

cent_test.df_uncor %>%
  ggplot() +
  geom_bar(aes(x=dx_tested, fill=dx_sig_fdr)) +
  facet_wrap(~model)
  
uncor_d_plt <- cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=pheno, y= d, fill=model), position="dodge") +
  theme(axis.text.x = element_blank()) +
  facet_wrap(~dx_tested, scales="free_y") 
#ggplotly(uncor_d_plt)

cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE & pheno %in% global_list) %>%
  ggplot() +
  geom_col(aes(x=dx_tested, y= case.control_resi), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~pheno)

cent_test.df_uncor %>% 
  mutate(effect_dir = sign(d)) %>%
  filter(dx_sig_fdr==TRUE & model == "full") %>%
  group_by(dx_tested, pheno_cat, effect_dir) %>%
  summarise(n=n())
```


```{r}
for (dx_name in dx_levels) {

  # Filter and rename
  dx.df_uncor <- cent_test.df_uncor %>%
    dplyr::filter(dx_tested == dx_name & model !="diff") %>%
    dplyr::rename(sig_fdr = dx_sig_fdr) %>%
    mutate(model = case_when(
      model == "full" ~ "Sex-Moderated Model",
      model == "null" ~ "Sex-Intercept Model"
    ))

  # Compute symmetric range for color scale
  d_range <- range(dx.df_uncor$d, na.rm = TRUE)
  sym_range <- c(-max(abs(d_range)), max(abs(d_range)))

  # Build plot
  dx_plt <- plot_cv_brain(df = dx.df_uncor,
                          fill_var = case.control_resi,
                          fill_name = "S",
                          fill_color = "ggthemes::Gold-Purple Diverging",
                          fill_limits = sym_range,
                          facet_var = "model",
                          height_list = c(1.03, 4.5, .93),
                          direction = -1,
                          keep_levels = TRUE,
                          return_plt = TRUE)
  
  #add title
  title <- ggdraw() + 
  draw_label(
    paste(dx_name, "effects on z-score"),
    x = 0,
    hjust = 0
  ) +
  theme(
    # add margin on the left of the drawing canvas,
    # so title is aligned with left edge of first plot
    plot.margin = margin(0, 0, 0, 7)
  )
  
dx_plt <- plot_grid(
  title, dx_plt,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.1, 1)
)

  # Save
  dx_lower <- tolower(dx_name)
  ggplot2::ggsave(
    paste0("figs/supplement/", dx_lower, "_totalFALSE_ttest_brainplt.png"),
    dx_plt, dpi = 600, height=9.35, width=7.1, bg = "white"
  )

  message("Saved plots for: ", dx_name)
}
```

brain plots across dx for just full model

```{r}
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

dx_all <- cent_test.df_uncor %>%
  dplyr::filter(model == "full") %>%
  filter(!(pheno %in% vent_list)) %>%
  dplyr::rename(sig_fdr = dx_sig_fdr)

# Compute symmetric range for color scale
range(dx_all$case.control_resi, na.rm = TRUE)
sym_range <- c(-.5, .5)

dx_brain_plt <- plot_cv_brain(df = dx_all, 
                          fill_var = case.control_resi, 
                          fill_name = "S", 
                          fill_color = "ggthemes::Gold-Purple Diverging", 
                          fill_limits = sym_range,
                          facet_var= "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE, 
                          include_global = FALSE,
                          keep_levels = TRUE,
                          direction = -1,
                          rel_widths = c(6,.5))

saveRDS(dx_brain_plt, "figs/ggobjs/dx_totalFALSE_brain_plt.rds")
```

boxplot

```{r}
####get sig plots
y_pos <- cent_df %>%
  group_by(dx_tested, pheno) %>%
  summarise(y_pos = max(std_score_full, na.rm = TRUE) + 0.6, .groups = "drop")

pvals <- cent_test.df_uncor %>%
  filter(pheno %in% global_list) %>%
  filter(model=="full") %>%
  mutate(dx_sig_fdr = ifelse(dx_sig_fdr==TRUE, "*", NA))

pvals <- left_join(pvals, y_pos, by = c("dx_tested", "pheno"))

dx_bxplt <- ggplot(cent_df, aes(x=pheno, y=std_score_full, color=control)) +
  geom_hline(yintercept = 0, color = "darkgray", linetype = "dashed") +
  geom_boxplot(outlier.size = .5) +
  geom_text(
    data = pvals,
    aes(x = pheno, y = y_pos, label = dx_sig_fdr),
    inherit.aes = FALSE,
    size = 4
  ) +
  facet_wrap(~dx_tested, nrow=2) +
  scale_x_discrete(labels = c("total.SA" = "Total SA", 
                              "mean.CT" = "Mean CT",
                              "CBV" = "Cerebellum",
                              "CSF" = "CSF",
                              "GMV" = "Cortical GM",
                              "sGMV" = "Subcortical GM",
                              "WMV" = "WM")) +
  scale_color_manual(
  values = c('PT' = "#D9B523FF", 'CN' = "#6C568CFF"),
  labels = c("CN" = "Control", "PT" = "Patient"),
  name = NULL
) +
  theme_linedraw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        axis.title.x=element_blank(),
        legend.position = "bottom") +
  labs(y="Centile Z-score")

saveRDS(dx_bxplt, "figs/ggobjs/dx_totalFALSE_bxplt.rds")
```

###Total-Size Corrected

save csv with size corrected and uncorrected results

```{r}
cent_test_csv <- cent_test.df_full %>%
  filter((total == FALSE & model != "diff") | (total == TRUE & model == "full")) %>%
  select(total, model, dx_tested, pheno_cat, pheno, t_stat, d, case.control_resi, mean_df1, mean_df2, n_cn, n_pt, p_value, dx_p.val_fdr, dx_sig_fdr) %>%
  rename(total_size_corrected = total,
         idp_category = pheno_cat,
         idp = pheno,
         p.val = p_value,
         significant_fdr = dx_sig_fdr,
         mean_zscore_control = mean_df1,
         mean_zscore_case = mean_df2,
         n_control = n_cn,
         n_case = n_pt,
         p.val_fdr = dx_p.val_fdr
         ) %>%
  mutate(model = ifelse(model=="full", "sex_with_age_moderator", "sex_intercept_only"))

fwrite(cent_test_csv, "figs/supplement/case_control_ttests.csv")
```
Also plot across dx

brain plots across dx for just full model

```{r}
dx_all <- cent_test.df_full %>%
  dplyr::filter(model == "full" & total == TRUE) %>%
  dplyr::rename(sig_fdr = dx_sig_fdr)

# Compute symmetric range for color scale
range(dx_all$case.control_resi, na.rm = TRUE)
sym_range <- c(-.45, .45)

dx_brain_plt_total <- plot_cv_brain(df = dx_all, 
                          fill_var = case.control_resi, 
                          fill_name = "S", 
                          fill_color = "ggthemes::Gold-Purple Diverging", 
                          fill_limits = sym_range,
                          facet_var= "dx_tested",
                          return_plt = TRUE, 
                          include_global = TRUE,
                          keep_levels = TRUE,
                          direction = -1,
                          rel_widths = c(6,.5))

ggsave("figs/supplement/dx_totalTRUE_brain_plt.png", dx_brain_plt_total, width=17, height=8, dpi=600, bg = "white")
```


###Change

How do results differ across full and null models?

```{r}
cent_test_change <- cent_test.df_uncor %>%
  filter(model != "diff") %>%
  group_by(dx_tested, pheno_cat, pheno, label) %>% 
  #first make sure model converged in both
  filter(n() == 2) %>%
  #then filter to significant vals
  filter(dx_sig_fdr == "TRUE") %>%
  summarise(
    sig_mod = case_when(
      n() == 1 ~ model,                          # significant in only one model
      n() == 2 & length(unique(sign(d))) > 1 ~ "opposite",  # both sig, opposite signs
      n() == 2 ~ "both"                          # both sig, same sign
    ),
    .groups = "drop"
  ) %>%
  mutate(sig_mod = factor(sig_mod, levels = c("both", "opposite", "full", "null")))

table(cent_test_change$sig_mod)

cent_disagree_plt <- cent_test_change %>%
  filter(sig_mod != "both") %>%
  ggplot() +
    geom_bar(aes(x=pheno_cat, fill=sig_mod)) +
    scale_fill_manual(labels = c("full" = "Sex-Moderated", "null" = "Sex-Intercept Only"),
                        values = c("full" = "#216936FF",
                                   "null" = "#A8BFAAFF")) +
    facet_wrap(~ dx_tested, nrow=1) +
    theme_linedraw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), legend.position = "bottom") +
    labs(y="Incongruous IDPs", x="", fill="Brain Chart \nModel")
saveRDS(cent_disagree_plt, "figs/ggobjs/cent_disagree_bar_plt.rds")
```

Male-biased disorders -> false positives, female-biased disorders -> false negatives? severe disorders with prominent neuroanatomical effects are well-captured by both models

plot on brains

```{r}
cent_test_change$sig_fdr <- TRUE #all sig by definition

dx_brain_change_plt <- plot_cv_brain(df = cent_test_change, 
                          fill_var = sig_mod, 
                          fill_name = "Model", 
                          fill_color=NULL,
                          fill_values = c("both" = "#393470", 
                                         "full" = "#216936FF" , 
                                        "null" = "#A8BFAAFF"),
                          fill_limits = c("both", "full", "null"),
                          facet_var= "dx_tested",
                          keep_levels = TRUE,
                          return_plt = TRUE, 
                          include_global = TRUE,
                          rel_widths = c(6,.5)) 
ggsave("figs/supplement/dx_brain_change_plt.png", dx_brain_change_plt, width=17, height=8, dpi=600, bg = "white")
```


##Exteremeness
Are patients more or less likely to have extreme centiles?

```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "*total*_extcent_fisher.csv"))

# Read all matching CSVs
ext_cent_fish <- lapply(paths, fread)
names(ext_cent_fish) <- paths

ext_cent_fish_df <- rbindlist(ext_cent_fish, idcol = "file") %>%
  mutate(total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)")) %>%
  group_by(total, model) %>%
  mutate(
    p_fdr = p.adjust(
      fisher_p,
      method = "fdr"
    )) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p_fdr <=0.05, TRUE, FALSE),
         pct_CN = (n_CN/n_total_CN)*100,
         pct_PT = (n_PT/n_total_PT)*100,
         pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  mutate(pct_diff = pct_PT - pct_CN) #pos -> cases overrepresented, neg -> cases under-represented

table(ext_cent_fish_df$sig_fdr, ext_cent_fish_df$dx_tested)

ext_cent_fish_df %>%
  filter(sig_fdr==TRUE) %>%
  group_by(total, model, dx_tested) %>%
  dplyr::count()
```

```{r}
ext_cent_fish_df %>%
  filter(sig_fdr==TRUE & total==FALSE & model == "full") %>%
  ggplot() +
  geom_histogram(aes(x=pct_diff, fill=dx_tested), position="identity", alpha=.5) +
  facet_wrap(~ext)
```

plot on brains

```{r}
for (dx_name in dx_levels) {

  # Filter and rename
  dx.df_uncor <- ext_cent_fish_df %>%
    dplyr::filter(dx_tested == dx_name & total == FALSE)

  # Compute symmetric range for color scale
  d_range <- range(dx.df_uncor$pct_diff, na.rm = TRUE)
  sym_range <- c(-max(abs(d_range)), max(abs(d_range)))

  # Split by model
  dx.df_full <- dx.df_uncor %>% dplyr::filter(model == "full")
  dx.df_null <- dx.df_uncor %>% dplyr::filter(model == "null")

  # Build plot components (without auto-assembling full figure)
  full_parts <- plot_cv_brain(df = dx.df_full,
                          fill_var = pct_diff,
                          fill_name = "Percent\nDiff",
                          fill_color = "grDevices::Blue-Red 3",
                          fill_limits = sym_range,
                          height_list = c(1.03, 4.5, .93),
                          facet_var = "ext",
                          keep_levels = TRUE,
                          return_plt = FALSE)

  null_parts <- plot_cv_brain(df = dx.df_null,
                          fill_var = pct_diff,
                          fill_name = "Percent\nDiff",
                          fill_color = "grDevices::Blue-Red 3",
                          fill_limits = sym_range,
                          height_list = c(1.03, 4.5, .93),
                          facet_var = "ext",
                          return_plt = FALSE,
                          keep_levels = TRUE
  )

  # Helper to assemble one brain panel (without legend) + visible title
  assemble_panel <- function(parts, title_text, heights = c(1.03, 4.5, .93)) {
    panel <- cowplot::plot_grid(
      parts[[1]], parts[[2]], parts[[3]],
      ncol = 1, nrow = 3,
      rel_heights = heights,
      align = "v", axis = "lr", greedy = TRUE
    )

    title_plot <- cowplot::ggdraw() +
      cowplot::draw_label(
        title_text,
        x = 0.5, y = 0.5, hjust = 0.5, vjust = 0.5,
        fontface = "bold", size = 13
      )

    cowplot::plot_grid(
      title_plot, panel,
      ncol = 1,
      rel_heights = c(0.12, 1)
    )
  }

  plt_full_titled <- assemble_panel(
    full_parts,
    paste(dx_name, "effects from sex-moderated model")
  )

  plt_null_titled <- assemble_panel(
    null_parts,
    paste(dx_name, "effects from sex-intercept model")
  )

  # Use one shared legend (same scale, so either legend works)
  shared_legend <- full_parts[[4]]

  # Combine both panels + single legend on right
  dx_plt <- cowplot::plot_grid(
    cowplot::plot_grid(plt_full_titled, plt_null_titled, ncol = 2),
    shared_legend,
    ncol = 2,
    rel_widths = c(12, 1)
  )

  # Save
  dx_lower <- tolower(dx_name)
  ggplot2::ggsave(
    paste0("figs/supplement/", dx_lower, "_totalFALSE_ext_brainplt.png"),
    dx_plt, dpi = 600, height = 8.5, width = 12, bg = "white"
  )

  message("Saved plots for: ", dx_name)
  
}
```

brainplots across dx

```{r}
dx_all <- ext_cent_fish_df %>%
  dplyr::filter(model == "full" & total == FALSE) %>%
  mutate(dx_ext = paste(dx_tested, ext, sep="_"))

# Compute symmetric range for color scale
range(dx_all$pct_diff, na.rm = TRUE)
sym_range <- c(-32.1, 32.1)

dx_high <- dx_all %>% filter(ext=="high")
dx_low <- dx_all %>% filter(ext=="low")

high_plt <- plot_cv_brain(df = dx_high, 
                          fill_var = pct_diff, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE, 
                          keep_levels = TRUE,
                          include_global = TRUE,
                          rel_widths = c(6,.5))
upper_combo <- plot_grid(
  high_plt[[1]], high_plt[[2]], high_plt[[3]],
  ncol = 1,
  align = "v",
  rel_heights = c(1.03, 4.5, .93)
)

upper_combo_labeled <- ggdraw() +
  draw_plot(upper_combo, x = 0.02, y = 0, width = 0.97, height = 1) +
  draw_grob(
    grid::rectGrob(gp = grid::gpar(fill = "#EBEBEB", col = NA)),
    x = 0, y = 0.01, width = 0.02, height = 0.95
  ) +
  draw_label("Centiles > 95%", x = 0.01, y = 0.5, angle = 90, size = 10)

low_plt <- plot_cv_brain(df = dx_low, 
                          fill_var = pct_diff, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE,
                         keep_levels = TRUE,
                          include_global = TRUE,
                          rel_widths = c(6,.5))

lower_combo <- plot_grid(
  low_plt[[1]], low_plt[[2]], low_plt[[3]],
  ncol = 1,
  align = "v",
  rel_heights = c(1.03, 4.5, .93)
)

lower_combo_labeled <- ggdraw() +
  draw_plot(lower_combo, x = 0.02, y = 0, width = 0.97, height = 1) +
  draw_grob(
    grid::rectGrob(gp = grid::gpar(fill = "#EBEBEB", col = NA)),
    x = 0, y = 0.01, width = 0.02, height = 0.95
  ) +
  draw_label("Centiles < 5%", x = 0.01, y = 0.5, angle = 90, size = 10)

ext_brain_plt <- plot_grid(
  plot_grid(
  upper_combo_labeled,
  lower_combo_labeled,
  ncol=1,
  align="v"),
  low_plt[[4]],
  ncol=2,
  rel_widths=c(10,1)
)

ggsave("figs/supplement/extreme_dx_totalFALSE_brainplt.png", ext_brain_plt, width=18, height=16, dpi=600, bg = "white")
```

###Total-Size Corrected

save csv with size corrected and uncorrected results

```{r}
ext_test_csv <- ext_cent_fish_df %>%
  filter(total == FALSE | (total == TRUE & model == "full")) %>%
  select(total, model, dx_tested, ext, pheno_cat, pheno, pct_CN, pct_PT, pct_diff, fisher_stat, fisher_p, p_fdr, sig_fdr) %>%
  rename(idp_category = pheno_cat,
         idp = pheno,
         p.val = fisher_p,
         significant_fdr = sig_fdr,
         percent_control = pct_CN,
         percent_case = pct_PT,
         percent_diff = pct_diff,
         p.val_fdr = p_fdr,
         centile_group = ext,
         total_size_corrected = total
         ) %>%
  mutate(model = ifelse(model=="full", "sex_with_age_moderator", "sex_intercept_only"),
         centile_group = ifelse(centile_group=="high", "above_95th", "below_5th"))

fwrite(ext_test_csv, "figs/supplement/case_control_proportion_tests.csv")
```

plot

```{r}
dx_all <- ext_cent_fish_df %>%
  dplyr::filter(model == "full" & total == TRUE) %>%
  mutate(dx_ext = paste(dx_tested, ext, sep="_"))

# Compute symmetric range for color scale
range(dx_all$pct_diff, na.rm = TRUE)
sym_range <- c(-32.1, 32.1)

dx_high <- dx_all %>% filter(ext=="high")
dx_low <- dx_all %>% filter(ext=="low")

high_plt <- plot_cv_brain(df = dx_high, 
                          fill_var = pct_diff, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE, 
                          keep_levels = TRUE,
                          include_global = TRUE,
                          rel_widths = c(6,.5))
upper_combo <- plot_grid(
  high_plt[[1]], high_plt[[2]], high_plt[[3]],
  ncol = 1,
  align = "v",
  rel_heights = c(1.03, 4.5, .93)
)

upper_combo_labeled <- ggdraw() +
  draw_plot(upper_combo, x = 0.02, y = 0, width = 0.97, height = 1) +
  draw_grob(
    grid::rectGrob(gp = grid::gpar(fill = "#EBEBEB", col = NA)),
    x = 0, y = 0.01, width = 0.02, height = 0.95
  ) +
  draw_label("Centiles > 95%", x = 0.01, y = 0.5, angle = 90, size = 10)

low_plt <- plot_cv_brain(df = dx_low, 
                          fill_var = pct_diff, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE,
                         keep_levels = TRUE,
                          include_global = TRUE,
                          rel_widths = c(6,.5))

lower_combo <- plot_grid(
  low_plt[[1]], low_plt[[2]], low_plt[[3]],
  ncol = 1,
  align = "v",
  rel_heights = c(1.03, 4.5, .93)
)

lower_combo_labeled <- ggdraw() +
  draw_plot(lower_combo, x = 0.02, y = 0, width = 0.97, height = 1) +
  draw_grob(
    grid::rectGrob(gp = grid::gpar(fill = "#EBEBEB", col = NA)),
    x = 0, y = 0.01, width = 0.02, height = 0.95
  ) +
  draw_label("Centiles < 5%", x = 0.01, y = 0.5, angle = 90, size = 10)

ext_brain_plt <- plot_grid(
  plot_grid(
  upper_combo_labeled,
  lower_combo_labeled,
  ncol=1,
  align="v"),
  low_plt[[4]],
  ncol=2,
  rel_widths=c(10,1)
)

ggsave("figs/supplement/extreme_dx_totalTRUE_brainplt.png", ext_brain_plt, width=18, height=16, dpi=600, bg = "white")
```

###Change

How do results differ across full and null models?

```{r}
ext_test_change <- ext_cent_fish_df %>%
  filter(total == FALSE) %>%
  group_by(dx_tested, pheno_cat, pheno, label, ext) %>% 
  #first make sure model converged in both
  filter(n() == 2) %>%
  #then filter to significant vals
  filter(sig_fdr == "TRUE") %>%
  summarise(
    sig_mod = case_when(
      n() == 1 ~ model,                          # significant in only one model
      n() == 2 & length(unique(sign(pct_diff))) > 1 ~ "opposite",  # both sig, opposite signs
      n() == 2 ~ "both"                          # both sig, same sign
    ),
    .groups = "drop"
  ) %>%
  mutate(sig_mod = factor(sig_mod, levels = c("both", "opposite", "full", "null")))

table(ext_test_change$sig_mod)

ext_test_change %>%
  filter(sig_mod != "both") %>%
  group_by(dx_tested, ext, sig_mod) %>%
  dplyr::count()

ext_test_change %>%
  filter(sig_mod != "both") %>%
  ggplot() +
    geom_bar(aes(x=pheno_cat, fill=sig_mod)) +
        scale_fill_manual(labels = c("full" = "Sex-Moderated", "null" = "Sex-Intercept Only"),
                        values = c("full" = "#216936FF",
                                   "null" = "#A8BFAAFF")) +
    facet_grid(ext ~ dx_tested) +
    theme_linedraw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(y="Incongruous IDPs", x="", fill="Brain Chart Model")
```

plot on brains

```{r}
ext_test_change$sig_fdr <- TRUE

#add back NS phenos
lab_df <- ext_cent_fish_df %>% 
  select(pheno, pheno_cat, label, dx_tested) %>%
  distinct()

change_high <- ext_test_change %>% filter(ext=="high")
change_low <- ext_test_change %>% filter(ext=="low")

change_high_nas <- full_join(change_high, lab_df)
change_low_nas <- full_join(change_low, lab_df)

high_plt <- plot_cv_brain(df = change_high_nas, 
                          fill_var = sig_mod, 
                          fill_name = "Model", 
                          fill_color=NULL,
                           fill_values = c("both" = "#393470", 
                                         "full" = "#216936FF" , 
                                        "null" = "#A8BFAAFF"),
                          fill_limits = c("both", "full", "null"),
                          facet_var= "dx_tested",
                          return_plt = FALSE,
                          keep_levels = TRUE,
                          include_global = TRUE,
                          rel_widths = c(6,.5))
upper_combo <- plot_grid(
  high_plt[[1]], high_plt[[2]], high_plt[[3]],
  ncol = 1,
  align = "v",
  rel_heights = c(1.03, 4.5, .93)
)

upper_combo_labeled <- ggdraw() +
  draw_plot(upper_combo, x = 0.02, y = 0, width = 0.97, height = 1) +
  draw_grob(
    grid::rectGrob(gp = grid::gpar(fill = "#EBEBEB", col = NA)),
    x = 0, y = 0.01, width = 0.02, height = 0.95
  ) +
  draw_label("Centiles > 95%", x = 0.01, y = 0.5, angle = 90, size = 10)

low_plt <- plot_cv_brain(df = change_low_nas, 
                          fill_var = sig_mod, 
                          fill_name = "Model", 
                          fill_color=NULL,
                           fill_values = c("both" = "#393470", 
                                         "full" = "#216936FF" , 
                                        "null" = "#A8BFAAFF"),
                          fill_limits = c("both", "full", "null"),
                          facet_var= "dx_tested",
                          return_plt = FALSE, 
                          include_global = TRUE,
                          keep_levels = TRUE,
                          rel_widths = c(6,.5))

#lower plot
lower_combo <- plot_grid(
  low_plt[[1]], low_plt[[2]], low_plt[[3]],
  ncol = 1,
  align = "v",
  rel_heights = c(1.03, 4.5, .93)
)

lower_combo_labeled <- ggdraw() +
  draw_plot(lower_combo, x = 0.02, y = 0, width = 0.97, height = 1) +
  draw_grob(
    grid::rectGrob(gp = grid::gpar(fill = "#EBEBEB", col = NA)),
    x = 0, y = 0.01, width = 0.02, height = 0.95
  ) +
  draw_label("Centiles < 5%", x = 0.01, y = 0.5, angle = 90, size = 10)

ext_brain_plt_change <- plot_grid(
  plot_grid(
  upper_combo_labeled,
  lower_combo_labeled,
  ncol=1,
  align="v"),
  low_plt[[4]],
  ncol=2,
  rel_widths=c(10,1)
)

ggsave("figs/supplement/extreme_dx_brainplt_change.png", ext_brain_plt_change, width=18, height=16, dpi=600, bg = "white")
```
