---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, ggcorrplot, sf)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

#ggseg Prep

```{r}
### dk for cortical ###
#write out new dk atlas w/o medial wall
dk_clean <- dk
dk_clean$data <- dk_clean$data %>%
  dplyr::filter(!is.na(region) & region !="corpus callosum")

#shrink for outlines
dk_region_gap <- 0.3        # Space between regions
dk_border_inset <- 0.3      # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
dk_fill <- dk_clean
dk_fill$data <- dk_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -dk_region_gap) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
dk_border <- dk_clean
dk_border$data <- dk_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(dk_region_gap + dk_border_inset)) %>%
  as_tibble() %>%
  st_as_sf()

### aseg for subcortical ###
#shrink for outlines
aseg_region_gap <- 0        # Space between regions
aseg_border_inset <- 1e-5     # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
aseg_fill <- aseg
aseg_fill$data <- aseg$data %>%
  st_as_sf() %>%
  st_buffer(dist = 0, 
            joinStyle = "MITRE",
            nQuadSegs=500) %>%
  st_buffer(0) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
aseg_border <- aseg
aseg_border$data <- aseg$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(aseg_region_gap + aseg_border_inset), nQuadSegs=500) %>%
  as_tibble() %>%
  st_as_sf()

### aseg for global ###
aseg_all_clean <- aseg_all
aseg_all_clean$data <- aseg_all_clean$data %>%
  dplyr::filter(label != "Brain-Stem")

#shrink for outlines
glob_region_gap <- 1e-10        # Space between regions
glob_border_inset <- 1e-10      # Additional inset for border

# Create shrunk atlas for fills (creates gaps between regions)
glob_fill <- aseg_all_clean
glob_fill$data <- aseg_all_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -glob_region_gap) %>%
  as_tibble() %>%
  st_as_sf()

# Create further shrunk atlas for borders (inside the fill)
glob_border <- aseg_all_clean
glob_border$data <- aseg_all_clean$data %>%
  st_as_sf() %>%
  st_buffer(dist = -(glob_region_gap + glob_border_inset)) %>%
  as_tibble() %>%
  st_as_sf()
```

trying to write funs for easier plotting
```{r}
plot_cv_brain <- function(df, fill_var, fill_name, fill_color, fill_limits, height_list){

#cortex
cortex_plt <- df %>%
  filter(pheno_cat == "Regional Vol" | pheno_cat == "Regional SA" | pheno_cat == "Regional CT") %>%
  group_by(cv_sample, pheno_cat) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = dk_fill,
             position = position_brain(hemi ~ side),
            aes(fill={{fill_var}}),#
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = dk_border,
                    position = position_brain(hemi ~ side),
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)

subcort_plt <- df %>%
  filter(pheno_cat == "Subcortical Vol") %>%
  group_by(cv_sample, pheno_cat) %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = aseg_fill,
            aes(fill={{fill_var}}),
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = aseg_border,
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)


#no buffer for now
glob_plt <- df %>%
  select(-any_of(c("geometry", "tissue_class"))) %>%
  group_by(cv_sample, pheno_cat) %>%
  filter(pheno_cat == "Global Vol") %>%
  ggplot() +
  theme_linedraw(base_size=12) +
  #fill
  ggseg::geom_brain(atlas = aseg_all_clean,
            aes(fill={{fill_var}}), #
            color=NA,
            size=0) +
  #lines
  ggseg::geom_brain(atlas = aseg_all_clean,
                    aes(color = sig_fdr, size = sig_fdr),
                    fill = NA) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.text.x = element_blank(), plot.margin = margin(0, 0, 0, 0)) +
  paletteer::scale_fill_paletteer_c(fill_color, name = fill_name, limits = fill_limits, direction = -1, na.value="gray90") +
  scale_color_manual(values = c("FALSE" = "white",
                                "TRUE" = "black"),
                     na.value = NA,
                     guide = 'none') +
  scale_size_manual(values = c("FALSE" = .5,
                               "TRUE" = .2),
                    na.value = 0,
                    guide = 'none') +
  facet_grid(pheno_cat ~ cv_sample)

#testing new spacing
# glob_plt <- glob_plt +
#   theme(plot.margin = margin(0, 0, 5, 0))

cortex_plt <- cortex_plt +
  theme(plot.margin = margin(5, 0, 5, 0))

full_plt_obj <- glob_plt / cortex_plt / subcort_plt +
  plot_layout(guides = "collect", heights = height_list)

# full_plt_obj <- glob_plt /plot_spacer()/ cortex_plt / plot_spacer()/ subcort_plt +
#   plot_layout(guides = 'collect', heights = unit(height_list, c('in', 'in', 'in')))

return(full_plt_obj)
}
```

#Defining integral funs

```{r}
#fun for derivative
get_deriv <- function(x, age) {
  c(NA, diff(x) / diff(age))
}

#riemann_stieltjes
rs_integral <- function(x, y, g = function(x) log10(x)) {
  stopifnot(
    length(x) == length(y),
    length(x) >= 2
  )

  # sort just in case
  ord <- order(x)
  x <- x[ord]
  y <- y[ord]

  gx <- g(x)

  # monotonicity check (important!)
  if (any(diff(gx) < 0)) {
    stop("g(x) must be non-decreasing for a Riemann–Stieltjes integral")
  }

  sum(y[-length(y)] * diff(gx))
}

#euclidean distance fun from https://www.geeksforgeeks.org/r-language/how-to-calculate-euclidean-distance-in-r/
euclidean_dist <- function(vect1, vect2) sqrt(sum((vect1 - vect2)^2))
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")


LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  select(!file)
```

```{r, eval=FALSE}
LR.df %>%
  group_by(cv_sample, total, weighted) %>%
  summarise(n_phenos = n_unique(pheno))

A_totalFALSE <- LR.df %>%
  filter(cv_sample == "A" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalFALSE)

A_totalTRUE <- LR.df %>%
  filter(cv_sample == "A" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalTRUE)

B_totalFALSE <- LR.df %>%
  filter(cv_sample == "B" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalFALSE)

B_totalTRUE <- LR.df %>%
  filter(cv_sample == "B" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalTRUE)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
         pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
)

table(LR.df2$effect, LR.df2$total)
```



```{r}
# sub_list
# unique(aseg$data$label)

LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$tissue_class), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas

check scales for plotting

```{r}
LR.df_plot %>%
  group_by(effect) %>%
  summarise(max_resi = max(resi))
```


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE)

LR.df_uncor %>%
  group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n()) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_uncor_novent <- LR.df_uncor %>%
  filter(!(pheno %in% vent_list))
range(LR.df_uncor_novent$resi)
full_plt_uncor <- plot_cv_brain(LR.df_uncor_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.1), c(1, 5, 1.1))

ggsave("figs/resi_uncorrected.png", full_plt_uncor, width=7, height=8.5, dpi=600)
```

#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE)

table(LR.df_tot$sig_fdr)

LR.df_tot %>%
group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n()) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_tot_novent <- LR.df_tot %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_novent$resi)
full_plt_tot <- plot_cv_brain(LR.df_tot_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.1), c(1, 5, 1.1))

ggsave("figs/resi_tbv.png", full_plt_tot, width=7, height=8.5, dpi=600)
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE)

table(LR.df_tot_sex$sig_fdr, LR.df_tot_sex$cv_sample)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)

LR.df_tot_sex %>%
  filter(sig_fdr == TRUE) %>%
  group_by(cv_sample) %>%
  summarise(min_resi = min(sig_resi),
            max_resi = max(sig_resi),
            mean_resi = mean(sig_resi))
```

```{r}
#removing ventricles for visualization:
LR.df_tot_sex_novent <- LR.df_tot_sex %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_sex_novent$resi)
full_plt_tot_sex <- plot_cv_brain(LR.df_tot_sex_novent, resi, "RESI", "grDevices::Greens 3", c(0,0.2), c(1, 5, 1.1))

ggsave("figs/resi_tbv_allsex.png", full_plt_tot_sex, width=7, height=8.5, dpi=600)
```


##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

sexdiffs.df_all <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         pheno = stringr::str_remove(basename(file), "_(uniform_)?sexdiffs\\.csv")) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    age_days = un_log(logAge_days))

#split into regular df and uniform df for correlations
sexdiffs.df_uniform <- sexdiffs.df_all[grepl("uniform", sexdiffs.df_all$file), ]
sexdiffs.df <- sexdiffs.df_all[!grepl("uniform", sexdiffs.df_all$file), ]

rm(output_fread_sd) #save space
rm(sexdiffs.df_all)
```

Fix scaling, derivatives

```{r}
sexdiffs.df_sc <- sexdiffs.df %>%
  group_by(total, weighted, cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    #tmp fix for sigma as CV, getting ln CV ratio
    logcv_M_div_F = log(sigma_Male/sigma_Female),
    #chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age)
```

Plot

```{r}
sexdiffs_tbv <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==TRUE)

sexdiffs <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==FALSE)

for (df in list(sexdiffs_tbv, sexdiffs)){
  t <- as.character(unique(df$total))
  print(t)
  
  #median centile
  cent_df <- df %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  cent_plt <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
  n <- paste0("figs/sexdiffs_tbv",t,"_cent.png")
  ggsave(n, cent_plt)
  print(cent_plt)
  
  cent_plt2 <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Scaled 50th Centile Sex-Difference")
  
  print(cent_plt2)
  n <- paste0("figs/sexdiffs_tbv",t,"_cent_scaled.png")
  ggsave(n, cent_plt2)
  #ggplotly(cent_plt)

  #sigma
  sig_df <- df %>%
    select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  sig_plt <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="log CV Ratio")
  print(sig_plt)
  n <- paste0("figs/sexdiffs_tbv",t,"_cv.png")
  ggsave(n, sig_plt)
  
  #derivatives
  deriv_df <- df %>%
    select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  deriv_plt <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="50th Centile Difference Derivative")
  print(deriv_plt)
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv.png")
  ggsave(n, deriv_plt)
  
  deriv_plt2 <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="Scaled 50th Centile Difference Derivative")
  print(deriv_plt2)
  
  n <- paste0("figs/sexdiffs_tbv",t,"_deriv_scaled.png")
  ggsave(n, deriv_plt2)
  #ggplotly(deriv_plt)
}
```

compare derivatives with chain rule
```{r}
sexdiffs_tbv %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("log_lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="50th Centile Difference Derivative")

sexdiffs_tbv %>%
  #filter(pheno_cat == "Subcortical Vol" & age_days > (365.25*40)) %>%
    ggplot() +
    geom_line(aes(x=age_days, y=deriv_M_minus_F_z.unlog, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("lifespan_fetal", sexdiffs_tbv$age_days) +
    labs(y="50th Centile Difference Derivative")
```

###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

Bootstrap!

```{r}
#calculate integral of deriv of diff^2 (all in z-scores)
sexdiffs_tbv_sq <- sexdiffs_tbv %>%
  mutate(sq_centile_diff_z = (centile_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_deriv_z = get_deriv(sq_centile_diff_z, logAge_days)) %>%
  mutate(sq_deriv_z.unlog = sq_deriv_z*deriv_age) %>% #chain rule to convert derivative
  ungroup()


#CHECK THAT THESE PLOTS MAKE SENSE!!
sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv_z, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv_sq$logAge_days) +
    labs(y="Change in Magnitude Sex Diff")

sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=age_days, y=sq_deriv_z.unlog, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("lifespan_fetal", sexdiffs_tbv_sq$age_days) +
    labs(y="Change in Magnitude Sex Diff")

#try getting integral
diff_int <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv_z),
            sq_int_z.unlog = rs_integral(logAge_days, sq_deriv_z))
```


try finding points at which pt derivative crosses 0 increasing vs decreasing

```{r}
cent_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv_z.unlog > 0 ~ "diverging",
    sq_deriv_z.unlog == 0 ~ "constant",
    sq_deriv_z.unlog < 0 ~ "converging",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- cent_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv_z, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

cent_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", cent_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)
```
Get max male and female biases and plot on centile differences

```{r}
peak_diffs_sum <- cent_df %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(centile_M_minus_F_z, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(centile_M_minus_F)),
            female_peak = min_gc(centile_M_minus_F_z, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(centile_M_minus_F))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")


peak_diffs_sum %>%
  ggplot() +
  geom_histogram(aes(x=peak_age, fill=sex), position="identity", alpha=.4) +
  format_x_axis("log_lifespan_fetal", peak_diffs_sum$peak_age) +
  facet_grid(pheno_cat~cv_sample, scales="free_y")

cent_df %>%
    #filter(pheno %in% consistent_agesex) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% consistent_agesex), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
```

differences tend to peak around beginning and end of life - likely d/t smaller sample sizes at extreme ages

plot convergence/divergence on brain

```{r}
#removing ventricles for visualization:
diff_int_novent <- full_join(LR.df_tot_sex, diff_int) %>%
  filter(!(pheno %in% vent_list))
range(diff_int_novent$sq_int_z.unlog)

full_converge_plt <- plot_cv_brain(diff_int_novent, sq_int_z.unlog, "Change in \nMagnitude \nSex Diff", "ggthemes::Orange-Blue Diverging", c(-.21, .21), c(2, 5, 2))

ggsave("figs/convergence_tbvTRUE.png", full_converge_plt, width=7, height=8, dpi=600)
```

####Sigma

*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of lnCVR squared: (not sure this totally makes sense tbh)
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#quick dirty code to get abs diff
sigma_df <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==TRUE) %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_cv_diff = (logcv_M_div_F)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_cv_deriv = get_deriv(sq_cv_diff, logAge_days)) %>%
  ungroup() %>%
  mutate(sq_cv_deriv.unlog = sq_cv_deriv*deriv_age)

sigma_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_cv_deriv, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sigma_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Variability")
```

###Cross-sample Consistency
Get across-sample euclidean distance
```{r}
dist_df <- data.frame()
for (pheno_val in pheno_list){
  df <- sexdiffs_tbv %>%
    filter(pheno==pheno_val) %>%
    arrange(logAge_days)
  
  dfA <- df %>%
    filter(cv_sample=="A")
  
  dfB <- df %>%
    filter(cv_sample=="B")
  
  if (nrow(dfA) == 0 | nrow(dfB) == 0) next
  
  cent_dist <- euclidean_dist(dfA$centile_M_minus_F_z, dfB$centile_M_minus_F_z)
  sigma_dist <- euclidean_dist(dfA$logcv_M_div_F, dfB$logcv_M_div_F)
  deriv_dist <- euclidean_dist(na.omit(dfA$deriv_M_minus_F_z), na.omit(dfB$deriv_M_minus_F_z))
  
  #rm nas for integral
  dfA <- dfA %>%
    select(logAge_days, deriv_M_minus_F_z) %>%
    na.omit()
  
  dfB <- dfB %>%
    select(logAge_days, deriv_M_minus_F_z) %>%
    na.omit()
  
  deriv_j <- trapz(dfA$logAge_days, abs(dfA$deriv_M_minus_F_z - dfB$deriv_M_minus_F_z))
  
  result <- data.frame(
    "pheno" = pheno_val,
    "centile_dist" = cent_dist,
    "sigma_dist" = sigma_dist,
    "deriv_dist" = deriv_dist,
    "deriv_trapz" = deriv_j
  )
  dist_df <- bind_rows(dist_df, result)
}

dist_df <- dist_df %>%
  mutate(centile_sim = 1/(1+centile_dist),
         sigma_sim = 1/(1+sigma_dist),
         deriv_sim = 1/(1+deriv_dist)
         )

thresholds <- dist_df %>%
  summarize(centile_dist_mad = mad(centile_dist),
            centile_dist_median = median(centile_dist),
            sigma_dist_mad = mad(sigma_dist),
            sigma_dist_median = median(sigma_dist),
            deriv_dist_mad = mad(deriv_dist),
            deriv_dist_median = median(deriv_dist)) %>%
  mutate(centile_dist_threshold = centile_dist_median+(1*centile_dist_mad),
         sigma_dist_threshold = sigma_dist_median+(1*sigma_dist_mad),
         deriv_dist_threshold = deriv_dist_median+(1*deriv_dist_mad))

ggplot(dist_df) +
  geom_histogram(aes(x=centile_sim), alpha=.5, fill="purple") +
  geom_histogram(aes(x=sigma_sim), alpha=.5, fill="green") +
  geom_histogram(aes(x=deriv_sim), alpha=.5, fill="orange")

dist_df %>%
  summarise(n_pass_cent = sum(centile_dist <= thresholds$centile_dist_threshold),
            n_pass_sigma = sum(sigma_dist <= thresholds$sigma_dist_threshold),
            n_pass_deriv = sum(deriv_dist <= thresholds$deriv_dist_threshold),
            n = n()) %>%
  mutate(pct_pass_cent = (n_pass_cent/n)*100,
         pct_pass_sigma = (n_pass_sigma/n)*100,
         pct_pass_deriv = (n_pass_deriv/n)*100)
  
```


```{r}
sexdiffs_tbv2 <- full_join(sexdiffs_tbv, dist_df)

cent_dist_plt <- sexdiffs_tbv2 %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z, centile_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    gghighlight(centile_dist <= thresholds$centile_dist_threshold, calculate_per_facet = TRUE) +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

cent_dist_plt

cent_dist_plt_sig <- sexdiffs_tbv2 %>%
    select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample, sigma_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(sigma_dist <= thresholds$sigma_dist_threshold, calculate_per_facet = TRUE) +
    theme(legend.position = "none") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Sex-Difference in Variability (lnCVR)")

cent_dist_plt_sig

cent_dist_plt_d <- sexdiffs_tbv2 %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F_z, deriv_dist) %>%
  #filter(deriv_dist <= thresholds$deriv_dist_threshold) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(deriv_dist <= thresholds$deriv_dist_threshold, calculate_per_facet = TRUE) +
    theme(legend.position = "none") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Sex-Difference in Growth")

cent_dist_plt_d

cent_dist_plt_d2 <- sexdiffs_tbv2 %>%
    select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F_z, deriv_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=deriv_dist)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

#ggplotly(cent_dist_plt_d2)
```

compare consistent trajectories to regions with consistent sex/ consistent age*sex effects

```{r}
dist_df2 <- dist_df %>%
  mutate(consistent_sex = ifelse(pheno %in% consistent_sex, TRUE, FALSE),
         consistent_agesex = ifelse(pheno %in% consistent_agesex, TRUE, FALSE))
  
ggplot(dist_df2) +
  geom_jitter(aes(x=1, y=deriv_dist, color=consistent_agesex), alpha=.4) +
  theme_bw()

ggplot(dist_df) +
  geom_jitter(aes(x=1, y=cent_dist), alpha=.4, color="red") +
  #gghighlight(pheno %in% consistent_sex) +
  theme_bw()
```

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)
Calculating across age, not log-age, for interpretability

```{r}
dominance_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = rs_integral(logAge_days, centile_M_minus_F_z),
            sigma_int.unlog =  rs_integral(logAge_days, logcv_M_div_F),
            age_range = max(age_days) - min(age_days)) %>%
  # rescaling integrals to 100 years
  mutate(centile_int_norm.unlog = (centile_int.unlog/age_range)*365.25*100,
         sigma_int_norm.unlog = (sigma_int.unlog/age_range)*365.25*100) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5)) +
  coord_flip()

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()
```


```{r}
dominance_plt_df <- full_join(LR.df_tot_sex, dominance_df %>% select(!pheno_cat))
dominance_plt_df %>%
  mutate(dominance = ifelse(sig_fdr==TRUE, dominance, "NA")) %>%
  group_by(cv_sample) %>%
  summarise(F_dom = sum(dominance=="Female")/n(),
            M_dom = sum(dominance=="Male")/n())
```

plot dominance on brain

```{r}
#removing ventricles for visualization:
dominance_plt_df_novent <- dominance_plt_df %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_novent$centile_int_norm.unlog)

full_dom_plt <- plot_cv_brain(dominance_plt_df_novent, centile_int_norm.unlog, "Normalized Sex\nDifference", "ggthemes::Orange-Blue Diverging", c(-0.07, 0.07), c(2, 5, 2))

ggsave("figs/dominance_tbvTRUE.png", full_dom_plt, width=7, height=8, dpi=600)
```

How much of the lifespan are phenos male vs female dominant?

```{r}
#define function
prop_above <- function(x, y, threshold=0){
  crosses <- find_threshold_crosses(y, x, threshold, return="x", return_endpoints = FALSE)
  crosses <- c(crosses, min(x), max(x)) %>% 
    sort()
  #overall age modeled
  total_range <- max(x) - min(x)
  
  ranges <- diff(crosses)
  
  #if y > threshold at min(x) -> sum odd number ranges
  odd_indices_logical <- (1:length(ranges)) %% 2 == 1
  
  if (y[1] > threshold) {
    items_at_odd_indices <- ranges[odd_indices_logical]
    time_above <- sum(items_at_odd_indices)
  } else {
    items_at_even_indices <- ranges[!odd_indices_logical]
    time_above <- sum(items_at_even_indices)
  }
  
  prop_above <- time_above/total_range
  
  return(prop_above)
}

# #test function
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="sGMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
# 
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="WMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
```

```{r}
prop_above_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, centile_M_minus_F_z)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_df$cv_sample, prop_above_df$class)

prop_above_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_df_novent <- full_join(LR.df_tot_sex, prop_above_df %>% select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_plt <- plot_cv_brain(prop_above_df_novent, prop_male, "Proportion Male-Dominant", "ggthemes::Orange-Blue Diverging", c(0, 1), c(2, 5, 2))

ggsave("figs/prop_male_tbvTRUE.png", full_above_plt, width=7, height=8, dpi=600)
```


####Sigma

```{r}
range(dominance_plt_df_novent$sigma_int_norm.unlog)

full_dom_plt_sigma <- plot_cv_brain(dominance_plt_df_novent, sigma_int_norm.unlog, "Normalized Sigma\nSex Difference", "ggthemes::Orange-Blue Diverging", c(-.03, .03), c(2, 5, 2))

ggsave("figs/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, width=7, height=8, dpi=600)
```

```{r}
prop_above_sigma_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, logcv_M_div_F)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_sigma_df$cv_sample, prop_above_sigma_df$class)

prop_above_sigma_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_sigma_df_novent <- full_join(LR.df_tot_sex, prop_above_sigma_df %>% select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_sig_plt <- plot_cv_brain(prop_above_sigma_df_novent, prop_male, "Proportion Male-Dominant\nSigma", "ggthemes::Orange-Blue Diverging", c(0, 1), c(2, 5, 2))

ggsave("figs/prop_male_sigma_tbvTRUE.png", full_above_sig_plt, width=7, height=8, dpi=600)
```



###Shape Differences
*At what points in the lifespan are males and females growing differently?*
difference of first derivatives!
can also get magnitude differences

```{r}
deriv_df_sum <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(int2_z = trapz(logAge_days, deriv_M_minus_F_z),
            int2_z.unlog = rs_integral(logAge_days, deriv_M_minus_F_z))
```

int > 0 -> males grow faster than females
int < 0 -> females grow faster
int = 0 -> same rate

```{r}
deriv_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(faster_rate = case_when(
    deriv_M_minus_F_z.unlog > 0 ~ "Males",
    deriv_M_minus_F_z.unlog == 0 ~ "Same",
    deriv_M_minus_F_z.unlog < 0 ~ "Females",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- deriv_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = deriv_M_minus_F_z.unlog, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

deriv_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=faster_rate)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", deriv_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))
```


###Correlation

```{r}
sexdiffs.df_uniform.sc <- sexdiffs.df_uniform %>%
  filter(total=="TRUE" & weighted==FALSE) %>%
  group_by(cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(
    #tmp fix for sigma as CV, getting ln CV ratio
    logcv_M_div_F = log(sigma_Male/sigma_Female),
    #chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age)

df_cor_cent <- sexdiffs.df_uniform.sc %>%
  dplyr::select(logAge_days, age_days, centile_M_minus_F_z, deriv_M_minus_F_z.unlog, pheno, cv_sample)

cor_A <- df_cor_cent %>%
  filter(cv_sample == "A") %>%
  pivot_wider(id_cols=logAge_days, names_from=pheno, values_from=centile_M_minus_F_z) %>%
  arrange(logAge_days) %>%
  dplyr::select(any_of(pheno_list)) %>%
  cor(use="pairwise.complete.obs")

cor_B <- df_cor_cent %>%
  filter(cv_sample == "B") %>%
  pivot_wider(id_cols=logAge_days, names_from=pheno, values_from=centile_M_minus_F_z) %>%
  arrange(logAge_days) %>%
  dplyr::select(any_of(pheno_list)) %>%
  cor(use="pairwise.complete.obs")
```

plot

```{r}
ggcorrplot(cor_A, outline.col = "white", type = "upper")
ggcorrplot(cor_B, outline.col = "white", type = "lower")

ggcorrplot(cor_A, outline.col = "white", type = "upper")
ggcorrplot(cor_B, outline.col = "white", type = "lower")
```

attempt clustering

```{r}
wide_A <- sexdiffs.df_uniform.sc %>%
  filter(cv_sample == "A") %>%
  pivot_wider(id_cols=age_days, names_from=pheno, values_from=deriv_M_minus_F_z.unlog)

skim(wide_A)

wide_A_filt <- wide_A %>%
  dplyr::select(!starts_with("SUBC.CC")) %>%
  na.omit()

matrix_A <- wide_A_filt  %>%
  arrange(age_days) %>%
  dplyr::select(any_of(pheno_list)) %>%
  t()

list_A <- tslist(matrix_A)

A_clust <- tsclust(list_A, distance="dtw_basic", k=2L)
plot(A_clust)
A_clust <- tsclust(list_A, distance="dtw_basic", k=3L)
plot(A_clust)
A_clust <- tsclust(list_A, distance="dtw_basic", k=4L)
plot(A_clust)
A_clust <- tsclust(list_A, distance="dtw_basic", k=5L)
plot(A_clust)
```

```{r}
basis <- create.bspline.basis(rangeval = range(wide_A_filt$age_days, na.rm = TRUE), 
                               nbasis = 20)
time_points <- wide_A_filt$age_days
brain_fd <- Data2fd(argvals = time_points, 
                    y = t(matrix_A), 
                    basisobj = basis)

# Perform functional PCA
# nharm = number of principal components to compute
n_components <- 10  # Adjust as needed
fpca_result <- pca.fd(fdobj = brain_fd, 
                      nharm = n_components)

variance_explained <- fpca_result$varprop
cumulative_variance <- cumsum(variance_explained)

# Create scree plot data
scree_data <- data.frame(
  PC = 1:length(variance_explained),
  Variance = variance_explained * 100,
  Cumulative = cumulative_variance * 100
)

ggplot(scree_data, aes(x = PC)) +
  geom_line(aes(y = Variance), color = "blue", linewidth = 1) +
  geom_point(aes(y = Variance), color = "blue", size = 3) +
  geom_line(aes(y = Cumulative), color = "red", linewidth = 1, linetype = "dashed") +
  geom_point(aes(y = Cumulative), color = "red", size = 3) +
  labs(title = "Functional PCA Scree Plot",
       x = "Principal Component",
       y = "Variance Explained (%)") +
  theme_minimal()

fpca_scores <- fpca_result$scores
region_names <- intersect(names(wide_A_filt), pheno_list)
rownames(fpca_scores) <- region_names
scores_subset <- fpca_scores[, 1:3, drop = FALSE]

max_k <- min(10, nrow(scores_subset) - 1)
wss <- sapply(1:max_k, function(k) {
  kmeans(scores_subset, centers = k, nstart = 25)$tot.withinss
})

p_elbow <- ggplot(data.frame(k = 1:max_k, wss = wss), aes(x = k, y = wss)) +
  geom_line(color = "blue", linewidth = 1) +
  geom_point(color = "blue", size = 3) +
  labs(title = "Elbow Method for Optimal k",
       x = "Number of Clusters",
       y = "Total Within-Cluster Sum of Squares") +
  theme_minimal() +
  scale_x_continuous(breaks = 1:max_k)

print(p_elbow)

# Perform k-means clustering (adjust k as needed)
k <- 3  # Number of clusters - adjust based on elbow plot
set.seed(123)
kmeans_result <- kmeans(scores_subset, centers = k, nstart = 25)
clusters <- kmeans_result$cluster
```

```{r}
# Get fitted values from the functional data object
fitted_values <- eval.fd(time_points, brain_fd)  # Returns matrix: cols = regions, rows = time points

# Transpose to match brain_matrix_clean format (rows = regions, cols = time points)
fitted_matrix <- t(fitted_values)

# Calculate residuals
residuals_matrix <- matrix_A - fitted_matrix

# 2. CALCULATE FIT STATISTICS FOR EACH REGION
# --------------------------------------------------

fit_stats <- data.frame(
  Region = region_names,
  R_squared = numeric(length(region_names)),
  RMSE = numeric(length(region_names)),
  MAE = numeric(length(region_names)),
  Max_Abs_Error = numeric(length(region_names))
)

for (i in 1:nrow(matrix_A)) {
  observed <- matrix_A[i, ]
  fitted <- fitted_matrix[i, ]
  residuals <- residuals_matrix[i, ]
  
  # R-squared
  ss_total <- sum((observed - mean(observed))^2)
  ss_residual <- sum(residuals^2)
  fit_stats$R_squared[i] <- 1 - (ss_residual / ss_total)
  
  # Root Mean Square Error
  fit_stats$RMSE[i] <- sqrt(mean(residuals^2))
  
  # Mean Absolute Error
  fit_stats$MAE[i] <- mean(abs(residuals))
  
  # Maximum absolute error
  fit_stats$Max_Abs_Error[i] <- max(abs(residuals))
}

# Print summary statistics
cat("\n=== OVERALL FIT STATISTICS ===\n")
cat(sprintf("Mean R-squared: %.4f (SD: %.4f)\n", 
            mean(fit_stats$R_squared), sd(fit_stats$R_squared)))
cat(sprintf("Median R-squared: %.4f\n", median(fit_stats$R_squared)))
cat(sprintf("Min R-squared: %.4f (Region: %s)\n", 
            min(fit_stats$R_squared), 
            fit_stats$Region[which.min(fit_stats$R_squared)]))
cat(sprintf("Max R-squared: %.4f (Region: %s)\n", 
            max(fit_stats$R_squared), 
            fit_stats$Region[which.max(fit_stats$R_squared)]))
cat(sprintf("\nMean RMSE: %.6f\n", mean(fit_stats$RMSE)))
cat(sprintf("Mean MAE: %.6f\n", mean(fit_stats$MAE)))

# Show worst-fitting regions
cat("\n=== WORST-FITTING REGIONS (Lowest R-squared) ===\n")
print(fit_stats %>% arrange(R_squared) %>% head(10))

# Show best-fitting regions
cat("\n=== BEST-FITTING REGIONS (Highest R-squared) ===\n")
print(fit_stats %>% arrange(desc(R_squared)) %>% head(10))

# 3. VISUALIZE FIT: HISTOGRAM OF R-SQUARED VALUES
# --------------------------------------------------

p_rsq_hist <- ggplot(fit_stats, aes(x = R_squared)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean(fit_stats$R_squared), 
             color = "red", linetype = "dashed", linewidth = 1) +
  annotate("text", x = mean(fit_stats$R_squared), y = Inf, 
           label = sprintf("Mean = %.3f", mean(fit_stats$R_squared)),
           vjust = 2, color = "red", size = 4) +
  labs(title = "Distribution of R-squared Values Across Regions",
       x = "R-squared",
       y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

print(p_rsq_hist)

# 4. VISUALIZE FIT: OBSERVED VS FITTED FOR SELECTED REGIONS
# --------------------------------------------------

# Select a few regions to visualize (worst, median, best fitting)
worst_region <- fit_stats$Region[which.min(fit_stats$R_squared)]
best_region <- fit_stats$Region[which.max(fit_stats$R_squared)]
median_region <- fit_stats$Region[order(fit_stats$R_squared)[round(nrow(fit_stats)/2)]]

selected_regions <- c(worst_region, median_region, best_region)

# Create plot data
plot_data_fit <- data.frame()

for (region in selected_regions) {
  idx <- which(region_names == region)
  rsq <- fit_stats$R_squared[idx]
  
  temp_df <- data.frame(
    Region = paste0(region, sprintf(" (R² = %.3f)", rsq)),
    age_days = time_points,
    Observed = matrix_A[idx, ],
    Fitted = fitted_matrix[idx, ]
  ) %>%
    pivot_longer(cols = c(Observed, Fitted), 
                 names_to = "Type", 
                 values_to = "Value")
  
  plot_data_fit <- rbind(plot_data_fit, temp_df)
}

p_fit_examples <- ggplot(plot_data_fit, aes(x = age_days, y = Value, color = Type)) +
  geom_line(linewidth = 1, alpha = 0.8) +
  geom_point(data = plot_data_fit %>% filter(Type == "Observed"), 
             size = 1.5, alpha = 0.4) +
  facet_wrap(~Region, scales = "free_y", ncol = 1) +
  scale_color_manual(values = c("Observed" = "black", "Fitted" = "red")) +
  labs(title = "Observed vs Fitted Values: Example Regions",
       subtitle = "Worst, Median, and Best Fitting Regions",
       x = "Age (days)",
       y = "Value") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        legend.position = "bottom")

print(p_fit_examples)

# 5. VISUALIZE RESIDUALS
# --------------------------------------------------

# Plot residuals over time
plot_data_residuals <- data.frame()

for (i in 1:nrow(residuals_matrix)) {
  temp_df <- data.frame(
    Region = region_names_clean[i],
    age_days = time_points,
    Residual = residuals_matrix[i, ]
  )
  plot_data_residuals <- rbind(plot_data_residuals, temp_df)
}

p_residuals <- ggplot(plot_data_residuals, 
                      aes(x = age_days, y = Residual, group = Region)) +
  geom_line(alpha = 0.2, color = "steelblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", linewidth = 1) +
  labs(title = "Residuals for All Regions",
       subtitle = "Difference between observed and fitted values",
       x = "Age (days)",
       y = "Residual") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

print(p_residuals)

# Residual vs Fitted plot (to check for patterns)
plot_data_resid_vs_fitted <- data.frame(
  Fitted = as.vector(fitted_matrix),
  Residual = as.vector(residuals_matrix)
)

p_resid_vs_fitted <- ggplot(plot_data_resid_vs_fitted, 
                             aes(x = Fitted, y = Residual)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", linewidth = 1) +
  geom_smooth(method = "loess", color = "darkblue", se = TRUE) +
  labs(title = "Residual vs Fitted Values",
       subtitle = "Should show no systematic pattern",
       x = "Fitted Value",
       y = "Residual") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

print(p_resid_vs_fitted)

# 6. VISUALIZE ALL REGIONS: GRID OF OBSERVED VS FITTED
# --------------------------------------------------

# For a more comprehensive view, plot all regions (or a random sample if too many)
n_regions_to_plot <- min(16, length(region_names_clean))  # Plot up to 16 regions

if (length(region_names_clean) > n_regions_to_plot) {
  set.seed(123)
  sample_regions <- sample(region_names_clean, n_regions_to_plot)
} else {
  sample_regions <- region_names_clean
}

plot_data_grid <- data.frame()

for (region in sample_regions) {
  idx <- which(region_names_clean == region)
  rsq <- fit_stats$R_squared[idx]
  
  temp_df <- data.frame(
    Region = paste0(region, sprintf("\n(R² = %.3f)", rsq)),
    age_days = time_points,
    Observed = brain_matrix_clean[idx, ],
    Fitted = fitted_matrix[idx, ]
  ) %>%
    pivot_longer(cols = c(Observed, Fitted), 
                 names_to = "Type", 
                 values_to = "Value")
  
  plot_data_grid <- rbind(plot_data_grid, temp_df)
}

p_grid <- ggplot(plot_data_grid, aes(x = age_days, y = Value, color = Type)) +
  geom_line(linewidth = 0.7, alpha = 0.6) +
  geom_point(data = plot_data_grid %>% filter(Type == "Observed"), 
             size = 0.8, alpha = 0.4) +
  facet_wrap(~Region, scales = "free_y", ncol = 4) +
  scale_color_manual(values = c("Observed" = "black", "Fitted" = "red")) +
  labs(title = "Observed vs Fitted: Sample of Regions",
       x = "Age (days)",
       y = "Value") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        legend.position = "bottom",
        strip.text = element_text(size = 7))

print(p_grid)

# 7. EXPORT FIT STATISTICS
# --------------------------------------------------

# Save fit statistics to CSV (optional)
# write.csv(fit_stats, "functional_fit_statistics.csv", row.names = FALSE)

# Return fit statistics
cat("\n=== FIT CHECK COMPLETE ===\n")
cat("Check the plots above to assess fit quality.\n")
cat("Good fit indicators:\n")
cat("  - R-squared > 0.95 for most regions\n")
cat("  - No systematic patterns in residual plots\n")
cat("  - Small RMSE relative to data scale\n")
```

plot

```{r}
plot_data_pcs <- data.frame(
  PC1 = fpca_scores[, 1],
  PC2 = fpca_scores[, 2],
  Cluster = factor(clusters),
  Region = region_names
)

p_clusters <- ggplot(plot_data_pcs, aes(x = PC1, y = PC2, color = Cluster, label = Region)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_text(size = 2.5, hjust = -0.1, vjust = 0.5, check_overlap = TRUE) +
  labs(title = "Brain Regions Clustered in fPCA Space",
       x = sprintf("PC1 (%.1f%% variance)", variance_explained[1] * 100),
       y = sprintf("PC2 (%.1f%% variance)", variance_explained[2] * 100)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

print(p_clusters)

result_df <- data.frame(pheno = region_names, cluster = factor(clusters))
plot_data_trajectories <- df_cor_cent %>%
  filter(cv_sample == "A") %>%
  inner_join(result_df) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)))

ggplot(plot_data_trajectories, 
                         aes(x = logAge_days, y = deriv_M_minus_F_z.unlog, group = pheno, color = cluster)) +
  geom_line(alpha = 0.5, linewidth = 0.7) +
  facet_wrap(~cluster, scales = "free_y") +
  labs(title = "Brain Region Trajectories by Cluster",
       x = "Age (days)",
       y = "Value") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        legend.position = "none")

```


```{r}
pc1_scores <- fpca_scores[, 1]
range(pc1_scores)
# Create a dataframe matching regions to their PC1 scores
pc1_lookup <- data.frame(
  pheno = region_names,
  PC1_Score = pc1_scores
)

# Reshape data for plotting and add PC1 scores
plot_data_pc1 <- full_join(plot_data_trajectories, pc1_lookup)

# Create the plot with continuous color scale based on PC1
ggplot(plot_data_pc1, aes(x = age_days, y = deriv_M_minus_F_z.unlog, group = pheno, color = PC1_Score)) +
  geom_line(alpha = 0.7, linewidth = 0.8) +
  scale_color_gradient2(low = "blue", mid = "gray90", high = "red", 
                        midpoint = 0,
                        name = "PC1 Score") +
  labs(title = "Brain Region Trajectories Colored by PC1 Loading",
       subtitle = sprintf("PC1 explains %.1f%% of variance", variance_explained[1] * 100),
       x = "Age (days)",
       y = "Value") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        legend.position = "right")

```

```{r}
# --------------------------------------------------
# VISUALIZE WHAT PC1 REPRESENTS
# --------------------------------------------------

# Plot the first principal component function
par(mfrow = c(1, 2))

# PC1 function
plot(fpca_result$harmonics[1], 
     main = sprintf("PC1 Function\n(%.1f%% variance)", variance_explained[1] * 100),
     xlab = "Age (days)",
     ylab = "PC1",
     lwd = 2,
     col = "darkblue")
abline(h = 0, lty = 2, col = "gray50")

# PC2 function for comparison
plot(fpca_result$harmonics[2], 
     main = sprintf("PC2 Function\n(%.1f%% variance)", variance_explained[2] * 100),
     xlab = "Age (days)",
     ylab = "PC2",
     lwd = 2,
     col = "darkred")
abline(h = 0, lty = 2, col = "gray50")

par(mfrow = c(1, 1))
```

