---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, sf, cowplot, subcortexVisualizationR)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
#remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)
```

#ggseg plotting functions and atlas definitions
```{r}
source("code/plot_cv_brain.R")
```

#Defining integral funs

```{r}
#fun for derivative
get_deriv <- function(x, age) {
  c(NA, diff(x) / diff(age))
}

#riemann_stieltjes
rs_integral <- function(x, y, g = function(x) log10(x)) {
  stopifnot(
    length(x) == length(y),
    length(x) >= 2
  )

  # sort just in case
  ord <- order(x)
  x <- x[ord]
  y <- y[ord]

  gx <- g(x)

  # monotonicity check (important!)
  if (any(diff(gx) < 0)) {
    stop("g(x) must be non-decreasing for a Riemann–Stieltjes integral")
  }

  sum(y[-length(y)] * diff(gx))
}

#euclidean distance fun from https://www.geeksforgeeks.org/r-language/how-to-calculate-euclidean-distance-in-r/
euclidean_dist <- function(vect1, vect2) sqrt(sum((vect1 - vect2)^2))
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv"))

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

length(output_fread)

LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight")) %>%
  dplyr::select(!file)
```

```{r, eval=FALSE}
LR.df %>%
  group_by(cv_sample, total, weighted) %>%
  summarise(n_phenos = n_unique(pheno))

A_totalFALSE <- LR.df %>%
  filter(cv_sample == "A" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalFALSE)

A_totalTRUE <- LR.df %>%
  filter(cv_sample == "A" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalTRUE)

B_totalFALSE <- LR.df %>%
  filter(cv_sample == "B" & total == "FALSE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalFALSE)

B_totalTRUE <- LR.df %>%
  filter(cv_sample == "B" & total == "TRUE" & weighted == "FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalTRUE)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
    pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)))

table(LR.df2$effect, LR.df2$total)
```



```{r}
LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(total, weighted, cv_sample) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$label), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE)

LR.df_uncor %>%
  group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n()) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_uncor_novent <- LR.df_uncor %>%
  filter(!(pheno %in% vent_list))
range(LR.df_uncor_novent$resi)
full_plt_uncor <- plot_cv_brain(LR.df_uncor_novent, resi, "RESI", "grDevices::Greens 2", c(0,0.1), c(1.03, 4.5, .93))

ggsave("figs/supplement/resi_uncorrected.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

```{r}
#test plotting just split A
LR.df_uncor_novent_B <- LR.df_uncor_novent %>%
  filter(cv_sample=="B")
B_plt_uncor <- plot_cv_brain(LR.df_uncor_novent_B, resi, "RESI", "grDevices::Greens 2", c(0,0.1), c(1, , 1.5), return_plt=FALSE)
saveRDS(B_plt_uncor, "figs/ggobjs/agebysex_B.rds")
```

#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE)

table(LR.df_tot$sig_fdr)

LR.df_tot %>%
group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n()) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_tot_novent <- LR.df_tot %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_novent$resi)

full_plt_tot <- plot_cv_brain(LR.df_tot_novent, resi, "RESI", "grDevices::Greens 2", c(0,0.1), c(1.03, 4.5, .93))

ggsave("figs/supplement/resi_tbv.png", full_plt_tot, dpi=600, height=8.5, width=7.1, bg = "white")
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE)

table(LR.df_tot_sex$sig_fdr, LR.df_tot_sex$cv_sample)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)

LR.df_tot_sex %>%
  filter(sig_fdr == TRUE) %>%
  group_by(cv_sample) %>%
  summarise(min_resi = min(sig_resi),
            max_resi = max(sig_resi),
            mean_resi = mean(sig_resi))
```

```{r}
#removing ventricles for visualization:
LR.df_tot_sex_novent <- LR.df_tot_sex %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_sex_novent$resi)

full_plt_tot_sex <- plot_cv_brain(LR.df_tot_sex_novent, resi, "RESI", "grDevices::Greens 2", c(0,0.2), c(1.03, 4.5, .93))

ggsave("figs/supplement/resi_tbv_allsex.png", full_plt_tot_sex, dpi=600, height=8.5, width=7.1, bg = "white")
```

plot just in B
```{r}
#removing ventricles for visualization:
LR.df_tot_sex_noventB <- LR.df_tot_sex_novent %>%
  filter(cv_sample=="A")
full_plt_tot_sex_B <- plot_cv_brain(LR.df_tot_sex_noventB, resi, "RESI", "grDevices::Greens 2", c(0,0.2), c(1, 5, 1.1), return_plt = FALSE)

saveRDS(full_plt_tot_sex_B, "figs/ggobjs/full_plt_tot_sex_B.rds")
```


Note phenos with consistent sex and/or age*sex effects, per LR tests

```{r}
#regions w sig sex effects in both samples
consistent_sex <- LR.df_tot_sex %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  filter(sig_fdr_A == TRUE & sig_fdr_B == TRUE) %>%
  pull(pheno)
 
#consistent age*sex effects
consistent_agesex <- LR.df_tot %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  filter(sig_fdr_A == TRUE & sig_fdr_B == TRUE) %>%
  pull(pheno)

#just within each sample
sig_agesex <- LR.df_tot %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  filter(sig_fdr==TRUE) %>%
  pull(samp_pheno)

sig_sex <- LR.df_tot_sex %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  filter(sig_fdr==TRUE) %>%
  pull(samp_pheno)
```

##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

sexdiffs.df_all <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         pheno = stringr::str_remove(basename(file), "_(uniform_)?sexdiffs\\.csv"),
         uniform = ifelse(grepl("uniform", file), TRUE, FALSE)) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    age_days = un_log(logAge_days)) %>%
  dplyr::select(!file)

sexdiffs.df <- sexdiffs.df_all %>% filter(uniform==FALSE)
sexdiffs.df.u <- sexdiffs.df_all %>% filter(uniform==TRUE)


rm(output_fread_sd) #save space
rm(sexdiffs.df_all)
```

chain rule to fix derivative scaling

```{r}
sexdiffs.df_sc <- sexdiffs.df %>%
  group_by(total, weighted, cv_sample, pheno) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(#chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age,
         deriv_logcv_M_div_F.unlog = deriv_logcv_M_div_F*deriv_age)

sexdiffs_tbv <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==TRUE)

sexdiffs <- sexdiffs.df_sc %>%
  filter(weighted==FALSE & total==FALSE)
```

Plot

```{r}
for (df in list(sexdiffs_tbv, sexdiffs)){
  t <- as.character(unique(df$total))
  print(t)
  
  #median centile
  cent_df <- df %>%
    dplyr::select(centile_M_minus_F, pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  cent_plt <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_cent.png")
  ggsave(n, cent_plt)
  print(cent_plt)
  
  cent_plt2 <- cent_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Scaled 50th Centile Sex-Difference")
  
  print(cent_plt2)
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_cent_scaled.png")
  ggsave(n, cent_plt2)

  #sigma
  sig_df <- df %>%
    dplyr::select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  sig_plt <- sig_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sig_df$logAge_days) +
    labs(y="log CV Ratio")
  print(sig_plt)
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_cv.png")
  ggsave(n, sig_plt)
  
  #derivatives
  deriv_df <- df %>%
    dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F) %>%
    na.omit() %>%
    arrange(pheno, logAge_days)
  
  deriv_plt <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="50th Centile Difference Derivative")
  print(deriv_plt)
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_deriv.png")
  ggsave(n, deriv_plt)
  
  deriv_plt2 <- deriv_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="Scaled 50th Centile Difference Derivative")
  print(deriv_plt2)
  
  n <- paste0("figs/misc/sexdiffs_tbv",t,"_deriv_scaled.png")
  ggsave(n, deriv_plt2)
}
```

*try median

interactive plots
```{r, eval=FALSE}
cent_plt <- sexdiffs_tbv %>%
  dplyr::select(centile_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    labs(y="Scaled 50th Centile Sex-Difference")
ggplotly(cent_plt)

deriv_plt <- sexdiffs_tbv %>%
  dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
    labs(y="Scaled 50th Centile Difference Derivative")
ggplotly(deriv_plt)
```

plot just sGMV B
```{r}
sgmv_centB <- sexdiffs_tbv %>%
  filter(cv_sample=="B" & pheno=="sGMV") %>%
  dplyr::select(centile_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
  geom_hline(yintercept=0, color="gray", linetype = "dashed") +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno), color="#4C7F58FF") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="50th Centile\nSex-Difference") +
    theme_linedraw() +
    theme(legend.position = "none", axis.title.x= element_blank())

sgmv_derivB <- sexdiffs_tbv %>%
  filter(cv_sample=="B" & pheno=="sGMV") %>%
  dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
      geom_hline(yintercept=0, color="gray", linetype = "dashed") +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno), color="#4C7F58FF") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="Sex-Difference\nDerivative") +
    theme_linedraw() +
    theme(legend.position = "none")

sgmv_sigmaB <- sexdiffs_tbv %>%
  filter(cv_sample=="B" & pheno=="sGMV") %>%
  dplyr::select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
      geom_hline(yintercept=0, color="gray", linetype = "dashed") +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno), color="#4C7F58FF") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="log CV Ratio") +
    theme_linedraw() +
    theme(legend.position = "none", axis.title.x= element_blank())

sGMV_B_plots <- list(sgmv_centB, sgmv_sigmaB, sgmv_derivB)
saveRDS(sGMV_B_plots, "figs/ggobjs/sGMV_B_trajectories.rds")
```

```{r}
sexdiffs_tbv_glob <- sexdiffs_tbv %>%
  mutate(age_days = age_days-280) %>%
  filter(pheno %in% global_list) %>%
  #rescale to be percentage max
  group_by(cv_sample, pheno) %>%
  #mutate(centile_M_minus_F_z = centile_M_minus_F_z / max(abs(centile_M_minus_F_z)) * 100) %>%
ungroup()

sexdiffs_tbv_glob_sums <- sexdiffs_tbv_glob %>%
  summarize(#age at peak sex diff
            max_sex_diff = extr_val(centile_M_minus_F_z, which_max_gc(abs(centile_M_minus_F_z))),
            age_max_sex_diff = extr_val(age_days, which_max_gc(abs(centile_M_minus_F_z))),
            #age at min sex diff
            min_sex_diff = extr_val(centile_M_minus_F_z, which_min_gc(abs(centile_M_minus_F_z))),
            age_min_sex_diff = extr_val(age_days, which_min_gc(abs(centile_M_minus_F_z))),
            #age peak change in sex diff
            max_deriv = extr_val(centile_M_minus_F_z, which_max_gc(abs(deriv_M_minus_F_z))),
            age_max_deriv = extr_val(age_days, which_max_gc(abs(deriv_M_minus_F_z))),
            #age at male peak
            max_M = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Male_z)),
            age_max_M = extr_val(age_days, which_max_gc(median_centile_Male_z)),
            #age at female peak
            max_F = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Female_z)),
            age_max_F = extr_val(age_days, which_max_gc(median_centile_Female_z)),
            #age at male peak growth
            max_deriv_M = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Male_z)),
            age_max_deriv_M = extr_val(age_days, which_max_gc(deriv_median_centile_Male_z)),
            #age at female peak growth
            max_deriv_F = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Female_z)),
            age_max_deriv_F = extr_val(age_days, which_max_gc(deriv_median_centile_Female_z)),
            .by = c(cv_sample, pheno)
            )

#add age at peak rate divergence!!

#pivot and rejoin
values_long <- sexdiffs_tbv_glob_sums %>%
  select(!starts_with("age_")) %>%
  pivot_longer(cols = -c(pheno, cv_sample), names_to = "metric", values_to = "value")

ages_long <- sexdiffs_tbv_glob_sums %>%
  select(pheno, cv_sample, starts_with("age_")) %>%
  pivot_longer(cols = starts_with("age_"), names_to = "metric", values_to = "age") %>%
  mutate(metric = gsub("^age_", "", metric))

sexdiffs_tbv_glob_sums_long <- left_join(values_long, ages_long, by = c("pheno", "cv_sample", "metric"))

cent_plt_pct <- sexdiffs_tbv_glob %>%
  dplyr::select(cv_sample, centile_M_minus_F_z, pheno, age_days) %>%
  ggplot() +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_line(aes(x=age_days, y=centile_M_minus_F_z, group=cv_sample, color=cv_sample)) +
  geom_point(
    data = sexdiffs_tbv_glob_sums_long,
    aes(x = age, y = value, color = cv_sample, shape = metric),
    size = 3
  ) +
  scale_shape_manual(
    values = c(
      "max_sex_diff" = 19,
      "min_sex_diff" = 15,
      "max_deriv" = 18,
      "max_M"        = 2,
      "max_F"        = 6,
      "max_deriv_M"  = 3,
      "max_deriv_F"  = 4
    )
  ) +
  theme_linedraw() +
  format_x_axis("lifespan", sexdiffs_tbv_glob$age_days) +
  facet_wrap(~pheno, scales="free_y") +
  labs(y="Normalized Sex-Difference")
cent_plt_pct
saveRDS(cent_plt_pct, "figs/ggobjs/sGMV_B_trajectories.rds")
```

Try averaging across splits

```{r}
sum_cols <- setdiff(names(sexdiffs.df.u), c("logAge_days", "cv_sample", "total", "weighted", "pheno", "uniform", "pheno_cat", "age_days"))

glob_mean <- sexdiffs.df.u %>%
  filter(pheno %in% global_list & total==TRUE) %>%
  mutate(age_days = round(age_days, digits=0)) %>%
  group_by(pheno, age_days) %>%
  summarise(across(all_of(sum_cols), mean, na.rm = TRUE)) %>%
  #rescale
  group_by(pheno) %>%
  mutate(centile_M_minus_F_z = centile_M_minus_F_z / max(abs(centile_M_minus_F_z)) * 100) %>%
  ungroup()

glob_mean_sums <- glob_mean %>%
  summarize(#age at peak sex diff
            max_sex_diff = extr_val(centile_M_minus_F_z, which_max_gc(abs(centile_M_minus_F_z))),
            age_max_sex_diff = extr_val(age_days, which_max_gc(abs(centile_M_minus_F_z))),
            #age at min sex diff
            min_sex_diff = extr_val(centile_M_minus_F_z, which_min_gc(abs(centile_M_minus_F_z))),
            age_min_sex_diff = extr_val(age_days, which_min_gc(abs(centile_M_minus_F_z))),
            #age peak change in sex diff
            max_deriv = extr_val(centile_M_minus_F_z, which_max_gc(abs(deriv_M_minus_F_z))),
            age_max_deriv = extr_val(age_days, which_max_gc(abs(deriv_M_minus_F_z))),
            #age at male peak
            max_M = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Male_z)),
            age_max_M = extr_val(age_days, which_max_gc(median_centile_Male_z)),
            #age at female peak
            max_F = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Female_z)),
            age_max_F = extr_val(age_days, which_max_gc(median_centile_Female_z)),
            #age at male peak growth
            max_deriv_M = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Male_z)),
            age_max_deriv_M = extr_val(age_days, which_max_gc(deriv_median_centile_Male_z)),
            #age at female peak growth
            max_deriv_F = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Female_z)),
            age_max_deriv_F = extr_val(age_days, which_max_gc(deriv_median_centile_Female_z)),
            .by = pheno
            )

#add age at peak rate divergence!!

#pivot and rejoin
values_long <- glob_mean_sums %>%
  select(!starts_with("age_")) %>%
  pivot_longer(cols = -c(pheno), names_to = "metric", values_to = "value")

ages_long <- glob_mean_sums %>%
  select(pheno, starts_with("age_")) %>%
  pivot_longer(cols = starts_with("age_"), names_to = "metric", values_to = "age") %>%
  mutate(metric = gsub("^age_", "", metric))

glob_mean_sums_long <- left_join(values_long, ages_long, by = c("pheno", "metric"))

cent_plt_pct_mean <- glob_mean %>%
  ggplot() +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(aes(x=age_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
  geom_point(
    data = glob_mean_sums_long,
    aes(x = age, y = value, color = pheno, shape = metric),
    size = 3
  ) +
  scale_shape_manual(
    values = c(
      "max_sex_diff" = 19,
      "min_sex_diff" = 15,
      "max_deriv" = 18,
      "max_M"        = 2,
      "max_F"        = 6,
      "max_deriv_M"  = 3,
      "max_deriv_F"  = 4
    )
  ) +
  theme_linedraw() +
  format_x_axis("lifespan", sexdiffs_tbv_glob$age_days) +
  facet_wrap(~pheno, scales="free_y") +
  labs(y="% Maximum Sex-Difference")
cent_plt_pct_mean
```

```{r}
biggest_agesex_resi <- LR.df_tot %>%
  mintyr:::top_perc(perc = 0.1,                # Select top 10%
         trait = c("resi"),  # Column to analyze
         by = "cv_sample",
         keep_data = TRUE)

biggest_agesex_resi <- biggest_agesex_resi$resi_0.1$data %>% ungroup() %>% select(!pheno_cat)
biggest_agesex_resi$top10 <- TRUE

biggest_sex_resi <- LR.df_tot_sex %>%
  mintyr:::top_perc(perc = 0.1,                # Select top 10%
         trait = c("resi"),  # Column to analyze
         by = "cv_sample",
         keep_data = TRUE)

biggest_sex_resi <- biggest_sex_resi$resi_0.1$data %>% ungroup() %>% select(!pheno_cat)
biggest_sex_resi$top10 <- TRUE

#highlight trajectories of phenos with large effects
top_sex_plt <- full_join(sexdiffs_tbv, biggest_sex_resi) %>%
  dplyr::select(centile_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, top10) %>%
  mutate(top10 = ifelse(is.na(top10), "no", "yes")) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(top10 == "yes", calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    labs(title="Male-Female Trajectory Differences in Phenos\nwith Strongest Sex Effects (Top 10%)", 
         y="Scaled 50th Centile Sex-Difference")

top_agesex_plt <- full_join(sexdiffs_tbv, biggest_agesex_resi) %>%
  dplyr::select(deriv_M_minus_F_z, pheno, logAge_days, pheno_cat, cv_sample, top10) %>%
  mutate(top10 = ifelse(is.na(top10), "no", "yes")) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
  gghighlight(top10 == "yes", calculate_per_facet = TRUE) +
    labs(title="Male-Female Rate-of-Growth Differences in Phenos\nwith Strongest Age*Sex Effects (Top 10%)", 
         y="Scaled 50th Centile Difference Derivative")

ggsave("figs/misc/top_sexdiffs_centiles.png", top_sex_plt)
ggsave("figs/misc/top_agesexdiffs_derivs.png", top_agesex_plt)

```

compare derivatives with chain rule
```{r}
sexdiffs_tbv %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("log_lifespan_fetal", sexdiffs_tbv$logAge_days) +
    labs(y="50th Centile Difference Derivative")

sexdiffs_tbv %>%
  #filter(pheno_cat == "Subcortical Vol" & age_days > (365.25*40)) %>%
    ggplot() +
    geom_line(aes(x=age_days, y=deriv_M_minus_F_z.unlog, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none") +
   format_x_axis("lifespan_fetal", sexdiffs_tbv$age_days) +
    labs(y="50th Centile Difference Derivative")
```

playing around with second deriv

```{r, eval=FALSE}
sexdiffs_tbv %>%
group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(deriv2 = get_deriv(deriv_M_minus_F_z, logAge_days)) %>%
  ggplot() +
  geom_line(aes(x=logAge_days, y=deriv2, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    theme(legend.position = "none")
```

###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#calculate integral of deriv of diff^2 (all in z-scores)
sexdiffs_tbv_sq <- sexdiffs_tbv %>%
  mutate(sq_centile_diff_z = (centile_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_deriv_z = get_deriv(sq_centile_diff_z, logAge_days)) %>%
  mutate(sq_deriv_z.unlog = sq_deriv_z*deriv_age) %>% #chain rule to convert derivative
  ungroup()

#CHECK THAT THESE PLOTS MAKE SENSE!!
sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv_z, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x:::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sexdiffs_tbv_sq$logAge_days) +
    labs(y="Change in Magnitude Sex Diff")

sexdiffs_tbv_sq %>%
    ggplot() +
    geom_line(aes(x=age_days, y=sq_deriv_z.unlog, group=pheno, color=pheno)) +
  geom_hline(yintercept = 0) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv_z))+
    ggh4x::facet_grid2(pheno_cat~cv_sample, scales = "free_y", independent = "y") +
    #facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("lifespan_fetal", sexdiffs_tbv_sq$age_days) +
    labs(y="Change in Magnitude Sex Diff")
```

```{r}
#try getting integral
convergence_int <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv_z),
            sq_int_z.unlog = rs_integral(logAge_days, sq_deriv_z)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging")) #agree regardless of whether log-age or non-log-age is used
```

right cerebellum cortex is clearly an outlier

try finding points at which pt derivative crosses 0 increasing vs decreasing

```{r}
cent_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv_z.unlog > 0 ~ "diverging",
    sq_deriv_z.unlog == 0 ~ "constant",
    sq_deriv_z.unlog < 0 ~ "converging",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- cent_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv_z, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

cent_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", cent_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)

table(cent_df$trajectory)
```

Get max male and female biases and plot on centile differences

```{r}
peak_diffs_sum <- cent_df %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(centile_M_minus_F_z, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(centile_M_minus_F)),
            female_peak = min_gc(centile_M_minus_F_z, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(centile_M_minus_F))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")


peak_diffs_sum %>%
  ggplot() +
  geom_histogram(aes(x=peak_age, fill=sex), position="identity", alpha=.4) +
  format_x_axis("log_lifespan_fetal", peak_diffs_sum$peak_age) +
  facet_grid(pheno_cat~cv_sample, scales="free_y")

cent_df %>%
    filter(pheno %in% consistent_agesex) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% consistent_agesex), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")
```

differences tend to peak around beginning and end of life - likely d/t smaller sample sizes at extreme ages

####Sigma

*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of lnCVR squared: (not sure this totally makes sense tbh)
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#quick dirty code to get abs diff
sigma_df <- sexdiffs_tbv %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_cv_diff = (logcv_M_div_F)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_cv_deriv = get_deriv(sq_cv_diff, logAge_days)) %>%
  ungroup() %>%
  mutate(sq_cv_deriv.unlog = sq_cv_deriv*deriv_age)

sigma_df %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_cv_deriv, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", sigma_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Variability")

#try getting integral
sigma_convergence_df <- sigma_df %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_cv_deriv),
            sq_int_z.unlog = rs_integral(logAge_days, sq_cv_deriv)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging"))
```

####Derivative

*Do males and females rates-of-growth tend to get more or less similar across the lifespan?*

Taking derivative of deriv squared: (not sure this totally makes sense tbh)
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#quick dirty code to get abs diff
deriv_df <- sexdiffs_tbv %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_deriv_diff = (deriv_M_minus_F_z)^2) %>%
  group_by(pheno, cv_sample) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_deriv2 = get_deriv(sq_deriv_diff, logAge_days)) %>%
  ungroup() %>%
  mutate(sq_deriv2.unlog = sq_deriv2*deriv_age)

sq_deriv2_plt <- deriv_df %>%
  #filter(pheno == "lh.CT.caudalanteriorcingulate") %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=sq_deriv2, group=pheno, color=pheno)) +
  #geom_smooth(aes(x=logAge_days, y=sq_deriv)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  #gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", deriv_df$logAge_days) +
    labs(y="Change in Magnitude Sex Diff in Velocity")
#ggplotly(sq_deriv2_plt)

#try getting integral
deriv_convergence_df <- deriv_df %>%
  group_by(pheno, cv_sample, pheno_cat)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv2)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging"))
```

###Cross-sample Consistency
Get across-sample euclidean distance
```{r}
dist_results <- vector("list", length(pheno_list))
for (i in seq_along(pheno_list)) {
  pheno_val <- pheno_list[[i]]
  df <- sexdiffs_tbv %>%
    filter(pheno==pheno_val) %>%
    arrange(logAge_days)
  
  dfA <- df %>%
    filter(cv_sample=="A")
  
  dfB <- df %>%
    filter(cv_sample=="B")
  
  if (nrow(dfA) == 0 || nrow(dfB) == 0) next
  
  cent_dist <- euclidean_dist(dfA$centile_M_minus_F_z, dfB$centile_M_minus_F_z)
  sigma_dist <- euclidean_dist(dfA$logcv_M_div_F, dfB$logcv_M_div_F)
  deriv_dist <- euclidean_dist(na.omit(dfA$deriv_M_minus_F_z), na.omit(dfB$deriv_M_minus_F_z))
  
  #rm nas for integral
  dfA <- dfA %>%
    dplyr::select(logAge_days, deriv_M_minus_F_z) %>%
    na.omit()
  
  dfB <- dfB %>%
    dplyr::select(logAge_days, deriv_M_minus_F_z) %>%
    na.omit()
  
  deriv_j <- trapz(dfA$logAge_days, abs(dfA$deriv_M_minus_F_z - dfB$deriv_M_minus_F_z))
  
  dist_results[[i]] <- data.frame(
    "pheno" = pheno_val,
    "centile_dist" = cent_dist,
    "sigma_dist" = sigma_dist,
    "deriv_dist" = deriv_dist,
    "deriv_trapz" = deriv_j
  )
}

dist_df <- dplyr::bind_rows(dist_results)

dist_df <- dist_df %>%
  mutate(centile_sim = 1/(1+centile_dist),
         sigma_sim = 1/(1+sigma_dist),
         deriv_sim = 1/(1+deriv_dist)
         )

thresholds <- dist_df %>%
  summarize(centile_dist_mad = mad(centile_dist),
            centile_dist_median = median(centile_dist),
            sigma_dist_mad = mad(sigma_dist),
            sigma_dist_median = median(sigma_dist),
            deriv_dist_mad = mad(deriv_dist),
            deriv_dist_median = median(deriv_dist)) %>%
  mutate(centile_dist_threshold = centile_dist_median+(1*centile_dist_mad),
         sigma_dist_threshold = sigma_dist_median+(1*sigma_dist_mad),
         deriv_dist_threshold = deriv_dist_median+(1*deriv_dist_mad))

ggplot(dist_df) +
  geom_histogram(aes(x=centile_sim), alpha=.5, fill="purple") +
  geom_histogram(aes(x=sigma_sim), alpha=.5, fill="green") +
  geom_histogram(aes(x=deriv_sim), alpha=.5, fill="orange")

dist_df %>%
  summarise(n_pass_cent = sum(centile_dist <= thresholds$centile_dist_threshold),
            n_pass_sigma = sum(sigma_dist <= thresholds$sigma_dist_threshold),
            n_pass_deriv = sum(deriv_dist <= thresholds$deriv_dist_threshold),
            n = n()) %>%
  mutate(pct_pass_cent = (n_pass_cent/n)*100,
         pct_pass_sigma = (n_pass_sigma/n)*100,
         pct_pass_deriv = (n_pass_deriv/n)*100)
  
```


```{r}
sexdiffs_tbv2 <- full_join(sexdiffs_tbv, dist_df)

# Define your breakpoints
low_value <- min(sexdiffs_tbv2$centile_dist, na.rm=T)
threshold <- thresholds$centile_dist_threshold
high_value <- max(sexdiffs_tbv2$centile_dist, na.rm=T)

cent_dist_plt <- sexdiffs_tbv2 %>%
    dplyr::select(pheno, logAge_days, pheno_cat, cv_sample, centile_M_minus_F_z, centile_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=centile_dist)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    theme_linedraw() +
    scale_color_gradientn(
  colors = c("#7EE036","#468815FF",  "#1B599DFF", "#1D2C60FF"),
  values = scales::rescale(c(low_value, threshold, threshold, high_value)),
  limits = c(low_value, high_value)
    ) +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

cent_dist_plt

#SIGMA
# Define your breakpoints
low_value <- min(sexdiffs_tbv2$sigma_dist, na.rm=T)
threshold <- thresholds$sigma_dist_threshold
high_value <- max(sexdiffs_tbv2$sigma_dist, na.rm=T)

cent_dist_plt_sig <- sexdiffs_tbv2 %>%
    dplyr::select(logcv_M_div_F, pheno, logAge_days, pheno_cat, cv_sample, sigma_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=sigma_dist)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    scale_color_gradientn(
  colors = c("#7EE036","#468815FF",  "#1B599DFF", "#1D2C60FF"),
  values = scales::rescale(c(low_value, threshold, threshold, high_value)),
  limits = c(low_value, high_value)
    ) +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Sex-Difference in Variability (lnCVR)")

cent_dist_plt_sig


#DERIV
low_value <- min(sexdiffs_tbv2$deriv_dist, na.rm=T)
threshold <- thresholds$deriv_dist_threshold
high_value <- max(sexdiffs_tbv2$deriv_dist, na.rm=T)

cent_dist_plt_d <- sexdiffs_tbv2 %>%
    dplyr::select(pheno, logAge_days, pheno_cat, cv_sample, deriv_M_minus_F_z, deriv_dist) %>%
    na.omit() %>%
    arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=deriv_dist)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    scale_color_gradientn(
  colors = c("#7EE036","#468815FF",  "#1B599DFF", "#1D2C60FF"),
  values = scales::rescale(c(low_value, threshold, threshold, high_value)),
  limits = c(low_value, high_value)
    ) +
    theme_linedraw() +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="Sex-Difference in Growth")

cent_dist_plt_d
```
re-assessing peak male-female differences only in phenos with consistent derivatives

```{r}
consistent_deriv <- dist_df %>% 
  filter(deriv_dist <= thresholds$deriv_dist_threshold) %>%
  pull(pheno)

keep_list <- intersect(consistent_deriv, consistent_agesex)

peak_diffs_sum %>%
  filter(pheno %in% keep_list) %>%
  mutate(peak_age_yrs = (un_log(peak_age)-280)/365.25) %>%
  ggplot() +
  geom_histogram(aes(x=peak_age_yrs, fill=sex), position="identity", alpha=.4) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  geom_vline(xintercept = 50, color='gray') +
  geom_vline(xintercept = 9, color='gray') +
  geom_vline(xintercept = 18, color='gray') +
  geom_vline(xintercept = 0, color='gray')

peak_plt <- sexdiffs_tbv %>%
    filter(pheno %in% keep_list) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_diffs_sum, pheno %in% keep_list), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference")

ggplotly(peak_plt)
```

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)
Calculating across age, not log-age, for interpretability

```{r}
dominance_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = rs_integral(logAge_days, centile_M_minus_F_z),
            sigma_int.unlog =  rs_integral(logAge_days, logcv_M_div_F),
            age_range = max(age_days) - min(age_days)) %>%
  # rescaling integrals to 100 years
  mutate(centile_int_norm.unlog = (centile_int.unlog/age_range)*365.25*100,
         sigma_int_norm.unlog = (sigma_int.unlog/age_range)*365.25*100) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5)) +
  coord_flip()

dominance_df %>%
  group_by(pheno_cat, cv_sample) %>% 
  count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()
```


```{r}
dominance_plt_df <- full_join(LR.df_tot_sex, dominance_df %>% dplyr::select(!pheno_cat))
dominance_plt_df %>%
  mutate(dominance = ifelse(sig_fdr==TRUE, dominance, "NA")) %>%
  group_by(cv_sample) %>%
  summarise(F_dom = sum(dominance=="Female")/n(),
            M_dom = sum(dominance=="Male")/n())
```

plot dominance on brain

```{r}
#removing ventricles for visualization:
dominance_plt_df_novent <- dominance_plt_df %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_novent$centile_int_norm.unlog)

full_dom_plt <- plot_cv_brain(dominance_plt_df_novent, centile_int_norm.unlog, "Normalized\nSex\nDifference", "ggthemes::Orange-Blue Diverging", c(-0.07, 0.07), c(1.03, 4.5, .93))

ggsave("figs/supplement/dominance_tbvTRUE.png", full_dom_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```

How much of the lifespan are phenos male vs female dominant?

```{r}
#define function
prop_above <- function(x, y, threshold=0){
  crosses <- find_threshold_crosses(y, x, threshold, return="x", return_endpoints = FALSE)
  crosses <- c(crosses, min(x), max(x)) %>% 
    sort()
  #overall age modeled
  total_range <- max(x) - min(x)
  
  ranges <- diff(crosses)
  
  #if y > threshold at min(x) -> sum odd number ranges
  odd_indices_logical <- (1:length(ranges)) %% 2 == 1
  
  if (y[1] > threshold) {
    items_at_odd_indices <- ranges[odd_indices_logical]
    time_above <- sum(items_at_odd_indices)
  } else {
    items_at_even_indices <- ranges[!odd_indices_logical]
    time_above <- sum(items_at_even_indices)
  }
  
  prop_above <- time_above/total_range
  
  return(prop_above)
}

# #test function
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="sGMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
# 
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="WMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
# 
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="A" & pheno=="GMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
# 
# test <- sexdiffs_tbv %>%
#   filter(cv_sample =="B" & pheno=="GMV")
# prop_above(test$age_days, test$centile_M_minus_F_z)
# ggplot(test) +
#   geom_line(aes(x=age_days, y=centile_M_minus_F_z))
```

```{r}
prop_above_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, centile_M_minus_F_z)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_df$cv_sample, prop_above_df$class)

prop_above_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_df_novent <- full_join(LR.df_tot_sex, prop_above_df %>% dplyr::select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_plt <- plot_cv_brain(prop_above_df_novent, prop_male, "Proportion\nMale-\nDominant", "ggthemes::Orange-Blue Diverging", c(0, 1), c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_tbvTRUE.png", full_above_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```


####Sigma

```{r}
range(dominance_plt_df_novent$sigma_int_norm.unlog)

full_dom_plt_sigma <- plot_cv_brain(dominance_plt_df_novent, sigma_int_norm.unlog, "Normalized\nSigma Sex\nDifference", "ggthemes::Orange-Blue Diverging", c(-.03, .03), c(1.03, 4.5, .93))

ggsave("figs/supplement/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, dpi=600, height=8.5, width=7.5, bg = "white")
```

```{r}
prop_above_sigma_df <- sexdiffs_tbv %>%
  group_by(pheno, cv_sample, pheno_cat) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, logcv_M_div_F)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male == 0 ~ "all_female",
    prop_male == 1 ~ "all_male",
    TRUE ~ "mixed")
  )

table(prop_above_sigma_df$cv_sample, prop_above_sigma_df$class)

prop_above_sigma_df %>%
  group_by(cv_sample) %>%
  summarise(n_female = sum(class=="all_female"),
            n_male = sum(class=="all_male"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_sigma_df_novent <- full_join(LR.df_tot_sex, prop_above_sigma_df %>% dplyr::select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

full_above_sig_plt <- plot_cv_brain(prop_above_sigma_df_novent, prop_male, "Proportion\nMale-\nDominant", "ggthemes::Orange-Blue Diverging", c(0, 1), c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_sigma_tbvTRUE.png", full_above_sig_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```


Prep for fig 3

```{r}
dominance_plt_df_B <- dominance_plt_df_novent %>%
  filter(cv_sample=="B")

#centiles
full_dom_pltB <- plot_cv_brain(dominance_plt_df_B, centile_int_norm.unlog, "Normalized\nSex\nDifference", "ggthemes::Orange-Blue Diverging", c(-0.07, 0.07), c(1.03, 4.5, .93), return_plt = FALSE)
full_dom_pltB[[4]] <- get_legend(
  ggplot(dominance_plt_df_B) +
    geom_point(aes(x=cv_sample, y=pheno, color=centile_int_norm.unlog)) +
     paletteer::scale_color_paletteer_c(
        "ggthemes::Orange-Blue Diverging",
        name = "Normalized\nSex Difference",
        limits = c(-0.07, 0.07),
        direction = -1,
        na.value = "gray90"
      )
  )
saveRDS(full_dom_pltB, "figs/ggobjs/cent_dom_B.rds")

#sigma
full_dom_plt_sigmaB <- plot_cv_brain(dominance_plt_df_B, sigma_int_norm.unlog, "Normalized\nSigma Sex\nDifference", "ggthemes::Orange-Blue Diverging", c(-0.07, 0.07), c(1.03, 4.5, .93), return_plt = FALSE)
saveRDS(full_dom_plt_sigmaB, "figs/ggobjs/sig_dom_B.rds")
```

###Shape Differences
*At what points in the lifespan are males and females growing differently?*
difference of first derivatives!
can also get magnitude differences

int > 0 -> males grow faster than females
int < 0 -> females grow faster
int = 0 -> same rate

```{r}
deriv_df <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(faster_rate = case_when(
    deriv_M_minus_F_z.unlog > 0 ~ "Males",
    deriv_M_minus_F_z.unlog == 0 ~ "Same",
    deriv_M_minus_F_z.unlog < 0 ~ "Females",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- deriv_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = deriv_M_minus_F_z.unlog, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

deriv_df %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=faster_rate)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", deriv_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))
```

do times with greatest differences in growth rate line up with puberty, menopause, etc? 

```{r}
peak_deriv_diffs <- sexdiffs_tbv %>%
  group_by(cv_sample, pheno, pheno_cat) %>%
  summarise(male_peak = max_gc(deriv_M_minus_F_z, na.rm = TRUE),
            male_peak_age = extr_val(logAge_days, which_max_gc(deriv_M_minus_F_z)),
            female_peak = min_gc(deriv_M_minus_F_z, na.rm = TRUE),
            female_peak_age = extr_val(logAge_days, which_min_gc(deriv_M_minus_F_z))
            ) %>%
  #remove peaks that don't correspond to actual bias
  mutate(male_peak = ifelse(male_peak>0, male_peak, NA),
         female_peak = ifelse(female_peak<0, female_peak, NA)) %>%
  pivot_longer(cols = matches("^(male|female)_"),
    names_to = c("sex", ".value"),
    names_pattern = "(male|female)_(.*)")

peak_deriv_hist <- peak_deriv_diffs %>%
  filter(pheno %in% keep_list) %>%
  mutate(peak_age_yrs = (un_log(peak_age)-280)/365.25) %>%
  ggplot() +
  geom_histogram(aes(x=peak_age_yrs, fill=sex), position="identity", alpha=.4) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  geom_vline(xintercept = 50, color='gray') +
  geom_vline(xintercept = 9, color='gray') +
  geom_vline(xintercept = 18, color='gray') +
  geom_vline(xintercept = 0, color='gray')

peak_deriv_hist

ggplotly(peak_deriv_hist)

yr_to_logAge_days <- function(x){
  log((x*365.25 + 280), base=10)
}

peak_plt_deriv <- sexdiffs_tbv %>%
    filter(pheno %in% keep_list) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=deriv_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    geom_point(data = dplyr::filter(peak_deriv_diffs, pheno %in% keep_list), 
             aes(x = peak_age, y = peak, color=pheno, shape=sex),
             size = 2) +
    theme_linedraw() +
    theme(legend.position = "none") +
    format_x_axis("log_lifespan_fetal", cent_df$logAge_days) +
    labs(y="50th Centile Sex-Difference") +
    geom_vline(xintercept = yr_to_logAge_days(50), color='gray') +
    geom_vline(xintercept = yr_to_logAge_days(9), color='gray') +
    geom_vline(xintercept = yr_to_logAge_days(18), color='gray') +
    geom_vline(xintercept = yr_to_logAge_days(0), color='gray')

peak_plt_deriv
```

###Grouping
- Male, becoming more biased
- Male, becoming less biased
- Female, becoming more biased
- Female, becoming less biased
- Male to Female
- Female to Male

using integral of deriv of centile diff squared (from convergence vs divergence analyses above)

```{r}
fifty_yrs <- log((50*365.25)+280, base=10)

#collect info
group_df <- sexdiffs_tbv %>%
  mutate(fifty_plus = ifelse(logAge_days >= fifty_yrs, "late", "early")) %>%
  arrange(age_days) %>%
  group_by(pheno, pheno_cat, fifty_plus, cv_sample) %>%
  summarise(centile_int.unlog = rs_integral(logAge_days, centile_M_minus_F_z),
            sigma_int.unlog = rs_integral(logAge_days, logcv_M_div_F)) %>%
  mutate(cent_bias = ifelse(centile_int.unlog > 0 , "M", "F"),
         sigma_bias = ifelse(sigma_int.unlog > 0, "M", "F")) %>%
  pivot_wider(id_cols=c(pheno, pheno_cat, cv_sample), names_from=fifty_plus, values_from=c(cent_bias, sigma_bias, centile_int.unlog, sigma_int.unlog))

#now categorize
group_sum <- full_join(group_df, convergence_int) %>%
  mutate(cent_group = case_when(
    cent_bias_early == "M" & cent_bias_late == "F" ~ "M_to_F",
    cent_bias_early == "F" & cent_bias_late == "M" ~ "F_to_M",
    cent_bias_early == "M" & cent_bias_late == "M" & overall_trajectory == "diverging" ~ "M_increasing",
    cent_bias_early == "M" & cent_bias_late == "M" & overall_trajectory == "converging" ~ "M_decreasing",
    cent_bias_early == "F" & cent_bias_late == "F" & overall_trajectory == "diverging" ~ "F_increasing",
    cent_bias_early == "F" & cent_bias_late == "F" & overall_trajectory == "converging" ~ "F_decreasing",
  )) %>%
  mutate(cent_group = factor(cent_group,
                             levels=c("M_increasing", "M_decreasing", "M_to_F", "F_to_M", "F_decreasing", "F_increasing"),
                             ordered=TRUE)) %>%
  dplyr::select(pheno, pheno_cat, cv_sample, cent_group)

group_sum %>%
  filter(pheno == "rh.GM.supramarginal")

sexdiffs_tbv %>%
  filter(pheno == "rh.GM.supramarginal" & cv_sample=="B") %>%
  arrange(age_days) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z)) +
  geom_vline(xintercept=un_log(fifty_yrs))
```

```{r}
group_color_scale <- scale_colour_manual(
  values = c("M_increasing"="#C06636FF", "M_decreasing"="#CE9344FF", "M_to_F" ="#E8B960FF", "F_to_M"="#7BA0B4FF", "F_decreasing"="#44728CFF", "F_increasing"="#17486FFF"),
  breaks = c("M_increasing", "M_decreasing", "M_to_F", "F_to_M", "F_decreasing", "F_increasing"),
  labels = c("Male Diverging", "Male Converging", "Male to Female", "Female to Male", "Female Converging", "Female Diverging"),
  name = "Group",
  na.value="gray90"
)

group_fill_scale <- scale_fill_manual(
  values = c("M_increasing"="#C06636FF", "M_decreasing"="#CE9344FF", "M_to_F" ="#E8B960FF", "F_to_M"="#7BA0B4FF", "F_decreasing"="#44728CFF", "F_increasing"="#17486FFF"),
  breaks = c("M_increasing", "M_decreasing", "M_to_F", "F_to_M", "F_decreasing", "F_increasing"),
  labels = c("Male Diverging", "Male Converging", "Male to Female", "Female to Male", "Female Converging", "Female Diverging"),
  name = "Group",
  na.value="gray90"
)
  

ggplot(group_sum) +
  geom_bar(aes(x=pheno_cat, fill=cent_group)) +
  facet_wrap(~cv_sample) + coord_flip() +
  gghighlight(pheno %in% consistent_agesex, calculate_per_facet = TRUE) +
  group_fill_scale

group_plt_df_all <- inner_join(LR.df_tot, group_sum)

group_bar_plt <- group_plt_df_all %>%
  group_by(pheno_cat, cv_sample) %>% 
  mutate(cent_group = as.character(cent_group)) %>%
  mutate(cent_group = ifelse(sig_fdr ==TRUE, cent_group, NA)) %>%
  count(cent_group, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=cent_group)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
  geom_text(aes(label = paste0(scales::percent(perc, accuracy = 0.1), "\n(", n, ")")), 
            position = position_fill(vjust = 0.5),
            size = 1.5) +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  theme_linedraw() +
  theme(axis.title= element_blank(), legend.position = "none",
        strip.text.x = element_blank(), strip.background.x = element_blank()) +
  group_fill_scale

group_bar_plt_total <- group_plt_df_all %>%
  group_by(cv_sample) %>% 
  mutate(cent_group = as.character(cent_group)) %>%
  mutate(cent_group = ifelse(sig_fdr ==TRUE, cent_group, NA)) %>%
  count(cent_group, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x="TOTAL", y=perc, fill=cent_group)) +
  geom_col(position="fill") +
  facet_wrap(~cv_sample) +
  geom_text(aes(label = paste0(scales::percent(perc, accuracy = 0.1), "\n(", n, ")")), 
            position = position_fill(vjust = 0.5),
            size = 1.5) +
  coord_flip() +
  theme_linedraw() +
  theme(axis.title= element_blank(), axis.text.x=element_blank()) +
  group_fill_scale

bar_legend <- get_legend(group_bar_plt_total)
group_bar_plt_total <- group_bar_plt_total + theme(legend.position = "none")

group_bar_plt_all <- plot_grid(
  plot_grid(
    group_bar_plt_total, 
    group_bar_plt,
    ncol = 1,
    rel_heights = c(1,4),
    align = "v",
    axis = "lr"),
  bar_legend,
  ncol=2,
  rel_widths = c(4,1)
)

ggsave("figs/supplement/group_bar_plt.png", group_bar_plt_all, dpi=600, height=6, width=10, bg = "white")
```

visualize trajectories

```{r}
sexdiffs_tbv_group <- full_join(sexdiffs_tbv, group_sum)
#median centile
cent_group_df <- sexdiffs_tbv_group %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  dplyr::select(cent_group, pheno, age_days, pheno_cat, cv_sample, centile_M_minus_F_z, samp_pheno) %>%
  na.omit() %>%
  arrange(pheno, age_days)
 
cent_group_plt <- cent_group_df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z, group=pheno, color=cent_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(samp_pheno %in% sig_agesex, calculate_per_facet = TRUE, use_direct_label=FALSE) +
  theme_linedraw() +
  format_x_axis("lifespan_fetal", cent_group_df$age_days) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust=1)) +
  group_color_scale +
  labs(y="Trajectory Sex-Difference")

ggsave("figs/supplement/centile_groupings.png", cent_group_plt, width=9, height=9, dpi=600)
#ggplotly(cent_group_plt)
```

also plot just B
```{r}
cent_group_pltB <- cent_group_df %>%
  filter(cv_sample=="B") %>%
  ggplot() +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z, group=pheno, color=cent_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(samp_pheno %in% sig_agesex, calculate_per_facet = TRUE, use_direct_label=FALSE) +
  theme_linedraw() +
  format_x_axis("lifespan_fetal", cent_group_df$age_days) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust=1), strip.text.x = element_blank(), strip.background.x = element_blank()) +
  group_color_scale +
  labs(y="50th Centile Sex-Difference")
saveRDS(cent_group_pltB, "figs/ggobjs/cent_groups_B.rds")
```

median lines with uniform points
```{r}
#add group info
uniform_groups <- full_join(group_plt_df_all, sexdiffs.df.u) %>%
  filter(sig_fdr==TRUE) %>%
  group_by(cv_sample, pheno_cat, cent_group, age_days) %>%
  summarise(median_traj = median(centile_M_minus_F_z, na.rm = TRUE),
            mean_traj = mean(centile_M_minus_F_z, na.rm = TRUE),
    .groups = "drop"
  )

med_trajectory_plt <- uniform_groups %>% 
  ggplot(aes(x = age_days, y = median_traj, color = cent_group)) +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(method = "gam") +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  group_color_scale +
  group_fill_scale +
  theme_linedraw() +
  format_x_axis("lifespan", uniform_groups$age_days) +
  labs(y="Median Sex-Difference")
ggsave("figs/supplement/med_centile_groupings.png", med_trajectory_plt, width=9, height=9, dpi=600)

#also just B
med_trajectory_pltB <- uniform_groups %>%
  filter(cv_sample=="B" & pheno_cat != "Ventricular Vol") %>%
  ggplot(aes(x = age_days, y = median_traj, color = cent_group)) +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(method = "gam") +
  facet_wrap(~pheno_cat, scales="free_y", ncol=1) +
  group_color_scale +
  group_fill_scale +
  theme_linedraw() +
  format_x_axis("lifespan", uniform_groups$age_days) +
  labs(y="Median Sex-Difference")
saveRDS(med_trajectory_pltB, "figs/ggobjs/cent_groups_median_B.rds")
```

trying again with clustering

```{r}
sexdiffs.df.u_centB <- full_join(LR.df_tot, sexdiffs.df.u) %>%
  filter(sig_fdr==TRUE & total==TRUE & weighted==FALSE & cv_sample=="B") %>%
  select(pheno, logAge_days, centile_M_minus_F_z)


```

visualize with significance from age*sex effects
````{r}
group_plt_df <- group_plt_df_all %>%
  filter(!(pheno %in% vent_list))

group_brain_plt <- plot_cv_brain(group_plt_df, cent_group, "Group", "ggthemes::Classic_Green_Orange_6", levels(group_plt_df$cent_group), c(1.03, 4.5, .93), return_plt = FALSE)

group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale

group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale

group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale

#recreate legends
legend <- get_legend(
  ggplot(group_plt_df) +
    geom_bar(aes(x=cv_sample, fill=cent_group)) + group_fill_scale
  )

plts <- cowplot::plot_grid(
    group_brain_plt[[1]],
    group_brain_plt[[2]],
    group_brain_plt[[3]],
    ncol = 1,
    nrow = 3,
    rel_heights = c(1.03, 4.5, .93),
    align = "v",
    axis = "lr",
    greedy = TRUE
  ) 
  
full_plt_obj <- plot_grid(plts, legend, ncol=2, rel_widths = c(3, 1))

ggsave("figs/supplement/group_brain_plt.png", full_plt_obj, dpi=600, height=8.5, width=8, bg = "white")
```

plot just B

```{r}
group_plt_dfB <- group_plt_df %>%
  filter(cv_sample=="B")

group_brain_pltB <- plot_cv_brain(group_plt_dfB, cent_group, "Group", "ggthemes::Classic_Green_Orange_6", levels(group_plt_dfB$cent_group), c(1.03, 4.5, .93), return_plt = FALSE)

group_brain_pltB[[1]] <- group_brain_pltB[[1]] + group_fill_scale

group_brain_pltB[[2]] <- group_brain_pltB[[2]] + group_fill_scale

group_brain_pltB[[3]] <- group_brain_pltB[[3]] + group_fill_scale

group_brain_pltB[[4]] <- get_legend(
  ggplot(group_plt_df) +
    geom_bar(aes(x=cv_sample, fill=cent_group)) + group_fill_scale
  )

saveRDS(group_brain_pltB, "figs/ggobjs/cent_groups_brain_B.rds")
```

####Sigma

```{r}
sigma_group_sum <- inner_join(group_df, sigma_convergence_df) %>%
  mutate(sigma_group = case_when(
    sigma_bias_early == "M" & sigma_bias_late == "F" ~ "M_to_F",
    sigma_bias_early == "F" & sigma_bias_late == "M" ~ "F_to_M",
    sigma_bias_early == "M" & sigma_bias_late == "M" & overall_trajectory == "diverging" ~ "M_increasing",
    sigma_bias_early == "M" & sigma_bias_late == "M" & overall_trajectory == "converging" ~ "M_decreasing",
    sigma_bias_early == "F" & sigma_bias_late == "F" & overall_trajectory == "diverging" ~ "F_increasing",
    sigma_bias_early == "F" & sigma_bias_late == "F" & overall_trajectory == "converging" ~ "F_decreasing",
  )) %>%
  mutate(sigma_group = factor(sigma_group,
                             levels=c("M_increasing", "M_decreasing", "M_to_F", "F_to_M", "F_decreasing", "F_increasing"),
                             ordered=TRUE)) %>%
  dplyr::select(pheno, pheno_cat, cv_sample, sigma_group)
```

```{r}
sexdiffs_tbv_sigma_group <- inner_join(sexdiffs_tbv, sigma_group_sum)

sigma_group_df <- sexdiffs_tbv_sigma_group %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  dplyr::select(sigma_group, pheno, age_days, pheno_cat, cv_sample, logcv_M_div_F, samp_pheno) %>%
  na.omit() %>%
  arrange(pheno, age_days)
 
sigma_group_plt <- sigma_group_df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=logcv_M_div_F, group=pheno, color=sigma_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(samp_pheno %in% sig_agesex, calculate_per_facet = TRUE, use_direct_label=FALSE) +
  theme_linedraw() +
  format_x_axis("lifespan_fetal", sigma_group_df$age_days) +
  group_color_scale +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust=1))+
  labs(y="log CV Ratio")
ggsave("figs/supplement/sigma_groupings.png", sigma_group_plt, width=9, height=9, dpi=600)

#ggplotly(sigma_group_plt)
```


```{r}
sigma_group_pltB <- sigma_group_df %>%
  filter(cv_sample=="B") %>%
  ggplot() +
  geom_line(aes(x=age_days, y=logcv_M_div_F, group=pheno, color=sigma_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  gghighlight(samp_pheno %in% sig_agesex, calculate_per_facet = TRUE, use_direct_label=FALSE) +
  theme_linedraw() +
  format_x_axis("lifespan_fetal", sigma_group_df$age_days) +
  group_color_scale +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust=1), strip.text.x = element_blank(), strip.background.x = element_blank())+
  labs(y="log CV Ratio")
saveRDS(sigma_group_pltB, "figs/ggobjs/sig_groups_B.rds")
```

median lines with uniform points
```{r}
#add group info
sigma_uniform_groups <- sigma_group_df %>%
  filter(samp_pheno %in% sig_agesex) %>%
  group_by(cv_sample, pheno_cat, sigma_group, age_days) %>%
  summarise(median_traj = median(logcv_M_div_F, na.rm = TRUE),
    .groups = "drop"
  )

med_trajectory_plt <- sigma_uniform_groups %>% 
  ggplot(aes(x = age_days, y = median_traj, color = sigma_group)) +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(method = "gam") +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  group_color_scale +
  group_fill_scale +
  theme_linedraw() +
  format_x_axis("lifespan", sigma_uniform_groups$age_days) +
  labs(y="Median Sex-Difference")
ggsave("figs/supplement/med_sigma_groupings.png", med_trajectory_plt, width=9, height=9, dpi=600)

#also just B
med_trajectory_pltB <- sigma_uniform_groups %>%
  filter(cv_sample=="B" & pheno_cat != "Ventricular Vol") %>%
  ggplot(aes(x = age_days, y = median_traj, color = sigma_group)) +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(method = "gam") +
  facet_wrap(~pheno_cat, scales="free_y", ncol=1) +
  group_color_scale +
  group_fill_scale +
  theme_linedraw() +
  format_x_axis("lifespan", sigma_uniform_groups$age_days) +
  labs(y="Median Sex-Difference")
saveRDS(med_trajectory_pltB, "figs/ggobjs/sigma_groups_median_B.rds")
```

visualize with significance from age*sex effects

````{r}
sigma_group_plt_df <- inner_join(LR.df_tot, sigma_group_df) %>%
  filter(!(pheno %in% vent_list))

sigma_group_plt <- plot_cv_brain(sigma_group_plt_df, sigma_group, "Group", "ggthemes::Classic_Green_Orange_6", levels(sigma_group_plt_df$sigma_group), c(1.03, 4.5, .93), return_plt = FALSE)

sigma_group_plt[[1]] <- sigma_group_plt[[1]] + group_fill_scale

sigma_group_plt[[2]] <- sigma_group_plt[[2]] + group_fill_scale

sigma_group_plt[[3]] <- sigma_group_plt[[3]] + group_fill_scale

sig.plts <- cowplot::plot_grid(
    sigma_group_plt[[1]],
    sigma_group_plt[[2]],
    sigma_group_plt[[3]],
    ncol = 1,
    nrow = 3,
    rel_heights = c(1.03, 4.5, .93),
    align = "v",
    axis = "lr",
    greedy = TRUE
  ) 
  
sig.full_plt_obj <- plot_grid(sig.plts, legend, ncol=2, rel_widths = c(3, 1))

ggsave("figs/supplement/sigma_group_brain_plt.png", sig.full_plt_obj, dpi=600, height=8.5, width=8, bg = "white")
```

####Derivative

```{r}
deriv_group_df <- sexdiffs_tbv %>%
  na.omit() %>%
  mutate(fifty_plus = ifelse(logAge_days >= fifty_yrs, "late", "early")) %>%
  arrange(age_days) %>%
  group_by(pheno, pheno_cat, fifty_plus, cv_sample) %>%
  summarise(deriv_int.unlog = rs_integral(logAge_days, deriv_M_minus_F_z)) %>%
  mutate(deriv_bias = ifelse(deriv_int.unlog > 0 , "M", "F")) %>%
  pivot_wider(id_cols=c(pheno, pheno_cat, cv_sample), names_from=fifty_plus, values_from=c(deriv_bias, deriv_int.unlog))

deriv_group_sum <- inner_join(deriv_group_df, deriv_convergence_df) %>%
  mutate(deriv_group = case_when(
    deriv_bias_early == "M" & deriv_bias_late == "F" ~ "M_to_F",
    deriv_bias_early == "F" & deriv_bias_late == "M" ~ "F_to_M",
    deriv_bias_early == "M" & deriv_bias_late == "M" & overall_trajectory == "diverging" ~ "M_increasing",
    deriv_bias_early == "M" & deriv_bias_late == "M" & overall_trajectory == "converging" ~ "M_decreasing",
    deriv_bias_early == "F" & deriv_bias_late == "F" & overall_trajectory == "diverging" ~ "F_increasing",
    deriv_bias_early == "F" & deriv_bias_late == "F" & overall_trajectory == "converging" ~ "F_decreasing",
  )) %>%
  mutate(deriv_group = factor(deriv_group,
                             levels=c("M_increasing", "M_decreasing", "M_to_F", "F_to_M", "F_decreasing", "F_increasing"),
                             ordered=TRUE)) %>%
  dplyr::select(pheno, pheno_cat, cv_sample, deriv_group)
```

```{r}
sexdiffs_tbv_deriv_group <- inner_join(sexdiffs_tbv, deriv_group_sum)

deriv_group_df <- sexdiffs_tbv_deriv_group %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  dplyr::select(deriv_group, pheno, age_days, pheno_cat, cv_sample, deriv_M_minus_F_z, samp_pheno) %>%
  na.omit() %>%
  arrange(pheno, age_days)
 
deriv_group_plt <- deriv_group_df %>%
  ggplot() +
  geom_line(aes(x=age_days, y=deriv_M_minus_F_z, group=pheno, color=deriv_group)) +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  theme_linedraw() +
  format_x_axis("lifespan_fetal", deriv_group_df$age_days) +
    gghighlight(samp_pheno %in% sig_agesex, calculate_per_facet = TRUE, use_direct_label=FALSE) +
  group_color_scale +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust=1))+
  labs(y="Male-Female Derivative")

#testing
test_plt <- deriv_group_df %>%
  filter(pheno %in% ct_list) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=deriv_M_minus_F_z, group=pheno, color=deriv_group)) +
  facet_grid(deriv_group~cv_sample, scales="free_y") +
  theme_linedraw() +
  format_x_axis("lifespan_fetal", deriv_group_df$age_days) +
    gghighlight(samp_pheno %in% sig_agesex, calculate_per_facet = TRUE, use_direct_label=FALSE) +
  group_color_scale +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust=1))+
  labs(y="Male-Female Derivative")

ggplotly(test_plt)
```


median lines with uniform points
```{r}
#add group info
deriv_uniform_groups <- deriv_group_df %>%
  filter(samp_pheno %in% sig_agesex) %>%
  group_by(cv_sample, pheno_cat, deriv_group, age_days) %>%
  summarise(median_traj = median(deriv_M_minus_F_z, na.rm = TRUE),
    .groups = "drop"
  )

med_trajectory_plt <- deriv_uniform_groups %>% 
  ggplot(aes(x = age_days, y = median_traj, color = deriv_group)) +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(method = "gam") +
  facet_grid(pheno_cat~cv_sample, scales="free_y") +
  group_color_scale +
  group_fill_scale +
  theme_linedraw() +
  format_x_axis("lifespan", deriv_uniform_groups$age_days) +
  labs(y="Derivative Sex-Difference")
ggsave("figs/supplement/med_deriv_groupings.png", med_trajectory_plt, width=9, height=9, dpi=600)

#also just B
med_trajectory_pltB <- deriv_uniform_groups %>%
  filter(cv_sample=="B" & pheno_cat != "Ventricular Vol") %>%
  ggplot(aes(x = age_days, y = median_traj, color = deriv_group)) +
  geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
  geom_smooth(method = "gam") +
  facet_wrap(~pheno_cat, scales="free_y", ncol=1) +
  group_color_scale +
  group_fill_scale +
  theme_linedraw() +
  format_x_axis("lifespan", deriv_uniform_groups$age_days) +
  labs(y="Derivative Sex-Difference")
saveRDS(med_trajectory_pltB, "figs/ggobjs/deriv_groups_median_B.rds")
```


````{r}
deriv_group_plt_df <- inner_join(LR.df_tot, deriv_group_df) %>%
  filter(!(pheno %in% vent_list))

deriv_group_plt <- plot_cv_brain(deriv_group_plt_df, deriv_group, "Group", "ggthemes::Classic_Green_Orange_6", levels(deriv_group_plt_df$deriv_group), c(1.03, 4.5, .93), return_plt = FALSE)

deriv_group_plt[[1]] <- deriv_group_plt[[1]] + group_fill_scale

deriv_group_plt[[2]] <- deriv_group_plt[[2]] + group_fill_scale

deriv_group_plt[[3]] <- deriv_group_plt[[3]] + group_fill_scale

sig.plts <- cowplot::plot_grid(
    deriv_group_plt[[1]],
    deriv_group_plt[[2]],
    deriv_group_plt[[3]],
    ncol = 1,
    nrow = 3,
    rel_heights = c(1.03, 4.5, .93),
    align = "v",
    axis = "lr",
    greedy = TRUE
  ) 
  
sig.full_plt_obj <- plot_grid(sig.plts, legend, ncol=2, rel_widths = c(3, 1))

ggsave("figs/supplement/deriv_group_brain_plt.png", sig.full_plt_obj, dpi=600, height=8.5, width=8, bg = "white")
```

plot just B

```{r}
deriv_group_plt_dfB <- deriv_group_plt_df %>%
  filter(cv_sample=="B")

deriv_group_pltB <- plot_cv_brain(deriv_group_plt_dfB, deriv_group, "Group", "ggthemes::Classic_Green_Orange_6", levels(deriv_group_plt_dfB$cent_group), c(1.03, 4.5, .93), return_plt = FALSE)

deriv_group_pltB[[1]] <- deriv_group_pltB[[1]] + group_fill_scale

deriv_group_pltB[[2]] <- deriv_group_pltB[[2]] + group_fill_scale

deriv_group_pltB[[3]] <- deriv_group_pltB[[3]] + group_fill_scale

deriv_group_pltB[[4]] <- get_legend(
  ggplot(deriv_group_plt_dfB) +
    geom_bar(aes(x=cv_sample, fill=deriv_group)) + group_fill_scale +
    theme(legend.position="bottom")
  )

saveRDS(deriv_group_pltB, "figs/ggobjs/deriv_groups_brain_B.rds")
```