---
title: "Review Test Models"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, sf, cowplot, subcortexVisualizationR, rstatix, freesurferformats, mgcv)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*/"
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

# install.packages("remotes")
#remotes::install_github("BGDlab/ggsegTissue", force=TRUE)
library(ggsegTissue)

#source spin test
source("https://raw.githubusercontent.com/frantisekvasa/rotate_parcellation/refs/heads/master/R/perm.sphere.p.R")
source("https://raw.githubusercontent.com/frantisekvasa/rotate_parcellation/refs/heads/master/R/rotate.parcellation.R")
fs_path <- "/mnt/isilon/bgdlab_processing/code/freesurfer-6.0/subjects/fsaverage/"
```

```{r}
source("code/plot_cv_brain.R")
```

#Defining integral funs

```{r}
#fun for derivative
get_deriv <- function(x, age) {
  c(NA, diff(x) / diff(age))
}

#riemann_stieltjes - updating default fun based on claude/convo with taki
rs_integral <- function(x, y, g = function(x) 10^(x)) {
  stopifnot(
    length(x) == length(y),
    length(x) >= 2
  )

  # sort just in case
  ord <- order(x)
  x <- x[ord]
  y <- y[ord]

  gx <- g(x)

  # monotonicity check (important!)
  if (any(diff(gx) < 0)) {
    stop("g(x) must be non-decreasing for a Riemann–Stieltjes integral")
  }

  sum(y[-length(y)] * diff(gx))
}

#euclidean distance fun from https://www.geeksforgeeks.org/r-language/how-to-calculate-euclidean-distance-in-r/
euclidean_dist <- function(vect1, vect2) sqrt(sum((vect1 - vect2)^2))
```

#Read & Prep CSVs
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "model_sums/*LRtest.csv")) %>% unique()

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")

length(output_fread)

LR.df <- bind_rows(output_fread, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted  = stringr::str_detect(file, "weight"),
         age2plus = stringr::str_detect(file, "age2plus"))%>%
  dplyr::select(!file)
```

```{r, eval=FALSE}
LR.df %>%
  group_by(cv_sample, total, weighted) %>%
  summarise(n_phenos = n_unique(pheno))

A_totalFALSE <- LR.df %>%
  filter(cv_sample == "A" & total == "FALSE" & weighted == "FALSE" & age2plus=="FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalFALSE)

A_totalTRUE <- LR.df %>%
  filter(cv_sample == "A" & total == "TRUE" & weighted == "FALSE" & age2plus=="FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, A_totalTRUE)

B_totalFALSE <- LR.df %>%
  filter(cv_sample == "B" & total == "FALSE" & weighted == "FALSE" & age2plus=="FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalFALSE)

B_totalTRUE <- LR.df %>%
  filter(cv_sample == "B" & total == "TRUE" & weighted == "FALSE" & age2plus=="FALSE") %>%
  pull(pheno) %>%
  unique()
setdiff(pheno_list, B_totalTRUE)
```

Get RESI effect size

```{r}
LR.df2 <- LR.df %>%
  mutate(resi = fsq2S(fsq),
    pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)))

table(LR.df2$effect, LR.df2$total)
```



```{r}
LR.df_plot <- LR.df2 %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  group_by(total, weighted, cv_sample, age2plus) %>% #apply FDR correction within each combo
  mutate(
    p.val_fdr = p.adjust(
      p_val,
      method = "fdr"
    )
  ) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p.val_fdr <=0.05, TRUE, FALSE)) %>%
  mutate(sig_fsq = ifelse(sig_fdr==TRUE, fsq, NA),
         sig_resi = ifelse(sig_fdr==TRUE, resi, NA))

LR.df_plot$cv_sample <- factor(LR.df_plot$cv_sample, levels = c("A", "B"))

#check cortical
setdiff(unique(dk$data$label), unique(LR.df_plot$label)) #corpus callosum is subdivided - remove from dk plots

#check subcortical
setdiff(unique(aseg$data$label), unique(LR.df_plot$label)) # all included

#check global
setdiff(unique(aseg_all$data$label), unique(LR.df_plot$tissue_class)) # all included

#see what's not included anywhere
setdiff(unique(LR.df_plot$label), c(unique(dk$data$label), unique(aseg$data$label)))
```

"Left-Accumbens-area", "Right-Accumbens-area", "Left-Cerebellum-Cortex" & "Left-Cerebellum-White-Matter" aren't in field of view for aseg atlas


#Non-Corrected Model

```{r}
LR.df_uncor <- LR.df_plot %>%
  dplyr::filter(total==FALSE & weighted == FALSE & age2plus=="FALSE")

LR.df_uncor %>%
  group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T),
            mean_fsq = mean(sig_fsq, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_uncor) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_uncor$sig_fsq, na.rm=T)
range(LR.df_uncor$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_uncor_novent <- LR.df_uncor %>%
  filter(!(pheno %in% vent_list))
range(LR.df_uncor_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_uncor_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)

ggsave("figs/supplement/resi_uncorrected.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

```{r}
#test plotting just split B
LR.df_uncor_novent_B <- LR.df_uncor_novent %>%
  filter(cv_sample=="B")
B_plt_uncor <- plot_cv_brain(df = LR.df_uncor_novent_B, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = FALSE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)
saveRDS(B_plt_uncor, "figs/ggobjs/agebysex_B.rds")
```
###Weighted
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==FALSE) %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total == FALSE & age2plus == FALSE) %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(weighted) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  filter(total==FALSE & weighted==TRUE & age2plus==FALSE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL) #keeping same scale from unweighted models

ggsave("figs/supplement/resi_uncorrected_weight.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

##CT in ages 2+
Sensitivity analyses

```{r}
LR.df_plot %>%
  dplyr::filter(weighted==FALSE & total==FALSE & pheno_cat == "Regional CT") %>%
  group_by(cv_sample, age2plus) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total == FALSE & weighted==FALSE) %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(age2plus) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```

similar but slightly larger S, slightly fewer significant regions

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  filter(total==FALSE & weighted==FALSE & age2plus==TRUE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          #height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = FALSE,
                          include_subcort = FALSE,
                          #rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL) #keeping same scale from unweighted models

ggsave("figs/supplement/resi_uncorrected_age2plus.png", full_plt_uncor, dpi=600, height=3, width=5, bg = "white")
```


#TBV-Corrected Models

##Age*Sex Effect

```{r}
LR.df_tot <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted == FALSE & age2plus == FALSE)

table(LR.df_tot$sig_fdr)

LR.df_tot %>%
group_by(cv_sample) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

ggplot(LR.df_tot) +
  geom_histogram(aes(x=sig_resi, fill=pheno_cat), position='identity', alpha=0.6) +
  facet_wrap(~cv_sample)

range(LR.df_tot$sig_fsq, na.rm=T)
range(LR.df_tot$sig_resi, na.rm=T)
```

```{r}
#removing ventricles for visualization:
LR.df_tot_novent <- LR.df_tot %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_novent$resi)

full_plt_tot <- plot_cv_brain(df = LR.df_tot_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)

ggsave("figs/supplement/resi_tbv.png", full_plt_tot, dpi=600, height=8.5, width=7.1, bg = "white")
```

Note phenos with age*sex effects, per LR tests
```{r}
#consistent age*sex effects in both CV splits
consistent_agesex <- LR.df_tot %>%
  pivot_wider(id_cols=c(pheno, total, weighted, pheno_cat, label), names_from="cv_sample", values_from = c(chi, df, p_val, fsq, effect, resi, p.val_fdr, sig_fdr, sig_fsq, sig_resi)) %>%
  filter(sig_fdr_A == TRUE & sig_fdr_B == TRUE) %>%
  pull(pheno)

#just within each sample
sig_agesex <- LR.df_tot %>%
  mutate(samp_pheno = paste(cv_sample, pheno, sep="_")) %>%
  filter(sig_fdr==TRUE) %>%
  pull(samp_pheno)
```

###Weighted
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & age2plus==FALSE) %>%
  group_by(cv_sample, weighted) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_s = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & age2plus==FALSE) %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(weighted) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```
more regions are significant in weighted analyses

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted==TRUE & age2plus==FALSE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL) #keeping same scale from unweighted models

ggsave("figs/supplement/resi_tbv_weight.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

###CT in ages 2+
Sensitivity analyses with inverse weighting by euler

```{r}
#find phenos where both 2+ and full CT models are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted==FALSE & pheno_cat=="Regional CT") %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(age2plus) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age" & weighted==FALSE & age2plus==TRUE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.1),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = FALSE, 
                          include_subcort = FALSE,
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL) #keeping same scale from unweighted models

ggsave("figs/supplement/resi_tbv_age2plus.png", full_plt_uncor, dpi=600, height=3, width=5, bg = "white")
```

##Sex Effect

```{r}
LR.df_tot_sex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE & age2plus==FALSE)

table(LR.df_tot_sex$sig_fdr, LR.df_tot_sex$cv_sample)

ggplot(LR.df_tot_sex) +
  geom_boxplot(aes(y=sig_resi, fill=pheno_cat)) +
  facet_wrap(~cv_sample)

range(LR.df_tot_sex$sig_fsq, na.rm=T)
range(LR.df_tot_sex$sig_resi, na.rm=T)

LR.df_tot_sex %>%
  filter(sig_fdr == TRUE) %>%
  group_by(cv_sample) %>%
  summarise(min_resi = min(sig_resi),
            max_resi = max(sig_resi),
            mean_resi = mean(sig_resi))
```

```{r}
#removing ventricles for visualization:
LR.df_tot_sex_novent <- LR.df_tot_sex %>%
  filter(!(pheno %in% vent_list))
range(LR.df_tot_sex_novent$resi)

full_plt_tot_sex <- plot_cv_brain(df = LR.df_tot_sex_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)

ggsave("figs/supplement/resi_tbv_allsex.png", full_plt_tot_sex, dpi=600, height=8.5, width=7.1, bg = "white")
```

plot just in B
```{r}
#removing ventricles for visualization:
LR.df_tot_sex_noventB <- LR.df_tot_sex_novent %>%
  filter(cv_sample=="A")
full_plt_tot_sex_B <- plot_cv_brain(df = LR.df_tot_sex_noventB, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = FALSE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL,
                          legend_position = "bottom")
  
saveRDS(full_plt_tot_sex_B, "figs/ggobjs/full_plt_tot_sex_B.rds")
```

###Weighted
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & age2plus == FALSE) %>%
  group_by(cv_sample, weighted) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_resi = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both weighted and unweighted are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & age2plus == FALSE) %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(weighted) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```

RESIs and n significant regions are basically identical across weighted and unweighted models 

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted==TRUE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = TRUE, 
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)
  
ggsave("figs/supplement/resi_tbv_allsex_weight.png", full_plt_uncor, dpi=600, height=8.5, width=7.1, bg = "white")
```

###CT in ages 2+
Sensitivity analyses with inverse weighting by euler

```{r}
LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted == FALSE & pheno_cat == "Regional CT") %>%
  group_by(cv_sample, age2plus) %>%
  summarise(n_sig = sum(sig_fdr==TRUE),
            n=n(),
            mean_resi = mean(sig_resi, na.rm=T)) %>%
  mutate(pct_sig = (n_sig/n)*100)

#find phenos where both 2+ and full CT models are significant
LR.df_agree <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & weighted==FALSE & pheno_cat=="Regional CT") %>%
  group_by(cv_sample, pheno) %>%
  filter(
    all(sig_fdr == TRUE) &        # all rows must be sig
    n_distinct(age2plus) == 2     # both levels of weighted must be present
  ) %>%
  ungroup()

LR.df_agree %>%
  group_by(cv_sample) %>%
  summarise(n=n()/2)
```

RESIs and n significant regions are basically identical

```{r}
#removing ventricles for visualization:
LR.df_w_novent <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all" & age2plus==TRUE & weighted==FALSE) %>%
  filter(!(pheno %in% vent_list))
range(LR.df_w_novent$resi)
full_plt_uncor <- plot_cv_brain(df = LR.df_w_novent, 
                          fill_var = resi, 
                          fill_name = "S", 
                          fill_color =  "grDevices::Greens 2", 
                          fill_limits =  c(0,0.2),
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          include_global = FALSE,
                          include_subcort = FALSE,
                          rel_widths = c(3, .6),
                          direction = -1,
                          fill_values = NULL)
  
ggsave("figs/supplement/resi_tbv_allsex_age2plus.png", full_plt_uncor, dpi=600, height=3, width=5, bg = "white")
```

##Save Table
save out csv with all significance tests

```{r}
LR_table <- LR.df_plot %>%
  select(total, effect, weighted, age2plus, pheno_cat, pheno, cv_sample, chi, df, p_val, fsq, resi, p.val_fdr, sig_fdr) %>%
  rename(total_size_covar = total,
         effect_tested = effect,
         surfhole_weighted = weighted,
         idp_category = pheno_cat,
         idp = pheno,
         p.val = p_val,
         significant_fdr = sig_fdr) %>%
  mutate(effect_tested = ifelse(effect_tested=="sex_age", "age-varying sex component", "any sex effect"))

fwrite(LR_table, "figs/supplement/significance_test_results.csv")
```

##Trajectories
Read in sexdiff csvs
```{r}
sex_diffs_paths <- Sys.glob(file.path(test_path, "cent_csvs/*_sexdiffs.csv"))
sex_diffs_paths <- sex_diffs_paths[stringr::str_detect(sex_diffs_paths, "totalTRUE")] #only total-size-corrected

stopifnot(length(sex_diffs_paths) <= (length(pheno_list)*2*2*2) + (68*2)) #two splits, weighted & unweighted, uniform and not uniform, plus age2plus CT (68 phenos * 2 splits)

# Read all matching CSVs
output_fread_sd <- lapply(sex_diffs_paths, fread)
length(output_fread_sd)
names(output_fread_sd) <- sex_diffs_paths

sexdiffs_tbv_all <- bind_rows(output_fread_sd, .id="file") %>%
  mutate(cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"),
         weighted = stringr::str_detect(file, "weight"),
         age2plus = stringr::str_detect(file, "age2plus"),
         pheno = stringr::str_remove(basename(file), "_(uniform_)?sexdiffs\\.csv"),
         uniform = ifelse(grepl("uniform", file), TRUE, FALSE)) %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% vent_list ~ "Ventricular Vol",
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_)),
    age_days = un_log(logAge_days)) %>%
  dplyr::select(!file) %>%
  group_by(total, weighted, cv_sample, pheno, age2plus) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(#chain rule per taki's suggestion
    deriv_age = get_deriv(logAge_days, age_days)) %>%
  ungroup() %>%
  mutate(deriv_M_minus_F_z.unlog = deriv_M_minus_F_z*deriv_age,
         deriv_logcv_M_div_F.unlog = deriv_logcv_M_div_F*deriv_age)

sexdiffs.df <- sexdiffs_tbv_all %>% filter(uniform==FALSE)
sexdiffs.df.u <- sexdiffs_tbv_all %>% 
  filter(uniform==TRUE & age2plus == FALSE)


rm(output_fread_sd) #save space
```

write out separate dfs
```{r}
sexdiffs_tbv <- sexdiffs_tbv_all %>%
  filter(weighted==FALSE & total==TRUE & age2plus == FALSE)
```

###Plots

Average across splits to make summary plots for global phenos

```{r}
sum_cols <- setdiff(names(sexdiffs.df.u), c("logAge_days", "cv_sample", "total", "weighted", "pheno", "uniform", "pheno_cat", "age_days"))

glob_mean <- sexdiffs.df.u %>%
  filter(pheno %in% global_list & total==TRUE) %>%
  group_by(pheno, age_days) %>%
  #filtering outliers
  filter(age_days >= 370) %>%
  summarise(across(all_of(sum_cols), mean)) %>%
  #rescale
  group_by(pheno) %>%
  mutate(centile_M_minus_F_z = centile_M_minus_F_z / max(abs(centile_M_minus_F_z)) * 100) %>%
  ungroup()

glob_mean_sums <- glob_mean %>%
  summarize(#age at peak sex diff
            max_sex_diff = extr_val(centile_M_minus_F_z, which_max_gc(abs(centile_M_minus_F_z))),
            age_max_sex_diff = extr_val(age_days, which_max_gc(abs(centile_M_minus_F_z))),
            #age peak change in sex diff
            max_deriv = extr_val(centile_M_minus_F_z, which_max_gc(abs(deriv_M_minus_F_z))),
            age_max_deriv = extr_val(age_days, which_max_gc(abs(deriv_M_minus_F_z))),
            #age at male peak
            max_M = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Male_z)),
            age_max_M = extr_val(age_days, which_max_gc(median_centile_Male_z)),
            #age at female peak
            max_F = extr_val(centile_M_minus_F_z, which_max_gc(median_centile_Female_z)),
            age_max_F = extr_val(age_days, which_max_gc(median_centile_Female_z)),
            #age at male peak growth
            max_deriv_M = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Male_z)),
            age_max_deriv_M = extr_val(age_days, which_max_gc(deriv_median_centile_Male_z)),
            #age at female peak growth
            max_deriv_F = extr_val(centile_M_minus_F_z, which_max_gc(deriv_median_centile_Female_z)),
            age_max_deriv_F = extr_val(age_days, which_max_gc(deriv_median_centile_Female_z)),
            .by = pheno
            )

#pivot and rejoin
values_long <- glob_mean_sums %>%
  select(!starts_with("age_")) %>%
  pivot_longer(cols = -c(pheno), names_to = "metric", values_to = "value")

ages_long <- glob_mean_sums %>%
  select(pheno, starts_with("age_")) %>%
  pivot_longer(cols = starts_with("age_"), names_to = "metric", values_to = "age") %>%
  mutate(metric = gsub("^age_", "", metric))

glob_mean_sums_long <- left_join(values_long, ages_long, by = c("pheno", "metric"))

#get age at min sex-diff separately to account for phenos crossing 0 multiple times
# All zero crossings per pheno = "min sex-bias" points.
# At each crossing, take whichever of the two flanking rows has smaller |value|.
# Trajectories with no crossing fall back to the single closest-to-zero row.
min_sex_diff_rows <- glob_mean %>%
  arrange(pheno, age_days) %>%
  group_by(pheno) %>%
  mutate(
    sign_now  = sign(centile_M_minus_F_z),
    sign_next = sign(lead(centile_M_minus_F_z)),
    crosses   = !is.na(sign_next) & sign_now != 0 & sign_next != 0 & sign_now != sign_next,
    # mark the closer-to-zero row of each adjacent crossing pair
    pick_here = crosses & abs(centile_M_minus_F_z) <= abs(lead(centile_M_minus_F_z)),
    pick_next = lag(crosses & abs(centile_M_minus_F_z) >  abs(lead(centile_M_minus_F_z)),
                    default = FALSE),
    keep      = pick_here | pick_next,
    any_cross = any(keep, na.rm = TRUE),
    # fallback: if no crossing, keep the single global min |value|
    keep      = if (any_cross[1]) keep
                else abs(centile_M_minus_F_z) == min(abs(centile_M_minus_F_z), na.rm = TRUE)
  ) %>%
  filter(keep) %>%
  ungroup() %>%
  transmute(
    pheno,
    metric = "min_sex_diff",
    value  = centile_M_minus_F_z,
    age    = age_days
  )

glob_mean_sums_long <- bind_rows(glob_mean_sums_long, min_sex_diff_rows)

#reformat x to keep birth label
reformat_x <- function(x_values, simplified=FALSE){
  tickLabels <- c()
  tickMarks <- c()
  if(simplified==TRUE){
    for (year in c(0, 25, 50, 75, 100)) {
     tickMarks <- append(tickMarks, year * 365.25 + 280)
    }
    tickLabels <- c("Birth", "25", "50", "75", "100")
  } else {
    for (year in seq(0, 100, by = 10)) {
      tickMarks <- append(tickMarks, year * 365.25 + 280)
    }
    tickLabels <- c("Birth", "10", "20", "30", "40", "50", "60", "70", "80", "90", "100")
  }
  
  unit_lab <- "(years)"
  list(scale_x_continuous(breaks = tickMarks, labels = tickLabels), labs(x = paste("Age at Scan", unit_lab)))
}

cent_plt_pct_mean <- glob_mean %>%
  ggplot() +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z, group=pheno, color=pheno), linewidth=1.5) +
  geom_point(
    data = glob_mean_sums_long,
    aes(x = age, y = value, color = pheno, shape = metric),
    size = 4
  ) +
  scale_shape_manual(
    values = c(
      "max_sex_diff" = 19,
      "min_sex_diff" = 15,
      "max_deriv"    = 18,
      "max_M"        = 2,
      "max_F"        = 6,
      "max_deriv_M"  = 3,
      "max_deriv_F"  = 4
    ),
    labels = c(
      "max_sex_diff" = "Max Sex Bias",
      "min_sex_diff" = "Min Sex Bias",
      "max_deriv"    = "Max Change Sex Bias",
      "max_M"        = "Max Male",
      "max_F"        = "Max Female",
      "max_deriv_M"  = "Max Change Male",
      "max_deriv_F"  = "Max Change Female"
    ),
    name = "Metric"
  ) +
  scale_color_discrete(
    labels = c(
      "total.SA" = "Total SA",
      "mean.CT"  = "Mean CT",
      "CBV"      = "Cerebellum",
      "CSF"      = "Ventricles",
      "GMV"      = "Cortical GM",
      "sGMV"     = "Subcortical GM",
      "WMV"      = "WM"
    ),
    name = "Phenotype"
  ) +
  theme_minimal() +
  reformat_x(glob_mean$age_days) +
  labs(y="% Maximum Sex-Difference")
saveRDS(cent_plt_pct_mean, "figs/ggobjs/glob_trajectories.rds")
ggsave("figs/glob_trajectories.png", cent_plt_pct_mean, dpi=600, width=11, height=5)
```

look at sGMV

```{r}
glob_mean_sums_long %>% 
  filter(pheno=="sGMV") %>%
  mutate(age=(age-280)/365.25)
```

###Convergence vs Divergence

*Do sex-differences tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of centile diffs squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
#calculate integral of deriv of diff^2 (all in z-scores)
sexdiffs_tbv_sq <- sexdiffs_tbv_all %>%
  mutate(sq_centile_diff_z = (centile_M_minus_F_z)^2) %>% #square so that positive -> more different, not more male
  group_by(pheno, cv_sample, weighted, age2plus) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_deriv_z = get_deriv(sq_centile_diff_z, logAge_days)) %>% #
  mutate(sq_deriv_z.unlog = sq_deriv_z*deriv_age) %>% #chain rule to convert derivative
  ungroup()

#try getting integral
convergence_int <- sexdiffs_tbv_sq %>%
  group_by(pheno, cv_sample, pheno_cat, weighted, age2plus)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_deriv_z),
            sq_int_z.unlog = rs_integral(logAge_days, sq_deriv_z)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging")) ##Should this be defined on `sq_int_z` or `sq_int_z.unlog`???

#checking
ggplot(convergence_int) +
  geom_point(aes(x=sq_int_z, y= sq_int_z.unlog), alpha=.3)

convergence_int %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(sq_int_z <0 & sq_int_z.unlog >0)

sexdiffs_tbv_sq %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(pheno=="SUBC.Left.Hippocampus" & cv_sample=="B") %>%
  ggplot() +
    geom_line(aes(x=logAge_days, y=sq_centile_diff_z))

sexdiffs_tbv_sq %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(pheno=="SUBC.Left.Hippocampus" & cv_sample=="B") %>%
  ggplot() +
    geom_line(aes(x=age_days, y=sq_centile_diff_z))

convergence_int %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(pheno=="SUBC.Left.Hippocampus" & cv_sample=="B")
```

finding points at which pt derivative crosses 0 increasing vs decreasing

```{r}
cent_df <- sexdiffs_tbv_sq %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  group_by(pheno, cv_sample)%>%
  arrange(age_days, .by_group = TRUE) %>%
  na.omit() %>%
  mutate(trajectory = case_when(
    sq_deriv_z.unlog > 0 ~ "diverging",
    sq_deriv_z.unlog == 0 ~ "constant",
    sq_deriv_z.unlog < 0 ~ "converging",
  )) %>% 
  ungroup()

#order by age first increasing
pheno_order_mean <- cent_df %>%
  group_by(pheno, cv_sample) %>%
  summarize(above_0 = first_above(y = sq_deriv_z, x = age_days, 
                                  threshold = 0, return = "x")) %>%
  group_by(pheno) %>%
  summarise(ord = mean(above_0), .groups = "drop") %>%
  arrange(ord) %>%
  pull(pheno)

cent_df %>%
  filter(age2plus==FALSE & weighted==FALSE) %>%
  mutate(pheno = factor(pheno, levels = pheno_order_mean)) %>%
  ggplot() +
  geom_line(aes(x=age_days, y=pheno, color=trajectory)) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y", space="free_y", switch="y") +
  format_x_axis("lifespan_fetal", cent_df$age_days) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        strip.placement='outside',
        strip.background.y=element_blank(), 
        strip.text.y.left = element_text(angle = 0))

#just look at phenos with consistent age*sex effects across both cvs
cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_histogram(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)

cent_df %>%
  filter(pheno %in% consistent_agesex) %>%
ggplot() +
  geom_density(aes(x=age_days, fill=trajectory), position="identity", alpha=.4) +
  facet_grid(pheno_cat ~ cv_sample, scales="free_y") +
  format_x_axis("lifespan_fetal", cent_df$age_days)
```


####Sigma
*Do sex-differences in variance (sigma) tend to increase, decrease, or remain constant across the lifespan?*

Taking derivative of lnCVR squared:
- positive derivative -> increasing sex diffs
- negative derivative -> decreasing sex diffs
- 0 derivative -> constant sex diffs

```{r}
sigma_df <- sexdiffs_tbv_all %>%
  arrange(pheno, logAge_days) %>%
  mutate(sq_cv_diff = (logcv_M_div_F)^2) %>%
  group_by(pheno, cv_sample, weighted, age2plus) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  mutate(sq_cv_deriv = get_deriv(sq_cv_diff, logAge_days)) %>%
  ungroup() %>%
  mutate(sq_cv_deriv.unlog = sq_cv_deriv*deriv_age)

#try getting integral
sigma_convergence_df <- sigma_df %>%
  group_by(pheno, cv_sample, pheno_cat, weighted, age2plus)%>%
  arrange(logAge_days, .by_group = TRUE) %>%
  na.omit() %>%
  summarise(sq_int_z = trapz(logAge_days, sq_cv_deriv),
            sq_int_z.unlog = rs_integral(logAge_days, sq_cv_deriv)) %>%
  mutate(overall_trajectory = ifelse(sq_int_z <0 ,"converging", "diverging"))
```

###Sex-Dominance
*Which regions on average are bigger/more variable in males vs females and at what points in the lifespan?*
Now look at sex-dominance with integrals (M-F > or < 0)
Calculating across age, not log-age, for interpretability

```{r}
dominance_df <- sexdiffs_tbv_all %>%
  group_by(pheno, cv_sample, pheno_cat, weighted, age2plus) %>%
  arrange(logAge_days, .by_group = TRUE) %>%
  summarise(centile_int.unlog = rs_integral(logAge_days, centile_M_minus_F_z),
            sigma_int.unlog =  rs_integral(logAge_days, logcv_M_div_F),
            age_range = max(age_days) - min(age_days)) %>%
  # rescaling integrals to 100 years
  mutate(centile_int_norm.unlog = (centile_int.unlog/age_range)*365.25*100,
         sigma_int_norm.unlog = (sigma_int.unlog/age_range)*365.25*100) %>%
  mutate(dominance = case_when(
    centile_int.unlog == 0 ~ "Balanced",
    centile_int.unlog < 0 ~ "Female",
    centile_int.unlog > 0 ~ "Male"
  ),
  dominance_norm = case_when(
    centile_int_norm.unlog == 0 ~ "Balanced",
    centile_int_norm.unlog < 0 ~ "Female",
    centile_int_norm.unlog > 0 ~ "Male"
  ),
  sigma_dominance = case_when(
    sigma_int.unlog == 0 ~ "Balanced",
    sigma_int.unlog < 0 ~ "Female",
    sigma_int.unlog > 0 ~ "Male"
  ),
  sigma_dominance_norm = case_when(
    sigma_int_norm.unlog == 0 ~ "Balanced",
    sigma_int_norm.unlog < 0 ~ "Female",
    sigma_int_norm.unlog > 0 ~ "Male"
  ))

dominance_df %>%
  filter(pheno_cat=="Regional CT") %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(pct_male=sum(dominance=="Male")/n()*100,
            pct_female=sum(dominance=="Female")/n()*100)

dominance_df %>%
  filter(age2plus==FALSE) %>%
  group_by(pheno_cat, cv_sample, weighted) %>% 
  dplyr::count(dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=dominance)) +
  geom_col(position="fill") +
  facet_grid(weighted~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5)) +
  coord_flip()
```

restrict to phenos with significant sex effects

```{r}
LR.df_sig <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_all")

LR.df_sigagesex <- LR.df_plot %>%
  dplyr::filter(total==TRUE & effect=="sex_age")

dominance_plt_df <- inner_join(LR.df_sig, dominance_df %>% dplyr::select(!pheno_cat))
dominance_plt_df %>%
  mutate(dominance = ifelse(sig_fdr==TRUE, dominance, "NA")) %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(F_dom = sum(dominance=="Female")/n(),
            M_dom = sum(dominance=="Male")/n())
```

plot dominance on brain

```{r}
#removing ventricles for visualization:
dominance_plt_df_novent <- dominance_plt_df %>%
  filter(total==TRUE & weighted==FALSE & age2plus==FALSE) %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_novent$centile_int_norm.unlog)

full_dom_plt <- plot_cv_brain(df=dominance_plt_df_novent, 
                              fill_var=centile_int_norm.unlog, 
                              fill_name="Sex Diff\n(Std.)", 
                              fill_color="ggthemes::Orange-Blue Diverging", 
                              fill_limits= c(-12000, 12000), 
                              height_list=c(1.03, 4.5, .93), 
                              direction=1)

ggsave("figs/supplement/dominance_tbvTRUE.png", full_dom_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```
Also plot weighted and age2plus for sensitivity

```{r}
#removing ventricles for visualization:
dominance_plt_df_weight <- dominance_plt_df %>%
  filter(total==TRUE & weighted==TRUE & age2plus==FALSE) %>%
  filter(!(pheno %in% vent_list))
range(dominance_plt_df_weight$centile_int_norm.unlog)

full_dom_plt_weighted <- plot_cv_brain(df=dominance_plt_df_weight, 
                              fill_var=centile_int_norm.unlog, 
                              fill_name="Sex Diff\n(Std.)", 
                              fill_color="ggthemes::Orange-Blue Diverging", 
                              fill_limits= c(-12000, 12000), #keep color scale consistent
                              height_list=c(1.03, 4.5, .93), 
                              direction=1)

ggsave("figs/supplement/dominance_tbvTRUE_weighted.png", full_dom_plt_weighted, dpi=600, height=8.5, width=7.5, bg = "white")
```

```{r}
#removing ventricles for visualization:
dominance_plt_df_age2plus <- dominance_plt_df %>%
  filter(total==TRUE & weighted==FALSE & age2plus==TRUE) %>%
  filter(pheno %in% ct_list)
range(dominance_plt_df_age2plus$centile_int_norm.unlog)

full_dom_plt_age2plus <- plot_cv_brain(df=dominance_plt_df_age2plus, 
                              fill_var=centile_int_norm.unlog, 
                              fill_name="Sex Diff\n(Std.)", 
                              fill_color="ggthemes::Orange-Blue Diverging", 
                              fill_limits= c(-12000, 12000), #keep color scale consistent
                              height_list=c(1.03, 4.5, .93),
                              include_global = FALSE,
                              include_subcort = FALSE,
                              direction=1)

ggsave("figs/supplement/dominance_tbvTRUE_age2plus.png", full_dom_plt_age2plus, dpi=600, height=8.5, width=7.5, bg = "white")
```

How much of the lifespan are phenos male vs female dominant?

```{r}
#define function
prop_above <- function(x, y, threshold=0){
  crosses <- find_threshold_crosses(y, x, threshold, return="x", return_endpoints = FALSE)
  crosses <- c(crosses, min(x), max(x)) %>% 
    sort()
  #overall age modeled
  total_range <- max(x) - min(x)
  
  ranges <- diff(crosses)
  
  #if y > threshold at min(x) -> sum odd number ranges
  odd_indices_logical <- (1:length(ranges)) %% 2 == 1
  
  if (y[1] > threshold) {
    items_at_odd_indices <- ranges[odd_indices_logical]
    time_above <- sum(items_at_odd_indices)
  } else {
    items_at_even_indices <- ranges[!odd_indices_logical]
    time_above <- sum(items_at_even_indices)
  }
  
  prop_above <- time_above/total_range
  
  return(prop_above)
}

test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="CBV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="WMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="A" & pheno=="GMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="B" & pheno=="GMV")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))

test <- sexdiffs_tbv %>%
  filter(cv_sample =="B" & pheno=="lh.SA.pericalcarine")
prop_above(test$age_days, test$centile_M_minus_F_z)
ggplot(test) +
  geom_line(aes(x=age_days, y=centile_M_minus_F_z))
```

```{r}
prop_above_df <- sexdiffs_tbv_all %>%
  group_by(pheno, cv_sample, pheno_cat, weighted, age2plus) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, centile_M_minus_F_z)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male < .05 ~ "F",
    prop_male > .95 ~ "M",
    TRUE ~ "mixed"),
    strict_class = case_when(
    prop_male == 0 ~ "F",
    prop_male == 1 ~ "M",
    TRUE ~ "mixed")
  )

#add back sex-effect significance, remove ventricles for visualization
prop_above_df_novent <- inner_join(LR.df_sig, prop_above_df %>% dplyr::select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

prop_above_df_novent_main <- prop_above_df_novent %>%
  filter(weighted==FALSE & age2plus==FALSE)

full_above_plt <- plot_cv_brain(df=prop_above_df_novent_main, 
                                fill_var=prop_male, 
                                fill_name = "Proportion\nMale-\nDominant",
                                fill_color = "ggthemes::Orange-Blue Diverging", 
                                fill_limits=c(0, 1), 
                                height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_tbvTRUE.png", full_above_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```

plot weighted & age 2+ sensitivity
```{r}
prop_above_df_novent_weighted <- prop_above_df_novent %>%
  filter(weighted==TRUE & age2plus==FALSE)

full_above_plt <- plot_cv_brain(df=prop_above_df_novent_weighted, 
                                fill_var=prop_male, 
                                fill_name = "Proportion\nMale-\nDominant",
                                fill_color = "ggthemes::Orange-Blue Diverging", 
                                fill_limits=c(0, 1), 
                                height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_tbvTRUE_weighted.png", full_above_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```

```{r}
prop_above_df_novent_age2plus <- prop_above_df_novent %>%
  filter(weighted==FALSE & age2plus==TRUE)

full_above_plt <- plot_cv_brain(df=prop_above_df_novent_age2plus, 
                                fill_var=prop_male, 
                                fill_name = "Proportion\nMale-\nDominant",
                                fill_color = "ggthemes::Orange-Blue Diverging", 
                                fill_limits=c(0, 1), 
                                include_global = FALSE,
                                include_subcort = FALSE,
                                height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_tbvTRUE_age2plus.png", full_above_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```

####Sigma

```{r}
dominance_df %>%
  filter(age2plus==FALSE) %>%
  group_by(pheno_cat, cv_sample, weighted) %>% 
  dplyr::count(sigma_dominance, pheno_cat, cv_sample) %>% 
  mutate(perc = n / sum(n)) %>% 
  ggplot(aes(x=pheno_cat, y=perc, fill=sigma_dominance)) +
  geom_col(position="fill") +
  facet_grid(weighted~cv_sample) +
    geom_text(aes(label = scales::percent(perc)), position = position_fill(vjust = 0.5))+
  coord_flip()


dominance_plt_df %>%
  #restrict to significant sex effects
  mutate(sigma_dominance = ifelse(sig_fdr==TRUE, sigma_dominance, "NA")) %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(pct_male=sum(sigma_dominance=="Male")/n()*100,
            pct_female=sum(sigma_dominance=="Female")/n()*100)
```
plot

```{r}
range(dominance_plt_df_novent$sigma_int_norm.unlog)

full_dom_plt_sigma <- plot_cv_brain(df=dominance_plt_df_novent, 
                                    fill_var=sigma_int_norm.unlog, 
                                    fill_name="Sex Diff\n(Std.)", 
                                    fill_color = "ggthemes::Orange-Blue Diverging", 
                                    fill_limits =  c(-3500, 3500),
                                    height_list= c(1.03, 4.5, .93))

ggsave("figs/supplement/sigma_dominance_tbvTRUE.png", full_dom_plt_sigma, dpi=600, height=8.5, width=7.5, bg = "white")
```

also plot sensitivity analyses (weighted and ages 2+)

```{r}
range(dominance_plt_df_weight$sigma_int_norm.unlog)

full_dom_plt_sigma_weight <- plot_cv_brain(df=dominance_plt_df_weight, 
                                    fill_var=sigma_int_norm.unlog, 
                                    fill_name="Sex Diff\n(Std.)", 
                                    fill_color = "ggthemes::Orange-Blue Diverging", 
                                    fill_limits =  c(-3500, 3500),
                                    height_list= c(1.03, 4.5, .93))

ggsave("figs/supplement/sigma_dominance_tbvTRUE_weight.png", full_dom_plt_sigma_weight, dpi=600, height=8.5, width=7.5, bg = "white")

range(dominance_plt_df_age2plus$sigma_int_norm.unlog)

full_dom_plt_sigma_age2plus <- plot_cv_brain(df=dominance_plt_df_age2plus, 
                                    fill_var=sigma_int_norm.unlog, 
                                    fill_name="Sex Diff\n(Std.)", 
                                    fill_color = "ggthemes::Orange-Blue Diverging", 
                                    fill_limits =  c(-3500, 3500),
                                    include_global = FALSE,
                                    include_subcort = FALSE,
                                    height_list= c(1.03, 4.5, .93))

ggsave("figs/supplement/sigma_dominance_tbvTRUE_age2plus.png", full_dom_plt_sigma_age2plus, dpi=600, height=4, width=7.5, bg = "white")
```

proportion male-dominant

```{r}
prop_above_sigma_df <- sexdiffs_tbv_all %>%
  group_by(pheno, cv_sample, pheno_cat, weighted, age2plus) %>%
  arrange(age_days, .by_group = TRUE) %>%
  summarise(prop_male = prop_above(age_days, logcv_M_div_F)) %>%
  ungroup() %>%
  mutate(class = case_when(
    prop_male < .05 ~ "F",
    prop_male > .95 ~ "M",
    TRUE ~ "mixed"),
    strict_class = case_when(
    prop_male == 0 ~ "F",
    prop_male == 1 ~ "M",
    TRUE ~ "mixed")
  )

table(prop_above_sigma_df$cv_sample, prop_above_sigma_df$class)

prop_above_sigma_df %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(n_female = sum(class=="F"),
            n_male = sum(class=="M"),
            n_mixed = sum(class=="mixed"),
            n=n()) %>%
  mutate(pct_female = (n_female/n)*100,
         pct_male = (n_male/n)*100,
         pct_mix = (n_mixed/n)*100)
```

```{r}
#add back sex-effect significance, remove ventricles for visualization
prop_above_sigma_df_novent <- inner_join(LR.df_sig, prop_above_sigma_df %>% dplyr::select(!pheno_cat)) %>%
  filter(!(pheno %in% vent_list))

prop_above_sigma_df_novent_main <- prop_above_sigma_df_novent %>% filter(weighted==FALSE & age2plus==FALSE)

full_above_sig_plt <- plot_cv_brain(df=prop_above_sigma_df_novent_main, 
                                    fill_var=prop_male, 
                                    fill_name="Proportion\nMale-\nDominant", 
                                    fill_color="ggthemes::Orange-Blue Diverging", 
                                    fill_limits= c(0, 1), 
                                    height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_sigma_tbvTRUE.png", full_above_sig_plt, dpi=600, height=8.5, width=7.5, bg = "white")
```
plot sensitivity analyses

```{r}
prop_above_sigma_df_novent_weighted <- prop_above_sigma_df_novent %>% filter(weighted==TRUE & age2plus==FALSE)

full_above_sig_plt_weighted <- plot_cv_brain(df=prop_above_sigma_df_novent_weighted, 
                                    fill_var=prop_male, 
                                    fill_name="Proportion\nMale-\nDominant", 
                                    fill_color="ggthemes::Orange-Blue Diverging", 
                                    fill_limits= c(0, 1), 
                                    height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_sigma_tbvTRUE_weighted.png", full_above_sig_plt_weighted, dpi=600, height=8.5, width=7.5, bg = "white")

prop_above_sigma_df_novent_age2plus <- prop_above_sigma_df_novent %>% filter(weighted==FALSE & age2plus==TRUE)

full_above_sig_plt_age2plus <- plot_cv_brain(df=prop_above_sigma_df_novent_age2plus, 
                                    fill_var=prop_male, 
                                    fill_name="Proportion\nMale-\nDominant", 
                                    fill_color="ggthemes::Orange-Blue Diverging", 
                                    fill_limits= c(0, 1),
                                    include_global = FALSE,
                                    include_subcort = FALSE,
                                    height_list=c(1.03, 4.5, .93))

ggsave("figs/supplement/prop_male_sigma_tbvTRUE_age2plus.png", full_above_sig_plt_age2plus, dpi=600, height=4, width=7.5, bg = "white")
```

Prep for fig 2

```{r}
dominance_plt_df_B <- dominance_plt_df_novent %>%
  filter(cv_sample=="B" & weighted==FALSE & age2plus==FALSE)

#centiles
range(dominance_plt_df_B$centile_int_norm.unlog)
full_dom_pltB <- plot_cv_brain(df=dominance_plt_df_B, 
                               fill_var=centile_int_norm.unlog, 
                               fill_name="Sex Difference (Std.)", 
                               fill_color="ggthemes::Orange-Blue Diverging", 
                               fill_limits=c(-12000, 12000),
                               height_list=c(1.03, 4.5, .93), 
                               return_plt = FALSE,
                               legend_position = "bottom")
saveRDS(full_dom_pltB, "figs/ggobjs/cent_dom_B.rds")

#sigma
range(dominance_plt_df_B$sigma_int_norm.unlog)
full_dom_plt_sigmaB <- plot_cv_brain(df=dominance_plt_df_B, 
                                     fill_var=sigma_int_norm.unlog, 
                                     fill_name="Sex Difference (Std.)", 
                                     fill_color="ggthemes::Orange-Blue Diverging", 
                                     fill_limits=c(-3500, 3500), 
                                     height_list=c(1.03, 4.5, .93), 
                                     return_plt = FALSE,
                                      legend_position = "bottom")
saveRDS(full_dom_plt_sigmaB, "figs/ggobjs/sig_dom_B.rds")
```

####Spin Test

exploratory check whether sex-dominance in trajectories is associated with sex-dominance in variance

setup based on code from Val and claude AI

```{r}
#read in files
sphere.lh <- read.fs.surface(file.path(fs_path, "/surf/lh.sphere"))
sphere.rh <- read.fs.surface(file.path(fs_path, "/surf/rh.sphere"))
annot.lh <- freesurfer:::read_annotation(file.path(fs_path, "/label/lh.aparc.annot"))
annot.rh <- freesurfer:::read_annotation(file.path(fs_path, "/label/rh.aparc.annot"))

# Compute centroid (except for medial wall)
rm <- which(annot.lh$colortable$label %in% c("unknown", "corpuscallosum"))
#left hemisphere
lhcoords <- t(sapply(c(1:length(annot.lh$colortable$label))[-rm], function(x) 
  colMeans(sphere.lh$vertices[which(annot.lh$label == annot.lh$colortable$code[x]), ])
))
rownames(lhcoords) <- annot.lh$colortable$label[-rm]
#right hemisphere
rhcoords <- t(sapply(c(1:length(annot.rh$colortable$label))[-rm], function(x) 
  colMeans(sphere.rh$vertices[which(annot.rh$label == annot.rh$colortable$code[x]), ])
))
rownames(rhcoords) <- annot.rh$colortable$label[-rm]
```

spin permutations
```{r}
perm_file <- "spin_perms_DK68_n10000.rds"
 
if (file.exists(perm_file)) {
  message("Loading existing permutations from ", perm_file)
  perm.id <- readRDS(perm_file)
} else {
  message("Generating 10,000 spin permutations (this may take a few minutes)...")
  perm.id <- rotate.parcellation(lhcoords, rhcoords, nrot = 10000, method = "hungarian")
  saveRDS(perm.id, perm_file)
  message("Permutations saved to ", perm_file)
}
```

prep dominance dfs

```{r}
spin_prep <- dominance_df %>%
  ungroup() %>%
  filter(age2plus==FALSE & weighted==FALSE) %>%
  filter(pheno %in% c(sa_list, ct_list, vol_list)) %>%
  select(pheno, cv_sample, pheno_cat, centile_int_norm.unlog, sigma_int_norm.unlog) %>%
  mutate(hemi = stringr::str_extract(pheno, "(lh|rh)"),
         pheno = str_extract(pheno, "(?<=\\.)[^.]+$"))

#check pheno names match
stopifnot(all(unique(spin_prep$pheno[spin_prep$hemi == "lh"]) %in% rownames(lhcoords)))
stopifnot(all(unique(spin_prep$pheno[spin_prep$hemi == "rh"]) %in% rownames(rhcoords)))
#check region counts = 34
region_counts <- spin_prep %>%
  group_by(cv_sample, pheno_cat, hemi) %>%
  summarise(n = n(), .groups = "drop") %>%
  filter(n != 34)
```

run spin test
```{r}
spin_out <- spin_prep %>%
  group_by(cv_sample, pheno_cat) %>%
  group_modify(~ {
    # Order LH and RH rows to match coordinate matrices
    lh_data <- .x %>%
      filter(hemi == "lh") %>%
      arrange(match(pheno, rownames(lhcoords)))

    rh_data <- .x %>%
      filter(hemi == "rh") %>%
      arrange(match(pheno, rownames(rhcoords)))

    # Subset coordinate matrices to only regions present in this group
    lhcoords_sub <- lhcoords[rownames(lhcoords) %in% lh_data$pheno, , drop = FALSE]
    rhcoords_sub <- rhcoords[rownames(rhcoords) %in% rh_data$pheno, , drop = FALSE]

    # Re-order data to match subsetted coordinate matrices
    lh_data <- lh_data %>% arrange(match(pheno, rownames(lhcoords_sub)))
    rh_data <- rh_data %>% arrange(match(pheno, rownames(rhcoords_sub)))

    # Generate group-specific permutations if regions are missing
    n_regions <- nrow(lh_data) + nrow(rh_data)
    if (n_regions < 68) {
      message("Group ", unique(.x$cv_sample[1]), " / ", unique(.x$pheno_cat[1]),
              ": only ", n_regions, " regions — generating group-specific permutations")
      perm.id_group <- rotate.parcellation(lhcoords_sub, rhcoords_sub,
                                           nrot = 10000, method = "hungarian")
    } else {
      perm.id_group <- perm.id
    }

    # Build map vectors
    map_centile <- c(lh_data$centile_int_norm.unlog, rh_data$centile_int_norm.unlog)
    map_sigma   <- c(lh_data$sigma_int_norm.unlog,   rh_data$sigma_int_norm.unlog)

    obs_r  <- cor(map_centile, map_sigma, method = "spearman")
    p_spin <- perm.sphere.p(map_centile, map_sigma, perm.id_group, corr.type = "spearman")

    tibble(obs_r = obs_r, p_spin = p_spin, n_regions = n_regions)
  }) %>%
  ungroup()

spin_out %>%
  mutate(sig = ifelse(p_spin <0.05, TRUE, FALSE))
```


###Grouping
grouping by M vs F, converging vs diverging

```{r}
prop_above_df_main <- prop_above_df %>% filter(weighted==FALSE & age2plus==FALSE)

ggplot(prop_above_df_main) +
  geom_histogram(aes(x=prop_male)) +
  geom_vline(xintercept = .05) +
  geom_vline(xintercept = .95)

# > 95% lifespan in one sex
prop_above_df %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(n_F=sum(class=="F"),
            n_M=sum(class=="M"),
            n_tot=n())

#100% lifespan in one sex
prop_above_df %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(n_F=sum(strict_class=="F"),
            n_M=sum(strict_class=="M"),
            n_tot=n())

#add age*sex significance info
prop_above_df2 <- inner_join(prop_above_df, LR.df_sigagesex)

#add info about convergence vs divergence
prop_above_df3 <- inner_join(prop_above_df2, convergence_int) %>%
  mutate(class = paste(class, overall_trajectory, sep="_"))

prop_above_df3 %>%
  filter(sig_fdr==TRUE) %>%
  group_by(weighted, age2plus, cv_sample, class) %>%
  summarise(n = n()) %>%
  ggplot() +
  geom_col(aes(x=cv_sample, y=n, fill=class), position="dodge") +
  facet_grid(weighted~age2plus)
```

what's going on with the mixed phenotypes?

```{r}
#add back trajectory
group_cd <- full_join(sexdiffs_tbv_all, prop_above_df3)

mixed_plt <- group_cd %>%
  filter(weighted==FALSE & age2plus==FALSE) %>% #ignore sensitivity for now
  filter(sig_fdr==TRUE & (class=="mixed_converging" | class=="mixed_diverging")) %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  theme(legend.position="none")
ggplotly(mixed_plt)

#try simple grouping by sign at start and end (Lena's method)
group_cd_mix <- group_cd %>%
  filter(class=="mixed_converging" | class=="mixed_diverging") %>%
  select(logAge_days, cv_sample, pheno, centile_M_minus_F_z, weighted, age2plus) %>%
  arrange(logAge_days) %>%
  group_by(weighted, age2plus, cv_sample, pheno) %>%
  slice(c(1,n())) %>%
  mutate(sex = ifelse(centile_M_minus_F_z > 0 , "M", "F")) %>%
  summarise(switch = paste(first(sex), last(sex), sep = "_to_"), .groups = "drop")
table(group_cd_mix$switch, group_cd_mix$cv_sample)

mixed_plt2 <- full_join(group_cd, group_cd_mix) %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(class=="mixed_converging" | class=="mixed_diverging" & sig_fdr==TRUE) %>%
  #filter(pheno != "rh.CT.rostralanteriorcingulate") %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=centile_M_minus_F_z, group=pheno, color=pheno)) +
    facet_grid(switch~cv_sample, scales="free_y") +
  theme(legend.position="none")

ggplotly(mixed_plt2)
```

```{r}
group_traj_final <- full_join(group_cd, group_cd_mix) %>%
  mutate(cent_group=case_when(
    class=="mixed_converging" | class=="mixed_diverging" ~ switch,
    TRUE ~ class
  )) %>%
  #regroup U trajectories
  rename(cent_group_full = cent_group) %>%
  mutate(cent_group = case_when(
    cent_group_full == "F_to_F" | cent_group_full == "M_to_M" ~ "Complex",
    TRUE ~ cent_group_full
  )) %>%
  mutate(parent_group = case_when(
    sig_fdr==FALSE ~ "Stable",
    cent_group == "F_diverging" | cent_group == "M_diverging" ~ "Diverging",
    cent_group == "F_converging" | cent_group == "M_converging" ~ "Converging",
    cent_group == "M_to_F" | cent_group == "F_to_M" ~ "Switch",
    TRUE ~ str_to_title(cent_group)
  ))

group_final <- group_traj_final %>%
  select(cv_sample, total, weighted, age2plus, pheno, pheno_cat, class,
         prop_male, strict_class, resi, sig_fdr, overall_trajectory,
         switch, cent_group, label, parent_group) %>%
  distinct() %>%
  mutate(cent_group = if_else(sig_fdr == FALSE, "NS", cent_group))
```

```{r}
#GET FINAL COUNTS
#group
group_final %>%
  filter(age2plus==FALSE) %>%
  group_by(weighted, cv_sample, cent_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)

#parent groups
group_final %>%
  filter(age2plus==FALSE) %>%
  group_by(weighted, cv_sample, parent_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)
```

plot formatting prep

```{r}
group_values <- c("F_diverging"="#9E3D22FF", "F_converging"="#C06636FF","F_to_M" ="#E8B960FF", 
             "M_to_F" = "#528AA9", "M_converging"="#426f89","M_diverging"="#154063","Complex" = "gray40",
             "NS" = "gray75")

group_color_scale <- scale_colour_manual(
  values = group_values,
  breaks = c("F_diverging", "F_converging", "F_to_M", 
             "M_to_F", "M_converging", "M_diverging", 
             "Complex", "NS"),
  labels = c("Female Diverging", "Female Converging", "Female to Male Bias", 
             "Male to Female Bias", "Male Converging", "Male Diverging", 
             "Complex", "Stable"),
  name = "Group",
  na.value="gray90"
)

group_fill_scale <- scale_fill_manual(
  values = group_values,
  breaks = c("F_diverging", "F_converging", "F_to_M", 
             "M_to_F", "M_converging", "M_diverging", 
             "Complex", "NS"),
  labels = c("Female Diverging", "Female Converging", "Female to Male Bias", 
             "Male to Female Bias", "Male Converging", "Male Diverging", 
             "Complex", "Stable"),
  name = "Group",
  na.value="gray90"
)

#alternate version with different spacing for labels
group_fill_scale2 <- scale_fill_manual(
  values = group_values,
  breaks = c("F_diverging", "F_converging", "F_to_M", 
             "M_to_F", "M_converging", "M_diverging", 
             "Complex", "NS"),
  labels = c("Female\nDiverging", "Female\nConverging", "Female to\nMale Bias", 
             "Male to\nFemale Bias", "Male\nConverging", "Male\nDiverging", 
             "Complex", "Stable"),
  name = "Group",
  na.value="gray90"
)

bottom_group_legend <- get_legend(
  ggplot(group_final) +
    geom_bar(aes(x=cv_sample, fill=cent_group)) +
    group_fill_scale2 +
    theme(legend.position = "bottom",
          legend.text = element_text(margin = margin(r = .5, unit = "cm"))) +
    guides(fill = guide_legend(nrow = 1)),
  legend="bottom"
  )
```

Plot group outputs for each value of weighted==T/F

```{r}
for (w in c(FALSE, TRUE)) {
  plot_suffix <- if (w) "_weighted" else ""

  gf  <- group_final %>% filter(weighted == w & age2plus == FALSE)
  gtf <- group_traj_final %>% filter(weighted == w & age2plus == FALSE)

###BARPLOTS OF COUNTS###
  group_bar_plt_total <- gf %>%
    group_by(cv_sample) %>% 
    mutate(cent_group = as.character(cent_group)) %>%
    mutate(cent_group = ifelse(sig_fdr ==TRUE, cent_group, NA)) %>%
    mutate(cent_group = factor(cent_group, levels = names(group_values))) %>%  
    dplyr::count(cent_group, cv_sample) %>% 
    mutate(perc = n / sum(n)) %>% 
    ggplot(aes(x="TOTAL", y=perc, fill=cent_group)) +
    geom_col(position="fill") +
    facet_wrap(~cv_sample) +
    geom_text(aes(label = n), 
              position = position_fill(vjust = 0.5),
              size = 1.5) +
    scale_y_continuous(labels=scales::percent) +
    labs(y="Percentage") +
    theme_linedraw() +
    theme(legend.position = "none",
          axis.title.x= element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    group_fill_scale
  
  group_bar_plt <- gf %>%
    #filter(sig_fdr==TRUE & pheno_cat != "Global Vol") %>%
    ggplot() +
    geom_bar(aes(x=parent_group, fill=cent_group)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    group_fill_scale +
    guides(x = guide_axis(angle = 45)) +
    labs(y="Count") +
    theme_linedraw() +
    theme(axis.title.x = element_blank(),
          legend.position = "none")
  
  group_bar_plt_all <- plot_grid(
      group_bar_plt_total, 
      group_bar_plt,
      ncol = 2,
      rel_widths = c(1,4),
      align = "hv",
      axis = "lr",
      labels=c("A", "B"))
  
  ####BRAIN PLOTS###
  group_final_novent <- gf %>%
  filter(!(pheno %in% vent_list))

  group_brain_plt <- plot_cv_brain(group_final_novent, cent_group, "Group", fill_limits=levels(group_final$cent_group), height_list=c(1.03, 4.5, .93), return_plt = FALSE, fill_values=group_values)
  
  group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale
  
  group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale
  
  group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale
  
  plts <- cowplot::plot_grid(
      group_brain_plt[[1]],
      group_brain_plt[[2]],
      group_brain_plt[[3]],
      ncol = 1,
      nrow = 3,
      rel_heights = c(1.03, 4.5, .93),
      align = "v",
      axis = "lr",
      greedy = TRUE
    ) 
    
  ###TRAJECTORIES###
  group_traj_full <- gtf %>%
    filter(sig_fdr==TRUE & cent_group != "Complex") %>%
    arrange(pheno, age_days) %>%
    ggplot(aes(x=age_days, y=centile_M_minus_F_z, color=cent_group)) +
    geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
    geom_line(aes(group=pheno), alpha=.1) +
    facet_grid(parent_group ~ cv_sample, scales="free_y") +
    geom_smooth(method = "gam") +
    group_color_scale +
    reformat_x(gtf$age_days) +
    theme_linedraw() +
    theme(legend.position="none") +
    labs(y="Scaled Median Sex Difference")
  
  #save combined group plots for supplement
  sub_plt <- plot_grid(NULL, plts,
                        group_traj_full,
                        nrow=1, rel_widths =c(.1,1,2),
                        labels=c("C", "", "D"))
  
  group_plot_supplement <- plot_grid(group_bar_plt_all,
                            sub_plt,
                            bottom_group_legend,
                            nrow=3,
                            ncol=1,
                            rel_heights = c(1, 1.1, .1))
  ggsave(paste0("figs/supplement/groups_all_plt", plot_suffix, ".png"), group_plot_supplement, dpi=300, height=16, width=16, bg = "white")
  
  #clear space
  rm(group_bar_plt_total, group_bar_plt, group_bar_plt_all,
   group_final_novent, group_brain_plt, plts,
   group_traj_full, sub_plt, group_plot_supplement, gf, gtf)
  gc()
}
```


```{r}
#just plot B
barplt_final <- group_final %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(cv_sample=="B" & pheno_cat != "Ventricular Vol") %>%
  mutate(parent_group = factor(case_when(
      cent_group == "NS" ~ "Stable",
      cent_group == "F_diverging" | cent_group == "M_diverging" ~ "Diverging",
      cent_group == "F_converging" | cent_group == "M_converging" ~ "Converging",
      cent_group == "M_to_F" | cent_group == "F_to_M" ~ "Switch",
      TRUE ~ str_to_title(cent_group)
    ), levels = c("Diverging", "Converging", "Switch", "Stable", "Complex"),
       ordered = TRUE)
  ) %>%
  dplyr::count(pheno_cat, parent_group, cent_group) %>%
  group_by(pheno_cat) %>%
  mutate(pct = n / sum(n)) %>%
  ungroup() %>%
  ggplot() +
  geom_bar(aes(x=parent_group, y=pct, fill=cent_group), stat="identity") +
  facet_wrap(~pheno_cat, nrow=1) +
  scale_y_continuous(labels=scales::percent) +
  group_fill_scale +
  guides(x = guide_axis(angle = 45)) +
  labs(y="Percentage") +
  theme_linedraw() +
  theme(axis.title.x = element_blank(), legend.position = "bottom")
saveRDS(barplt_final, "figs/ggobjs/cent_groups_barplt_B.rds")

##BRAIN PLOT
group_finalB <- group_final %>% 
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(cv_sample=="B")
group_brain_plt <- plot_cv_brain(group_finalB, 
                                 cent_group, 
                                 "Group", 
                                 fill_limits=levels(group_finalB$cent_group), 
                                 height_list=c(1.03, 4.5, .93), 
                                 return_plt = FALSE, 
                                 fill_values=group_values)

group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale

group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale

group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale

#recreate legends
group_brain_plt[[4]] <- bottom_group_legend
saveRDS(group_brain_plt, "figs/ggobjs/cent_groups_brain_B.rds")

##TRAJECTORIES
group_traj_plt_B <- group_traj_final %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(sig_fdr==TRUE & cv_sample=="B" & cent_group != "complex") %>%
  arrange(pheno, age_days) %>%
  ggplot(aes(x=age_days, y=centile_M_minus_F_z, color=cent_group)) +
  geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
  #geom_line(aes(group=pheno), alpha=.2) +
  facet_wrap(~parent_group, nrow=3) +
  geom_smooth(method = "gam", se=FALSE) +
  group_color_scale +
  reformat_x(group_traj_final$age_days) +
  theme_linedraw() +
  theme(legend.position="none") +
  labs(y="Scaled Median Sex Difference")
saveRDS(group_traj_plt_B, "figs/ggobjs/cent_groups_traj_B.rds")
```

####Sigma

```{r}
prop_above_sigma_df_main <- prop_above_sigma_df %>% filter(weighted==FALSE & age2plus==FALSE)

ggplot(prop_above_sigma_df_main) +
  geom_histogram(aes(x=prop_male)) +
  geom_vline(xintercept = .05) +
  geom_vline(xintercept = .95)

# > 95% lifespan in one sex
prop_above_sigma_df %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(n_F=sum(class=="F"),
            n_M=sum(class=="M"),
            n_tot=n())

#100% lifespan in one sex
prop_above_sigma_df %>%
  group_by(cv_sample, weighted, age2plus) %>%
  summarise(n_F=sum(strict_class=="F"),
            n_M=sum(strict_class=="M"),
            n_tot=n())

#add age*sex significance info
prop_above_sigma_df2 <- inner_join(prop_above_sigma_df, LR.df_sigagesex)

#add info about convergence vs divergence
prop_above_sigma_df3 <- inner_join(prop_above_sigma_df2, convergence_int) %>%
  mutate(class = paste(class, overall_trajectory, sep="_"))

prop_above_sigma_df3 %>%
  filter(sig_fdr==TRUE) %>%
  group_by(weighted, age2plus, cv_sample, class) %>%
  summarise(n = n()) %>%
  ggplot() +
  geom_col(aes(x=cv_sample, y=n, fill=class), position="dodge") +
  facet_grid(weighted~age2plus)
```

what's going on with the mixed phenotypes?

```{r}
#add back trajectory
sigma_group_cd <- full_join(sexdiffs_tbv_all, prop_above_sigma_df3)

sigma_group_cd %>%
  filter(weighted==FALSE & age2plus==FALSE) %>% #ignore sensitivity for now
  filter(sig_fdr==TRUE & (class=="mixed_converging" | class=="mixed_diverging")) %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
  theme(legend.position="none")

#try simple grouping by sign at start and end (Lena's method)
sigma_group_cd_mix <- sigma_group_cd %>%
  filter(class=="mixed_converging" | class=="mixed_diverging") %>%
  select(logAge_days, cv_sample, pheno, logcv_M_div_F, weighted, age2plus) %>%
  arrange(logAge_days) %>%
  group_by(weighted, age2plus, cv_sample, pheno) %>%
  slice(c(1,n())) %>%
  mutate(sex = ifelse(logcv_M_div_F > 0 , "M", "F")) %>%
  summarise(switch = paste(first(sex), last(sex), sep = "_to_"), .groups = "drop")
table(sigma_group_cd_mix$switch, sigma_group_cd_mix$cv_sample)

sigma_mixed_plt2 <- full_join(sigma_group_cd, sigma_group_cd_mix) %>%
  filter(weighted==FALSE & age2plus==FALSE) %>%
  filter(class=="mixed_converging" | class=="mixed_diverging" & sig_fdr==TRUE) %>%
  arrange(pheno, logAge_days) %>%
    ggplot() +
    geom_line(aes(x=logAge_days, y=logcv_M_div_F, group=pheno, color=pheno)) +
    facet_grid(switch~cv_sample, scales="free_y") +
  theme(legend.position="none")

#ggplotly(sigma_mixed_plt2)
```
sample A rh.CT.lateraloccipital is the only feature thats wiggly, doesn't just switch M-F or vice versa

```{r}
group_sig_traj_final <- full_join(sigma_group_cd, sigma_group_cd_mix) %>%
  mutate(sig_group=case_when(
    class=="mixed_converging" | class=="mixed_diverging" ~ switch,
    TRUE ~ class
  )) %>%
  mutate(sig_group = case_when(
    sig_group == "F_to_F" | sig_group == "M_to_M" ~ "Complex",
    TRUE ~ sig_group
  )) %>%
  mutate(parent_group = case_when(
    sig_fdr == FALSE ~ "Stable",
    sig_group == "F_diverging" | sig_group == "M_diverging" ~ "Diverging",
    sig_group == "F_converging" | sig_group == "M_converging" ~ "Converging",
    sig_group == "M_to_F" | sig_group == "F_to_M" ~ "Switch",
    TRUE ~ str_to_title(sig_group)
  ))

group_sig_final <- group_sig_traj_final %>%
  select(cv_sample, total, weighted, age2plus, pheno, pheno_cat, class,
         prop_male, strict_class, resi, sig_fdr, overall_trajectory,
         switch, sig_group, label, parent_group) %>%
  distinct() %>%
  mutate(sig_group = if_else(sig_fdr == FALSE, "NS", sig_group))
```

```{r}
#GET FINAL COUNTS
#group
group_sig_final %>%
  filter(age2plus==FALSE) %>%
  group_by(weighted, cv_sample, sig_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)

#parent groups
group_sig_final %>%
  filter(age2plus==FALSE) %>%
  group_by(weighted, cv_sample, parent_group) %>%
  summarise(n = n()) %>%
  mutate(pct = (n / sum(n))*100)
```


Plot group outputs for each value of weighted==T/F

```{r}
for (w in c(FALSE, TRUE)) {
  plot_suffix <- if (w) "_weighted" else ""

  gf  <- group_sig_final %>% filter(weighted == w & age2plus == FALSE)
  gtf <- group_sig_traj_final %>% filter(weighted == w & age2plus == FALSE)

###BARPLOTS OF COUNTS###
  group_bar_plt_total <- gf %>%
    group_by(cv_sample) %>% 
    mutate(sig_group = as.character(sig_group)) %>%
    mutate(sig_group = ifelse(sig_fdr ==TRUE, sig_group, NA)) %>%
    mutate(sig_group = factor(sig_group, levels = names(group_values))) %>%  
    dplyr::count(sig_group, cv_sample) %>% 
    mutate(perc = n / sum(n)) %>% 
    ggplot(aes(x="TOTAL", y=perc, fill=sig_group)) +
    geom_col(position="fill") +
    facet_wrap(~cv_sample) +
    geom_text(aes(label = n), 
              position = position_fill(vjust = 0.5),
              size = 1.5) +
    scale_y_continuous(labels=scales::percent) +
    labs(y="Percentage") +
    theme_linedraw() +
    theme(legend.position = "none",
          axis.title.x= element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank()) +
    group_fill_scale
  
  group_bar_plt <- gf %>%
    #filter(sig_fdr==TRUE & pheno_cat != "Global Vol") %>%
    ggplot() +
    geom_bar(aes(x=parent_group, fill=sig_group)) +
    facet_grid(pheno_cat~cv_sample, scales="free_y") +
    group_fill_scale +
    guides(x = guide_axis(angle = 45)) +
    labs(y="Count") +
    theme_linedraw() +
    theme(axis.title.x = element_blank(),
          legend.position = "none")
  
  group_bar_plt_all <- plot_grid(
      group_bar_plt_total, 
      group_bar_plt,
      ncol = 2,
      rel_widths = c(1,4),
      align = "hv",
      axis = "lr",
      labels=c("A", "B"))
  
  ####BRAIN PLOTS###
  group_final_novent <- gf %>%
  filter(!(pheno %in% vent_list))

  group_brain_plt <- plot_cv_brain(group_final_novent, sig_group, "Group", fill_limits=levels(group_final$sig_group), height_list=c(1.03, 4.5, .93), return_plt = FALSE, fill_values=group_values)
  
  group_brain_plt[[1]] <- group_brain_plt[[1]] + group_fill_scale
  
  group_brain_plt[[2]] <- group_brain_plt[[2]] + group_fill_scale
  
  group_brain_plt[[3]] <- group_brain_plt[[3]] + group_fill_scale
  
  plts <- cowplot::plot_grid(
      group_brain_plt[[1]],
      group_brain_plt[[2]],
      group_brain_plt[[3]],
      ncol = 1,
      nrow = 3,
      rel_heights = c(1.03, 4.5, .93),
      align = "v",
      axis = "lr",
      greedy = TRUE
    ) 
    
  ###TRAJECTORIES###
  group_traj_full <- gtf %>%
    filter(sig_fdr==TRUE & sig_group != "Complex") %>%
    arrange(pheno, age_days) %>%
    ggplot(aes(x=age_days, y=logcv_M_div_F, color=sig_group)) +
    geom_hline(yintercept = 0, linetype='dashed', color='grey20') +
    geom_line(aes(group=pheno), alpha=.1) +
    facet_grid(parent_group ~ cv_sample, scales="free_y") +
    geom_smooth(method = "gam") +
    group_color_scale +
    reformat_x(gtf$age_days) +
    theme_linedraw() +
    theme(legend.position="none") +
    labs(y="Scaled Median Sex Difference")
  
  #save combined group plots for supplement
  sub_plt <- plot_grid(NULL, plts,
                        group_traj_full,
                        nrow=1, rel_widths =c(.1,1,2),
                        labels=c("C", "", "D"))
  
  group_plot_supplement <- plot_grid(group_bar_plt_all,
                            sub_plt,
                            bottom_group_legend,
                            nrow=3,
                            ncol=1,
                            rel_heights = c(1, 1.1, .1))
  ggsave(paste0("figs/supplement/groups_sigma_all_plt", plot_suffix, ".png"), group_plot_supplement, dpi=300, height=16, width=16, bg = "white")
  
  #clear space
  rm(group_bar_plt_total, group_bar_plt, group_bar_plt_all,
   group_final_novent, group_brain_plt, plts,
   group_traj_full, sub_plt, group_plot_supplement, gf, gtf)
  gc()
}
```

###Data-Driven Clustering

```{r}
clust_fun <- function(df, clust_var, k){
  medoids<-NULL
  #filter df
  df <- df %>%
    select(pheno, logAge_days, clust_var, pheno_cat, age_days)
  
  #pivot wide - rows = pheno levels, cols = time points
  ts_wide <- df %>%
    arrange(pheno, age_days) %>%
    select(pheno, age_days, clust_var) %>%
    pivot_wider(
      names_from  = age_days,
      values_from = clust_var
    ) %>%
    filter(!if_all(.cols = !pheno, ~is.na(.))) %>%
    select(where(~ !any(is.na(.))))
  
  # Store pheno labels, then convert to matrix
  pheno_labels <- ts_wide$pheno
  ts_matrix <- ts_wide %>%
    select(-pheno) %>%
    as.matrix()

  rownames(ts_matrix) <- pheno_labels
  
  #drop any missing vals
  ts_matrix_complete <- ts_matrix[, colSums(is.na(ts_matrix)) == 0]
  ts_matrix_complete <- ts_matrix_complete[rowSums(is.na(ts_matrix_complete)) == 0, ]
  
  #euclidean distance
  dist_matrix <- dist(ts_matrix_complete, method = "euclidean")
  
  #Cluster
  set.seed(42)
  
  # Elbow plot using average silhouette width
    elbow_plt <- factoextra:::fviz_nbclust(ts_matrix_complete, cluster:::pam, method = "silhouette", linecolor = "gray10")
    print(elbow_plt)
  
    pm <- cluster:::pam(dist_matrix, k = k, diss = TRUE)
    clusters <- setNames(pm$clustering, rownames(ts_matrix_complete))
    
    medoids <- pm$medoids

  # Add cluster assignments back to original data
  cluster_df <- tibble(
    pheno   = names(clusters),
    cluster = clusters
  )
  
  data_with_clusters <- df %>%
    left_join(cluster_df, by = "pheno")
  
  out <- list(data_with_clusters, elbow_plt, medoids)
  return(out)
}

plot_clust <- function(cent_clust, clust_var, df){
  data_with_clusters <- cent_clust[[1]]
  medoids <- cent_clust[[3]]
  data_with_clusters <- data_with_clusters %>%
    mutate(cluster=factor(cluster, ordered=TRUE))
  
  #plot time series by cluster
  plt_traj <- data_with_clusters %>%
    ggplot(aes(x = age_days, y = !!sym(clust_var),
               group = pheno, colour = cluster)) +
    geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
    geom_line(alpha = 0.3) +
    paletteer::scale_colour_paletteer_d("rockthemes::deelite") +
    facet_wrap(~cluster, nrow=2) +
    labs(
      title = "Pheno Time Series by Cluster",
      colour = "Cluster"
    ) +
    reformat_x(data_with_clusters$age_days, simplified=TRUE) +
    theme_minimal()
  print(plt_traj)

#plot medoids
  med_plt <- data_with_clusters %>%
    filter(pheno %in% medoids) %>%
    #mutate(cluster = factor(paste("Cluster", cluster, "Medoid Trajectory"))) %>%
    ggplot(aes(x = age_days, y = !!sym(clust_var),
                group = pheno, color = cluster)) +
    geom_hline(yintercept = 0,linetype='dashed',color='grey20')+
    geom_line(linewidth=1.5) +
    paletteer::scale_colour_paletteer_d("rockthemes::deelite") +
    facet_wrap(~pheno, nrow=2) +
    labs(title = "Cluster Medoid", colour = "Cluster"
    ) +
  reformat_x(data_with_clusters$age_days, simplified=TRUE) +
      theme_minimal()
    print(med_plt)
  
  #plot on brains
  brain_data <- df %>%
    select(pheno, pheno_cat, label, sig_fdr, cv_sample) %>%
    distinct()
  
  brain_cluster <- data_with_clusters %>%
    select(pheno, cluster) %>%
    distinct()
  
  brain_df <- full_join(brain_data, brain_cluster)
  brain_plt <- plot_cv_brain(df = brain_df, 
                          fill_var = cluster, 
                          fill_name = "Cluster", 
                          fill_color = "rockthemes::deelite",
                          facet_var= "cv_sample",
                          height_list=c(1.03, 4.5, .93), 
                          return_plt = TRUE, 
                          legend_position = "bottom",
                          include_global = TRUE, 
                          rel_widths = c(3, .6))
 
#plot list
plt_list <- c(plt_traj, med_plt, brain_plt)
return(plt_list)
}
```

prep dfs
```{r}
#add age-by-sex significance info
LR.df_tot.u <- inner_join(LR.df_sigagesex, sexdiffs.df.u) %>%
  filter(sig_fdr==TRUE & total==TRUE)

LR.df_tot.uA <- LR.df_tot.u %>%
  filter(cv_sample=="A" & weighted==FALSE)

LR.df_tot.uB <- LR.df_tot.u %>%
  filter(cv_sample=="B" & weighted==FALSE)

#also save out weighted for sensitivity analyses
LR.df_tot.uA_w <- LR.df_tot.u %>%
  filter(cv_sample=="A" & weighted==TRUE)

LR.df_tot.uB_w <- LR.df_tot.u %>%
  filter(cv_sample=="B" & weighted==TRUE)
```

cluster and plot - wrapping for simplicity
```{r}
for (df in list(LR.df_tot.uA, LR.df_tot.uB, LR.df_tot.uA_w, LR.df_tot.uB_w)){
  cv_sample <- unique(df[["cv_sample"]])
  w <- unique(df[["weighted"]])
  
  #Cluster
  cent_clust <- clust_fun(df, "centile_M_minus_F_z", 2)
  #save outputs
  rds_name <- paste0("figs/ggobjs/median_kmedoid_clust_", cv_sample, "_weighted", w, ".rds")
  saveRDS(cent_clust, rds_name)
  
  #get elbow plt
  elbow <- cent_clust[[2]]
  print(elbow)
  
  #Plot
  cent_clust_plt <- plot_clust(cent_clust, "centile_M_minus_F_z", df)
  
  #update plots
  cent_clust_plt[[1]] <- cent_clust_plt[[1]] +
  labs(y="Median Sex Difference (Std.)") +
  theme(legend.position="none")
  
  clust_legend <- get_legend(cent_clust_plt[[2]])
  cent_clust_plt[[2]] <- cent_clust_plt[[2]] + 
    labs(y="Median Sex Difference (Std.)") +
    theme(legend.position="none")

  cent_clust_wrap <- plot_grid( 
    plot_grid(
      plot_grid(
      elbow,
      cent_clust_plt[[2]],
      nrow=2,
      rel_heights=c(1,1),
      labels = c("A", "B")
    ),
    cent_clust_plt[[1]],
    NULL,
  cent_clust_plt[[3]],
  ncol=4,
  rel_widths = c(2,2,.02,1.7),
  labels=c("","C", "D","")),
  clust_legend,
  nrow=1,
  rel_widths=c(10,1))
  
  file_name <- paste0("figs/supplement/cent_kmedoid_", cv_sample, "_weighted", w, ".png")
  ggsave(file_name, cent_clust_wrap, dpi=600, height=8, width=13, bg = "white")
}
```

####Sigma

```{r}
for (df in list(LR.df_tot.uA, LR.df_tot.uB, LR.df_tot.uA_w, LR.df_tot.uB_w)){
  cv_sample <- unique(df[["cv_sample"]])
  w <- unique(df[["weighted"]])
  
  #Cluster
  sig_clust <- clust_fun(df, "logcv_M_div_F", 2)
  #save outputs
  rds_name <- paste0("figs/ggobjs/sigma_kmedoid_clust_", cv_sample, "_weighted", w, ".rds")
  saveRDS(sig_clust, rds_name)
  
  #get elbow plt
  elbow <- sig_clust[[2]]
  print(elbow)
  
  #Plot
  sig_clust_plt <- plot_clust(sig_clust, "logcv_M_div_F", df)
  
  #update plots
  sig_clust_plt[[1]] <- sig_clust_plt[[1]] +
  labs(y="Variance Ratio") +
  theme(legend.position="none")
  
  clust_legend <- get_legend(sig_clust_plt[[2]])
  sig_clust_plt[[2]] <- sig_clust_plt[[2]] + 
    labs(y="Variance Ratio") +
    theme(legend.position="none")

  sig_clust_wrap <- plot_grid( 
    plot_grid(
      plot_grid(
      elbow,
      sig_clust_plt[[2]],
      nrow=2,
      rel_heights=c(1,1),
      labels = c("A", "B")
    ),
    sig_clust_plt[[1]],
    NULL,
  sig_clust_plt[[3]],
  ncol=4,
  rel_widths = c(2,2,.02,1.7),
  labels=c("","C", "D","")),
  clust_legend,
  nrow=1,
  rel_widths=c(10,.5))
  
  file_name <- paste0("figs/supplement/sigma_kmedoid_", cv_sample, "_weighted", w, ".png")
  ggsave(file_name, sig_clust_wrap, dpi=600, height=8, width=13, bg = "white")
}
```
