---
title: "Review Centile Tests"
output: html_document
date: "2024-10-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, rlang, ggrepel, gamlss, ggseg, ggsegExtra, RESI, paletteer, patchwork, plotly, pracma, gcplyr, legendry, gghighlight, forcats, RESI, ggpubr, cowplot, EnvStats, binom)

#devtools::install_github("BGDlab/gamlssTools", force=TRUE)
library(gamlssTools)

sa_list <- readRDS("./pheno_lists/cortical_surf.rds")
ct_list <- readRDS("./pheno_lists/cortical_thickness.rds")
vol_list <- readRDS("./pheno_lists/cortical_vols.rds")
global_list <- readRDS("./pheno_lists/global_vols.rds")
sub_list <- readRDS("./pheno_lists/subcortical_vols.rds")
pheno_list <- c(sa_list, ct_list, vol_list, global_list, sub_list)
#just looking at non-TBV-corrected results for now
test_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*totalFALSE*/"

dx_tested <- c("SCZ", "ALZ", "ASD", "MDD", "GAD", "ADHD")

# install.packages("remotes")
remotes::install_github("BGDlab/ggsegTissue")
library(ggsegTissue)
#remotes::install_github("ggseg/ggsegExtra")
source("code/plot_cv_brain.R")
```

#Read & Prep CSVs

just visualize global centiles

```{r}
glob_path <- "/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test/*global*totalFALSE*/"
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(glob_path, "cent_csvs/*_*_*cent.csv")) %>%
  grep("CN|PT", ., value = TRUE)

# Read all matching CSVs
cent_df_list <- lapply(paths, function(p) {
  fread(p) %>%
    mutate(
      cv_sample = stringr::str_extract(p, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
      total = stringr::str_extract(p, "(?<=total)(TRUE|FALSE)"),
      weighted = stringr::str_detect(p, "weight"),
      control = stringr::str_extract(p, "(CN|PT)"),
      dx_tested  = sub("CN_", "", dx_recode)
    ) %>%
    select(!c(sexMale_x_logAge))
})

length(cent_df_list)
#stringr::str_extract(paths, "(?<=cv_sample_)[A-Za-z]+(?=_test)")
cent_df <- rbindlist(cent_df_list, fill=TRUE)

rm(cent_df_list)
```

Case-Control results
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path(test_path, "cent_csvs/*cent_pt*_casecontrol_test.csv"))

#expect 1 row per pheno/cv_sample/dx (or fewer for missing variables)
n_expected <- length(dx_tested)*length(pheno_list)*2
stopifnot(length(paths) <= n_expected)

# Read all matching CSVs
output_fread <- lapply(paths, fread)
length(output_fread)
names(output_fread) <- paths


cent_test.df <- bind_rows(output_fread, .id="file") %>%
  mutate(pheno = sapply(basename(file), function(f) {
           matched <- pheno_list[startsWith(f, pheno_list)]
           if (length(matched) > 0) matched[which.max(nchar(matched))] else NA_character_
         }),
         cv_sample = stringr::str_extract(file, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
         total = stringr::str_extract(file, "(?<=total)(TRUE|FALSE)"), #just looking at FALSE for now
         weighted  = stringr::str_detect(file, "weight")) %>%
  select(-case.control_est) %>%
  distinct() %>%
  mutate(pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  select(!file) %>%
  group_by(total, weighted, cv_sample, model) %>% #apply FDR correction within each combo ??apply fdr correction within each model??
  mutate(
    dx_p.val_fdr = p.adjust(
      case.control_p.val,
      method = "fdr"
    ),
    #flipping sign so that neg effect means pts < CN
    case.control_d = -case.control_d) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(
    dx_sig_fdr = ifelse(dx_p.val_fdr <= 0.05, TRUE, FALSE),
    case.control_resi = d2S(d = case.control_d, pi = case.control_pi)
  ) %>%
  #get back sign from d - pending approval from simon
  mutate(case.control_resi = ifelse(case.control_d < 0, -case.control_resi, case.control_resi)) %>%
  ungroup()
```

##Case-Control Diffs

```{r}
#full model
cent_test.df_uncor <- cent_test.df %>%
  filter(total == "FALSE" & weighted == "FALSE")

#expect 2 rows per path, 1 full model and 1 null model
stopifnot(nrow(cent_test.df_uncor) <= n_expected*2)

cent_test.df_uncor %>%
  ggplot() +
  geom_bar(aes(x=dx, fill=dx_sig_fdr)) +
  facet_grid(model~cv_sample)
  
uncor_d_plt <- cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=pheno, y= case.control_d, fill=model), position="dodge") +
  theme(axis.text.x = element_blank()) +
  facet_grid(dx~cv_sample, scales="free_y") 
#ggplotly(uncor_d_plt)

cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE & pheno %in% global_list) %>%
  ggplot() +
  geom_col(aes(x=dx, y= case.control_resi, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~pheno)

cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE & pheno %in% global_list) %>%
  ggplot() +
  geom_col(aes(x=dx, y= case.control_d, fill=cv_sample), position="dodge") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  facet_grid(model~pheno)
```

```{r}
cent_plt_df <- cent_test.df_uncor %>%
  filter(dx_sig_fdr == TRUE) %>%
  arrange(pheno_cat, pheno) %>%
  mutate(pheno_nested = interaction(pheno, pheno_cat, sep = "__"))

cent_plt_uncor <- ggplot(cent_plt_df) +
  geom_col(aes(x = pheno_nested, y = case.control_d, fill = model), position = "dodge") +
  guides(x = legendry::guide_axis_nested(
  key = key_range_auto(sep = "__"),
  levels_text = list(
    element_blank(),        # inner (pheno) - hide
    element_text(size = 8, angle = 45, vjust = 0.5)  # outer (pheno_cat) - show
  )
)) +
  theme_bw() +
  theme(
    axis.ticks.x    = element_blank(),
    legend.position = "right",
    axis.title.x = element_blank()
  ) +
  scale_fill_discrete(name = "Sex-Effect", labels = c("Age-Varying", "Intercept Only")) +
  facet_grid(dx~cv_sample, scales = "free_y") +
  labs(y = "Cohen's D")
ggsave("figs/supplement/cent_test_col_uncorrected_NEW.png", cent_plt_uncor, dpi = 600, width = 14, height=7)

cent_plt_resi_uncor <- ggplot(cent_plt_df) +
  geom_col(aes(x = pheno_nested, y = case.control_resi, fill = model), position = "dodge") +
  guides(x = legendry::guide_axis_nested(
  key = key_range_auto(sep = "__"),
  levels_text = list(
    element_blank(),        # inner (pheno) - hide
    element_text(size = 8, angle = 45, vjust = 0.5)  # outer (pheno_cat) - show
  )
)) +
  theme_bw() +
  theme(
    axis.ticks.x    = element_blank(),
    legend.position = "right",
    axis.title.x = element_blank()
  ) +
  scale_fill_discrete(name = "Sex-Effect", labels = c("Age-Varying", "Intercept Only")) +
  facet_grid(dx~cv_sample, scales = "free_y") +
  labs(y = "RESI")
ggsave("figs/misc/cent_test_col_resi_uncorrected_NEW.png", cent_plt_resi_uncor, dpi = 600, width = 14, height=7)
```

```{r}
dx_levels <- unique(cent_test.df_uncor$dx)

for (dx_name in dx_levels) {

  # Filter and rename
  dx.df_uncor <- cent_test.df_uncor %>%
    dplyr::filter(dx == dx_name) %>%
    dplyr::rename(sig_fdr = dx_sig_fdr)

  # Compute symmetric range for color scale
  d_range <- range(dx.df_uncor$case.control_d, na.rm = TRUE)
  sym_range <- c(-max(abs(d_range)), max(abs(d_range)))

  # Split by model
  dx.df_full <- dx.df_uncor %>% dplyr::filter(model == "full")
  dx.df_null <- dx.df_uncor %>% dplyr::filter(model == "null")

  # Build plot components (without auto-assembling full figure)
  full_parts <- plot_cv_brain(df = dx.df_full,
                          fill_var = case.control_d,
                          fill_name = "Cohen's D",
                          fill_color = "ggthemes::Gold-Purple Diverging",
                          fill_limits = sym_range,
                          height_list = c(1.03, 4.5, .93),
                          direction = -1,
                          return_plt = FALSE)

  null_parts <- plot_cv_brain(df = dx.df_null,
                          fill_var = case.control_d,
                          fill_name = "Cohen's D",
                          fill_color = "ggthemes::Gold-Purple Diverging",
                          fill_limits = sym_range,
                          height_list = c(1.03, 4.5, .93),
                          direction = -1,
                          return_plt = FALSE)

  # Helper to assemble one brain panel (without legend) + visible title
  assemble_panel <- function(parts, title_text, heights = c(1.03, 4.5, .93)) {
    panel <- cowplot::plot_grid(
      parts[[1]], parts[[2]], parts[[3]],
      ncol = 1, nrow = 3,
      rel_heights = heights,
      align = "v", axis = "lr", greedy = TRUE
    )

    title_plot <- cowplot::ggdraw() +
      cowplot::draw_label(
        title_text,
        x = 0.5, y = 0.5, hjust = 0.5, vjust = 0.5,
        fontface = "bold", size = 13
      )

    cowplot::plot_grid(
      title_plot, panel,
      ncol = 1,
      rel_heights = c(0.12, 1)
    )
  }

  plt_full_titled <- assemble_panel(
    full_parts,
    paste(dx_name, "effects from sex-moderated model")
  )

  plt_null_titled <- assemble_panel(
    null_parts,
    paste(dx_name, "effects from sex-intercept model")
  )

  # Use one shared legend (same scale, so either legend works)
  shared_legend <- full_parts[[4]]

  # Combine both panels + single legend on right
  dx_plt <- cowplot::plot_grid(
    cowplot::plot_grid(plt_full_titled, plt_null_titled, ncol = 2),
    shared_legend,
    ncol = 2,
    rel_widths = c(6, 1)
  )

  # Save
  dx_lower <- tolower(dx_name)
  ggplot2::ggsave(
    paste0("figs/supplement/", dx_lower, "_uncorrected_NEW.png"),
    dx_plt, dpi = 600, height = 8.5, width = 13, bg = "white"
  )

  message("Saved plots for: ", dx_name)
}
```


```{r}
cent_df_uncor <- cent_df %>%
  filter(total == "FALSE" & weighted == "FALSE")

cent_df_full_long <- cent_df_uncor %>%
  select(logAge_days, cv_sample, control, dx_tested, control, sex, ends_with("std_score_full")) %>%
  pivot_longer(cols = ends_with("std_score_full"),
    names_to = "pheno",
    names_pattern = "^(.*)_std_score_full$",
    values_to = "std_score_full") %>%
  na.omit(std_score_full)

####get sig plots
y_pos <- cent_df_full_long %>%
  filter(pheno %in% global_list) %>%
  group_by(cv_sample, dx_tested, pheno) %>%
  summarise(y_pos = max(std_score_full, na.rm = TRUE) + 0.6, .groups = "drop")

pvals <- cent_test.df_uncor %>%
  filter(pheno %in% global_list) %>%
  rename(dx_tested = dx) %>%
  filter(model=="full") %>%
  mutate(dx_sig_fdr = ifelse(dx_sig_fdr==TRUE, "*", NA))

pvals <- left_join(pvals, y_pos, by = c("cv_sample", "dx_tested", "pheno"))

dx_bxplt <- ggplot(cent_df_full_long, aes(x=pheno, y=std_score_full, color=control)) +
  geom_boxplot(outlier.size = .5) +
  geom_text(
    data = pvals,
    aes(x = pheno, y = y_pos, label = dx_sig_fdr),
    inherit.aes = FALSE,
    size = 4
  ) +
  facet_grid(cv_sample ~ dx_tested) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(y="Standard Score")

ggsave("figs/supplement/dx_bxplt_NEW.png", dx_bxplt, height=7, width=9, dpi=600)
```
save just B

```{r}
dx_bxpltB <- cent_df_full_long %>%
  filter(cv_sample=="B") %>%
  ggplot(aes(x=pheno, y=std_score_full, color=control)) +
  geom_boxplot(outlier.size = .5) +
  geom_text(
    data = pvals %>% filter(cv_sample=="B"),
    aes(x = pheno, y = y_pos, label = dx_sig_fdr),
    inherit.aes = FALSE,
    size = 4
  ) +
  facet_wrap(~dx_tested, nrow=3) +
  scale_x_discrete(labels = c("total.SA" = "Total SA", 
                              "mean.CT" = "Mean CT",
                              "CBV" = "Cerebellum",
                              "CSF" = "CSF",
                              "GMV" = "Cortical GM",
                              "sGMV" = "Subcortical GM",
                              "WMV" = "WM")) +
  scale_color_discrete(labels = c("CN" = "Control", "PT" = "Patient"), name = NULL) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        axis.title.x=element_blank(),
        legend.position = "bottom") +
  labs(y="Standard Score")

saveRDS(dx_bxpltB, "figs/ggobjs/dx_bxpltB_NEW.rds")
```

brainplots across dx for just B

```{r}
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

dx_all <- cent_test.df_uncor %>%
  dplyr::filter(model == "full" & cv_sample == "B") %>%
  filter(!(pheno %in% vent_list)) %>%
  select(-cv_sample) %>%
  dplyr::rename(sig_fdr = dx_sig_fdr)

# Compute symmetric range for color scale
d_range <- range(dx_all$case.control_resi, na.rm = TRUE)
sym_range <- c(-.5, .5)

dx_brain_plt <- plot_cv_brain(df = dx_all, 
                          fill_var = case.control_resi, 
                          fill_name = "RESI", 
                          fill_color = "ggthemes::Gold-Purple Diverging", 
                          fill_limits = sym_range,
                          facet_var= "dx",
                          height_list=c(5,1.1), 
                          return_plt = TRUE, 
                          include_global = FALSE,
                          direction = -1,
                          rel_widths = c(6,.5))
ggsave("figs/tmp.png", dx_brain_plt, width=18, height=7, dpi=600, bg = "white")
saveRDS(dx_brain_plt, "figs/ggobjs/dx_brain_plt_NEW.rds")
```

###Replication
What phenos replicate across splits in the full model?
```{r}
cent_rep_df <- cent_test.df_uncor %>%
  select(dx, model, pheno, cv_sample, case.control_d, dx_sig_fdr, pheno_cat, label, case.control_resi) %>%
  mutate(d_sign = sign(case.control_d)) %>%
  group_by(dx, model, pheno_cat, pheno, label) %>%
  summarise(sig_fdr = ifelse(sum(dx_sig_fdr==TRUE) < 2, FALSE, TRUE),
            same_sign = ifelse(sum(dx_sig_fdr==TRUE) < 2, NA, n_distinct(d_sign) == 1),
            mean_d = mean(case.control_d),
            mean_resi = mean(case.control_resi)) %>%
  mutate(#mean_d = ifelse(same_sign==TRUE, mean_d, NA),
         #mean_resi = ifelse(same_sign==TRUE, mean_resi, NA),
         sig_fdr = ifelse(same_sign==TRUE & sig_fdr==TRUE, TRUE, FALSE))

cent_rep_df %>%
  group_by(dx, model, same_sign) %>%
  filter(!is.na(same_sign)) %>%
  summarise(n=n())

#1 pheno in ADHD is sig in opposite directions
cent_rep_df %>%
  filter(same_sign==FALSE)
```

less agreement in full model vs null model for ADHD and 1 pheno in opposite directions in both models; 1 pheno agree in GAD full but 0 in GAD null; slightly less agreement for other phenos in full model vs null

plot

```{r}
cent_rep_df_full <- cent_rep_df %>%
  filter(model=="full")
range(cent_rep_df_full$mean_resi, na.rm = TRUE)
sym_range <- c(-.5, .5)

dx_rep_brain_plt <- plot_cv_brain(df = cent_rep_df_full, 
                          fill_var = mean_resi, 
                          fill_name = "RESI", 
                          fill_color = "ggthemes::Gold-Purple Diverging", 
                          fill_limits = sym_range,
                          facet_var= "dx",
                          return_plt = TRUE, 
                          include_global = TRUE,
                          direction = -1,
                          rel_widths = c(6,.5))
ggsave("figs/supplement/dx_casecontrol_rep_NEW.png", dx_rep_brain_plt, width=18, height=8.5, dpi=600, bg = "white")
saveRDS(dx_rep_brain_plt, "figs/ggobjs/dx_casecontrol_rep_NEW.rds")
```

##Exteremeness
Are patients more or less likely to have extreme centiles?

```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test", "*_extcent_sum.csv"))

# Read all matching CSVs
ext_cent_sum <- lapply(paths, function(p) {
  fread(p) %>%
    mutate(cv_sample = stringr::str_extract(p, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
           dx_tested  = sub("CN_", "", dx_recode),
           dx = ifelse(grepl("^CN_", dx_recode), "CN", "PT"))
})

ext_cent_sum_df <- ext_cent_sum %>%
  reduce(full_join)

stopifnot(nrow(ext_cent_sum_df) <= n_expected*2*3*2*2) #2 models, 3 groups, 2 sexes, case v control
```

```{r}
ext_cent_means <- ext_cent_sum_df %>%
  filter(ext != "mid") %>%
  group_by(cv_sample, dx_tested, model, sex, ext, dx) %>%
  summarise(mean_pct = mean(pct, na.rm=T))

#signifigance testing
comparisons <- list(
  c("CN_full",  "PT_full"),
  c("CN_null",  "PT_null")
)

ttest_results <- ext_cent_sum_df %>%
  filter(ext != "mid") %>%
  mutate(dx_model = paste(dx, model, sep = "_")) %>%
  group_by(cv_sample, dx_tested, ext, sex) %>%
  summarise(
    map_dfr(comparisons, function(comp) {
      grp       <- cur_data()
      g1_vals   <- grp %>% filter(dx_model == comp[1]) %>% arrange(pheno) %>% pull(pct)
      g2_vals   <- grp %>% filter(dx_model == comp[2]) %>% arrange(pheno) %>% pull(pct)

      # shared phenotypes only (required for pairing)
      shared    <- intersect(
        grp %>% filter(dx_model == comp[1]) %>% pull(pheno),
        grp %>% filter(dx_model == comp[2]) %>% pull(pheno)
      )
      g1 <- grp %>% filter(dx_model == comp[1], pheno %in% shared) %>% arrange(pheno) %>% pull(pct)
      g2 <- grp %>% filter(dx_model == comp[2], pheno %in% shared) %>% arrange(pheno) %>% pull(pct)

      if (length(g1) < 2) {
        tibble(comparison = paste(comp, collapse = " vs "), p_value = NA_real_,
               statistic = NA_real_, mean_diff = NA_real_, n_pairs = length(g1))
      } else {
        res <- t.test(g1, g2, paired = FALSE)
        tibble(
          comparison = paste(comp, collapse = " vs "),
          p_value    = res$p.value,
          statistic  = as.numeric(res$statistic),
          mean_diff  = as.numeric(res$estimate),   # mean of (g1 - g2)
          n_pairs    = length(g1)
        )
      }
    }),
    .groups = "drop"
  ) %>%
  # FDR correction within cv_sample
  group_by(cv_sample) %>%
  mutate(
    p_fdr = p.adjust(p_value, method = "fdr"),
    sig   = case_when(
      p_fdr < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )


# ── 2. Build significance bracket positions ───────────────────────────────────
# Map each comparison to (left dx_model, right dx_model) x-positions within
# each facet. The boxplot uses position_dodge, so we need nudged x positions.

# dx_model levels that actually appear (adjust if your data differs)
dm_levels <- c("CN_full", "CN_null", "PT_full", "PT_null")

# x-position of each dx_model within a given dx_tested group
# ggplot dodge width default = 0.75; 4 groups → spacing ≈ 0.75/4 * index
dodge_positions <- function(levels, width = 0.75) {
  n   <- length(levels)
  seq(-(n - 1) / 2, (n - 1) / 2, length.out = n) * (width / n) * (n / (n - 1))
}

dm_offsets <- setNames(dodge_positions(dm_levels), dm_levels)

# dx_tested gets integer x positions (1, 2, …) in ggplot
dx_tested_levels <- unique(ext_cent_sum_df$dx_tested)
dx_x             <- setNames(seq_along(dx_tested_levels), dx_tested_levels)

comp_map <- tibble(
  comparison = c("PT_full vs PT_null", "CN_full vs PT_full", "CN_null vs PT_null"),
  left_dm    = c("PT_full",            "CN_full",            "CN_null"),
  right_dm   = c("PT_null",            "PT_full",            "PT_null"),
  bracket_y_offset = c(0, 3, 6)   # stack brackets vertically
)

sig_plot_df <- ttest_results %>%
  left_join(comp_map, by = "comparison") %>%
  mutate(
    x_base  = dx_x[dx_tested],
    x_left  = x_base + dm_offsets[left_dm],
    x_right = x_base + dm_offsets[right_dm]
  )


# ── 3. Plot ───────────────────────────────────────────────────────────────────

plot_df <- ext_cent_sum_df %>%
  filter(ext != "mid") %>%
  mutate(dx_model = paste(dx, model, sep = "_"))

# compute a per-facet y-max to anchor brackets just above the boxes
facet_ymax <- plot_df %>%
  group_by(cv_sample, sex, dx_tested, ext) %>%
  summarise(y_max = max(pct, na.rm = TRUE), .groups = "drop")

sig_plot_df <- sig_plot_df %>%
  left_join(facet_ymax, by = c("cv_sample", "sex", "dx_tested", "ext")) %>%
  mutate(
    y_bracket = y_max * 1.05 + bracket_y_offset,
    y_label   = y_bracket + 0.5
  )

ggplot(plot_df) +
  geom_hline(yintercept = 5, color = "darkgray", linetype = "dashed") +
  geom_boxplot(aes(x = dx_tested, y = pct, fill = dx_model), position = "dodge") +

  # bracket segments: left arm
  geom_segment(
    data = sig_plot_df,
    aes(x = x_left, xend = x_left, y = y_bracket - 0.3, yend = y_bracket),
    inherit.aes = FALSE, linewidth = 0.35
  ) +
  # bracket segments: right arm
  geom_segment(
    data = sig_plot_df,
    aes(x = x_right, xend = x_right, y = y_bracket - 0.3, yend = y_bracket),
    inherit.aes = FALSE, linewidth = 0.35
  ) +
  # bracket segments: horizontal bar
  geom_segment(
    data = sig_plot_df,
    aes(x = x_left, xend = x_right, y = y_bracket, yend = y_bracket),
    inherit.aes = FALSE, linewidth = 0.35
  ) +
  # significance label
  geom_text(
    data = sig_plot_df,
    aes(x = (x_left + x_right) / 2, y = y_label, label = sig),
    inherit.aes = FALSE, size = 3
  ) +
  ggh4x::facet_nested(ext ~ cv_sample + sex, scales = "free_y") +
  theme_linedraw() +
  scale_fill_manual(values = c("CN_full" = "#F9E011FF", "CN_null" ="#D9B523FF", "PT_full" ="#DDDAEFFF", "PT_null"="#796D9DFF"),
  name = "Group"
)

```
summarising across sex

```{r}
ext_cent_sum_df2 <- ext_cent_sum_df %>%
  filter(ext != "mid") %>%
  group_by(cv_sample, dx_tested, model, ext, dx, pheno) %>%
  summarise(n=sum(n),
            n_total=sum(n_total)) %>%
  mutate(pct = (n/n_total)*100)

ttest_results2 <- ext_cent_sum_df2 %>%
  mutate(dx_model = paste(dx, model, sep = "_")) %>%
  group_by(cv_sample, dx_tested, ext) %>%
  summarise(
    map_dfr(comparisons, function(comp) {
      grp       <- cur_data()
      g1_vals   <- grp %>% filter(dx_model == comp[1]) %>% arrange(pheno) %>% pull(pct)
      g2_vals   <- grp %>% filter(dx_model == comp[2]) %>% arrange(pheno) %>% pull(pct)

      # shared phenotypes only (required for pairing)
      shared    <- intersect(
        grp %>% filter(dx_model == comp[1]) %>% pull(pheno),
        grp %>% filter(dx_model == comp[2]) %>% pull(pheno)
      )
      g1 <- grp %>% filter(dx_model == comp[1], pheno %in% shared) %>% arrange(pheno) %>% pull(pct)
      g2 <- grp %>% filter(dx_model == comp[2], pheno %in% shared) %>% arrange(pheno) %>% pull(pct)

      if (length(g1) < 2) {
        tibble(comparison = paste(comp, collapse = " vs "), p_value = NA_real_,
               statistic = NA_real_, mean_diff = NA_real_, n_pairs = length(g1))
      } else {
        res <- t.test(g1, g2, paired = TRUE)
        tibble(
          comparison = paste(comp, collapse = " vs "),
          p_value    = res$p.value,
          statistic  = as.numeric(res$statistic),
          mean_diff  = as.numeric(res$estimate),   # mean of (g1 - g2)
          n_pairs    = length(g1)
        )
      }
    }),
    .groups = "drop"
  ) %>%
  # FDR correction within cv_sample
  group_by(cv_sample) %>%
  mutate(
    p_fdr = p.adjust(p_value, method = "fdr"),
    sig   = case_when(
      p_fdr < 0.05  ~ "*",
      TRUE          ~ "ns"
    )
  )
```

Now look pheno-wise
```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test", "*_extcent_fisher.csv"))

# Read all matching CSVs
ext_cent_fish <- lapply(paths, fread)

ext_cent_fish_df <- ext_cent_fish %>%
  reduce(full_join) %>%
  group_by(cv_sample, model) %>%
  mutate(
    p_fdr = p.adjust(
      fisher_p,
      method = "fdr"
    )) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p_fdr <=0.05, TRUE, FALSE),
         pct_CN = (n_CN/n_total_CN)*100,
         pct_PT = (n_PT/n_total_PT)*100,
         pheno_cat = factor(case_when(
    pheno %in% global_list ~ "Global Vol",
    pheno %in% vol_list ~ "Regional Vol",
    pheno %in% sub_list ~ "Subcortical Vol",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_))) %>%
  mutate(label = case_when(
    pheno %in% sub_list ~ gsub("^SUBC\\.", "", pheno),
    TRUE ~ gsub("\\.(SA|GM|CT)\\.", "_", pheno))) %>%
  mutate(label = gsub("\\.", "-", label)) %>%
  mutate(label = ifelse(label == "3rd-Ventricle", "x3rd-ventricle", label)) %>%
  mutate(label = ifelse(label == "4th-Ventricle", "x4th-ventricle", label)) %>%
  mutate(label = case_when(
    pheno == "sGMV" ~ "sGM",
    pheno == "GMV" ~ "cGM",
    pheno == "WMV" ~ "WM",
    pheno == "CSF" ~ "CSF",
    pheno == "CBV" ~ "Cerebellum",
    TRUE ~ label
  )) %>%
  mutate(pct_diff = pct_PT - pct_CN) #pos -> cases overrepresented, neg -> cases under-represented

table(ext_cent_fish_df$sig_fdr, ext_cent_fish_df$dx_tested)
```

```{r}
ext_cent_fish_df %>%
  filter(sig_fdr==TRUE) %>%
  ggplot() +
  geom_bar(aes(x=dx_tested, fill=model), position="dodge") +
  facet_grid(ext~cv_sample)

ext_cent_fish_df %>%
  filter(sig_fdr==TRUE) %>%
  ggplot() +
  geom_histogram(aes(x=pct_diff, fill=dx_tested), position="identity", alpha=.5) +
  facet_grid(ext~cv_sample)
```

plot on brains

```{r}
ext_cent_fish_df <- ext_cent_fish_df %>%
  mutate(samp_ext = paste(cv_sample, ext, sep="_"))
dx_levels <- unique(ext_cent_fish_df$dx_tested)

for (dx_name in dx_levels) {

  # Filter and rename
  dx.df_uncor <- ext_cent_fish_df %>%
    dplyr::filter(dx_tested == dx_name)

  # Compute symmetric range for color scale
  d_range <- range(dx.df_uncor$pct_diff, na.rm = TRUE)
  sym_range <- c(-max(abs(d_range)), max(abs(d_range)))

  # Split by model
  dx.df_full <- dx.df_uncor %>% dplyr::filter(model == "full")
  dx.df_null <- dx.df_uncor %>% dplyr::filter(model == "null")

  # Build plot components (without auto-assembling full figure)
  full_parts <- plot_cv_brain(df = dx.df_full,
                          fill_var = pct_diff,
                          fill_name = "Percent\nDifference",
                          fill_color = "grDevices::Blue-Red 3",
                          fill_limits = sym_range,
                          height_list = c(1.03, 4.5, .93),
                          facet_var = "samp_ext",
                          return_plt = FALSE)

  null_parts <- plot_cv_brain(df = dx.df_null,
                          fill_var = pct_diff,
                          fill_name = "Percent\nDifference",
                          fill_color = "grDevices::Blue-Red 3",
                          fill_limits = sym_range,
                          height_list = c(1.03, 4.5, .93),
                          facet_var = "samp_ext",
                          return_plt = FALSE
  )

  # Helper to assemble one brain panel (without legend) + visible title
  assemble_panel <- function(parts, title_text, heights = c(1.03, 4.5, .93)) {
    panel <- cowplot::plot_grid(
      parts[[1]], parts[[2]], parts[[3]],
      ncol = 1, nrow = 3,
      rel_heights = heights,
      align = "v", axis = "lr", greedy = TRUE
    )

    title_plot <- cowplot::ggdraw() +
      cowplot::draw_label(
        title_text,
        x = 0.5, y = 0.5, hjust = 0.5, vjust = 0.5,
        fontface = "bold", size = 13
      )

    cowplot::plot_grid(
      title_plot, panel,
      ncol = 1,
      rel_heights = c(0.12, 1)
    )
  }

  plt_full_titled <- assemble_panel(
    full_parts,
    paste(dx_name, "effects from sex-moderated model")
  )

  plt_null_titled <- assemble_panel(
    null_parts,
    paste(dx_name, "effects from sex-intercept model")
  )

  # Use one shared legend (same scale, so either legend works)
  shared_legend <- full_parts[[4]]

  # Combine both panels + single legend on right
  dx_plt <- cowplot::plot_grid(
    cowplot::plot_grid(plt_full_titled, plt_null_titled, ncol = 2),
    shared_legend,
    ncol = 2,
    rel_widths = c(12, 1)
  )

  # Save
  dx_lower <- tolower(dx_name)
  ggplot2::ggsave(
    paste0("figs/supplement/", dx_lower, "_ext_uncorrected_NEW.png"),
    dx_plt, dpi = 600, height = 8.5, width = 22, bg = "white"
  )

  message("Saved plots for: ", dx_name)
  
}
```

brainplots across dx for just B

```{r}
#visualizing ventricles separately
vent_list <- c("CSF", "SUBC.Left.Lateral.Ventricle", "SUBC.Right.Lateral.Ventricle", "SUBC.3rd.Ventricle", "SUBC.4th.Ventricle")

dx_all <- ext_cent_fish_df %>%
  dplyr::filter(model == "full" & cv_sample == "B") %>%
  filter(!(pheno %in% vent_list)) %>%
  select(-cv_sample) %>%
  mutate(dx_ext = paste(dx_tested, ext, sep="_"))

# Compute symmetric range for color scale
d_range <- range(dx_all$pct_diff, na.rm = TRUE)
sym_range <- c(-32, 32)

dx_high <- dx_all %>% filter(ext=="high")
dx_low <- dx_all %>% filter(ext=="low")

high_plt <- plot_cv_brain(df = dx_high, 
                          fill_var = pct_diff, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE, 
                          include_global = FALSE,
                          rel_widths = c(6,.5))
low_plt <- plot_cv_brain(df = dx_low, 
                          fill_var = pct_diff, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          height_list=c(5,1.1), 
                          return_plt = FALSE, 
                          include_global = FALSE,
                          rel_widths = c(6,.5))
saveRDS(high_plt, "figs/ggobjs/dx_exthigh_brain_plt_NEW.rds")
saveRDS(low_plt, "figs/ggobjs/dx_extlow_brain_plt_NEW.rds")
```

###Replication
What phenos replicate across splits in the full model?
```{r}
ext_rep_df <- ext_cent_fish_df %>%
  select(dx_tested, samp_ext, model, pheno, sig_fdr, pheno_cat, label, pct_diff) %>%
  rename(sig_fdr_single = sig_fdr) %>%
  separate(samp_ext, into = c("cv_sample", "ext"), sep = "_") %>%
  mutate(pct_sign = sign(pct_diff)) %>%
  group_by(dx_tested, model, ext, pheno_cat, pheno, label) %>%
  summarise(sig_fdr = ifelse(sum(sig_fdr_single==TRUE) < 2, FALSE, TRUE),
            same_sign = ifelse(sum(sig_fdr_single==TRUE) < 2, NA, n_distinct(pct_sign) == 1),
            mean_pct = mean(pct_diff)) %>%
  mutate(sig_fdr = ifelse(same_sign==TRUE & sig_fdr==TRUE, TRUE, FALSE)
        )

ext_rep_df %>%
  group_by(dx_tested, model, same_sign, ext) %>%
  filter(!is.na(same_sign)) %>%
  summarise(n=n())
```

all significant in same direction

plot

```{r}
ext_rep_full <- ext_rep_df %>%
  filter(model=="full") %>%
  mutate(dx_ext = paste(dx_tested, ext, sep="_"))

range(ext_rep_full$mean_pct, na.rm = TRUE)
sym_range <- c(-33, 33)

#split into high and low separately
dx_high <- ext_rep_full %>% filter(ext=="high")
dx_low <- ext_rep_full %>% filter(ext=="low")

high_plt <- plot_cv_brain(df = dx_high, 
                          fill_var = mean_pct, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          return_plt = FALSE, 
                          rel_widths = c(6,.5))
low_plt <- plot_cv_brain(df = dx_low, 
                          fill_var = mean_pct, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl", #scico::romaO
                          fill_limits = sym_range,
                          facet_var = "dx_tested",
                          return_plt = FALSE, 
                          rel_widths = c(6,.5))
saveRDS(high_plt, "figs/ggobjs/dx_exthigh_rep_NEW.rds")
saveRDS(low_plt, "figs/ggobjs/dx_extlow_rept_NEW.rds")

#also just quick visual for now
ext_rep_brain_plt <- plot_cv_brain(df = ext_rep_full, 
                          fill_var = mean_pct, 
                          fill_name = "Percent\nDifference",
                          fill_color = "pals::ocean.curl",
                          fill_limits = sym_range,
                          facet_var= "dx_ext",
                          return_plt = TRUE, 
                          include_global = TRUE,
                          direction = -1,
                          rel_widths = c(10,.5))
ggsave("figs/supplement/dx_ext_rep_NEW.png", ext_rep_brain_plt, width=30, height=8.5, dpi=600, bg = "white")
```


###Change
how many subjects change between full and null models?

```{r}
# Use Sys.glob to expand wildcard paths
paths <- Sys.glob(file.path("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts/cv_sample_?_test", "*_extcent_diffs.csv"))

# Read all matching CSVs
ext_cent_diff <- lapply(paths, function(p) {
  fread(p) %>%
    mutate(cv_sample = stringr::str_extract(p, "(?<=cv_sample_)[A-Za-z]+(?=_test)"),
           dx_tested  = sub("CN_", "", dx_recode),
           dx = ifelse(grepl("^CN_", dx_recode), "CN", "PT"))
})

ext_cent_diff_df <- ext_cent_diff %>%
  reduce(full_join)

ext_cent_diff_df %>%
  filter(dx=="PT") %>%
  select(dx_recode, sex, n_total) %>%
  distinct()
```

```{r}
ext_cent_diff_df %>%
  filter(dx=="PT") %>%
  ggplot() +
  geom_col(aes(x=dx_tested, y=mean_phenos, fill=sex), position="dodge") +
  facet_wrap(~cv_sample)

ext_cent_diff_df %>%
  filter(dx=="PT") %>%
  summarise(min_pct = min(pct_change),
            max_pct = max(pct_change))

ext_cent_diff_df %>%
  filter(dx=="PT") %>%
  ggplot() +
  geom_col(aes(x=dx_tested, y=pct_change, fill=sex), position="dodge") +
  facet_wrap(~cv_sample)

ext_cent_diff_df %>%
  filter(dx=="PT") %>%
  ggplot() +
  geom_col(aes(x=dx_tested, y=pct_change, fill=sex), position="dodge") +
  facet_wrap(~cv_sample)
```

test sex-differences in prop sample with changes
```{r}
run_fisher <- function(n_change_F, n_change_M, n_total_F, n_total_M) {
  twoSamplePermutationTestProportion(
    x = c(n_change_F, n_change_M),
    y = c(n_total_F, n_total_M),
    x.and.y = "Number Successes and Trials",
    alternative = "two.sided"
  )
}

fisher_df <- ext_cent_diff_df %>%
  filter(dx=="PT") %>%
  pivot_wider(id_cols = c(cv_sample, dx_tested), values_from=c(n_change, n_total), names_from=sex)

  #run stats
fisher_results <- pmap(
  list(fisher_df$n_change_F, fisher_df$n_change_M, fisher_df$n_total_F, fisher_df$n_total_M),
  run_fisher
)

fisher_final <- fisher_df %>%
  ungroup() %>%
  mutate(
    fisher_stat = map_dbl(fisher_results, ~ .x$statistic),
    fisher_p    = map_dbl(fisher_results, ~ .x$p.value)) %>%
  rowwise() %>%
  mutate(
    ci_lower_F = binom.confint(n_change_F, n_total_F, method="exact")$lower, #also getting conf int estimates
    ci_upper_F = binom.confint(n_change_F, n_total_F, method="exact")$upper,
    ci_lower_M = binom.confint(n_change_M, n_total_M, method="exact")$lower,
    ci_upper_M = binom.confint(n_change_M, n_total_M, method="exact")$upper
  ) %>%
  group_by(cv_sample) %>%
  mutate(
    p_fdr = p.adjust(
      fisher_p,
      method = "fdr"
    )) %>%
  ungroup() %>%
  mutate(sig_fdr = ifelse(p_fdr <= 0.05, TRUE, FALSE))

table(fisher_final$dx_tested, fisher_final$sig_fdr)
```
replot

```{r}
# pivot fisher_final CIs and p-values to long format for joining
ci_long <- fisher_final %>%
  select(cv_sample, dx_tested, ci_lower_F, ci_upper_F, ci_lower_M, ci_upper_M, p_fdr, sig_fdr) %>%
  pivot_longer(
    cols = c(ci_lower_F, ci_upper_F, ci_lower_M, ci_upper_M),
    names_to = c(".value", "sex"),
    names_pattern = "(ci_lower|ci_upper)_(F|M)"
  ) %>%
  mutate(sig_label = case_when(
    p_fdr < 0.001 ~ "***",
    p_fdr < 0.01  ~ "**",
    p_fdr < 0.05  ~ "*",
    TRUE ~ ""
  ))

plot_df <- ext_cent_diff_df %>%
  filter(dx == "PT") %>%
  left_join(ci_long, by = c("cv_sample", "dx_tested", "sex"))

# separate df to position brackets just above the highest error bar per facet/x group
bracket_df <- plot_df %>%
  filter(cv_sample=="B") %>%
  group_by(dx_tested) %>%
  summarise(
    bracket_y = max(ci_upper) * 1.05,  # just above highest error bar
    sig_label = first(sig_label),
    .groups = "drop"
  )

change_barplt <- plot_df %>%
  filter(cv_sample=="B") %>%
  ggplot(aes(x = dx_tested, y = pct_change/100, fill = sex)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    position = position_dodge(width = 0.9),
    width = 0.25
  ) +
  geom_text(
    data = bracket_df,
    aes(x = dx_tested, y = bracket_y, label = sig_label),
    inherit.aes = FALSE,
    size = 5
  ) +
  theme_linedraw() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(
    values = c('F' = "#c05f0d", 'M' = "#38688D"),
    breaks = c("F", "M"),
    labels = c("Female", "Male"),
    name = "Sex") +
  labs(y="Discordant Patients",
       x="Diagnosis")
saveRDS(change_barplt, "figs/ggobjs/change_barpltB_NEW.rds")
```
