---
title: "Spline Testing"
output: html_document
date: "2024-12-18"
---

Playing around with different spline configurations to see how they effect worm plots. Sticking to sex-stratified models in unharmonized global phenos (using study_site as fixed effect)

```{r warning=FALSE}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, stats, skimr, gamlss, gamlssTools, gamlss.add, mgcv)

source("code/gamlss_fit_funs.R")
```

look into csf
```{r}
df_A <- fread("data/cv_sample_A.csv", stringsAsFactors=TRUE)
range(df_A$GMV, na.rm=TRUE)
range(df_A$CSF, na.rm=TRUE)
log(1000, base=10)
```

load data
```{r}
df_B <- fread("data/cv_sample_B.csv", stringsAsFactors=TRUE)

df_WMV <- df_B %>%
  dplyr::select(all_of(c("WMV", "fs_version_GM", "logAge_days", "sexMale", "study_site", "sexMale_x_logAge", "age_days", "sexMale_x_age", "age_days"))) %>%
  na.omit() %>%
  trunc_coverage("logAge_days")

ggplot(df_WMV) +
  geom_histogram(aes(x=logAge_days))

ggplot(df_WMV) +
  geom_histogram(aes(x=age_days/365.25))
```


attempting frankensample - removing adults so distribution is less skewed by age
```{r}
ggplot(df_B) +
  geom_histogram(aes(x=logAge_days))

ggplot(df_B) +
  geom_histogram(aes(x=age_days/365.25))
```


```{r}
# Example data
set.seed(123)
df_full <- fread("data/v3_CN_cleaned.csv", stringsAsFactors=TRUE)

# Bin ages into 10 equal-width intervals
df_full$age_bin <- cut(df_full$logAge_days, breaks = 10)

ggplot(df_full) +
  geom_histogram(aes(x=logAge_days, fill=age_bin))
```

sample up to 1000 from each age bin

```{r}
balanced_df <- df_full %>%
  group_by(age_bin) %>%
  slice_sample(n = 500, replace = FALSE) %>%
  ungroup()

ggplot(balanced_df) +
  geom_histogram(aes(x=logAge_days, fill=age_bin))
```
```{r}
df <- df_B %>%
  dplyr::select("sGMV", "fs_version_GM", "logAge_days", "sexMale", "sexMale_x_logAge", "study_site") %>%
  na.omit()

df_test <- df %>%
  mutate(sGMV = sGMV + 1000) #try adding 1000 to keep from going negative

# linear_mod <- gamlss(sGMV ~ 0 + logAge_days + sexMale,
#     family="NO", data=df,
#     control=gamlss.control(n.cyc = 200))

# linear_mod <- lm(sGMV ~ 0 + logAge_days + sexMale + random(study_site), data=df)
# 
# ggplot(df, aes(x=logAge_days, y=sGMV, color=as.factor(sexMale)))+
#   geom_point() +
#   geom_line(aes(y = predict(linear_mod)))  # fitted line

#make_centile_fan(linear_mod, df, x_var="logAge_days", color_var="sexMale")

test_mod <- gamlss(sGMV ~ pb(sexMale_x_logAge, method = "GAIC", k = log(nrow(df)), control = pb.control(order = 3)) + 
    pb(logAge_days, method = "GAIC", k = log(nrow(df)), control = pb.control(order = 3)) + 
    sexMale + random(study_site),
    sigma.formula= ~pb(sexMale_x_logAge, method = "GAIC", k = log(nrow(df)), control = pb.control(order = 3)) + 
    pb(logAge_days, method = "GAIC", k = log(nrow(df)), control = pb.control(order = 3)) + 
    sexMale + random(study_site), 
    nu.formula=~logAge_days,
    family="BCCG", data=df_test
    #control=gamlss.control(n.cyc = 200, mu.step = 0.000000001, sigma.step= 0.0000000001, nu.step=0.00000000001),
    #start.from=linear_mod
    )

make_centile_fan(test_mod, df, x_var="logAge_days", color_var="sexMale")

```

```{r}
df2 <- fread("data/cv_sample_A_dfs/sGMV_totalFALSE_logPhenoFALSE_logAgeTRUE.csv", stringsAsFactors=TRUE)

#make_centile_fan(sGMV_none_int_mod, df2, x_var="logAge_days", color_var="sexMale")

df2_test <- df2 %>% mutate(sGMV = sGMV+1000)

test_mod2 <- gamlss(sGMV ~ ga(~s(sexMale_x_logAge, k=20, bs="ad") + 
                                s(logAge_days, k=20, bs="ad") + 
                                sexMale + random(study_site), method = 'ML'),
                    sigma.formula= ~ga(~s(sexMale_x_logAge, k=20, bs="ad") + 
                                s(logAge_days, k=20, bs="ad") + 
                                sexMale + random(study_site), method = 'ML'), 
    nu.formula=~logAge_days,
    family="BCCG", 
    data=df2_test,
    control=gamlss.control(n.cyc = 200)
    # start.from=linear_mod
    )
```

can also play around with order (m), method, etc

```{r}
df3 <- fread("data/cv_sample_A_dfs/GMV_totalFALSE_logPhenoTRUE_logAgeTRUE.csv", stringsAsFactors=TRUE, na.strings = "")
gmv_mod1 <- readRDS("cv_sample_A_train/global_vols_logPhenoTRUE_logAgeTRUE_pbmods/model_objs/GMV_both_pbage_BestMod.rds")
gmv_mod1$call$family <- gmv_mod1$family

make_centile_fan(gmv_mod1, df3, x_var="logAge_days", color_var="sexMale", desiredCentiles = c(0.5, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95),
                 special_term = "sexMale_x_logAge = sexMale * logAge_days",
                 remove_cent_effect="study_site",
                 remove_point_effect = "study_site")

unscale <- function(x){10^x - 5}

make_centile_fan(gmv_mod1, df3, x_var="logAge_days", color_var="sexMale", desiredCentiles = c(0.5, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95),
                 special_term = "sexMale_x_logAge = sexMale * logAge_days",
                 remove_cent_effect="study_site",
                 remove_point_effect = "study_site",
                 y_scale=unscale)
```

un-log pheno

```{r}


```
