---
title: "sensitivity_df_prep"
output: html_document
date: "2025-10-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Prep split-half dataframes for sensitivity analyses with stricter scan QC

```{r}
setwd("/mnt/isilon/bgdlab_processing/Margaret/sex_mod_braincharts")
```

```{r, warning=FALSE}}
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, dplyr)
```

load data cleaned data
```{r}
df_clean <- fread("./data/v3_CN_cleaned.csv", stringsAsFactors=T)
skimr::skim(df_clean$SurfaceHoles)
```

look just at data with euler

```{r}
df_euler <- df_clean %>%
  filter(!is.na(SurfaceHoles))
range(df_euler$SurfaceHoles)
range(df_euler$age_days)

df_euler %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=SurfaceHoles))
```

playing around with re-scaling euler w/in study
```{r}
df_euler2 <- df_euler %>%
  group_by(study) %>%
  mutate(SurfaceHoles_scaled = (SurfaceHoles - min(SurfaceHoles, na.rm = TRUE)) /
                               (max(SurfaceHoles, na.rm = TRUE) - min(SurfaceHoles, na.rm = TRUE))) %>%
  ungroup()

df_euler2 %>%
  ggplot() +
  geom_boxplot(aes(x=study, y=SurfaceHoles_scaled))

df_euler2 %>%
  ggplot() +
  geom_point(aes(x=logAge_days, y=SurfaceHoles_scaled))
df_euler2 %>%
  ggplot() +
  geom_point(aes(x=logAge_days, y=SurfaceHoles))
```

using weight-deriving formula from Kat's analyses
```{r}
df_euler2$weight <- (1 - df_euler2$SurfaceHoles_scaled) * (nrow(df_euler2) / sum(1 - df_euler2$SurfaceHoles_scaled))
sum(df_euler2$weight)
nrow(df_euler2)
stopifnot(all.equal(sum(df_euler2$weight), nrow(df_euler2)))

df_euler2 %>%
  ggplot() +
  geom_point(aes(x=weight, y=SurfaceHoles_scaled))
```

split data for cv modeling using code from `lbcc_eda.Rmd`

```{r}
valid_sites <- df_euler2 %>%
  mutate(sex=as.factor(sexMale)) %>%
  group_by(study_site, sex) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(study_site) %>%
  summarise(
    valid = all(c("0", "1") %in% sex) && all(n[match(c("0", "1"), sex, nomatch = 0)] >= 10),
    .groups = "drop"
  ) %>%
  filter(valid) %>%
  pull(study_site)

df_euler_sites <- df_euler2 %>%
  filter(study_site %in% valid_sites)
nrow(df_euler_sites)
nrow(df_euler2) - nrow(df_euler_sites)

df_euler_sites %>%
  ggplot() +
  geom_point(aes(x=logAge_days, y=SurfaceHoles_scaled))
```

divide
```{r}
set.seed(5678)
#add sex factor var
df_euler_sites2 <- df_euler_sites %>% mutate(sex=as.factor(sexMale))

split_samples <- splitstackshape::stratified(df_euler_sites2, group=c("study_site", "sex"), size=0.5, bothSets = TRUE)

samp_A <- split_samples$SAMP1
samp_B <- split_samples$SAMP2

ggplot(samp_A) +
  geom_histogram(aes(x=age_days, fill=sex))
ggplot(samp_B) +
  geom_histogram(aes(x=age_days, fill=sex))

samp_A$sample <- "A"
samp_B$sample <- "B"
full_df <- rbind(samp_A, samp_B)

ggplot(full_df)+
  geom_violin(aes(x=sample, y=age_days, fill=as.factor(sexMale))) +
  theme_linedraw() +
  labs(fill = "Male")

ggplot(full_df)+
  geom_histogram(aes(x=weight, fill=as.factor(sample)), position="identity", alpha=0.5)

ggplot(full_df)+
  geom_bar(aes(x=study_site, fill=sample), position = "fill") +
  theme_linedraw()
```

what if we use the original splits and just add back in Euler weights?
```{r}
og_sample_A <- fread("./data/cv_sample_A.csv", stringsAsFactors=T) %>% filter(!is.na(SurfaceHoles))
og_sample_B <- fread("./data/cv_sample_B.csv", stringsAsFactors=T) %>% filter(!is.na(SurfaceHoles))

og_sample_A_weight <- left_join(og_sample_A, df_euler_sites)
og_sample_B_weight <- left_join(og_sample_B, df_euler_sites)

og_sample_A_weight$sample <- "A"
og_sample_B_weight$sample <- "B"
og_splits <- rbind(og_sample_A_weight, og_sample_B_weight)
```

```{r}
ggplot(og_splits)+
  geom_violin(aes(x=sample, y=age_days, fill=as.factor(sexMale))) +
  theme_linedraw() +
  labs(fill = "Male")

ggplot(og_splits)+
  geom_histogram(aes(x=weight, fill=as.factor(sample)), position="identity", alpha=0.5)

table(og_splits$sample, og_splits$sex)
table(full_df$sample, full_df$sex)

ggplot(og_splits)+
  geom_bar(aes(x=study_site, fill=sample), position = "fill") +
  theme_linedraw()

ggplot(og_splits)+
  geom_bar(aes(x=sample, fill=study_site), position = "fill") +
  theme_linedraw() +
  theme(legend.position="none")
```

very comparable, using OG splits for consistency with original models

save
```{r}
fwrite(og_sample_A_weight, "./data/cv_sample_A_euler.csv")
fwrite(og_sample_B_weight, "./data/cv_sample_B_euler.csv")
```